### 2024-11-15 proposal_EDA/deGuzman_IDS702Projectproposal_template.qmd
update proposal templatre

---
title: "Project Proposal"
author: "Sizhe Chen, Naikyah Dhariwala, Peter de Guzman, Jennifer Li"
subtitle: "Due November 17 at 11:59pm"
format: pdf
---

#### Load Packages

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse)
```

# Dataset 1 (top choice)

**Data source:** The 2020 Collaborative Multi-racial Post-Election Survey (CMPS)

**Brief description:** The survey has an unweighted sample size of 17,545 respondents. Each row represents a unique respondent. The data is collected through an online survey with 800 unique questions. The survey's main focus is on attitudes about the 2020 Election and candidates, experiences with racism, policy attitudes, immigration, and personal experiences with civic engagement across many facets of American life.

The 2020 CMPS was offered in English, Spanish, Chinese (simplified), Chinese (traditional), Korean, Vietnamese, Arabic, Urdu, Farsi and Haitian Creole. The survey dataset includes Black, White, Latino and Asian respondents as well as additional oversamples of respondents from hard-to-reach populations including, Afro-Latinos, Black immigrants, Native Americans, Native Hawaiians, Muslims and people who identify as LGBTQ.

**Research question 1:** Are Americans who believe their race is important to their identity more likely to be interested in politics than those that are not?

-   Outcome variable (include the name/description and type of variable): The outcome variable is an ordinal variable labeled Q29. The survey question is "Some people are very interested in politics while other people can't stand politics how about you? are you...". The response options are in a five-point scale from "Not at all interested in politics" to "Very interested in politics."

**Research question 2:**

-   Outcome variable (include the name/description and type of variable):

**Load the data and provide a `glimpse()`**:

```{r data-1, warning=FALSE}
rm(list=ls())
load("/Users/pdeguz01/Documents/git/Data/2020 CMPS/DS0001/39096-0001-Data.rda")
data <- da39096.0001
data %>% slice_head(n = 100) %>% glimpse()
```

**Exploratory Plots:**

```{r plots-1}
wtd_tableQ29 <- data %>% count(Q29, wt=WEIGHT)
wtd_tableQ29 <- wtd_tableQ29 %>% mutate('pct' = (n/sum(n, na.rm=TRUE)))

ggplot(wtd_tableQ29, aes(x=Q29, y = pct)) +
  geom_bar(stat="identity", fill = "orange") +
  geom_text(aes(label=percent(pct)), vjust=1.6, color="white", size =3.5) +
  labs(title = "Reported Interest in Politics", 
       x = "Level of Interest", y = "Percent (Weighted)") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

contingency_table <- table(data$Q29, data$Q271)
contingency_df <- as.data.frame(contingency_table)
contingency_df <- contingency_df %>% mutate(Share = (Freq/sum(Freq)))

ggplot(contingency_df, aes(x = Var1, y = Var2, fill = Share)) +
  geom_tile() +
  labs(title="Heatmap of Interest in Politics and Racial Identity", x = "Interest in Politics", y = "Importance of Racial Identity") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_continuous(low = "white", high = "blue", labels = percent_format(scale = 100)) 
```

# Dataset 2

**Data source:**

**Brief description:**

**Research question 1:**

-   Outcome variable (include the name/description and type of variable):

**Research question 2:**

-   Outcome variable (include the name/description and type of variable):

**Load the data and provide a `glimpse()`**:

```{r data-2, warning=FALSE}

```

**Exploratory Plots:**

```{r plots-2}

```

#

# Dataset 3 (optional)

**Data source:**

**Brief description:**

**Research question 1:**

-   Outcome variable (include the name/description and type of variable):

**Research question 2:**

-   Outcome variable (include the name/description and type of variable):

**Load the data and provide a `glimpse()`**:

```{r data-3, warning=FALSE}

```

**Exploratory Plots:**

```{r plots-3}

```

#

# Team Charter

#### When will you meet as a team to work on the project components? Will these meetings be held in person or virtually?

#### What is your group policy on missing team meetings (e.g., how much advance notice should be provided)?

#### How will your team communicate (email, Slack, text messages)? What is your policy on appropriate response time (within a certain number of hours? Nights/weekends?)?

####

### 2024-11-15 proposal_EDA/cmps_eda.R

@@ -40,8 +40,6 @@ tile_table <-  data %>% group_by(Q29, Q271) %>%
   spread(Q29, n)
 
 contingency_table <- table(data$Q29, data$Q271)
-
-print(contingency_table)
 contingency_df <- as.data.frame(contingency_table)
 contingency_df <- contingency_df %>% mutate(Share = (Freq/sum(Freq)))
 

### 2024-11-16 proposal_EDA/deGuzman_IDS702Projectproposal_template.qmd
updating proposal draft

@@
 format: pdf
+header-includes:
+      - \usepackage{float}
+      - \floatplacement{table}{H}
 ---
@@
 library(tidyverse)
+library(ggplot2)
+library(formattable)
+library(scales)
 ```
@@
 
+The main predictor variable is "Q271" which asks “How important is being [RACIAL GROUP] to your identity?”.
+The response options are a five-point scale from "Not at all important" to "Extremely important".
+Other potential predictors include survey questions on respondent's self-perception of race and discrimination, and their own attitudes towards other racial groups.
+In addition, there are multiple variables that could be used as demographic controls such as age, gender, education, income, and party identification.
+The variables not shown in the output include responses to the rest of the 800 survey questions, covering topics such as policy attitudes, civic engagement, and political participation.
+
 **Research question 2:**
@@
 
-```{r data-1, warning=FALSE}
+```{r data-1, warning=FALSE, echo=FALSE}
 rm(list=ls())
@@
 data <- da39096.0001
-data %>% slice_head(n = 100) %>% glimpse()
+reordered_data <- data %>% relocate(Q29, Q271, WEIGHT, 
+                                    AGE1_HID, AGE2_HID, .after = DATE)
+reordered_data %>% slice_head(n = 50) %>% str()
 ```
@@
 
-```{r plots-1}
+```{r plots-1, echo=FALSE}
 wtd_tableQ29 <- data %>% count(Q29, wt=WEIGHT)
@@
 
-contingency_table <- table(data$Q29, data$Q271)
+data <- data %>% mutate(Q29_shortened = case_when(Q29 == "(1) Very interested in politics" ~ "Very interested",
+                                                   Q29 == "(2) Somewhat interested" ~ "Somewhat interested",
+                                                   Q29 == "(3) Not that interested in politics" ~ "Not that interested",
+                                                   Q29 == "(4) Not at all interested in politics" ~ "Not at all interested",
+                                                   .default = "Other"))
+contingency_table <- table(data$Q29_shortened, data$Q271)
 contingency_df <- as.data.frame(contingency_table)
@@
   geom_tile() +
-  labs(title="Heatmap of Interest in Politics and Racial Identity", x = "Interest in Politics", y = "Importance of Racial Identity") +
+  labs(title="Interest in Politics and Racial Identity", x = "Interest in Politics", y = "Importance of Racial Identity") +
   theme_minimal() + 
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 0.4, vjust = 0.5)) +
   scale_fill_continuous(low = "white", high = "blue", labels = percent_format(scale = 100)) 

### 2024-11-16 proposal_EDA/deGuzman_IDS702Projectproposal_template.qmd
updating proposals

@@
 
-**Research question 2:**
+**Variable Type**: Ordinal
 
--   Outcome variable (include the name/description and type of variable):
+**Research question 2:** Controlling for other demographic variables such as age and gender,How do experiences with racism influence the importance respondents place on policy issues related to racial and ethnic discrimination?
+
+- Outcome variable (include the name/description and type of variable): Stopping discrimination against racial/ethnic minorities (Q1R5) — a binary variable indicating whether respondents view it as an important issue.
+
+**Variable Type:** Binary.
+
+**Interaction Term Idea:** Include an interaction term between race/ethnicity (S2RACER) and experiences with racism to see if the influence of racism experiences differs across racial/ethnic groups.
 
@@
 
-```{r plots-1, echo=FALSE}
-wtd_tableQ29 <- data %>% count(Q29, wt=WEIGHT)
+```{r plots-1, echo=FALSE, warning=FALSE}
+
+data <- data %>% mutate(Q29_shortened = case_when(Q29 == "(1) Very interested in politics" ~ "Very interested",
+                                                   Q29 == "(2) Somewhat interested" ~ "Somewhat interested",
+                                                   Q29 == "(3) Not that interested in politics" ~ "Not that interested",
+                                                   Q29 == "(4) Not at all interested in politics" ~ "Not at all interested",
+                                                   .default = "Other"))
+wtd_tableQ29 <- data %>% count(Q29_shortened, wt=WEIGHT)
 wtd_tableQ29 <- wtd_tableQ29 %>% mutate('pct' = (n/sum(n, na.rm=TRUE)))
-
-ggplot(wtd_tableQ29, aes(x=Q29, y = pct)) +
+ggplot(wtd_tableQ29, aes(x=Q29_shortened, y = pct)) +
   geom_bar(stat="identity", fill = "orange") +
@@
 
-data <- data %>% mutate(Q29_shortened = case_when(Q29 == "(1) Very interested in politics" ~ "Very interested",
-                                                   Q29 == "(2) Somewhat interested" ~ "Somewhat interested",
-                                                   Q29 == "(3) Not that interested in politics" ~ "Not that interested",
-                                                   Q29 == "(4) Not at all interested in politics" ~ "Not at all interested",
-                                                   .default = "Other"))
 contingency_table <- table(data$Q29_shortened, data$Q271)
@@
   scale_fill_continuous(low = "white", high = "blue", labels = percent_format(scale = 100))
+
+data$Q1R5_cleaned <- as.numeric(sub("^\\((\\d+)\\).*", "\\1", data$Q1R5))
+
+data$Q1R5_cleaned <- factor(data$Q1R5_cleaned, levels = c(0, 1), labels = c("Not Important", "Important"))
+
+# Plot the distribution of Q1R5
+ggplot(data, aes(x = Q1R5_cleaned)) +
+  geom_bar(fill = "skyblue") +
+  labs(title = "Importance of Stopping Discrimination Against Racial/Ethnic Minorities",
+       x = "Importance Level",
+       y = "Count") +
+  theme_minimal()
+
+data$MAIN_ETH_QUOTA_numeric <- as.numeric(sub("^\\((\\d+)\\).*", "\\1", data$MAIN_ETH_QUOTA))
+data$Race_Ethnicity <- factor(data$MAIN_ETH_QUOTA_numeric,
+                              levels = 1:7,
+                              labels = c("White", "Latino", "Black", "AAPI", "ANAI", "MENA", "NH"))
+
+# Plot with short-term race labels on the x-axis and percentage labels on the bars
+ggplot(data, aes(x = Race_Ethnicity, fill = Q1R5_cleaned)) +
+  geom_bar(position = "fill") +
+  labs(title = "Importance of Stopping Discrimination by Race/Ethnicity",
+       x = "Race/Ethnicity",
+       y = "Proportion",
+       fill = "Importance Level",
+       caption = "Race/Ethnicity Abbreviations: ANAI = Native American, NH = Native Hawaiian, AAPI = Asian American/Pacific Islander, MENA = Middle Eastern/North African") +
+  scale_y_continuous(labels = scales::percent) +
+  theme_minimal() +
+  geom_text(aes(label = scales::percent(after_stat(count)/sum(after_stat(count)))),
+            stat = "count", 
+            position = position_fill(vjust = 0.5),
+            color = "white")
 ```

### 2024-11-16 proposal_EDA/stat research question.Rmd
updating proposals

--- 2024-11-15 proposal_EDA/deGuzman_IDS702Projectproposal_template.qmd
@@
 ---
-title: "Project Proposal"
-author: "Sizhe Chen, Naikyah Dhariwala, Peter de Guzman, Jennifer Li"
-subtitle: "Due November 17 at 11:59pm"
-format: pdf
+title: "stat final project"
+output: pdf_document
+date: "2024-11-15"
 ---
 
-#### Load Packages
-
-```{r load-packages, message = FALSE, warning = FALSE}
-library(tidyverse)
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
 ```
 
-# Dataset 1 (top choice)
+```{r}
+library(ggplot2)
+library(prettyR)
+```
 
-**Data source:** The 2020 Collaborative Multi-racial Post-Election Survey (CMPS)
+```{r}
+load("/Users/chensi/Desktop/MIDS/Fall 2024/IDS 702/ICPSR_39096/DS0001/39096-0001-Data.rda")  
+data <- da39096.0001
+```
 
-**Brief description:** The survey has an unweighted sample size of 17,545 respondents.
-Each row represents a unique respondent.
-The data is collected through an online survey with 800 unique questions.
-The survey's main focus is on attitudes about the 2020 Election and candidates, experiences with racism, policy attitudes, immigration, and personal experiences with civic engagement across many facets of American life.
+**Research question 2:** Controlling for other demographic variables such as age and gender,How do experiences with racism influence the importance respondents place on policy issues related to racial and ethnic discrimination?
+- Outcome variable (include the name/description and type of variable): Stopping discrimination against racial/ethnic minorities (Q1R5) — a binary variable indicating whether respondents view it as an important issue.
 
-The 2020 CMPS was offered in English, Spanish, Chinese (simplified), Chinese (traditional), Korean, Vietnamese, Arabic, Urdu, Farsi and Haitian Creole.
-The survey dataset includes Black, White, Latino and Asian respondents as well as additional oversamples of respondents from hard-to-reach populations including, Afro-Latinos, Black immigrants, Native Americans, Native Hawaiians, Muslims and people who identify as LGBTQ.
+**Variable Type:** Binary.
 
-**Research question 1:** Are Americans who believe their race is important to their identity more likely to be interested in politics than those that are not?
-
--   Outcome variable (include the name/description and type of variable): The outcome variable is an ordinal variable labeled Q29.
-The survey question is "Some people are very interested in politics while other people can't stand politics how about you?
-are you...".
-The response options are in a five-point scale from "Not at all interested in politics" to "Very interested in politics."
-
-**Research question 2:**
-
--   Outcome variable (include the name/description and type of variable):
-
-**Load the data and provide a `glimpse()`**:
-
-```{r data-1, warning=FALSE}
-rm(list=ls())
-load("/Users/pdeguz01/Documents/git/Data/2020 CMPS/DS0001/39096-0001-Data.rda")
-data <- da39096.0001
-data %>% slice_head(n = 100) %>% glimpse()
-```
+**Interaction Term Idea:** Include an interaction term between race/ethnicity (S2RACER) and experiences with racism to see if the influence of racism experiences differs across racial/ethnic groups.
 
@@
 
-```{r plots-1}
-wtd_tableQ29 <- data %>% count(Q29, wt=WEIGHT)
-wtd_tableQ29 <- wtd_tableQ29 %>% mutate('pct' = (n/sum(n, na.rm=TRUE)))
+```{r}
+data$Q1R5_cleaned <- as.numeric(sub("^\\((\\d+)\\).*", "\\1", data$Q1R5))
 
-ggplot(wtd_tableQ29, aes(x=Q29, y = pct)) +
-  geom_bar(stat="identity", fill = "orange") +
-  geom_text(aes(label=percent(pct)), vjust=1.6, color="white", size =3.5) +
-  labs(title = "Reported Interest in Politics", 
-       x = "Level of Interest", y = "Percent (Weighted)") +
-  theme_minimal() +
-  theme(plot.title = element_text(hjust = 0.5))
+data$Q1R5_cleaned <- factor(data$Q1R5_cleaned, levels = c(0, 1), labels = c("Not Important", "Important"))
 
-contingency_table <- table(data$Q29, data$Q271)
-contingency_df <- as.data.frame(contingency_table)
-contingency_df <- contingency_df %>% mutate(Share = (Freq/sum(Freq)))
+# Check the cleaned variable
+table(data$Q1R5_cleaned)
 
-ggplot(contingency_df, aes(x = Var1, y = Var2, fill = Share)) +
-  geom_tile() +
-  labs(title="Heatmap of Interest in Politics and Racial Identity", x = "Interest in Politics", y = "Importance of Racial Identity") +
-  theme_minimal() + 
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  scale_fill_continuous(low = "white", high = "blue", labels = percent_format(scale = 100)) 
-```
-
-# Dataset 2
-
-**Data source:**
-
-**Brief description:**
-
-**Research question 1:**
-
--   Outcome variable (include the name/description and type of variable):
-
-**Research question 2:**
-
--   Outcome variable (include the name/description and type of variable):
-
-**Load the data and provide a `glimpse()`**:
-
-```{r data-2, warning=FALSE}
+# Plot the distribution of Q1R5
+ggplot(data, aes(x = Q1R5_cleaned)) +
+  geom_bar(fill = "skyblue") +
+  labs(title = "Importance of Stopping Discrimination Against Racial/Ethnic Minorities",
+       x = "Importance Level",
+       y = "Count") +
+  theme_minimal()
 
@@
 
-**Exploratory Plots:**
+```{r}
+data$MAIN_ETH_QUOTA_numeric <- as.numeric(sub("^\\((\\d+)\\).*", "\\1", data$MAIN_ETH_QUOTA))
+data$Race_Ethnicity <- factor(data$MAIN_ETH_QUOTA_numeric,
+                              levels = 1:7,
+                              labels = c("White", "Latino", "Black", "AAPI", "ANAI", "MENA", "NH"))
+table(data$Race_Ethnicity)
 
-```{r plots-2}
-
+# Plot with short-term race labels on the x-axis and percentage labels on the bars
+ggplot(data, aes(x = Race_Ethnicity, fill = Q1R5_cleaned)) +
+  geom_bar(position = "fill") +
+  labs(title = "Importance of Stopping Discrimination by Race/Ethnicity",
+       x = "Race/Ethnicity",
+       y = "Proportion",
+       fill = "Importance Level",
+       caption = "Race/Ethnicity Abbreviations: ANAI = Native American, NH = Native Hawaiian, AAPI = Asian American/Pacific Islander, MENA = Middle Eastern/North African") +
+  scale_y_continuous(labels = scales::percent) +
+  theme_minimal() +
+  geom_text(aes(label = scales::percent(..count../sum(..count..))),
+            stat = "count", 
+            position = position_fill(vjust = 0.5),
+            color = "white")
 ```
-
-#
-
-# Dataset 3 (optional)
-
-**Data source:**
-
-**Brief description:**
-
-**Research question 1:**
-
--   Outcome variable (include the name/description and type of variable):
-
-**Research question 2:**
-
--   Outcome variable (include the name/description and type of variable):
-
-**Load the data and provide a `glimpse()`**:
-
-```{r data-3, warning=FALSE}
-
-```
-
-**Exploratory Plots:**
-
-```{r plots-3}
-
-```
-
-#
-
-# Team Charter
-
-#### When will you meet as a team to work on the project components? Will these meetings be held in person or virtually?
-
-#### What is your group policy on missing team meetings (e.g., how much advance notice should be provided)?
-
-#### How will your team communicate (email, Slack, text messages)? What is your policy on appropriate response time (within a certain number of hours? Nights/weekends?)?
-
-####

### 2024-12-02 Naki_BLR.qmd
First

---
title: "Naki_BLR"
format: html
editor: visual
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

The `echo: false` option disables the printing of code (only output is displayed).

# Loading dataset

```{r}
#| echo: false
# Set the file path
file_path <- "/Users/nakiyahdhariwala/StatsProject/39096-0001-Data.rda"

# Load the RDA file
load(file_path)

# List objects loaded into the environment
ls()

# Access the dataset (replace `da39096.0001` with the actual name found in `ls()`)
dataset <- da39096.0001  # Use the correct object name

```

The `echo: false` option disables the printing of code (only output is displayed).

```{r}
# Exploring the data

#colnames(dataset)
head(dataset)
```

# Variable selection

```{r}
required_cols <- c("Q1R5", # Outcome Variable; Stopping discrimination against racial/ethnic minorities (Binary: 0 = Not important, 1 = Important).
                   "S2_RACE_PRIME",   # Primary Race
                   "MAIN_ETH_QUOTA",  # Ethnic Quota
                   "S3",              # Orientation
                   "S3B",             # Gender
                   #"S4",              # Current state of residence
                   "S5_AGE",          # Age categories
                   "S13",             # Highest level of education
                   "S14",             # Community lived in

                   # Variables capturing experiences with racism
                   # Create stacked bar charts for this section to identify which variables we shoul look into further

                   "Q627",            # Treated unfairly or discriminated against

                   "Q629R1",          # Unfair treatment due to racial background/ethnicity
                   "Q629R2",          # Unfair treatment due to skin color
                   "Q629R3",          # Unfair treatment due to gender
                   "Q629R4",          # Unfair treatment due to sexuality/orientation
                   "Q629R5",          # Unfair treatment due to immigration status
                   "Q629R6",          # Unfair treatment due to religion
                   "Q629R7",          # Unfair treatment due to accent

                   "Q633",            # Discrimination impact on life satisfaction

                   # Religion-related variables
                   "Q58R1",            # Catholic
                   "Q58R2",            # Protestant
                   "Q58R3",            # Other Christian
                   "Q58R4",            # Jewish
                   "Q58R5",            # Muslim
                   "Q58R6",            # Hindu
                   "Q58R7",            # Buddhist
                   "Q58R8",            # Mormon/LDS
                   "Q58R9",            # Folk Religion
                   "Q58R10",           # Daoist
                   "Q58R11",           # Shamanism
                   "Q58R12",           # Ancestral Religion
                   "Q58R13",           # Spiritual but not religious
                   "Q58R14",           # Atheist
                   "Q58R15")            # Agnostic

# Create a new dataframe with selected variables
dataset_new <- dataset[, required_cols]
head(dataset_new)

```

# Convert variables

```{r}
# Convert the required columns to factors
dataset_new[required_cols] <- lapply(dataset_new[required_cols], as.factor)

# Verify that the columns have been successfully converted to factors
str(dataset_new[required_cols])
head(dataset_new)
```

```{r}
# Convert individual religion columns to numeric
for (i in 1:15) {
  col_name <- paste0("Q58R", i, "_numeric")
  dataset_new[[col_name]] <- as.numeric(sub("^\\((\\d+)\\).*", "\\1", dataset_new[[paste0("Q58R", i)]]))
}

# Create a new Religion variable
dataset_new$Religion <- "None"
dataset_new$Religion[dataset_new$Q58R1_numeric == 1] <- "Catholic"
dataset_new$Religion[dataset_new$Q58R2_numeric == 1] <- "Protestant"
dataset_new$Religion[dataset_new$Q58R3_numeric == 1] <- "Other Christian"
dataset_new$Religion[dataset_new$Q58R4_numeric == 1] <- "Jewish"
dataset_new$Religion[dataset_new$Q58R5_numeric == 1] <- "Muslim"
dataset_new$Religion[dataset_new$Q58R6_numeric == 1] <- "Hindu"
dataset_new$Religion[dataset_new$Q58R7_numeric == 1] <- "Buddhist"
dataset_new$Religion[dataset_new$Q58R8_numeric == 1] <- "Mormon/LDS"
dataset_new$Religion[dataset_new$Q58R9_numeric == 1] <- "Folk Religion"
dataset_new$Religion[dataset_new$Q58R10_numeric == 1] <- "Daoist"
dataset_new$Religion[dataset_new$Q58R11_numeric == 1] <- "Shamanism"
dataset_new$Religion[dataset_new$Q58R12_numeric == 1] <- "Ancestral Religion"
dataset_new$Religion[dataset_new$Q58R13_numeric == 1] <- "Spiritual but not religious"
dataset_new$Religion[dataset_new$Q58R14_numeric == 1] <- "Atheist"
dataset_new$Religion[dataset_new$Q58R15_numeric == 1] <- "Agnostic"

# Check if an individual reported multiple religious affiliations 
dataset_new$multiple_religions <- rowSums(dataset_new[, grep("_numeric", names(dataset_new))] == 1)
check <- table(dataset_new$multiple_religions)

# Handle multiple religious affiliations
dataset_new$Religion[rowSums(dataset_new[, grep("_numeric", names(dataset_new))] == 1) > 1] <- "Multiple Religions"

# Convert Religion to a factor
dataset_new$Religion <- factor(dataset_new$Religion,
                               levels = c("Catholic", "Protestant", "Other Christian", "Jewish", "Muslim",
                                          "Hindu", "Buddhist", "Mormon/LDS", "Folk Religion",
                                          "Daoist", "Shamanism", "Ancestral Religion",
                                          "Spiritual but not religious", "Atheist", "Agnostic", 
                                          "Multiple Religions", "None"))

# Create a table of counts for the new Religion variable
table(dataset_new$Religion)

```

```{r}
# Remove redundant religion columns (Q58R1 to Q58R15 and their numeric versions)
redundant_cols <- c(
  "Q58R1", "Q58R2", "Q58R3", "Q58R4", "Q58R5", "Q58R6", "Q58R7", "Q58R8", 
  "Q58R9", "Q58R10", "Q58R11", "Q58R12", "Q58R13", "Q58R14", "Q58R15",
  "Q58R1_numeric", "Q58R2_numeric", "Q58R3_numeric", "Q58R4_numeric", 
  "Q58R5_numeric", "Q58R6_numeric", "Q58R7_numeric", "Q58R8_numeric", 
  "Q58R9_numeric", "Q58R10_numeric", "Q58R11_numeric", "Q58R12_numeric", 
  "Q58R13_numeric", "Q58R14_numeric", "Q58R15_numeric"
)

# Remove redundant columns from the dataset
dataset_clean <- dataset_new[, !(names(dataset_new) %in% redundant_cols)]

# Take a look at the updated dataset
head(dataset_clean)
```

# Missing values

```{r}
# Check the number of missing values for each variable
missing_summary <- colSums(is.na(dataset_clean))
missing_summary

# Finding proportion of Missing Data
missing_proportion <- colSums(is.na(dataset_clean)) / nrow(dataset_clean)
missing_proportion

# Finding total missing data proportion
overall_missing <- sum(is.na(dataset_clean)) / (nrow(dataset_clean) * ncol(dataset_clean))
overall_missing

# Omiting missing data (Not gonna do it for now)
# dataset_clean <- na.omit(dataset_clean)
# head(dataset_clean)
```

We can see that 9180 values are missing from every column that relates to questions of unfair treatment. Upon further investigation we find that missing values account for approximately 52% of the total column data. For now, we decide not to do anything and proceed further with the analysis

```{r}
head(dataset_new)
head(dataset_clean)
```

# Cleaned version

Make one chart showing the outcome distribution

```{r}
# (Outcome) Distribution of the proportion of people who view "Stopping discrimination against racial/ethnic minorities" as an important issue
table(dataset_clean$Q1R5) / nrow(dataset_clean)

table(dataset_clean$S2_RACE_PRIME)
table(dataset_clean$MAIN_ETH_QUOTA)

```
```{r}
# List of columns to be transformed
cols <- c("Q629R1", "Q629R2", "Q629R3", "Q629R4", "Q629R5", "Q629R6", "Q629R7")

# Convert factors to numeric (0 or 1) based on levels
dataset_clean$Q629R1_racialbackground <- ifelse(as.numeric(dataset_clean$Q629R1) == 1, 1, 0)
dataset_clean$Q629R2_skincolor <- ifelse(as.numeric(dataset_clean$Q629R2) == 1, 1, 0)
dataset_clean$Q629R3_gender <- ifelse(as.numeric(dataset_clean$Q629R3) == 1, 1, 0)
dataset_clean$Q629R4_orientation <- ifelse(as.numeric(dataset_clean$Q629R4) == 1, 1, 0)
dataset_clean$Q629R5_immigration <- ifelse(as.numeric(dataset_clean$Q629R5) == 1, 1, 0)
dataset_clean$Q629R6_religion <- ifelse(as.numeric(dataset_clean$Q629R6) == 1, 1, 0)
dataset_clean$Q629R7_accent <- ifelse(as.numeric(dataset_clean$Q629R7) == 1, 1, 0)

# Check the distribution of the newly created binary columns
for (col in c("Q629R1_racialbackground", "Q629R2_skincolor", "Q629R3_gender", 
              "Q629R4_orientation", "Q629R5_immigration", "Q629R6_religion", 
              "Q629R7_accent")) {
  print(col)  # Print the column name
  print(table(dataset_clean[[col]]))  # Access the column dynamically
}
```

```{r}
# Checking for levels in the model
level <- sapply(dataset_clean, function(x) length(unique(x)))
level

dataset_clean$interaction_term <- interaction(dataset_clean$S2_RACE_PRIME, dataset_clean$MAIN_ETH_QUOTA)

model <- glm(Q1R5 ~ S2_RACE_PRIME +   # Primary Race
                    MAIN_ETH_QUOTA +  # Ethnic Quota
                   S3 +              # Orientation
                   S3B +             # Gender
                   S5_AGE +          # Age categories
                   S13 +             # Highest level of education
                   S14 +             # Community lived in

                   # Variables capturing experiences with racism
                  Q627 +           # Treated unfairly or discriminated against
                  Religion +
               interaction_term,
                  #Q629R1,          # Unfair treatment due to racial background/ethnicity
                  #Q629R2 +          # Unfair treatment due to skin color
                  #Q629R3 +          # Unfair treatment due to gender
                  #Q629R4 +          # Unfair treatment due to sexuality/orientation
                  #Q629R5 +          # Unfair treatment due to immigration status
                  #Q629R6 +          # Unfair treatment due to religion
                  #Q629R7 +          # Unfair treatment due to accent
                  # 
                  #Q633,             # Discrimination impact on life satisfaction
                   data = dataset_clean, family = binomial)            
summary(model)

#prop.table(table(dataset_clean$Q629R1))
#prop.table(table(dataset_clean$Q629R2))
# Repeat for other variables
```

```{r}
step_model <- step(model, direction = "both")
step_model
```

### 2024-12-02 Naki_EDA_Exploration.qmd
First

---
title: "EDA_Naki"
format: html
editor: visual
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

# Loading dataset

```{r}
#| echo: false
# Set the file path
file_path <- "/Users/nakiyahdhariwala/StatsProject/39096-0001-Data.rda"

# Load the RDA file
load(file_path)

# List objects loaded into the environment
ls()

# Access the dataset (replace `da39096.0001` with the actual name found in `ls()`)
dataset <- da39096.0001  # Use the correct object name

```

The `echo: false` option disables the printing of code (only output is displayed).

```{r}
# Exploring the data

#colnames(dataset)
head(dataset)
```

# Variable selection

```{r}
required_cols <- c("Q1R5", # Outcome Variable; Stopping discrimination against racial/ethnic minorities (Binary: 0 = Not important, 1 = Important).
                   "S2_RACE_PRIME",   # Primary Race
                   # "MAIN_ETH_QUOTA",  # Ethnic Quota
                   "S3",              # Orientation
                   "S3B",             # Gender
                   "S4",              # Current state of residence
                   "S5_AGE",          # Age categories
                   "S13",             # Highest level of education
                   "S14",             # Community lived in

                   # Variables capturing experiences with racism
                   # Create stacked bar charts for this section to identify which variables we shoul look into further 
                   "Q627",            # Treated unfairly or discriminated against
                   # "Q628",            # Location of discrimination
                   "Q629R1",          # Unfair treatment due to racial background/ethnicity
                   "Q629R2",          # Unfair treatment due to skin color
                   "Q629R3",          # Unfair treatment due to gender
                   "Q629R4",          # Unfair treatment due to sexuality/orientation
                   "Q629R5",          # Unfair treatment due to immigration status
                   "Q629R6",          # Unfair treatment due to religion
                   "Q629R7",          # Unfair treatment due to accent

                   "Q633",            # Discrimination impact on life satisfaction
                   # Religion-related variables
                   "Q58R1",            # Catholic
                   "Q58R2",            # Protestant
                   "Q58R3",            # Other Christian
                   "Q58R4",            # Jewish
                   "Q58R5",            # Muslim
                   "Q58R6",            # Hindu
                   "Q58R7",            # Buddhist
                   "Q58R8",            # Mormon/LDS
                   "Q58R9",            # Folk Religion
                   "Q58R10",           # Daoist
                   "Q58R11",           # Shamanism
                   "Q58R12",           # Ancestral Religion
                   "Q58R13",           # Spiritual but not religious
                   "Q58R14",           # Atheist
                   "Q58R15")            # Agnostic

# Create a new dataframe with selected variables
dataset_new <- dataset[, required_cols]
head(dataset_new)

```

# Convert variables

```{r}
# Convert the required columns to factors
dataset_new[required_cols] <- lapply(dataset_new[required_cols], as.factor)

# Verify that the columns have been successfully converted to factors
str(dataset_new[required_cols])
head(dataset_new)
```

```{r}
# Convert individual religion columns to numeric
for (i in 1:15) {
  col_name <- paste0("Q58R", i, "_numeric")
  dataset_new[[col_name]] <- as.numeric(sub("^\\((\\d+)\\).*", "\\1", dataset_new[[paste0("Q58R", i)]]))
}

# Create a new Religion variable
dataset_new$Religion <- "None"
dataset_new$Religion[dataset_new$Q58R1_numeric == 1] <- "Catholic"
dataset_new$Religion[dataset_new$Q58R2_numeric == 1] <- "Protestant"
dataset_new$Religion[dataset_new$Q58R3_numeric == 1] <- "Other Christian"
dataset_new$Religion[dataset_new$Q58R4_numeric == 1] <- "Jewish"
dataset_new$Religion[dataset_new$Q58R5_numeric == 1] <- "Muslim"
dataset_new$Religion[dataset_new$Q58R6_numeric == 1] <- "Hindu"
dataset_new$Religion[dataset_new$Q58R7_numeric == 1] <- "Buddhist"
dataset_new$Religion[dataset_new$Q58R8_numeric == 1] <- "Mormon/LDS"
dataset_new$Religion[dataset_new$Q58R9_numeric == 1] <- "Folk Religion"
dataset_new$Religion[dataset_new$Q58R10_numeric == 1] <- "Daoist"
dataset_new$Religion[dataset_new$Q58R11_numeric == 1] <- "Shamanism"
dataset_new$Religion[dataset_new$Q58R12_numeric == 1] <- "Ancestral Religion"
dataset_new$Religion[dataset_new$Q58R13_numeric == 1] <- "Spiritual but not religious"
dataset_new$Religion[dataset_new$Q58R14_numeric == 1] <- "Atheist"
dataset_new$Religion[dataset_new$Q58R15_numeric == 1] <- "Agnostic"

# Convert Religion to a factor
dataset_new$Religion <- factor(dataset_new$Religion,
                               levels = c("Catholic", "Protestant", "Other Christian", "Jewish", "Muslim",
                                          "Hindu", "Buddhist", "Mormon/LDS", "Folk Religion",
                                          "Daoist", "Shamanism", "Ancestral Religion",
                                          "Spiritual but not religious", "Atheist", "Agnostic", "None"))

# Create a table of counts for the new Religion variable
table(dataset_new$Religion)

```

```{r}
# Remove redundant religion columns (Q58R1 to Q58R15 and their numeric versions)
redundant_cols <- c(
  "Q58R1", "Q58R2", "Q58R3", "Q58R4", "Q58R5", "Q58R6", "Q58R7", "Q58R8", 
  "Q58R9", "Q58R10", "Q58R11", "Q58R12", "Q58R13", "Q58R14", "Q58R15",
  "Q58R1_numeric", "Q58R2_numeric", "Q58R3_numeric", "Q58R4_numeric", 
  "Q58R5_numeric", "Q58R6_numeric", "Q58R7_numeric", "Q58R8_numeric", 
  "Q58R9_numeric", "Q58R10_numeric", "Q58R11_numeric", "Q58R12_numeric", 
  "Q58R13_numeric", "Q58R14_numeric", "Q58R15_numeric"
)

# Remove redundant columns from the dataset
dataset_clean <- dataset_new[, !(names(dataset_new) %in% redundant_cols)]

# Take a look at the updated dataset
head(dataset_clean)
```

# Missing values

```{r}
# Check the number of missing values for each variable
missing_summary <- colSums(is.na(dataset_clean))
missing_summary

# Finding proportion of Missing Data
missing_proportion <- colSums(is.na(dataset_clean)) / nrow(dataset_clean)
missing_proportion

# Finding total missing data proportion
overall_missing <- sum(is.na(dataset_new)) / (nrow(dataset_new) * ncol(dataset_new))
overall_missing

# Omiting missing data (Not gonna do it for now)
# dataset_clean <- na.omit(dataset_clean)
# head(dataset_clean)
```

We can see that 9180 values are missing from every column that relates to questions of unfair treatment. Upon further investigation we find that missing values account for approximately 52% of the total column data. For now, we decide not to do anything and proceed further with the analysis

# Cleaned version

```{r}
# (Outcome) Distribution of the proportion of people who view "Stopping discrimination against racial/ethnic minorities" as an important issue
table(dataset_new$Q1R5) / nrow(dataset_new)
```

```{r}
#Summarize counts and proportions for every variable

#lapply(dataset_new[, -1], table)

```

```{r}
# Function to create a summary table with counts and proportions
create_summary_table <- function(variable, var_name) {
  counts <- table(variable)
  if (length(counts) == 0) {
    return(data.frame(Variable = character(), Level = character(), Count = numeric(), Proportion = numeric()))
  }
  proportions <- prop.table(counts)
  summary_df <- data.frame(
    Variable = var_name,
    Level = names(counts),
    Count = as.numeric(counts),
    Proportion = round(as.numeric(proportions), 4)
  )
  return(summary_df)
}

# Apply the function to each variable of interest in the cleaned dataset
categorical_vars <- names(dataset_clean)

# Initialize an empty data frame to store all summary tables
full_summary_table <- data.frame(Variable = character(), Level = character(), Count = numeric(), Proportion = numeric())

# Iterate through each variable and generate the summary table
for (var in categorical_vars) {
  summary_table <- create_summary_table(dataset_clean[[var]], var)
  full_summary_table <- rbind(full_summary_table, summary_table)
}

# Print the full summary table
full_summary_table

```

| **Variable** | **Level** | **Count** | **Proportion** |
|------------------|--------------------|------------------|------------------|
| Q1R5 | \(0\) NO TO: Stopping discrimination against racial/ethnic minorities | 5728 | 0.6848 |
| Q1R5 | \(1\) Stopping discrimination against racial/ethnic minorities | 2637 | 0.3152 |
| S2_RACE_PRIME | \(1\) White | 1310 | 0.1566 |
| S2_RACE_PRIME | \(2\) Hispanic or Latino | 1836 | 0.2195 |
| S2_RACE_PRIME | \(3\) Black or African American | 2733 | 0.3267 |
| S2_RACE_PRIME | \(4\) Asian American | 1834 | 0.2192 |
| S2_RACE_PRIME | \(5\) American Indian/Native American | 357 | 0.0427 |
| S2_RACE_PRIME | \(6\) Arab, Middle Eastern or North African | 145 | 0.0173 |
| S2_RACE_PRIME | \(7\) Native Hawaiian | 39 | 0.0047 |
| S2_RACE_PRIME | \(8\) Not Hawaiian, but other Pacific Islander | 111 | 0.0133 |
| MAIN_ETH_QUOTA | \(1\) White | 955 | 0.1142 |
| MAIN_ETH_QUOTA | \(2\) Latino | 2007 | 0.2399 |
| MAIN_ETH_QUOTA | \(3\) Black | 2578 | 0.3082 |
| MAIN_ETH_QUOTA | \(4\) AAPI | 1898 | 0.2269 |
| MAIN_ETH_QUOTA | \(5\) Native American | 552 | 0.066 |
| MAIN_ETH_QUOTA | \(6\) MENA | 176 | 0.021 |
| MAIN_ETH_QUOTA | \(7\) Native Hawaiian | 199 | 0.0238 |
| S3 | \(01\) Straight, that is, not gay or lesbian | 7037 | 0.8412 |
| S3 | \(02\) Gay or lesbian | 391 | 0.0467 |
| S3 | \(03\) Bisexual | 587 | 0.0702 |
| S3 | \(04\) Something else (Specify) | 207 | 0.0247 |
| S3 | \(88\) Refused | 90 | 0.0108 |
| S3 | \(99\) I don | 53 | 0.0063 |
| S3B | \(1\) Man | 3388 | 0.405 |
| S3B | \(2\) Woman | 4844 | 0.5791 |
| S3B | \(3\) Non-binary | 109 | 0.013 |
| S3B | \(4\) Something else (Specify) | 24 | 0.0029 |
| S4 | \(01\) Alaska | 18 | 0.0022 |
| S4 | \(02\) Alabama | 120 | 0.0143 |
| S4 | \(03\) Arizona | 207 | 0.0247 |
| S4 | \(04\) Arkansas | 74 | 0.0088 |
| S4 | \(05\) California | 1456 | 0.1741 |
| S4 | \(06\) Colorado | 136 | 0.0163 |
| S4 | \(07\) Connecticut | 77 | 0.0092 |
| S4 | \(08\) Delaware | 28 | 0.0033 |
| S4 | \(09\) District of Columbia | 43 | 0.0051 |
| S4 | \(10\) Florida | 677 | 0.0809 |
| S4 | \(11\) Georgia | 349 | 0.0417 |
| S4 | \(12\) Hawaii | 84 | 0.01 |
| S4 | \(13\) Idaho | 15 | 0.0018 |
| S4 | \(14\) Illinois | 333 | 0.0398 |
| S4 | \(15\) Indiana | 99 | 0.0118 |
| S4 | \(16\) Iowa | 27 | 0.0032 |
| S4 | \(17\) Kansas | 39 | 0.0047 |
| S4 | \(18\) Kentucky | 74 | 0.0088 |
| S4 | \(19\) Louisiana | 112 | 0.0134 |
| S4 | \(20\) Maine | 8 | 0.001 |
| S4 | \(21\) Maryland | 208 | 0.0249 |
| S4 | \(22\) Massachusetts | 142 | 0.017 |
| S4 | \(23\) Michigan | 182 | 0.0218 |
| S4 | \(24\) Minnesota | 79 | 0.0094 |
| S4 | \(25\) Mississippi | 68 | 0.0081 |
| S4 | \(26\) Missouri | 89 | 0.0106 |
| S4 | \(27\) Montana | 8 | 0.001 |
| S4 | \(28\) Nebraska | 24 | 0.0029 |
| S4 | \(29\) Nevada | 92 | 0.011 |
| S4 | \(30\) New Hampshire | 13 | 0.0016 |
| S4 | \(31\) New Jersey | 248 | 0.0296 |
| S4 | \(32\) New Mexico | 56 | 0.0067 |
| S4 | \(33\) New York | 593 | 0.0709 |
| S4 | \(34\) North Carolina | 258 | 0.0308 |
| S4 | \(35\) North Dakota | 11 | 0.0013 |
| S4 | \(36\) Ohio | 235 | 0.0281 |
| S4 | \(37\) Oklahoma | 88 | 0.0105 |
| S4 | \(38\) Oregon | 74 | 0.0088 |
| S4 | \(39\) Pennsylvania | 231 | 0.0276 |
| S4 | \(40\) Rhode Island | 16 | 0.0019 |
| S4 | \(41\) South Carolina | 99 | 0.0118 |
| S4 | \(42\) South Dakota | 13 | 0.0016 |
| S4 | \(43\) Tennessee | 135 | 0.0161 |
| S4 | \(44\) Texas | 862 | 0.103 |
| S4 | \(45\) Utah | 46 | 0.0055 |
| S4 | \(46\) Vermont | 7 | 0.0008 |
| S4 | \(47\) Virginia | 210 | 0.0251 |
| S4 | \(48\) Washington | 202 | 0.0241 |
| S4 | \(49\) West Virginia | 19 | 0.0023 |
| S4 | \(50\) Wisconsin | 78 | 0.0093 |
| S4 | \(51\) Wyoming | 3 | 0.0004 |
| S5_AGE | \(1\) 16-17 | 0 | 0.0 |
| S5_AGE | \(2\) 18-29 | 2150 | 0.257 |
| S5_AGE | \(3\) 30-39 | 1639 | 0.1959 |
| S5_AGE | \(4\) 40-49 | 1376 | 0.1645 |
| S5_AGE | \(5\) 50-59 | 1289 | 0.1541 |
| S5_AGE | \(6\) 60-69 | 1289 | 0.1541 |
| S5_AGE | \(7\) 70+ | 622 | 0.0744 |
| S13 | \(1\) Grades 1-8 | 58 | 0.0069 |
| S13 | \(2\) Some High School, but did not graduate | 292 | 0.0349 |
| S13 | \(3\) High School graduate or GED | 1335 | 0.1596 |
| S13 | \(4\) Some college | 1834 | 0.2192 |
| S13 | \(5\) Associates, 2-year degree | 960 | 0.1148 |
| S13 | \(6\) Bachelors, 4-year degree | 2210 | 0.2642 |
| S13 | \(7\) Post-graduate degree | 1676 | 0.2004 |
| S14 | \(1\) Large urban area | 2517 | 0.3009 |
| S14 | \(2\) Large suburb near large city | 2793 | 0.3339 |
| S14 | \(3\) Small suburb near small town or city | 1331 | 0.1591 |
| S14 | \(4\) Small town or small city | 1130 | 0.1351 |
| S14 | \(5\) Rural area | 594 | 0.071 |
| Q627 | \(1\) Yes | 8365 | 1.0 |
| Q627 | \(2\) No | 0 | 0.0 |
| Q629R1 | \(0\) NO TO: Racial background or ethnicity | 2947 | 0.3523 |
| Q629R1 | \(1\) Racial background or ethnicity | 5418 | 0.6477 |
| Q629R2 | \(0\) NO TO: Skin color | 4040 | 0.483 |
| Q629R2 | \(1\) Skin color | 4325 | 0.517 |
| Q629R3 | \(0\) NO TO: Gender | 5340 | 0.6384 |
| Q629R3 | \(1\) Gender | 3025 | 0.3616 |
| Q629R4 | \(0\) NO TO: Sexuality or sexual orientation | 7219 | 0.863 |
| Q629R4 | \(1\) Sexuality or sexual orientation | 1146 | 0.137 |
| Q629R5 | \(0\) NO TO: Immigration status | 7718 | 0.9227 |
| Q629R5 | \(1\) Immigration status | 647 | 0.0773 |
| Q629R6 | \(0\) NO TO: Religion | 7299 | 0.8726 |
| Q629R6 | \(1\) Religion | 1066 | 0.1274 |
| Q629R7 | \(0\) NO TO: Accent, regardless of whether or not you have an accent | 6881 | 0.8226 |
| Q629R7 | \(1\) Accent, regardless of whether or not you have an accent | 1484 | 0.1774 |
| Q633 | \(1\) A lot | 997 | 0.1192 |
| Q633 | \(2\) Some | 2002 | 0.2393 |
| Q633 | \(3\) A little | 2964 | 0.3543 |
| Q633 | \(4\) Not at all | 1908 | 0.2281 |
| Q633 | \(5\) Don | 426 | 0.0509 |
| Q633 | \(6\) Refused | 68 | 0.0081 |
| Religion | Catholic | 1484 | 0.1774 |
| Religion | Protestant | 1901 | 0.2273 |
| Religion | Other Christian | 931 | 0.1113 |
| Religion | Jewish | 144 | 0.0172 |
| Religion | Muslim | 265 | 0.0317 |
| Religion | Hindu | 190 | 0.0227 |
| Religion | Buddhist | 231 | 0.0276 |
| Religion | Mormon/LDS | 50 | 0.0060 |
| Religion | Folk Religion | 15 | 0.0018 |
| Religion | Daoist | 11 | 0.0013 |
| Religion | Shamanism | 17 | 0.0020 |
| Religion | Ancestral Religion | 51 | 0.0061 |
| Religion | Spiritual but not religious | 787 | 0.0941 |
| Religion | Atheist | 444 | 0.0531 |
| Religion | Agnostic | 601 | 0.0718 |
| Religion | None | 1243 | 0.1486 |

# Visualization

```{r}
library(ggplot2)
library(dplyr)

```

# `Q627` vs `Q1R5`: Experience of Racism vs Stopping Discrimination

```{r}
ggplot(dataset_clean, aes(x = Q627, fill = Q1R5)) +
  geom_bar(position = "fill") +
  labs(
    title = "Q627 vs Q1R5",
    y = "Proportion",
    x = "Q627 (Experience of Racism)",
    fill = "Q1R5 (Stopping Discrimination)"
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal()

```

# `MAIN_ETH_QUOTA` vs `Q1R5`: Ethnic Quota vs Stopping Discrimination

```{r}
ggplot(dataset_clean, aes(x = MAIN_ETH_QUOTA, fill = Q1R5)) +
  geom_bar(position = "fill") +
  labs(
    title = "MAIN_ETH_QUOTA vs Q1R5",
    y = "Proportion",
    x = "MAIN_ETH_QUOTA (Ethnic Quota)",
    fill = "Q1R5 (Stopping Discrimination)"
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    plot.title = element_text(hjust = 0.5)
  )
```

# `S3B` vs `Q1R5`: Gender vs Stopping Discrimination

```{r}
ggplot(dataset_clean, aes(x = S3B, fill = Q1R5)) +
  geom_bar(position = "fill") +
  labs(
    title = "S3B vs Q1R5",
    y = "Proportion",
    x = "S3B (Gender)",
    fill = "Q1R5 (Stopping Discrimination)"
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    plot.title = element_text(hjust = 0.5)
  )
```

# `S2_RACE_PRIME` vs `Q1R5`: Primary Race vs Stopping Discrimination

```{r}
ggplot(dataset_clean, aes(x = S2_RACE_PRIME, fill = Q1R5)) +
  geom_bar(position = "fill") +
  labs(
    title = "S2_RACE_PRIME vs Q1R5",
    y = "Proportion",
    x = "S2_RACE_PRIME (Primary Race)",
    fill = "Q1R5 (Stopping Discrimination)"
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal()+
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    plot.title = element_text(hjust = 0.5)
  )

```

# `S5_AGE` vs `Q1R5`: Age vs Stopping Discrimination

```{r}
ggplot(dataset_clean, aes(x = S5_AGE, fill = Q1R5)) +
  geom_bar(position = "fill") +
  labs(
    title = "S5_AGE vs Q1R5",
    y = "Proportion",
    x = "S5_AGE (Age Group)",
    fill = "Q1R5 (Stopping Discrimination)"
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal()+
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    plot.title = element_text(hjust = 0.5)
  )

```

# `S14` vs `Q1R5`: Community Type vs Stopping Discrimination

```{r}
ggplot(dataset_clean, aes(x = S14, fill = Q1R5)) +
  geom_bar(position = "fill") +
  labs(
    title = "S14 vs Q1R5",
    y = "Proportion",
    x = "S14 (Community Type)",
    fill = "Q1R5 (Stopping Discrimination)"
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal()+
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    plot.title = element_text(hjust = 0.5)
  )

```

# Plot for `S13` (Education Level)

```{r}
ggplot(dataset_clean, aes(x = S13, fill = Q1R5)) +
  geom_bar(position = "fill") +
  labs(
    title = "S13 vs Q1R5",
    y = "Proportion",
    x = "S13 (Highest Level of Education)",
    fill = "Q1R5 (Stopping Discrimination)"
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    plot.title = element_text(hjust = 0.5)
  )

```

# Plot for `Q58` (Religion Variables)

```{r}
ggplot(dataset_clean, aes(x = Religion, fill = Q1R5)) +
  geom_bar(position = "fill") +
  labs(
    title = "Q58 vs Q1R5",
    y = "Proportion",
    x = "Religion",
    fill = "Q1R5 (Stopping Discrimination)"
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 7),
    plot.title = element_text(hjust = 0.5)
  )

```

# Plot for `Q633` (Impact of Discrimination on Life Satisfaction)

```{r}
ggplot(dataset_clean, aes(x = Q633, fill = Q1R5)) +
  geom_bar(position = "fill") +
  labs(
    title = "Q633 vs Q1R5",
    y = "Proportion",
    x = "Q633 (Discrimination Impact on Life Satisfaction)",
    fill = "Q1R5 (Stopping Discrimination)"
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    plot.title = element_text(hjust = 0.5)
  )
```

# Plot for `Q629` (Unfair Treatment Variables)

```{r}
# Create a sample dataset to use for plotting
plot_data <- data.frame(
  Variable = c(rep("Q629R1", 2), rep("Q629R2", 2), rep("Q629R3", 2), rep("Q629R4", 2), rep("Q629R5", 2), rep("Q629R6", 2), rep("Q629R7", 2)),
  Level = c("(0) NO TO: Racial background or ethnicity", "(1) Racial background or ethnicity",
            "(0) NO TO: Skin color", "(1) Skin color",
            "(0) NO TO: Gender", "(1) Gender",
            "(0) NO TO: Sexuality or sexual orientation", "(1) Sexuality or sexual orientation",
            "(0) NO TO: Immigration status", "(1) Immigration status",
            "(0) NO TO: Religion", "(1) Religion",
            "(0) NO TO: Accent, regardless of whether or not you have an accent", "(1) Accent, regardless of whether or not you have an accent"),
  Count = c(2947, 5418, 4040, 4325, 5340, 3025, 7219, 1146, 7718, 647, 7299, 1066, 6881, 1484),
  Proportion = c(0.3523, 0.6477, 0.483, 0.517, 0.6384, 0.3616, 0.863, 0.137, 0.9227, 0.0773, 0.8726, 0.1274, 0.8226, 0.1774)
)

# Plot for each variable
ggplot(plot_data, aes(x = Variable, y = Proportion, fill = Level)) +
  geom_bar(stat = "identity", position = "fill") +
  labs(title = "Proportions of Different Variables by Level",
       x = "Variable",
       y = "Proportion") +
  scale_y_continuous(labels = scales::percent) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# chi-square

```{r}
dataset_clean$S5_AGE <- droplevels(dataset_clean$S5_AGE)

# Create a summary table for all categorical variables
variables <- c("Q1R5", "S2_RACE_PRIME", "MAIN_ETH_QUOTA", "S3", "S3B", "S5_AGE", "S13", "Q627", "Q629R1", "Q629R2", "Q629R3", "Q629R4", "Q629R5", "Q629R6", "Q629R7", "Q633", "Religion")

# Chi-square tests for each variable
chi_square_results <- lapply(variables, function(var) {
  chisq.test(table(dataset_clean[[var]], dataset_clean$Q1R5))
})

# Print chi-square results
names(chi_square_results) <- variables
print(chi_square_results)

```

| Variable | Chi-Square Statistic | Degrees of Freedom (df) | P-Value | Test Type |
|---------------|---------------|---------------|---------------|---------------|

|               |        |     |            |                            |
|---------------|--------|-----|------------|----------------------------|
| S2_RACE_PRIME | 416.06 | 7   | \< 2.2e-16 | Pearson's Chi-squared test |

|                |        |     |            |                            |
|----------------|--------|-----|------------|----------------------------|
| MAIN_ETH_QUOTA | 383.38 | 6   | \< 2.2e-16 | Pearson's Chi-squared test |

|     |        |     |         |                            |
|-----|--------|-----|---------|----------------------------|
| S3  | 11.184 | 5   | 0.04786 | Pearson's Chi-squared test |

|     |        |     |           |                            |
|-----|--------|-----|-----------|----------------------------|
| S3B | 31.259 | 3   | 7.497e-07 | Pearson's Chi-squared test |

|        |        |     |        |                            |
|--------|--------|-----|--------|----------------------------|
| S5_AGE | 6.2621 | 5   | 0.2815 | Pearson's Chi-squared test |

|     |        |     |         |                            |
|-----|--------|-----|---------|----------------------------|
| S13 | 13.215 | 6   | 0.03975 | Pearson's Chi-squared test |

|      |     |     |     |                            |
|------|-----|-----|-----|----------------------------|
| Q627 | NaN | 1   | NA  | Pearson's Chi-squared test |

|  |  |  |  |  |
|---------------|---------------|---------------|---------------|---------------|
| Q629R1 | 114.83 | 1 | \< 2.2e-16 | Pearson's Chi-squared test with Yates' correction |

|  |  |  |  |  |
|---------------|---------------|---------------|---------------|---------------|
| Q629R2 | 72.719 | 1 | \< 2.2e-16 | Pearson's Chi-squared test with Yates' correction |

|        |         |     |        |                                                   |
|---------------|---------------|---------------|---------------|---------------|
| Q629R3 | 0.41214 | 1   | 0.5209 | Pearson's Chi-squared test with Yates' correction |

|        |        |     |        |                                                   |
|---------------|---------------|---------------|---------------|---------------|
| Q629R4 | 1.3169 | 1   | 0.2511 | Pearson's Chi-squared test with Yates' correction |

|        |       |     |           |                                                   |
|---------------|---------------|---------------|---------------|---------------|
| Q629R5 | 11.48 | 1   | 0.0007036 | Pearson's Chi-squared test with Yates' correction |

|  |  |  |  |  |
|---------------|---------------|---------------|---------------|---------------|
| Q629R6 | 20.113 | 1 | 7.301e-06 | Pearson's Chi-squared test with Yates' correction |

|        |        |     |        |                                                   |
|---------------|---------------|---------------|---------------|---------------|
| Q629R7 | 2.6287 | 1   | 0.1049 | Pearson's Chi-squared test with Yates' correction |

|      |        |     |           |                            |
|------|--------|-----|-----------|----------------------------|
| Q633 | 70.581 | 5   | 7.759e-14 | Pearson's Chi-squared test |

|          |        |     |           |                            |
|----------|--------|-----|-----------|----------------------------|
| Religion | 84.974 | 15  | 8.507e-12 | Pearson's Chi-squared test |

```{r}
# Model Fitting - Logistic Regression
# Logistic regression to predict stopping discrimination (Q1R5)

# Create interaction term between MAIN_ETH_QUOTA and Religion
dataset_clean$interaction_term <- interaction(dataset_clean$MAIN_ETH_QUOTA, dataset_clean$Religion)

```

```{r}
dataset_clean <- droplevels(dataset_clean)
dataset_clean <- dataset_clean %>%
  mutate(across(where(is.character), as.factor))
# Refitting the logistic regression model without Q627
model <- glm(Q1R5 ~ S2_RACE_PRIME + MAIN_ETH_QUOTA + S3 + S3B + S5_AGE + S13 + Q629R1 + Q633 + Religion + interaction_term,
             data = dataset_clean,
             family = binomial())

# Summary of the model
summary(model)
```

### **Interaction Terms**

1.  **Experiences with Racism × Perceptions of Discrimination**:

    -   Interaction between **Q627** (personal experience) and **Q619-Q626** (perceived discrimination toward groups).

2.  **Race/Ethnicity × Social Prioritization**:

    -   Interaction between **S2_RACE** and policy prioritization for racial equity issues (e.g., Q117-Q125).

3.  **Racism Experience × Leadership Perceptions**:

    -   Interaction between **Q627** and perceived efficacy of racial/ethnic leadership (Q159-Q161).

4.  **Interaction (Q627 × S2_RACE)**:

    -   Investigates how the impact of discrimination varies by racial/ethnic group, directly addressing intersectionality.

### **Key Recommendations**

1.  Focus on **Q627**, **Q629R1**, and **S2_RACE** as the core predictors.

2.  Use **interaction terms** (e.g., Q627 × S2_RACE) to explore differences across racial/ethnic groups.

3.  Include **Q633** to capture the subjective impact of discrimination.

4.  Use **perceptions of systemic discrimination (Q619-Q626)** to contextualize personal experiences.

```{r}
```

```{r echo: false}
# Creating an interaction term
# Create interaction term
dataset$Interaction <- interaction(dataset$Q627, dataset$S2_RACE)

                   # "S2_RACE_PRIME",   # Primary Race
                   # "MAIN_ETH_QUOTA",  # Ethnic Quota
                   # "S3",              # How you think of yourself
                   # "S3B",             # Gender
                   # "S4",              # Current state of residence
                   # # "S5",            # Year of birth (commented out)
                   # "S5_AGE",          # Age categories
                   # "S13",             # Highest level of education
                   # "S14",             # Community lived in

                   # # Variables capturing experiences with racism
                   # "Q627",            # Treated unfairly or discriminated against
                   # # "Q628",            # Location of discrimination
                   # "Q629R1",          # Unfair treatment due to racial background/ethnicity
                   # "Q629R2",          # Unfair treatment due to skin color
                   # "Q629R3",          # Unfair treatment due to gender
                   # "Q629R4",          # Unfair treatment due to sexuality/orientation
                   # "Q629R5",          # Unfair treatment due to immigration status
                   # "Q629R6",          # Unfair treatment due to religion
                   # "Q629R7",          # Unfair treatment due to accent
                   # "Q633",            # Discrimination impact on life satisfaction
                   # "Q619_Q626R2",     # Perceptions of discrimination against Blacks
                   # "Q619_Q626R3",     # Perceptions of discrimination against Asian Americans
                   # "Q619_Q626R4",     # Perceptions of discrimination against Native Americans
                   # "Q619_Q626R5",     # Perceptions of discrimination against Immigrants
                   # "Q619_Q626R6")    # Perceptions of discrimination against Latinos

# Logistic regression model
model <- glm(
  Q1R5 ~ S2_RACE_PRIME + MAIN_ETH_QUOTA + S3 + S3B + S4 + S5_AGE + S13 + S14 + 
   Q627 + Q629R1 + Q629R2 + Q629R3 + Q629R4 + Q629R5 + Q629R6 + Q629R7 +
    Q633 +
     Q619_Q626R2 + Q619_Q626R3 + Q619_Q626R4 + Q619_Q626R5 + Q619_Q626R6,
  family = binomial(link = "logit"),
  data = dataset_new
)

# Summarize model
summary(model)
```

```{r}
# Check levels of all variables used in the model
sapply(dataset[, c("S2_RACE_PRIME", "MAIN_ETH_QUOTA", "S3", "S3B", "S4", "S5_AGE", "S13", "S14", 
                   "Q627", "Q629R1", "Q629R2", "Q629R3", "Q629R4", "Q629R5", "Q629R6", "Q629R7",
                   "Q633", "Q619_Q626R2", "Q619_Q626R3", "Q619_Q626R4", "Q619_Q626R5", "Q619_Q626R6")], 
       function(x) if(is.factor(x)) levels(x) else unique(x))

# Identify single-level factors
single_level_vars <- names(Filter(function(x) length(unique(x)) <= 1, dataset_new))
print(single_level_vars)

```

```{r}
{r}
# Example: Stacked bar plot for Q627
library(ggplot2)
ggplot(dataset_new, aes(x = Q627, fill = Q1R5)) +
  geom_bar(position = "fill") +
  labs(title = "Q627 vs Q1R5", y = "Proportion", x = "Q627 (Experience of Racism)")

ggplot(dataset_new, aes(x = MAIN_ETH_QUOTA, fill = Q1R5)) +
  geom_bar(position = "fill") +
  labs(title = "Q627 vs Q1R5", y = "Proportion", x = "Q627 (Experience of Racism)")
```

### 2024-12-02 peter_EDA.qmd
adding files

---
title: "peter_EDA"
format: html
author: "Peter de Guzman"
editor: visual
---

Conducting EDA on the 2020 CMPS dataset for Research Question 1

# Loading data

# Selecting Variables

# Cleaning Variables

# Analysis of Missing Data

# Save cleaned data

# Creating summary table

# Running chi-square tests

# Fitting OLR Model

# Fitting MLR Model

### 2024-12-02 peter_plots.qmd
adding files

---
title: "peter_plots"
format: html
editor: visual
---

##

# Loading data

# plot 1:

# plot 2:

# plot 3:

### 2024-12-02 peter_EDA.qmd
updating EDA qmd

@@
 # Loading data
+```{r, echo=FALSE}
+file_path <- "/Users/pdeguz01/Documents/git/Data/2020 CMPS/DS0001/39096-0001-Data.rda"
+# Load data 
+load(file_path)
+data <- da39096.0001
+rm(da39096.0001)
+```
 
@@
 
+Need to include the outcome variable, which is an ordinal variable Q29.
+Main predictor variable is Q271 "How important is being [RACIAL GROUP] to your identity?"
+```{r}
+required_cols <- c("Q29", # Outcome Variable; interest in politics
+                   "Q271", # Predictor Variable; importance of race to identity
+                   "S3", # Orientation
+                   "S3B", # Gender
+                   "S5_AGE", # Age Category
+                   "S13", # Educational Attainment
+                   "S14", # Community Type
+                   "S7", # Nativity
+                   "Q21", # Party ID
+                   "S2_M", # Multiracial
+
+                   # Religion-related variables
+                   "Q58R1",            # Catholic
+                   "Q58R2",            # Protestant
+                   "Q58R3",            # Other Christian
+                   "Q58R4",            # Jewish
+                   "Q58R5",            # Muslim
+                   "Q58R6",            # Hindu
+                   "Q58R7",            # Buddhist
+                   "Q58R8",            # Mormon/LDS
+                   "Q58R9",            # Folk Religion
+                   "Q58R10",           # Daoist
+                   "Q58R11",           # Shamanism
+                   "Q58R12",           # Ancestral Religion
+                   "Q58R13",           # Spiritual but not religious
+                   "Q58R14",           # Atheist
+                   "Q58R15",            # Agnostic
+
+                   # Variables capturing experiences with racism
+                   "Q627", # Treated unfairly or discriminated against ,
+                   "Q628", # Location of discrimination
+                   "Q629R1" # Unfair treatment due to racial background/ethnicity
+                   )
+
+# Create new df
+data_filtered <- data[, required_cols]
+head(data_filtered)
+
+```
+
 # Cleaning Variables
 
+```{r}
+# Convert the required columns to factors
+data_filtered[required_cols] <- lapply(data_filtered[required_cols], as.factor)
+
+# Verify that the columns have been successfully converted to factors
+str(data_filtered[required_cols])
+head(data_filtered)
+```
+Cleaning and Recoding Religion Variables
+```{r}
+# Convert individual religion columns to numeric
+for (i in 1:15) {
+  col_name <- paste0("Q58R", i, "_numeric")
+  dataset_new[[col_name]] <- as.numeric(sub("^\\((\\d+)\\).*", "\\1", dataset_new[[paste0("Q58R", i)]]))
+}
+
+# Create a new Religion variable
+dataset_new$Religion <- "None"
+dataset_new$Religion[dataset_new$Q58R1_numeric == 1] <- "Catholic"
+dataset_new$Religion[dataset_new$Q58R2_numeric == 1] <- "Protestant"
+dataset_new$Religion[dataset_new$Q58R3_numeric == 1] <- "Other Christian"
+dataset_new$Religion[dataset_new$Q58R4_numeric == 1] <- "Jewish"
+dataset_new$Religion[dataset_new$Q58R5_numeric == 1] <- "Muslim"
+dataset_new$Religion[dataset_new$Q58R6_numeric == 1] <- "Hindu"
+dataset_new$Religion[dataset_new$Q58R7_numeric == 1] <- "Buddhist"
+dataset_new$Religion[dataset_new$Q58R8_numeric == 1] <- "Mormon/LDS"
+dataset_new$Religion[dataset_new$Q58R9_numeric == 1] <- "Folk Religion"
+dataset_new$Religion[dataset_new$Q58R10_numeric == 1] <- "Daoist"
+dataset_new$Religion[dataset_new$Q58R11_numeric == 1] <- "Shamanism"
+dataset_new$Religion[dataset_new$Q58R12_numeric == 1] <- "Ancestral Religion"
+dataset_new$Religion[dataset_new$Q58R13_numeric == 1] <- "Spiritual but not religious"
+dataset_new$Religion[dataset_new$Q58R14_numeric == 1] <- "Atheist"
+dataset_new$Religion[dataset_new$Q58R15_numeric == 1] <- "Agnostic"
+
+# Convert Religion to a factor
+dataset_new$Religion <- factor(dataset_new$Religion,
+                               levels = c("Catholic", "Protestant", "Other Christian", "Jewish", "Muslim",
+                                          "Hindu", "Buddhist", "Mormon/LDS", "Folk Religion",
+                                          "Daoist", "Shamanism", "Ancestral Religion",
+                                          "Spiritual but not religious", "Atheist", "Agnostic", "None"))
+
+# Create a table of counts for the new Religion variable
+table(dataset_new$Religion)
+
+```
+
+Removing depracated Religion variables
+```{r}
+# Remove redundant religion columns (Q58R1 to Q58R15 and their numeric versions)
+redundant_cols <- c(
+  "Q58R1", "Q58R2", "Q58R3", "Q58R4", "Q58R5", "Q58R6", "Q58R7", "Q58R8", 
+  "Q58R9", "Q58R10", "Q58R11", "Q58R12", "Q58R13", "Q58R14", "Q58R15",
+  "Q58R1_numeric", "Q58R2_numeric", "Q58R3_numeric", "Q58R4_numeric", 
+  "Q58R5_numeric", "Q58R6_numeric", "Q58R7_numeric", "Q58R8_numeric", 
+  "Q58R9_numeric", "Q58R10_numeric", "Q58R11_numeric", "Q58R12_numeric", 
+  "Q58R13_numeric", "Q58R14_numeric", "Q58R15_numeric"
+)
+
+# Remove redundant columns from the dataset
+dataset_clean <- dataset_new[, !(names(dataset_new) %in% redundant_cols)]
+
+# Take a look at the updated dataset
+head(dataset_clean)
+```
+
+Cleaning and Recoding Other Variables
+
 # Analysis of Missing Data
+```{r}
+
+```
 
 # Save cleaned data
+```{r}
+
+```
 
 # Creating summary table
+```{r}
 
-# Running chi-square tests
-
-# Fitting OLR Model
-
-# Fitting MLR Model
+```

### 2024-12-03 peter_EDA.qmd
updating EDA

--- 2024-12-02 peter_EDA.qmd
@@
 
+# Packages
+
+```{r}
+library(tidyverse)
+library(dplyr)
+```
+
 # Loading data
+
+```{r, echo=FALSE}
+rm(list=ls())
+file_path <- "/Users/pdeguz01/Documents/git/Data/2020 CMPS/DS0001/39096-0001-Data.rda"
+# Load data 
+load(file_path)
+data <- da39096.0001
+rm(da39096.0001)
+```
 
@@
 
+Need to include the outcome variable, which is an ordinal variable Q29. Main predictor variable is Q271 "How important is being \[RACIAL GROUP\] to your identity?"
+
+```{r}
+required_cols <- c("Q29", # Outcome Variable; interest in politics
+                   "Q271", # Predictor Variable; importance of race to identity
+                   "S2_RACE_PRIME", # Race
+                   "S3", # Orientation
+                   "S3B", # Gender
+                   "S5_AGE", # Age Category
+                   "S13", # Educational Attainment
+                   "S14", # Community Type
+                   "S7", # Nativity
+                   "Q21", # Party ID
+                   "S2_M", # Multiracial
+
+                   # Religion-related variables
+                   "Q58R1",            # Catholic
+                   "Q58R2",            # Protestant
+                   "Q58R3",            # Other Christian
+                   "Q58R4",            # Jewish
+                   "Q58R5",            # Muslim
+                   "Q58R6",            # Hindu
+                   "Q58R7",            # Buddhist
+                   "Q58R8",            # Mormon/LDS
+                   "Q58R9",            # Folk Religion
+                   "Q58R10",           # Daoist
+                   "Q58R11",           # Shamanism
+                   "Q58R12",           # Ancestral Religion
+                   "Q58R13",           # Spiritual but not religious
+                   "Q58R14",           # Atheist
+                   "Q58R15",            # Agnostic
+
+                   # Variables capturing experiences with racism
+                   "Q627", # Treated unfairly or discriminated against ,
+                   "Q628", # Location of discrimination
+                   "Q629R1" # Unfair treatment due to racial background/ethnicity
+                   )
+
+# Create new df
+data_filtered <- data[, required_cols]
+head(data_filtered)
+
+```
+
 # Cleaning Variables
+
+```{r}
+# Convert the required columns to factors
+data_filtered[required_cols] <- lapply(data_filtered[required_cols], as.factor)
+
+# Verify that the columns have been successfully converted to factors
+str(data_filtered[required_cols])
+head(data_filtered)
+```
+
+Cleaning and Recoding Religion Variables
+
+```{r}
+# Convert individual religion columns to numeric
+for (i in 1:15) {
+  col_name <- paste0("Q58R", i, "_numeric")
+  data_filtered[[col_name]] <- as.numeric(sub("^\\((\\d+)\\).*", "\\1", data_filtered[[paste0("Q58R", i)]]))
+}
+
+# Create a new Religion variable
+data_filtered$Religion <- "None"
+data_filtered$Religion[data_filtered$Q58R1_numeric == 1] <- "Catholic"
+data_filtered$Religion[data_filtered$Q58R2_numeric == 1] <- "Protestant"
+data_filtered$Religion[data_filtered$Q58R3_numeric == 1] <- "Other Christian"
+data_filtered$Religion[data_filtered$Q58R4_numeric == 1] <- "Jewish"
+data_filtered$Religion[data_filtered$Q58R5_numeric == 1] <- "Muslim"
+data_filtered$Religion[data_filtered$Q58R6_numeric == 1] <- "Hindu"
+data_filtered$Religion[data_filtered$Q58R7_numeric == 1] <- "Buddhist"
+data_filtered$Religion[data_filtered$Q58R8_numeric == 1] <- "Mormon/LDS"
+data_filtered$Religion[data_filtered$Q58R9_numeric == 1] <- "Folk Religion"
+data_filtered$Religion[data_filtered$Q58R10_numeric == 1] <- "Daoist"
+data_filtered$Religion[data_filtered$Q58R11_numeric == 1] <- "Shamanism"
+data_filtered$Religion[data_filtered$Q58R12_numeric == 1] <- "Ancestral Religion"
+data_filtered$Religion[data_filtered$Q58R13_numeric == 1] <- "Spiritual but not religious"
+data_filtered$Religion[data_filtered$Q58R14_numeric == 1] <- "Atheist"
+data_filtered$Religion[data_filtered$Q58R15_numeric == 1] <- "Agnostic"
+
+# Convert Religion to a factor
+data_filtered$Religion <- factor(data_filtered$Religion,
+                               levels = c("Catholic", "Protestant", "Other Christian", "Jewish", "Muslim",
+                                          "Hindu", "Buddhist", "Mormon/LDS", "Folk Religion",
+                                          "Daoist", "Shamanism", "Ancestral Religion",
+                                          "Spiritual but not religious", "Atheist", "Agnostic", "None"))
+
+# Create a table of counts for the new Religion variable
+table(data_filtered$Religion)
+
+```
+
+Removing deprecated Religion variables
+
+```{r}
+# Remove redundant religion columns (Q58R1 to Q58R15 and their numeric versions)
+redundant_cols <- c(
+  "Q58R1", "Q58R2", "Q58R3", "Q58R4", "Q58R5", "Q58R6", "Q58R7", "Q58R8", 
+  "Q58R9", "Q58R10", "Q58R11", "Q58R12", "Q58R13", "Q58R14", "Q58R15",
+  "Q58R1_numeric", "Q58R2_numeric", "Q58R3_numeric", "Q58R4_numeric", 
+  "Q58R5_numeric", "Q58R6_numeric", "Q58R7_numeric", "Q58R8_numeric", 
+  "Q58R9_numeric", "Q58R10_numeric", "Q58R11_numeric", "Q58R12_numeric", 
+  "Q58R13_numeric", "Q58R14_numeric", "Q58R15_numeric"
+)
+
+# Remove redundant columns from the dataset
+dataset_clean <- data_filtered[, !(names(data_filtered) %in% redundant_cols)]
+
+# Take a look at the updated dataset
+head(dataset_clean)
+```
+
+Cleaning and Recoding Other Variables
+
+```{r}
+dataset_clean <- dataset_clean %>% mutate(educ = case_when(S13 == "(1) Grades 1-8" ~ "No college experience",
+                                                           S13 == "(2) Some High School, but did not graduate" ~ "No college experience",
+                                                           S13 == "(3) High School graduate or GED" ~ "No college experience",
+                                                           S13 == "(4) Some college" ~ "Some college experience",
+                                                           S13 == "(5) Associates, 2-year degree" ~ "College graduate",
+                                                           S13 == "(6) Bachelors, 4-year degree" ~ "College graduate",
+                                                           S13 == "(7) Post-graduate degree" ~ "College graduate"),
+                                          sex_orientation = case_when(S3 == "(01) Straight, that is, not gay or lesbian" ~ "Straight",
+                                                                     S3 == "(02) Gay or lesbian" ~ "Queer",
+                                                                     S3 == "(03) Bisexual" ~ "Queer",
+                                                                     S3 == "(04) Something else (Specify)" ~ "Queer",
+                                                                     S3 == "(88) Refused" ~ "Refused/Don't Know",
+                                                                     S3 == "(99) I don" ~ "Refused/Don't Know"),
+                                          rural = case_when(S14 == "(1) Large urban area" ~ "Urban/suburban",
+                                                            S14 == "(2) Large suburb near large city" ~ "Urban/suburban",
+                                                            S14 == "(3) Small suburb near small town or city" ~ "Urban/suburban",
+                                                            S14 == "(4) Small town or small city" ~ "Urban/suburban",
+                                                            S14 == "(5) Rural area" ~ "Rural"))
+```
 
@@
 
+```{r}
+# Check the number of missing values for each variable
+missing_summary <- colSums(is.na(dataset_clean))
+missing_summary
+
+# Finding proportion of Missing Data
+missing_proportion <- colSums(is.na(dataset_clean)) / nrow(dataset_clean)
+missing_proportion
+
+# Finding total missing data proportion
+overall_missing <- sum(is.na(dataset_clean)) / (nrow(dataset_clean) * ncol(dataset_clean))
+overall_missing
+```
+
+There is overall 8.3% missing values across all cells in this dataset. This is largely driven by two variables:
+
+-   Q628: (IF S7=2 or 3 & Q627=1\] Did you experience that discrimination here in the United States, or in your home country?)
+
+-   and Q629R1: In your opinion, were you unfairly treated because of your (Racial background or ethnicity)
+
+For Q628, there are 15,709 (89.53%) missing values.
+For Q629R1 there are 9,180 (52.32%) missing values.
+We may need to consider not including these variables, because they may reduce our usable sample size.
+
 # Save cleaned data
+
+```{r}
+#save(dataset_clean, file = "/Users/pdeguz01/Documents/git/Data/2020 CMPS/Output/2020cmps_clean.RData")
+
+```
+
+# Looking at Outcome and main Predictor variable
+
+```{r}
+# Outcome - Q29
+table(dataset_clean$Q29) / nrow(dataset_clean)
+# Main Predictor - Q271
+table(dataset_clean$Q271) / nrow(dataset_clean)
+```
+
+Table of Q29 for full sample - Interest in Politics
+
+| Level of Interest in Politics     | Percent of Total Respondents  |
+|-----------------------------------|-------------------------------|
+| Very interested in politics       | 24.54%                        |
+| Somewhat interested               | 43.39%                        |
+| Not that interested in politics   | 20.76%                        |
+| Not at all interested in politics | 11.31%                        |
+
+Table of Q271 for full sample - How important is being a member of \[RACIAL GROUP\] to your identity?
+
+| Importance of Race to Identity | Percent of Total Respondents |
+|--------------------------------|------------------------------|
+| Extremely Important            | 31.27%                       |
+| Very Important                 | 24.68%                       |
+| Moderately Important           | 21.40%                       |
+| Slightly Important             | 8.58%                        |
+| Not at all important           | 14.07%                       |
+
+# Counts and proportion for every variable
+
+```{r}
+lapply(dataset_clean[, -1], table)
+
+```
 
@@
 
-# Running chi-square tests
+```{r}
 
-# Fitting OLR Model
+# Function to create a summary table with counts and proportions
+create_summary_table <- function(variable, var_name) {
+  counts <- table(variable)
+  if (length(counts) == 0) {
+    return(data.frame(Variable = character(), Level = character(), Count = numeric(), Proportion = numeric()))
+  }
+  proportions <- prop.table(counts)
+  summary_df <- data.frame(
+    Variable = var_name,
+    Level = names(counts),
+    Count = as.numeric(counts),
+    Proportion = round(as.numeric(proportions), 4)
+  )
+  return(summary_df)
+}
 
-# Fitting MLR Model
+# Apply the function to each variable of interest in the cleaned dataset
+categorical_vars <- names(dataset_clean)
+
+# Initialize an empty data frame to store all summary tables
+full_summary_table <- data.frame(Variable = character(), Level = character(), Count = numeric(), Proportion = numeric())
+
+# Iterate through each variable and generate the summary table
+for (var in categorical_vars) {
+  summary_table <- create_summary_table(dataset_clean[[var]], var)
+  full_summary_table <- rbind(full_summary_table, summary_table)
+}
+
+# Print the full summary table
+full_summary_table
+
+```
+
+| **Variable** | **Level** | **Count** | **Proportion** |
+|------------------|-------------------|------------------|------------------|
+| Q29 | \(1\) Very interested in politics | 4306 | 25% |
+| Q29 | \(2\) Somewhat interested | 7612 | 43% |
+| Q29 | \(3\) Not that interested in politics | 3643 | 21% |
+| Q29 | \(4\) Not at all interested in politics | 1984 | 11% |
+| Q271 | \(1\) Extremely important | 5487 | 31% |
+| Q271 | \(2\) Very important | 4330 | 25% |
+| Q271 | \(3\) Moderately important | 3754 | 21% |
+| Q271 | \(4\) Slightly important | 1505 | 9% |
+| Q271 | \(5\) Not at all important | 2469 | 14% |
+| S2_RACE_PRIME | \(1\) White | 3923 | 22% |
+| S2_RACE_PRIME | \(2\) Hispanic or Latino | 3873 | 22% |
+| S2_RACE_PRIME | \(3\) Black or African American | 4613 | 26% |
+| S2_RACE_PRIME | \(4\) Asian American | 3826 | 22% |
+| S2_RACE_PRIME | \(5\) American Indian/Native American | 713 | 4% |
+| S2_RACE_PRIME | \(6\) Arab, Middle Eastern or North African | 226 | 1% |
+| S2_RACE_PRIME | \(7\) Native Hawaiian | 106 | 1% |
+| S2_RACE_PRIME | \(8\) Not Hawaiian, but other Pacific Islander | 255 | 1% |
+| S3 | \(01\) Straight, that is, not gay or lesbian | 15404 | 88% |
+| S3 | \(02\) Gay or lesbian | 594 | 3% |
+| S3 | \(03\) Bisexual | 937 | 5% |
+| S3 | \(04\) Something else (Specify) | 286 | 2% |
+| S3 | \(88\) Refused | 195 | 1% |
+| S3 | \(99\) I don | 129 | 1% |
+| S3B | \(1\) Man | 7431 | 42% |
+| S3B | \(2\) Woman | 9932 | 57% |
+| S3B | \(3\) Non-binary | 148 | 1% |
+| S3B | \(4\) Something else (Specify) | 34 | 0% |
+| S5_AGE | \(2\) 18-29 | 4026 | 23% |
+| S5_AGE | \(3\) 30-39 | 3614 | 21% |
+| S5_AGE | \(4\) 40-49 | 2783 | 16% |
+| S5_AGE | \(5\) 50-59 | 2717 | 15% |
+| S5_AGE | \(6\) 60-69 | 2745 | 16% |
+| S5_AGE | \(7\) 70 + | 1660 | 9% |
+| S13 | \(1\) Grades 1-8 | 126 | 1% |
+| S13 | \(2\) Some High School, but did not graduate | 596 | 3% |
+| S13 | \(3\) High School graduate or GED | 3227 | 18% |
+| S13 | \(4\) Some college | 3689 | 21% |
+| S13 | \(5\) Associates, 2-year degree | 2017 | 12% |
+| S13 | \(6\) Bachelors, 4-year degree | 4625 | 26% |
+| S13 | \(7\) Post-graduate degree | 3265 | 19% |
+| S7 | \(1\) United States | 13297 | 76% |
+| S7 | \(2\) Another country | 3962 | 23% |
+| S7 | \(3\) on the Island of Puerto Rico | 286 | 2% |
+| Q21 | \(1\) Republican | 3238 | 18% |
+| Q21 | \(2\) Democrat | 8318 | 47% |
+| Q21 | \(3\) Independent | 5174 | 29% |
+| Q21 | \(4\) Other party | 815 | 5% |
+| S2_M | \(1\) Yes | 4671 | 27% |
+| S2_M | \(2\) No | 12874 | 73% |
+| Q627 | \(1\) Yes | 8365 | 48% |
+| Q627 | \(2\) No | 9180 | 52% |
+| Q628 | \(1\) In the United States | 1404 | 76% |
+| Q628 | \(2\) In my country of origin | 153 | 8% |
+| Q628 | \(3\) In both places | 279 | 15% |
+| Q629R1 | \(0\) NO TO: Racial background or ethnicity | 2947 | 35% |
+| Q629R1 | \(1\) Racial background or ethnicity | 5418 | 65% |
+| Religion | Catholic | 3702 | 21% |
+| Religion | Protestant | 4045 | 23% |
+| Religion | Other Christian | 1826 | 10% |
+| Religion | Jewish | 264 | 2% |
+| Religion | Muslim | 524 | 3% |
+| Religion | Hindu | 491 | 3% |
+| Religion | Buddhist | 464 | 3% |
+| Religion | Mormon/LDS | 112 | 1% |
+| Religion | Folk Religion | 27 | 0% |
+| Religion | Daoist | 21 | 0% |
+| Religion | Shamanism | 26 | 0% |
+| Religion | Ancestral Religion | 77 | 0% |
+| Religion | Spiritual but not religious | 1288 | 7% |
+| Religion | Atheist | 820 | 5% |
+| Religion | Agnostic | 1084 | 6% |
+| Religion | None | 2774 | 16% |
+| educ | College graduate | 9907 | 56% |
+| educ | No college experience | 3949 | 23% |
+| educ | Some college experience | 3689 | 21% |
+| sex_orientaton | Queer | 1817 | 10% |
+| sex_orientation | Refused/Don't Know | 324 | 2% |
+| sex_orientation | Straight | 15404 | 88% |
+| rural | Rural | 1453 | 8% |
+| rural | Urban/suburban | 16092 | 92% |
+|  |  |  |  |
+|  |  |  |  |
+|  |  |  |  |
+|  |  |  |  |
+|  |  |  |  |
+|  |  |  |  |
+|  |  |  |  |
+|  |  |  |  |

### 2024-12-03 peter_EDA.qmd
updating recoding script

@@
                                                             S14 == "(4) Small town or small city" ~ "Urban/suburban",
-                                                            S14 == "(5) Rural area" ~ "Rural"))
+                                                            S14 == "(5) Rural area" ~ "Rural"),
+                                          white = case_when(S2_RACE_PRIME == "(1) White" ~ "White", 
+                                                            S2_RACE_PRIME == "(2) Hispanic or Latino" ~ "Non-White",
+                                                            S2_RACE_PRIME == "(3) Black or African American" ~ "Non-White",
+                                                            S2_RACE_PRIME == "(4) Asian American" ~ "Non-White",
+                                                            S2_RACE_PRIME == "(5) American Indian/Native American" ~ "Non-White",
+                                                            S2_RACE_PRIME == "(6) Arab, Middle Eastern or North African" ~ "Non-White",
+                                                            S2_RACE_PRIME == "(7) Native Hawaiian" ~ "Non-White",
+                                                            S2_RACE_PRIME == "(8) Not Hawaiian, but other Pacific Islander" ~ "Non-White"),
+                                          race_recoded = case_when(S2_RACE_PRIME == "(1) White" ~ "White",
+                                                                   S2_RACE_PRIME == "(2) Hispanic or Latino" ~ "Latino",
+                                                                   S2_RACE_PRIME == "(3) Black or African American" ~ "Black",
+                                                                   S2_RACE_PRIME == "(4) Asian American" ~ "Asian American",
+                                                                   S2_RACE_PRIME == "(5) American Indian/Native American" ~ "AIAN",
+                                                                   S2_RACE_PRIME == "(6) Arab, Middle Eastern or North African" ~ "Arab/MENA",
+                                                                   S2_RACE_PRIME == "(7) Native Hawaiian" ~ "NHPI",
+                                                                   S2_RACE_PRIME == "(8) Not Hawaiian, but other Pacific Islander" ~ "NHPI"))
+
 ```
@@
 ```{r}
-#save(dataset_clean, file = "/Users/pdeguz01/Documents/git/Data/2020 CMPS/Output/2020cmps_clean.RData")
+save(dataset_clean, file = "/Users/pdeguz01/Documents/git/Data/2020 CMPS/Output/2020cmps_clean.RData")
 
@@
 | Level of Interest in Politics     | Percent of Total Respondents |
-|-----------------------------------|-------------------------------|
+|-----------------------------------|------------------------------|
 | Very interested in politics       | 24.54%                       |

### 2024-12-03 peter_plots.qmd
updating recoding script

@@
 ## 
+# Packages
+```{r}
+library(tidyverse)
+library(janitor)
+library(ggplot2)
+library(scales)
+```
 
 # Loading data
+```{r}
+file_path = "/Users/pdeguz01/Documents/git/Data/2020 CMPS/Output/2020cmps_clean.RData"
+load(file_path)
+```
 
 # plot 1:
+Race and importance of being a member of racial group to identity (Q271)
+```{r}
+
+data_summary <- dataset_clean %>%
+  group_by(S2_RACE_PRIME, Q271) %>%
+  summarise(count = n()) %>%
+  mutate(proportion = count / sum(count) * 100)  # Calculate proportion for each race
+
+races <- unique(data_summary$S2_RACE_PRIME)
+
+# Loop through each race and generate a plot
+for (race_value in races) {
+  plot_data <- data_summary %>% filter(S2_RACE_PRIME == race_value)
+
+  p <- ggplot(plot_data, aes(x = Q271, y = proportion, fill = Q271)) +
+    geom_bar(stat = "identity", position = "dodge") +
+    labs(title = paste("Proportions for", race_value, "Respondents"), x = "Importance of Race to Identity", y = "Proportion", fill = "Rating") + 
+    scale_x_discrete(labels = c("1", "2", "3", "4", "5")) +
+    scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentages
+    theme_minimal() +
+    theme(
+      axis.text = element_text(size = 12),  # Customize axis text size
+      plot.title = element_text(size = 14, hjust = 0.5)  # Center title
+    )
+
+  # Save each plot as a separate image file (optional)
+  #ggsave(paste0("plot_", race_value, ".png"), plot = p, width = 6, height = 4)
+
+  # Print the plot (optional, if you want to view them in the R console)
+  print(p)
+}
+
+```
 
 # plot 2:
+Race and interest in politics (Q29)
+```{r}
+
+```
 
 # plot 3:
+White/non-white and importance of race to identity (Q271)
+
+# plot:
+White/non-white and interest in politics (Q29)
+
+# plot 3:
+College Experience and importance of being a member of racial group to identity (Q271)
+```{r}
+data_summary <- dataset_clean %>%
+  group_by(educ, Q271) %>%
+  summarise(count = n()) %>%
+  mutate(proportion = count / sum(count) * 100)  # Calculate proportion for each race
+
+plot_somecollege <- data_summary %>% filter(educ == "Some college experience") %>% ggplot(aes(x = Q271, y = proportion, fill = Q271)) +
+    geom_bar(stat = "identity", position = "dodge") +
+    labs(title = paste("Proportions for Respondents with Some college Experience"), x = "Importance of Race to Identity", y = "Proportion", fill = "Rating") + 
+    scale_x_discrete(labels = c("1", "2", "3", "4", "5")) +
+    scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentages
+    theme_minimal() +
+    theme(
+      axis.text = element_text(size = 12),  # Customize axis text size
+      plot.title = element_text(size = 14, hjust = 0.5)  # Center title
+    )
+
+print(plot_somecollege)
+
+plot_nocollege <- data_summary %>% filter(educ == "No college experience") %>% ggplot(aes(x = Q271, y = proportion, fill = Q271)) +
+    geom_bar(stat = "identity", position = "dodge") +
+    labs(title = paste("Proportions for Respondents with No college Experience"), x = "Importance of Race to Identity", y = "Proportion", fill = "Rating") + 
+    scale_x_discrete(labels = c("1", "2", "3", "4", "5")) +
+    scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentages
+    theme_minimal() +
+    theme(
+      axis.text = element_text(size = 12),  # Customize axis text size
+      plot.title = element_text(size = 14, hjust = 0.5)  # Center title
+    )
+
+print(plot_nocollege)
+
+```
+
+# plot 4:
+College Experience and interest in politics (Q29)
+```{r}
+
+```
+
+# plot 4:
+Age and importance of being a member of racial group to identity (Q271)
+```{r}
+
+```
+
+# plot 5:
+Age and interest in politics (Q29)
+```{r}
+
+```
+
+# plot 6:
+Rurality and importance of being a member of racial group to identity (Q271)
+```{r}
+
+```
+
+# plot 7:
+Rurality and interest in politics
+```{r}
+
+```

### 2024-12-03 peter_plots.qmd
updating recoding and plotting

--- 2024-12-02 peter_plots.qmd
@@
 ## 
+# Packages
+```{r}
+library(tidyverse)
+library(janitor)
+library(ggplot2)
+library(scales)
+```
 
 # Loading data
+```{r}
+file_path = "/Users/pdeguz01/Documents/git/Data/2020 CMPS/Output/2020cmps_clean.RData"
+load(file_path)
+```
 
 # plot 1:
+Race and importance of being a member of racial group to identity (Q271)
+```{r}
+
+data_summary <- dataset_clean %>%
+  group_by(race_recoded, Q271) %>%
+  summarise(count = n()) %>%
+  mutate(proportion = count / sum(count) * 100)  # Calculate proportion for each race
+
+races <- unique(data_summary$race_recoded)
+
+# Loop through each race and generate a plot
+for (race_value in races) {
+  plot_data <- data_summary %>% filter(race_recoded == race_value)
+
+  p <- ggplot(plot_data, aes(x = Q271, y = proportion, fill = Q271)) +
+    geom_bar(stat = "identity", position = "dodge") +
+    labs(title = paste("Proportions for", race_value, "Respondents"), x = "Importance of Race to Identity", y = "Proportion", fill = "Rating") + 
+    scale_x_discrete(labels = c("1", "2", "3", "4", "5")) +
+    scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentages
+    geom_text(aes(label = sprintf("%.0f%%", proportion)),  
+              position = position_dodge(width = 0.8),   
+              vjust = 1.5,                               
+              size = 4,                                  
+              color = "white") +  
+    theme_minimal() +
+    theme(
+      axis.text = element_text(size = 12),  # Customize axis text size
+      plot.title = element_text(size = 14, hjust = 0.5)  # Center title
+    )
+
+  # Save each plot as a separate image file (optional)
+  #ggsave(paste0("plot_", race_value, ".png"), plot = p, width = 6, height = 4)
+
+  # Print the plot (optional, if you want to view them in the R console)
+  print(p)
+}
+
+```
 
 # plot 2:
+Race and interest in politics (Q29)
+```{r}
+data_summary <- dataset_clean %>%
+  group_by(race_recoded, Q29) %>%
+  summarise(count = n()) %>%
+  mutate(proportion = count / sum(count) * 100)  # Calculate proportion for each race
+
+races <- unique(data_summary$race_recoded)
+
+# Loop through each race and generate a plot
+for (race_value in races) {
+  plot_data <- data_summary %>% filter(race_recoded == race_value)
+
+  p <- ggplot(plot_data, aes(x = Q29, y = proportion, fill = Q29)) +
+    geom_bar(stat = "identity", position = "dodge") +
+    labs(title = paste("Proportions for", race_value, "Respondents"), x = "Interest in Politics", y = "Proportion", fill = "Rating") + 
+    scale_x_discrete(labels = c("1", "2", "3", "4", "5")) +
+    scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentages
+    geom_text(aes(label = sprintf("%.0f%%", proportion)),  
+              position = position_dodge(width = 0.8),   
+              vjust = 1.5,                               
+              size = 4,                                  
+              color = "white") +  
+    theme_minimal() +
+    theme(
+      axis.text = element_text(size = 12),  # Customize axis text size
+      plot.title = element_text(size = 14, hjust = 0.5)  # Center title
+    )
+
+  # Save each plot as a separate image file (optional)
+  #ggsave(paste0("plot_", race_value, ".png"), plot = p, width = 6, height = 4)
+
+  # Print the plot (optional, if you want to view them in the R console)
+  print(p)
+}
+
+```
 
 # plot 3:
+White/Non-White and importance of race to identity (Q271)
+
+# plot:
+White/Non-White and interest in politics (Q29)
+
+# plot 3:
+College Experience and importance of being a member of racial group to identity (Q271)
+```{r}
+data_summary <- dataset_clean %>%
+  group_by(educ, Q271) %>%
+  summarise(count = n()) %>%
+  mutate(proportion = count / sum(count) * 100)  # Calculate proportion for each education group
+
+plot_collegegrad <- data_summary %>% filter(educ == "College graduate") %>% ggplot(aes(x = Q271, y = proportion, fill = Q271)) +
+    geom_bar(stat = "identity", position = "dodge") +
+    labs(title = paste("Proportions for Respondents that are College Graduates"), x = "Importance of Race to Identity", y = "Proportion", fill = "Rating") + 
+    scale_x_discrete(labels = c("1", "2", "3", "4", "5")) +
+    scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentages
+    geom_text(aes(label = sprintf("%.0f%%", proportion)),  
+              position = position_dodge(width = 0.8),   
+              vjust = 1.5,                               
+              size = 4,                                  
+              color = "white") +   
+    theme_minimal() +
+    theme(
+      axis.text = element_text(size = 12),  # Customize axis text size
+      plot.title = element_text(size = 14, hjust = 0.5)  # Center title
+    )
+
+print(plot_collegegrad)
+
+plot_somecollege <- data_summary %>% filter(educ == "Some college experience") %>% ggplot(aes(x = Q271, y = proportion, fill = Q271)) +
+    geom_bar(stat = "identity", position = "dodge") +
+    labs(title = paste("Proportions for Respondents with Some College Experience"), x = "Importance of Race to Identity", y = "Proportion", fill = "Rating") + 
+    scale_x_discrete(labels = c("1", "2", "3", "4", "5")) +
+    scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentages
+    geom_text(aes(label = sprintf("%.0f%%", proportion)),  
+              position = position_dodge(width = 0.8),   
+              vjust = 1.5,                               
+              size = 4,                                  
+              color = "white") +  
+    theme_minimal() +
+    theme(
+      axis.text = element_text(size = 12),  # Customize axis text size
+      plot.title = element_text(size = 14, hjust = 0.5)  # Center title
+    )
+
+print(plot_somecollege)
+
+plot_nocollege <- data_summary %>% filter(educ == "No college experience") %>% ggplot(aes(x = Q271, y = proportion, fill = Q271)) +
+    geom_bar(stat = "identity", position = "dodge") +
+    labs(title = paste("Proportions for Respondents with No College Experience"), x = "Importance of Race to Identity", y = "Proportion", fill = "Rating") + 
+    scale_x_discrete(labels = c("1", "2", "3", "4", "5")) +
+    scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentages
+    geom_text(aes(label = sprintf("%.0f%%", proportion)),  
+              position = position_dodge(width = 0.8),   
+              vjust = 1.5,                               
+              size = 4,                                  
+              color = "white") +  
+    theme_minimal() +
+    theme(
+      axis.text = element_text(size = 12),  # Customize axis text size
+      plot.title = element_text(size = 14, hjust = 0.5)  # Center title
+    )
+
+print(plot_nocollege)
+
+```
+
+# plot 4:
+College Experience and interest in politics (Q29)
+```{r}
+
+data_summary <- dataset_clean %>%
+  group_by(educ, Q29) %>%
+  summarise(count = n()) %>%
+  mutate(proportion = count / sum(count) * 100)  # Calculate proportion for each education group
+
+plot_collegegrad <- data_summary %>% filter(educ == "College graduate") %>% ggplot(aes(x = Q29, y = proportion, fill = Q29)) +
+    geom_bar(stat = "identity", position = "dodge") +
+    labs(title = paste("Proportions for Respondents that are College Graduates"), x = "Interest in Politics", y = "Proportion", fill = "Rating") + 
+    scale_x_discrete(labels = c("1", "2", "3", "4", "5")) +
+    scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentages
+    geom_text(aes(label = sprintf("%.0f%%", proportion)),  
+              position = position_dodge(width = 0.8),   
+              vjust = 1.5,                               
+              size = 4,                                  
+              color = "white") +   
+    theme_minimal() +
+    theme(
+      axis.text = element_text(size = 12),  # Customize axis text size
+      plot.title = element_text(size = 14, hjust = 0.5)  # Center title
+    )
+
+print(plot_collegegrad)
+
+plot_somecollege <- data_summary %>% filter(educ == "Some college experience") %>% ggplot(aes(x = Q29, y = proportion, fill = Q29)) +
+    geom_bar(stat = "identity", position = "dodge") +
+    labs(title = paste("Proportions for Respondents with Some College Experience"), x = "Interest in Politics", y = "Proportion", fill = "Rating") + 
+    scale_x_discrete(labels = c("1", "2", "3", "4", "5")) +
+    scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentages
+    geom_text(aes(label = sprintf("%.0f%%", proportion)),  
+              position = position_dodge(width = 0.8),   
+              vjust = 1.5,                               
+              size = 4,                                  
+              color = "white") +  
+    theme_minimal() +
+    theme(
+      axis.text = element_text(size = 12),  # Customize axis text size
+      plot.title = element_text(size = 14, hjust = 0.5)  # Center title
+    )
+
+print(plot_somecollege)
+
+plot_nocollege <- data_summary %>% filter(educ == "No college experience") %>% ggplot(aes(x = Q29, y = proportion, fill = Q29)) +
+    geom_bar(stat = "identity", position = "dodge") +
+    labs(title = paste("Proportions for Respondents with No College Experience"), x = "Interest in Politics", y = "Proportion", fill = "Rating") + 
+    scale_x_discrete(labels = c("1", "2", "3", "4", "5")) +
+    scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentages
+    geom_text(aes(label = sprintf("%.0f%%", proportion)),  
+              position = position_dodge(width = 0.8),   
+              vjust = 1.5,                               
+              size = 4,                                  
+              color = "white") +  
+    theme_minimal() +
+    theme(
+      axis.text = element_text(size = 12),  # Customize axis text size
+      plot.title = element_text(size = 14, hjust = 0.5)  # Center title
+    )
+
+print(plot_nocollege)
+
+```
+
+# plot 4:
+Age and importance of being a member of racial group to identity (Q271)
+```{r}
+data_summary <- dataset_clean %>%
+  group_by(S5_AGE, Q271) %>%
+  summarise(count = n()) %>%
+  mutate(proportion = count / sum(count) * 100)  # Calculate proportion for each race
+
+age_groups <- unique(data_summary$S5_AGE)
+
+# Loop through each age group and generate a plot
+for (age_value in age_groups) {
+  plot_data <- data_summary %>% filter(S5_AGE == age_value)
+
+  p <- ggplot(plot_data, aes(x = Q271, y = proportion, fill = Q271)) +
+    geom_bar(stat = "identity", position = "dodge") +
+    labs(title = paste("Proportions for Respondents Ages", age_value), x = "Importance of Race to Identity", y = "Proportion", fill = "Rating") + 
+    scale_x_discrete(labels = c("1", "2", "3", "4", "5")) +
+    scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentages
+    geom_text(aes(label = sprintf("%.0f%%", proportion)),  
+              position = position_dodge(width = 0.8),   
+              vjust = 1.5,                               
+              size = 4,                                  
+              color = "white") +  
+    theme_minimal() +
+    theme(
+      axis.text = element_text(size = 12),  # Customize axis text size
+      plot.title = element_text(size = 14, hjust = 0.5)  # Center title
+    )
+
+  # Save each plot as a separate image file (optional)
+  #ggsave(paste0("plot_", race_value, ".png"), plot = p, width = 6, height = 4)
+
+  # Print the plot (optional, if you want to view them in the R console)
+  print(p)
+}
+```
+
+# plot 5:
+Age and interest in politics (Q29)
+```{r}
+data_summary <- dataset_clean %>%
+  group_by(S5_AGE, Q29) %>%
+  summarise(count = n()) %>%
+  mutate(proportion = count / sum(count) * 100)  # Calculate proportion for each race
+
+age_groups <- unique(data_summary$S5_AGE)
+
+# Loop through each age group and generate a plot
+for (age_value in age_groups) {
+  plot_data <- data_summary %>% filter(S5_AGE == age_value)
+
+  p <- ggplot(plot_data, aes(x = Q29, y = proportion, fill = Q29)) +
+    geom_bar(stat = "identity", position = "dodge") +
+    labs(title = paste("Proportions for Respondents Ages", age_value), x = "Importance of Race to Identity", y = "Proportion", fill = "Rating") + 
+    scale_x_discrete(labels = c("1", "2", "3", "4", "5")) +
+    scale_y_continuous(labels = scales::percent_format(scale = 1)) +  
+    geom_text(aes(label = sprintf("%.0f%%", proportion)),  
+              position = position_dodge(width = 0.8),   
+              vjust = 1.5,                               
+              size = 4,                                  
+              color = "white") +  
+    theme_minimal() +
+    theme(
+      axis.text = element_text(size = 12), 
+      plot.title = element_text(size = 14, hjust = 0.5) 
+    )
+
+  # Save each plot as a separate image file 
+  #ggsave(paste0("plot_", race_value, ".png"), plot = p, width = 6, height = 4)
+
+  # Print the plot 
+  print(p)
+}
+
+```
+
+# plot 6:
+Rurality and importance of being a member of racial group to identity (Q271)
+```{r}
+
+```
+
+# plot 7:
+Rurality and interest in politics (Q29)
+```{r}
+
+data_summary <- dataset_clean %>%
+  group_by(rural, Q29) %>%
+  summarise(count = n()) %>%
+  mutate(proportion = count / sum(count) * 100)  # Calculate proportion for each race
+
+rural_groups <- unique(data_summary$rural)
+
+# Loop through each age group and generate a plot
+for (rural_value in rural_groups) {
+  plot_data <- data_summary %>% filter(rural == rural_value)
+
+  p <- ggplot(plot_data, aes(x = Q29, y = proportion, fill = Q29)) +
+    geom_bar(stat = "identity", position = "dodge") +
+    labs(title = paste("Proportions for Respondents in", rural_value, "Areas"), x = "Importance of Race to Identity", y = "Proportion", fill = "Rating") + 
+    scale_x_discrete(labels = c("1", "2", "3", "4", "5")) +
+    scale_y_continuous(labels = scales::percent_format(scale = 1)) +  
+    geom_text(aes(label = sprintf("%.0f%%", proportion)),  
+              position = position_dodge(width = 0.8),   
+              vjust = 1.5,                               
+              size = 4,                                  
+              color = "white") +  
+    theme_minimal() +
+    theme(
+      axis.text = element_text(size = 12), 
+      plot.title = element_text(size = 14, hjust = 0.5) 
+    )
+
+  # Save each plot as a separate image file 
+  #ggsave(paste0("plot_", race_value, ".png"), plot = p, width = 6, height = 4)
+
+  # Print the plot 
+  print(p)
+}
+
+```

### 2024-12-04 peter_plots.qmd
updating plots

@@
   # Save each plot as a separate image file (optional)
-  #ggsave(paste0("plot_", race_value, ".png"), plot = p, width = 6, height = 4)
+  #ggsave(paste0("plot_", race_value, "Q271.png"), plot = p, width = 6, height = 4)
 
@@
   # Save each plot as a separate image file (optional)
-  #ggsave(paste0("plot_", race_value, ".png"), plot = p, width = 6, height = 4)
+  #ggsave(paste0("plot_", race_value, "Q29.png"), plot = p, width = 6, height = 4)
 
@@
 White/Non-White and importance of race to identity (Q271)
+```{r}
+
+data_summary <- dataset_clean %>%
+  group_by(white, Q271) %>%
+  summarise(count = n()) %>%
+  mutate(proportion = count / sum(count) * 100)  # Calculate proportion for each education group
+
+plot_whiteQ271 <- data_summary %>% filter(white == "White") %>% ggplot(aes(x = Q271, y = proportion, fill = Q271)) +
+    geom_bar(stat = "identity", position = "dodge") +
+    labs(title = paste("Proportions for White Respondents"), x = "Importance of Race to Identity", y = "Proportion", fill = "Rating") + 
+    scale_x_discrete(labels = c("1", "2", "3", "4", "5")) +
+    scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentages
+    geom_text(aes(label = sprintf("%.0f%%", proportion)),  
+              position = position_dodge(width = 0.8),   
+              vjust = 1.5,                               
+              size = 4,                                  
+              color = "white") +   
+    theme_minimal() +
+    theme(
+      axis.text = element_text(size = 12),  # Customize axis text size
+      plot.title = element_text(size = 14, hjust = 0.5)  # Center title
+    )
+
+print(plot_whiteQ271)
+
+plot_nonwhiteQ271 <- data_summary %>% filter(white == "Non-White") %>% ggplot(aes(x = Q271, y = proportion, fill = Q271)) +
+    geom_bar(stat = "identity", position = "dodge") +
+    labs(title = paste("Proportions for Non-White Respondents"), x = "Importance of Race to Identity", y = "Proportion", fill = "Rating") + 
+    scale_x_discrete(labels = c("1", "2", "3", "4", "5")) +
+    scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentages
+    geom_text(aes(label = sprintf("%.0f%%", proportion)),  
+              position = position_dodge(width = 0.8),   
+              vjust = 1.5,                               
+              size = 4,                                  
+              color = "white") +   
+    theme_minimal() +
+    theme(
+      axis.text = element_text(size = 12),  # Customize axis text size
+      plot.title = element_text(size = 14, hjust = 0.5)  # Center title
+    )
+
+print(plot_nonwhiteQ271)
+
+```
 
@@
 White/Non-White and interest in politics (Q29)
+```{r}
+data_summary <- dataset_clean %>%
+  group_by(white, Q29) %>%
+  summarise(count = n()) %>%
+  mutate(proportion = count / sum(count) * 100)  # Calculate proportion for each education group
+
+plot_whiteQ29 <- data_summary %>% filter(white == "White") %>% ggplot(aes(x = Q29, y = proportion, fill = Q29)) +
+    geom_bar(stat = "identity", position = "dodge") +
+    labs(title = paste("Proportions for White Respondents"), x = "Interest in Politics", y = "Proportion", fill = "Rating") + 
+    scale_x_discrete(labels = c("1", "2", "3", "4", "5")) +
+    scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentages
+    geom_text(aes(label = sprintf("%.0f%%", proportion)),  
+              position = position_dodge(width = 0.8),   
+              vjust = 1.5,                               
+              size = 4,                                  
+              color = "white") +   
+    theme_minimal() +
+    theme(
+      axis.text = element_text(size = 12),  # Customize axis text size
+      plot.title = element_text(size = 14, hjust = 0.5)  # Center title
+    )
+
+print(plot_whiteQ29)
+
+plot_nonwhiteQ29 <- data_summary %>% filter(white == "Non-White") %>% ggplot(aes(x = Q29, y = proportion, fill = Q29)) +
+    geom_bar(stat = "identity", position = "dodge") +
+    labs(title = paste("Proportions for Non-White Respondents"), x = "Interest in Politics", y = "Proportion", fill = "Rating") + 
+    scale_x_discrete(labels = c("1", "2", "3", "4", "5")) +
+    scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentages
+    geom_text(aes(label = sprintf("%.0f%%", proportion)),  
+              position = position_dodge(width = 0.8),   
+              vjust = 1.5,                               
+              size = 4,                                  
+              color = "white") +   
+    theme_minimal() +
+    theme(
+      axis.text = element_text(size = 12),  # Customize axis text size
+      plot.title = element_text(size = 14, hjust = 0.5)  # Center title
+    )
+
+print(plot_nonwhiteQ29)
+
+```
 
@@
 
-print(plot_collegegrad)
+print(plot_collegegradQ271)
 
@@
 
-print(plot_somecollege)
+print(plot_somecollegeQ271)
 
@@
 
-print(plot_nocollege)
+print(plot_nocollegeQ271)
 
@@
 
-print(plot_collegegrad)
+print(plot_collegegradQ29)
 
@@
 
-print(plot_somecollege)
+print(plot_somecollegeQ29)
 
@@
 
-print(plot_nocollege)
+print(plot_nocollegeQ29)
 
@@
   # Save each plot as a separate image file (optional)
-  #ggsave(paste0("plot_", race_value, ".png"), plot = p, width = 6, height = 4)
+  #ggsave(paste0("plot_", age_value, "Q271.png"), plot = p, width = 6, height = 4)
 
@@
   # Save each plot as a separate image file 
-  #ggsave(paste0("plot_", race_value, ".png"), plot = p, width = 6, height = 4)
+  #ggsave(paste0("plot_", age_value, "Q29.png"), plot = p, width = 6, height = 4)
 
@@
 ```{r}
+data_summary <- dataset_clean %>%
+  group_by(rural, Q271) %>%
+  summarise(count = n()) %>%
+  mutate(proportion = count / sum(count) * 100)  # Calculate proportion for each race
 
+rural_groups <- unique(data_summary$rural)
+
+# Loop through each age group and generate a plot
+for (rural_value in rural_groups) {
+  plot_data <- data_summary %>% filter(rural == rural_value)
+
+  p <- ggplot(plot_data, aes(x = Q271, y = proportion, fill = Q271)) +
+    geom_bar(stat = "identity", position = "dodge") +
+    labs(title = paste("Proportions for Respondents in", rural_value, "Areas"), x = "Importance of Race to Identity", y = "Proportion", fill = "Rating") + 
+    scale_x_discrete(labels = c("1", "2", "3", "4", "5")) +
+    scale_y_continuous(labels = scales::percent_format(scale = 1)) +  
+    geom_text(aes(label = sprintf("%.0f%%", proportion)),  
+              position = position_dodge(width = 0.8),   
+              vjust = 1.5,                               
+              size = 4,                                  
+              color = "white") +  
+    theme_minimal() +
+    theme(
+      axis.text = element_text(size = 12), 
+      plot.title = element_text(size = 14, hjust = 0.5) 
+    )
+
+  # Save each plot as a separate image file 
+  #ggsave(paste0("plot_", rural_value, "Q271.png"), plot = p, width = 6, height = 4)
+
+  # Print the plot 
+  print(p)
+}
 ```
@@
     geom_bar(stat = "identity", position = "dodge") +
-    labs(title = paste("Proportions for Respondents in", rural_value, "Areas"), x = "Importance of Race to Identity", y = "Proportion", fill = "Rating") + 
+    labs(title = paste("Proportions for Respondents in", rural_value, "Areas"), x = "Interest in Politics", y = "Proportion", fill = "Rating") + 
     scale_x_discrete(labels = c("1", "2", "3", "4", "5")) +
@@
   # Save each plot as a separate image file 
-  #ggsave(paste0("plot_", race_value, ".png"), plot = p, width = 6, height = 4)
+  #ggsave(paste0("plot_", rural_value, "Q29.png"), plot = p, width = 6, height = 4)
 

### 2024-12-06 Naki_Main.qmd
new files

---
title: "Naki_Main"
format: html
editor: visual
---

```{r}

# Load the packages
library(ggplot2)
library(broom)
library(dplyr)
library(car)
library(caret)
```

# Introduction

The Collaborative Multi-Racial Post-Election Survey (CMPS), United States, 2020 is an extensive dataset designed to capture diverse perspectives across racial and ethnic groups in the United States. Respondents were asked to address a wide range of issues, including political attitudes, voting behavior, social concerns, and policy preferences.

# Preliminary Cleaning

## Loading the Dataset

```{r}
file_path <- "/Users/nakiyahdhariwala/StatsProject/39096-0001-Data.rda"

load(file_path)          # Load the RDA file
ls()                     # List objects loaded into the environment
dataset <- da39096.0001  # Access the dataset

head(dataset)
ncol(dataset)
nrow(dataset)
```

The dataset contains 17,545 rows, with each row representing an individual respondent's data, and 1,490 columns, each capturing a specific variable or question response. These variables include demographic information, opinions on political and social issues, and responses to survey questions.

This particular analysis focused on a single topic within the larger dataset. Specifically, our research question is: **How do personal experiences with racism, in combination with demographic factors such as race, gender, and education, influence the importance respondents place on policies aimed at addressing racial and ethnic discrimination, controlling for other social and political factors?**

```{r}
# Calculate proportions
proportions <- prop.table(table(dataset$Q1R5))
proportions
# Define custom labels
custom_labels <- c("Not important", "Important")

# Combine custom labels with percentages
pie_labels <- paste(custom_labels, round(proportions * 100, 2), "%")

# Plot the pie chart with custom labels
pie(proportions,
    labels = pie_labels,
    main = "Q1R5: Stopping discrimination against racial/ethnic minorities",
    col = c("lightblue", "salmon"))
```

The outcome variable Q1R5 (Stopping discrimination against racial/ethnic minorities) for this Research Question is a binary variable which captures whether respondents view stopping racial and ethnic discrimination as an important issue (i.e., Important OR Not Important). Therefore, we will make use Logistic regression as it is suitable for modeling binary outcome variables and allows for the inclusion of both predictor and control variables. While the distribution is skewed (74.25% say "No"), logistic regression is robust to imbalance in the outcome variable. However, extreme imbalance (e.g., 95%-5%) might require additional techniques like resampling or penalized logistic regression.

We can expect the analysis to reveal whether and to what extent experiences with racism predict the importance of racial/ethnic discrimination policy, controlling for demographic factors.

## Extracting relevant columns
```{r}
# Define the required columns
required_cols <- c(
  # Outcome Variable: Stopping discrimination against racial/ethnic minorities
  "Q1R5",

  # Demographic Variables
  "S2_RACE_PRIME",   # Primary Race
  "S3",              # SexualOrientation
  "S3B",             # Gender
  "S5_AGE",          # Age categories
  "S13",             # Highest level of education
  "S14",             # Community lived in

  # Variables capturing experiences with racism
  "Q629R1",          # Unfair treatment due to racial background/ethnicity
  "Q629R2",          # Unfair treatment due to skin color
  "Q629R3",          # Unfair treatment due to gender
  "Q629R4",          # Unfair treatment due to sexuality/orientation
  "Q629R5",          # Unfair treatment due to immigration status
  "Q629R6",          # Unfair treatment due to religion
  "Q629R7",          # Unfair treatment due to accent

  "Q627",            # Treated unfairly or discriminated against because of your race, ethnicity, gender, sexuality, being an immigrant, religious heritage or having an accent?
  "Q633",            # Discrimination impact on life satisfaction

  # Religion-related variables
  "Q58R1", "Q58R2", "Q58R3", "Q58R4", "Q58R5", "Q58R6", "Q58R7", 
  "Q58R8", "Q58R9", "Q58R10", "Q58R11", "Q58R12", "Q58R13", 
  "Q58R14", "Q58R15", # Various religious affiliations

  # Perceptions of discrimination
  "Q619_Q626R1",     # Perceptions of discrimination against Whites
  "Q619_Q626R2",     # Perceptions of discrimination against Blacks
  "Q619_Q626R3",     # Perceptions of discrimination against Asian Americans
  "Q619_Q626R4",     # Perceptions of discrimination against Native Americans
  "Q619_Q626R5",     # Perceptions of discrimination against Immigrants
  "Q619_Q626R6",     # Perceptions of discrimination against Latinos
  "Q619_Q626R7",     # Perceptions of discrimination against GaysAndLesbians
  "Q619_Q626R8"      # Perceptions of discrimination against Muslims
)
```

## Cleaning and factoring data for all columns
```{r}
# Create a new dataframe with selected variables
dataset_new <- dataset[, required_cols]

# Recode the outcome variable 'Q1R5' for readability
dataset_new$Q1R5 <- factor(
  ifelse(
    dataset_new$Q1R5 == "(0) NO TO: Stopping discrimination against racial/ethnic minorities", 
    "Not important",
    ifelse(dataset_new$Q1R5 == "(1) Stopping discrimination against racial/ethnic minorities", 
           "Important", 
           "Not specified")
  ),
  levels = c("Not important", "Important", "Not specified") # Define factor levels
)

# Define the columns for perceptions of discrimination
perception_cols <- c(
  "Q619_Q626R1",  # Perceptions of discrimination against Whites
  "Q619_Q626R2",  # Perceptions of discrimination against Blacks
  "Q619_Q626R3",  # Perceptions of discrimination against Asian Americans
  "Q619_Q626R4",  # Perceptions of discrimination against Native Americans
  "Q619_Q626R5",  # Perceptions of discrimination against Immigrants
  "Q619_Q626R6",  # Perceptions of discrimination against Latinos
  "Q619_Q626R7",  # Perceptions of discrimination against GaysAndLesbians
  "Q619_Q626R8",  # Perceptions of discrimination against Muslims
  "Q633"          # Discrimination impact on life satisfaction
)

# Recode 'Don' as 'Don't know' and ensure columns remain factors
dataset_new[perception_cols] <- lapply(dataset_new[perception_cols], function(x) {
  x <- as.character(x)  # Temporarily convert to character for recoding
  x <- ifelse(x == "(5) Don", "(5) Don't know", x)  # Recode 'Don' to 'Don't know'
  factor(x)  # Convert back to factor
})

# Verify the changes in perception columns
head(dataset_new[perception_cols])

# Define required columns to convert to factors (if not already defined)
# Assuming `required_cols` is previously defined; otherwise, specify explicitly
# e.g., required_cols <- c("Q1R5", "Q619_Q626R1", ...)
dataset_new[required_cols] <- lapply(dataset_new[required_cols], as.factor)

# Verify the structure of the dataset
str(dataset_new[required_cols])

# Display the first few rows to check the changes
head(dataset_new)
```

## Consolidating Religion to one column
```{r}
# Convert individual religion columns to numeric
for (i in 1:15) {
  col_name <- paste0("Q58R", i, "_numeric")
  dataset_new[[col_name]] <- as.numeric(sub("^\\((\\d+)\\).*", "\\1", dataset_new[[paste0("Q58R", i)]]))
}

# Create a new Religion variable
dataset_new$Religion <- "None"

# Assign religion based on the numeric values of each Q58R column
religion_mapping <- c(
  "Q58R1_numeric" = "Catholic", 
  "Q58R2_numeric" = "Protestant", 
  "Q58R3_numeric" = "Other Christian",
  "Q58R4_numeric" = "Jewish", 
  "Q58R5_numeric" = "Muslim", 
  "Q58R6_numeric" = "Hindu", 
  "Q58R7_numeric" = "Buddhist", 
  "Q58R8_numeric" = "Mormon/LDS", 
  "Q58R9_numeric" = "Folk Religion", 
  "Q58R10_numeric" = "Daoist", 
  "Q58R11_numeric" = "Shamanism", 
  "Q58R12_numeric" = "Ancestral Religion", 
  "Q58R13_numeric" = "Spiritual but not religious", 
  "Q58R14_numeric" = "Atheist", 
  "Q58R15_numeric" = "Agnostic"
)

# Loop through the mapping to assign values
for (col in names(religion_mapping)) {
  dataset_new$Religion[dataset_new[[col]] == 1] <- religion_mapping[[col]]
}

# Handle multiple religious affiliations
dataset_new$Religion[rowSums(dataset_new[, grep("_numeric", names(dataset_new))] == 1) > 1] <- "Multiple Religions"

# Convert Religion to a factor with specific levels
dataset_new$Religion <- factor(dataset_new$Religion,
                               levels = c("Catholic", "Protestant", "Other Christian", "Jewish", "Muslim",
                                          "Hindu", "Buddhist", "Mormon/LDS", "Folk Religion",
                                          "Daoist", "Shamanism", "Ancestral Religion",
                                          "Spiritual but not religious", "Atheist", "Agnostic", 
                                          "Multiple Religions", "None"))

# Create a table of counts for the new Religion variable
table(dataset_new$Religion)

# Remove redundant religion columns (Q58R1 to Q58R15 and their numeric versions)
redundant_cols <- c(
  paste0("Q58R", 1:15), paste0("Q58R", 1:15, "_numeric")
)

# Remove redundant columns from the dataset
dataset_clean <- dataset_new[, !(names(dataset_new) %in% redundant_cols)]

# Take a look at the updated dataset
head(dataset_clean)
```

We first consolidated multiple columns related to respondents' religious affiliations into a single, categorical variable called Religion. We also add another option called 'Multiple Religions' where respondents identified with more than one religion.

# Checking for missing data

```{r}
# Check the number of missing values for each variable
missing_summary <- sapply(dataset_clean, function(x) sum(is.na(x)))

# Create a data frame to display the missing values clearly
missing_df <- data.frame(Variable = names(missing_summary), Missing_Count = missing_summary)
print(missing_df)

# Extract the variables Q629R1 to Q629R7
racism_vars <- dataset_clean[, grep("^Q629R[1-7]$", names(dataset_clean))]

# Calculate the proportion of missing values for each variable
missing_proportions <- colSums(is.na(racism_vars)) / nrow(racism_vars)

# Convert to a data frame
missing_summary <- data.frame(Variable = names(missing_proportions),
                              Missing_Proportion = round(missing_proportions, 4))

print(missing_summary)
```

Variables Q629R1 - Q629R7 each have 9,180 missing values, accounting for over 52% of their data. Due to this high proportion of missingness, we decided to remove these variables from the analysis.

This decision is justified as these variables ("Q629R1" to "Q629R7") capture detailed experiences with unfair treatment but overlap significantly with the broader variable Q627 ("Treated unfairly or discriminated against"), which remains in the analysis. By retaining Q627, we ensure that the overall essence of experiences with racism is preserved, albeit in a more general and complete form.

Removing these variables enhances data quality by addressing issues of excessive missingness, reduces the risk of bias, and strengthens the reliability and interpretability of the analysis.

```{r}
# Remove the variables Q629R1 to Q629R7
dataset_clean <- dataset_clean[, !(names(dataset_clean) %in% c("Q629R1", "Q629R2", 
                                                               "Q629R3", "Q629R4", 
                                                               "Q629R5", "Q629R6", "Q629R7"))]

head(dataset_clean)
```

## Rename the remaining column headers for ease of understanding
```{r}
colnames(dataset_clean)
colnames(dataset_clean) <- c(
  "ImportanceOfStoppingDiscriminationAgainstMinorities",
  "PrimaryRace", 
  "Orientation", 
  "Gender", 
  "Age", 
  "HighestLevelOfEducation", 
  "CommunityLivedIn", 
  "TreatedUnfairlyOrDiscriminatedAgainst", 
  "DiscriminationImpactOnLifeSatisfaction",
  "PerceptionsOfDiscriminationAgainstWhites",
  "PerceptionsOfDiscriminationAgainstBlacks", 
  "PerceptionsOfDiscriminationAgainstAsianAmericans", 
  "PerceptionsOfDiscriminationAgainstNativeAmericans", 
  "PerceptionsOfDiscriminationAgainstImmigrants", 
  "PerceptionsOfDiscriminationAgainstLatinos", 
  "PerceptionsOfDiscriminationAgainstGaysAndLesbians",
  "PerceptionsOfDiscriminationAgainstMuslims",
  "Religion"
)

```

# Look for exploratory plots in the other documents

# Binomial Logistic Regression

## Model

```{r}
library(vcd)
library(tidyverse)
library(corrplot)
```

```{r}
model <- glm(ImportanceOfStoppingDiscriminationAgainstMinorities ~ PrimaryRace +
               Orientation + Gender + Age + 
               HighestLevelOfEducation + CommunityLivedIn +
               TreatedUnfairlyOrDiscriminatedAgainst + 
               DiscriminationImpactOnLifeSatisfaction + 
               PerceptionsOfDiscriminationAgainstWhites + 
               PerceptionsOfDiscriminationAgainstBlacks +
               PerceptionsOfDiscriminationAgainstAsianAmericans + 
               PerceptionsOfDiscriminationAgainstNativeAmericans + 
               PerceptionsOfDiscriminationAgainstImmigrants + 
               PerceptionsOfDiscriminationAgainstLatinos + 
               PerceptionsOfDiscriminationAgainstGaysAndLesbians + 
               PerceptionsOfDiscriminationAgainstMuslims +
               Religion,

        data = dataset_clean, family = "binomial")
```

### Check VIF values
```{r}
# Check multicollinearity using VIF
vif(model)
```

```{r}
## # Extract odds ratios
## odds_ratios <- exp(coef(model))
## 
## # Get confidence intervals for coefficients
## conf_intervals <- exp(confint(model))
## 
## # Extract p-values
## p_values <- summary(model)$coefficients[, 4]
## 
## # Combine everything into a single data frame
## results_model <- data.frame(
##   OddsRatio = odds_ratios,
##   LowerCI = conf_intervals[, 1],
##   UpperCI = conf_intervals[, 2],
##   PValue = p_values
## )
## 
## # Display results
## results_model
```

## Perform stepwise selection to select best parameters
```{r}
# Perform stepwise selection
model_stepwise <- step(model, 
                       direction = "both",
                       trace = 0)
```

### Check VIF Values
```{r}
# Check multicollinearity using VIF
vif(model_stepwise)
```

## Extract Odds for each variable in the stepwise model
```{r}
# Extract odds ratios
odds_ratios <- exp(coef(model_stepwise))

# Get confidence intervals for coefficients
conf_intervals <- exp(confint(model_stepwise))

# Extract p-values
p_values <- summary(model_stepwise)$coefficients[, 4]

# Combine everything into a single data frame
results_model_stepwise <- data.frame(
  OddsRatio = odds_ratios,
  LowerCI = conf_intervals[, 1],
  UpperCI = conf_intervals[, 2],
  PValue = p_values
)

# Display results
results_model_stepwise
```

## Plot showing odds
```{r}
# Filter out only significant variables (p-value < 0.05)
significant_data <- results_model_stepwise[results_model_stepwise$PValue < 0.051, ]

# Create the plot with inverted axes and CI shown as error bars
ggplot(significant_data, aes(y = rownames(significant_data), x = OddsRatio)) +
  geom_point(size = 4, color = "blue") +  # Plot the odds ratio as points
  geom_errorbarh(aes(xmin = LowerCI, xmax = UpperCI), height = 0.2, color = "black") +  # Add horizontal error bars
  labs(title = "Significant Variables with Confidence Intervals (Odds Ratio)",
       x = "Odds Ratio",
       y = "Variables") +
  theme(axis.text.y = element_text(angle = 0, hjust = 1))  # Keep y-axis labels horizontal
  theme(plot.margin = margin(1, 2, 1, 1, "cm"))  # Adjust margins for larger size

# Save the plot with larger dimensions
ggsave("significant_variables_plot.png", width = 12, height = 8)
```

```{r}
# Filter for significant variables (p-value < 0.05)
significant_data <- model_stepwise |> 
  filter(PValue < 0.05)

# Prepare data for plotting
# You might need to adjust this section to ensure the data is in long format with appropriate variable names
plot_data <- significant_data |>
  select(Variable, OddsRatio) |>
  mutate(Variable = factor(Variable, levels = unique(Variable)))

# Create the boxplot
ggplot(plot_data, aes(x = Variable, y = OddsRatio)) +
  geom_boxplot() +
  labs(title = "Boxplot of Significant Variables (Odds Ratio)",
       x = "Variables",
       y = "Odds Ratio") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

### 2024-12-06 Naki_MainPlots.qmd
new files

---
title: "Naki_MainPlots"
format: html
editor: visual
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this: The `echo: false` option disables the printing of code (only output is displayed).

```{r}
1 + 1
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r}
file_path <- "/Users/nakiyahdhariwala/StatsProject/39096-0001-Data.rda"

load(file_path)          # Load the RDA file
ls()                     # List objects loaded into the environment
dataset <- da39096.0001  # Access the dataset

head(dataset)
ncol(dataset)
nrow(dataset)
```

```{r}
# Calculate proportions
proportions <- prop.table(table(dataset$Q1R5))
proportions
# Define custom labels
custom_labels <- c("Not important", "Important")

# Combine custom labels with percentages
pie_labels <- paste(custom_labels, round(proportions * 100, 2), "%")

# Plot the pie chart with custom labels
pie(proportions,
    labels = pie_labels,
    main = "Q1R5: Stopping discrimination against racial/ethnic minorities",
    col = c("lightblue", "salmon"))
```

## Cleaning and factoring all columns in the df

```{r}
# Define the required columns
required_cols <- c(
  # Outcome Variable: Stopping discrimination against racial/ethnic minorities (Binary: 0 = Not important, 1 = Important)
  "Q1R5",

  # Demographic Variables
  "S2_RACE_PRIME",   # Primary Race
  "S3",              # SexualOrientation
  "S3B",             # Gender
  "S5_AGE",          # Age categories
  "S13",             # Highest level of education
  "S14",             # Community lived in

  # Variables capturing experiences with racism
  "Q629R1",          # Unfair treatment due to racial background/ethnicity
  "Q629R2",          # Unfair treatment due to skin color
  "Q629R3",          # Unfair treatment due to gender
  "Q629R4",          # Unfair treatment due to sexuality/orientation
  "Q629R5",          # Unfair treatment due to immigration status
  "Q629R6",          # Unfair treatment due to religion
  "Q629R7",          # Unfair treatment due to accent

  "Q627",            # Treated unfairly or discriminated against because of your race, ethnicity, gender, sexuality, being an immigrant, religious heritage or having an accent?
  "Q633",            # Discrimination impact on life satisfaction

  # Religion-related variables
  "Q58R1", "Q58R2", "Q58R3", "Q58R4", "Q58R5", "Q58R6", "Q58R7", 
  "Q58R8", "Q58R9", "Q58R10", "Q58R11", "Q58R12", "Q58R13", 
  "Q58R14", "Q58R15", # Various religious affiliations

  # Perceptions of discrimination
  "Q619_Q626R1",     # Perceptions of discrimination against Whites
  "Q619_Q626R2",     # Perceptions of discrimination against Blacks
  "Q619_Q626R3",     # Perceptions of discrimination against Asian Americans
  "Q619_Q626R4",     # Perceptions of discrimination against Native Americans
  "Q619_Q626R5",     # Perceptions of discrimination against Immigrants
  "Q619_Q626R6",     # Perceptions of discrimination against Latinos
  "Q619_Q626R7",     # Perceptions of discrimination against GaysAndLesbians
  "Q619_Q626R8"      # Perceptions of discrimination against Muslims
)

# Create a new dataframe with selected variables
dataset_new <- dataset[, required_cols]

# Recode the outcome variable 'Q1R5' for readability
dataset_new$Q1R5 <- ifelse(
  dataset_new$Q1R5 == "(0) NO TO: Stopping discrimination against racial/ethnic minorities", 
  "Not important",
  ifelse(dataset_new$Q1R5 == "(1) Stopping discrimination against racial/ethnic minorities", 
         "Important", 
         "Not specified")
)

# Define the columns for perceptions of discrimination
perception_cols <- c(
  "Q619_Q626R1",  # Perceptions of discrimination against Whites
  "Q619_Q626R2",  # Perceptions of discrimination against Blacks
  "Q619_Q626R3",  # Perceptions of discrimination against Asian Americans
  "Q619_Q626R4",  # Perceptions of discrimination against Native Americans
  "Q619_Q626R5",  # Perceptions of discrimination against Immigrants
  "Q619_Q626R6",  # Perceptions of discrimination against Latinos
  "Q619_Q626R7",  # Perceptions of discrimination against GaysAndLesbians
  "Q619_Q626R8",  # Perceptions of discrimination against Muslims

  "Q633"          # Discrimination impact on life satisfaction
)

# Recode 'Don' as 'Don't know' in the specified columns
dataset_new[perception_cols] <- lapply(dataset_new[perception_cols], function(x) {
  x <- ifelse(x == "Don", "Don't know", x)
  return(x)
})

# Verify the changes
head(dataset_new[perception_cols])

# Convert all required columns to factors
dataset_new[required_cols] <- lapply(dataset_new[required_cols], as.factor)

# Verify the structure of the dataset
str(dataset_new[required_cols])

# Display the first few rows to check the changes
head(dataset_new)
```

This code prepares the dataset for further analysis by selecting and cleaning up the relevant variables, ensuring they are in a suitable format (i.e., factors for categorical data).

**Q633 & Q619_Q626R1 - Q619_Q626R8 Category Key**

-   1 - A lot

-   2 - Some

-   3 - A little

-   4 - None at all

-   5 - Don't Know

## Changing Factor Key

```{r}
# Define the mapping from numeric values to the desired labels
perception_labels <- c("A lot", "Some", "A little", "None at all", "Don't know")

# Convert the factored perception columns to the correct labels
dataset_new[perception_cols] <- lapply(dataset_new[perception_cols], function(x) {
  # Convert the numeric factors (1:5) to the corresponding perception labels
  x <- factor(x, levels = 1:5, labels = perception_labels)
  return(x)
})

# Verify the changes
head(dataset_new[perception_cols])

# Check the structure
str(dataset_new[perception_cols])

```

## Consolidating Religion into one column

```{r}
# Convert individual religion columns to numeric
for (i in 1:15) {
  col_name <- paste0("Q58R", i, "_numeric")
  dataset_new[[col_name]] <- as.numeric(sub("^\\((\\d+)\\).*", "\\1", dataset_new[[paste0("Q58R", i)]]))
}

# Create a new Religion variable
dataset_new$Religion <- "None"

# Assign religion based on the numeric values of each Q58R column
religion_mapping <- c(
  "Q58R1_numeric" = "Catholic", 
  "Q58R2_numeric" = "Protestant", 
  "Q58R3_numeric" = "Other Christian",
  "Q58R4_numeric" = "Jewish", 
  "Q58R5_numeric" = "Muslim", 
  "Q58R6_numeric" = "Hindu", 
  "Q58R7_numeric" = "Buddhist", 
  "Q58R8_numeric" = "Mormon/LDS", 
  "Q58R9_numeric" = "Folk Religion", 
  "Q58R10_numeric" = "Daoist", 
  "Q58R11_numeric" = "Shamanism", 
  "Q58R12_numeric" = "Ancestral Religion", 
  "Q58R13_numeric" = "Spiritual but not religious", 
  "Q58R14_numeric" = "Atheist", 
  "Q58R15_numeric" = "Agnostic"
)

# Loop through the mapping to assign values
for (col in names(religion_mapping)) {
  dataset_new$Religion[dataset_new[[col]] == 1] <- religion_mapping[[col]]
}

# Handle multiple religious affiliations
dataset_new$Religion[rowSums(dataset_new[, grep("_numeric", names(dataset_new))] == 1) > 1] <- "Multiple Religions"

# Convert Religion to a factor with specific levels
dataset_new$Religion <- factor(dataset_new$Religion,
                               levels = c("Catholic", "Protestant", "Other Christian", "Jewish", "Muslim",
                                          "Hindu", "Buddhist", "Mormon/LDS", "Folk Religion",
                                          "Daoist", "Shamanism", "Ancestral Religion",
                                          "Spiritual but not religious", "Atheist", "Agnostic", 
                                          "Multiple Religions", "None"))

# Create a table of counts for the new Religion variable
table(dataset_new$Religion)

# Remove redundant religion columns (Q58R1 to Q58R15 and their numeric versions)
redundant_cols <- c(
  paste0("Q58R", 1:15), paste0("Q58R", 1:15, "_numeric")
)

# Remove redundant columns from the dataset
dataset_clean <- dataset_new[, !(names(dataset_new) %in% redundant_cols)]

# Take a look at the updated dataset
head(dataset_clean)

```

# Checking for missing data

```{r}
# Check the number of missing values for each variable
missing_summary <- sapply(dataset_clean, function(x) sum(is.na(x)))

# Create a data frame to display the missing values clearly
missing_df <- data.frame(Variable = names(missing_summary), Missing_Count = missing_summary)
print(missing_df)

# Extract the variables Q629R1 to Q629R7
racism_vars <- dataset_clean[, grep("^Q629R[1-7]$", names(dataset_clean))]

# Calculate the proportion of missing values for each variable
missing_proportions <- colSums(is.na(racism_vars)) / nrow(racism_vars)

# Convert to a data frame
missing_summary <- data.frame(Variable = names(missing_proportions),
                              Missing_Proportion = round(missing_proportions, 4))

print(missing_summary)
```

```{r}
# Remove the variables Q629R1 to Q629R7
dataset_clean <- dataset_clean[, !(names(dataset_clean) %in% c("Q629R1", "Q629R2", 
                                                               "Q629R3", "Q629R4", 
                                                               "Q629R5", "Q629R6", "Q629R7"))]

head(dataset_clean)
```

## Rename the remaining column headers for ease of understanding

```{r}
colnames(dataset_clean)
colnames(dataset_clean) <- c(
  "ImportanceOfStoppingDiscriminationAgainstMinorities",
  "PrimaryRace", 
  "Orientation", 
  "Gender", 
  "Age", 
  "HighestLevelOfEducation", 
  "CommunityLivedIn", 
  "TreatedUnfairlyOrDiscriminatedAgainst", 
  "DiscriminationImpactOnLifeSatisfaction",
  "PerceptionsOfDiscriminationAgainstWhites",
  "PerceptionsOfDiscriminationAgainstBlacks", 
  "PerceptionsOfDiscriminationAgainstAsianAmericans", 
  "PerceptionsOfDiscriminationAgainstNativeAmericans", 
  "PerceptionsOfDiscriminationAgainstImmigrants", 
  "PerceptionsOfDiscriminationAgainstLatinos", 
  "PerceptionsOfDiscriminationAgainstGaysAndLesbians",
  "PerceptionsOfDiscriminationAgainstMuslims",
  "Religion"
)
colnames(dataset_clean)
```

# Exploratory Plots Analysis

```{r}
# Load necessary libraries
library(ggplot2)

# Plot for "PrimaryRace"
ggplot(dataset_clean, aes(x = PrimaryRace, fill = ImportanceOfStoppingDiscriminationAgainstMinorities)) +
  geom_bar(position = "fill") +
  labs(title = "Distribution of Primary Race by Importance of Stopping Discrimination", x = "Primary Race", y = "Proportion") +
  scale_fill_manual(values = c("lightblue", "salmon")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for "Orientation"
ggplot(dataset_clean, aes(x = Orientation, fill = ImportanceOfStoppingDiscriminationAgainstMinorities)) +
  geom_bar(position = "fill") +
  labs(title = "Distribution of Orientation by Importance of Stopping Discrimination", x = "Orientation", y = "Proportion") +
  scale_fill_manual(values = c("lightblue", "salmon")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for "Gender"
ggplot(dataset_clean, aes(x = Gender, fill = ImportanceOfStoppingDiscriminationAgainstMinorities)) +
  geom_bar(position = "fill") +
  labs(title = "Distribution of Gender by Importance of Stopping Discrimination", x = "Gender", y = "Proportion") +
  scale_fill_manual(values = c("lightblue", "salmon")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for "Age"
ggplot(dataset_clean, aes(x = Age, fill = ImportanceOfStoppingDiscriminationAgainstMinorities)) +
  geom_bar(position = "fill") +
  labs(title = "Distribution of Age by Importance of Stopping Discrimination", x = "Age Category", y = "Proportion") +
  scale_fill_manual(values = c("lightblue", "salmon")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for "HighestLevelOfEducation"
ggplot(dataset_clean, aes(x = HighestLevelOfEducation, fill = ImportanceOfStoppingDiscriminationAgainstMinorities)) +
  geom_bar(position = "fill") +
  labs(title = "Distribution of Highest Level of Education by Importance of Stopping Discrimination", x = "Highest Level of Education", y = "Proportion") +
  scale_fill_manual(values = c("lightblue", "salmon")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for "CommunityLivedIn"
ggplot(dataset_clean, aes(x = CommunityLivedIn, fill = ImportanceOfStoppingDiscriminationAgainstMinorities)) +
  geom_bar(position = "fill") +
  labs(title = "Distribution of Community Lived In by Importance of Stopping Discrimination", x = "Community Lived In", y = "Proportion") +
  scale_fill_manual(values = c("lightblue", "salmon")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for "TreatedUnfairlyOrDiscriminatedAgainst"
ggplot(dataset_clean, aes(x = TreatedUnfairlyOrDiscriminatedAgainst, fill = ImportanceOfStoppingDiscriminationAgainstMinorities)) +
  geom_bar(position = "fill") +
  labs(title = "Distribution of Treated Unfairly by Importance of Stopping Discrimination", x = "Treated Unfairly", y = "Proportion") +
  scale_fill_manual(values = c("lightblue", "salmon")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for "DiscriminationImpactOnLifeSatisfaction"
ggplot(dataset_clean, aes(x = DiscriminationImpactOnLifeSatisfaction, fill = ImportanceOfStoppingDiscriminationAgainstMinorities)) +
  geom_bar(position = "fill") +
  labs(title = "Distribution of Discrimination Impact on Life Satisfaction by Importance of Stopping Discrimination", x = "Impact on Life Satisfaction", y = "Proportion") +
  scale_fill_manual(values = c("lightblue", "salmon")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for "PerceptionsOfDiscriminationAgainstWhites"
ggplot(dataset_clean, aes(x = PerceptionsOfDiscriminationAgainstWhites, fill = ImportanceOfStoppingDiscriminationAgainstMinorities)) +
  geom_bar(position = "fill") +
  labs(title = "Perceptions of Discrimination Against Whites by Importance of Stopping Discrimination", x = "Discrimination Against Whites", y = "Proportion") +
  scale_fill_manual(values = c("lightblue", "salmon")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for "PerceptionsOfDiscriminationAgainstBlacks"
ggplot(dataset_clean, aes(x = PerceptionsOfDiscriminationAgainstBlacks, fill = ImportanceOfStoppingDiscriminationAgainstMinorities)) +
  geom_bar(position = "fill") +
  labs(title = "Perceptions of Discrimination Against Blacks by Importance of Stopping Discrimination", x = "Discrimination Against Blacks", y = "Proportion") +
  scale_fill_manual(values = c("lightblue", "salmon")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for "PerceptionsOfDiscriminationAgainstAsianAmericans"
ggplot(dataset_clean, aes(x = PerceptionsOfDiscriminationAgainstAsianAmericans, fill = ImportanceOfStoppingDiscriminationAgainstMinorities)) +
  geom_bar(position = "fill") +
  labs(title = "Perceptions of Discrimination Against Asian Americans by Importance of Stopping Discrimination", x = "Discrimination Against Asian Americans", y = "Proportion") +
  scale_fill_manual(values = c("lightblue", "salmon")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for "PerceptionsOfDiscriminationAgainstNativeAmericans"
ggplot(dataset_clean, aes(x = PerceptionsOfDiscriminationAgainstNativeAmericans, fill = ImportanceOfStoppingDiscriminationAgainstMinorities)) +
  geom_bar(position = "fill") +
  labs(title = "Perceptions of Discrimination Against Native Americans by Importance of Stopping Discrimination", x = "Discrimination Against Native Americans", y = "Proportion") +
  scale_fill_manual(values = c("lightblue", "salmon")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for "PerceptionsOfDiscriminationAgainstImmigrants"
ggplot(dataset_clean, aes(x = PerceptionsOfDiscriminationAgainstImmigrants, fill = ImportanceOfStoppingDiscriminationAgainstMinorities)) +
  geom_bar(position = "fill") +
  labs(title = "Perceptions of Discrimination Against Immigrants by Importance of Stopping Discrimination", x = "Discrimination Against Immigrants", y = "Proportion") +
  scale_fill_manual(values = c("lightblue", "salmon")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for "PerceptionsOfDiscriminationAgainstLatinos"
ggplot(dataset_clean, aes(x = PerceptionsOfDiscriminationAgainstLatinos, fill = ImportanceOfStoppingDiscriminationAgainstMinorities)) +
  geom_bar(position = "fill") +
  labs(title = "Perceptions of Discrimination Against Latinos by Importance of Stopping Discrimination", x = "Discrimination Against Latinos", y = "Proportion") +
  scale_fill_manual(values = c("lightblue", "salmon")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for "PerceptionsOfDiscriminationAgainstGaysAndLesbians"
ggplot(dataset_clean, aes(x = PerceptionsOfDiscriminationAgainstGaysAndLesbians, fill = ImportanceOfStoppingDiscriminationAgainstMinorities)) +
  geom_bar(position = "fill") +
  labs(
    title = "Perceptions of Discrimination Against Gays and Lesbians by Importance of Stopping Discrimination", 
    x = "Discrimination Against Gays and Lesbians", 
    y = "Proportion"
  ) +
  scale_fill_manual(values = c("lightblue", "salmon")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for "PerceptionsOfDiscriminationAgainstMuslims"
ggplot(dataset_clean, aes(x = PerceptionsOfDiscriminationAgainstMuslims, fill = ImportanceOfStoppingDiscriminationAgainstMinorities)) +
  geom_bar(position = "fill") +
  labs(
    title = "Perceptions of Discrimination Against Muslims by Importance of Stopping Discrimination", 
    x = "Discrimination Against Muslims", 
    y = "Proportion"
  ) +
  scale_fill_manual(values = c("lightblue", "salmon")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for "Religion"
ggplot(dataset_clean, aes(x = Religion, fill = ImportanceOfStoppingDiscriminationAgainstMinorities)) +
  geom_bar(position = "fill") +
  labs(title = "Distribution of Religion by Importance of Stopping Discrimination", x = "Religion", y = "Proportion") +
  scale_fill_manual(values = c("lightblue", "salmon")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

# Interaction Terms Tests

```{r}
# Fit a logistic regression model
model <- glm(Q1R5 ~ S2_RACE_PRIME + # Primary Race
               S3 +                 # Orientation
               S3B +                # Gender
               S5_AGE +             # Age
               S13 +                # Highest level of education
               S14 +                # Community lived in
               Q627 +               # Treated unfairly or discriminated against
               Q633 +               # Discrimination impact on life satisfaction
               Q619_Q626R2 +        # Perceptions of discrimination against Blacks
               Q619_Q626R3 +        # Perceptions of discrimination against Asian Americans
               Q619_Q626R4 +        # Perceptions of discrimination against Native Americans
               Q619_Q626R5 +        # Perceptions of discrimination against Immigrants
               Q619_Q626R6 +        # Perceptions of discrimination against Latinos
               Religion +
               S3B * Q627,

            data = dataset_clean, family = binomial)

# summary(model)
```

##1.

```{r}
# Fit a logistic regression model with an interaction term
model_interaction <- glm(Q1R5 ~ S2_RACE_PRIME + # Primary Race
               S3 +                 # Orientation
               S3B +                # Gender
               S5_AGE +             # Age
               S13 +                # Highest level of education
               S14 +                # Community lived in
               Q627 +               # Treated unfairly or discriminated against
               Q633 +               # Discrimination impact on life satisfaction
               Q619_Q626R2 +        # Perceptions of discrimination against Blacks
               Q619_Q626R3 +        # Perceptions of discrimination against Asian Americans
               Q619_Q626R4 +        # Perceptions of discrimination against Native Americans
               Q619_Q626R5 +        # Perceptions of discrimination against Immigrants
               Q619_Q626R6 +        # Perceptions of discrimination against Latinos
               Religion +
               S3 * Q627,

            data = dataset_clean, family = binomial)

summary(model_interaction)

# Compare models
#AIC(model, model_interaction)
```

### 2024-12-06 OldFiles/Naki_EDA_Exploration.qmd
new files

--- 2024-12-02 Naki_EDA_Exploration.qmd
@@
 | **Variable** | **Level** | **Count** | **Proportion** |
-|------------------|--------------------|------------------|------------------|
+|----|----|----|----|
 | Q1R5 | \(0\) NO TO: Stopping discrimination against racial/ethnic minorities | 5728 | 0.6848 |
@@
 | Variable | Chi-Square Statistic | Degrees of Freedom (df) | P-Value | Test Type |
-|---------------|---------------|---------------|---------------|---------------|
+|----------|----------------------|-------------------------|---------|-----------|
 
@@
 |  |  |  |  |  |
-|---------------|---------------|---------------|---------------|---------------|
+|----|----|----|----|----|
 | Q629R1 | 114.83 | 1 | \< 2.2e-16 | Pearson's Chi-squared test with Yates' correction |
@@
 |  |  |  |  |  |
-|---------------|---------------|---------------|---------------|---------------|
+|----|----|----|----|----|
 | Q629R2 | 72.719 | 1 | \< 2.2e-16 | Pearson's Chi-squared test with Yates' correction |
@@
 |        |         |     |        |                                                   |
-|---------------|---------------|---------------|---------------|---------------|
+|--------|---------|-----|--------|---------------------------------------------------|
 | Q629R3 | 0.41214 | 1   | 0.5209 | Pearson's Chi-squared test with Yates' correction |
@@
 |        |        |     |        |                                                   |
-|---------------|---------------|---------------|---------------|---------------|
+|--------|--------|-----|--------|---------------------------------------------------|
 | Q629R4 | 1.3169 | 1   | 0.2511 | Pearson's Chi-squared test with Yates' correction |
@@
 |        |       |     |           |                                                   |
-|---------------|---------------|---------------|---------------|---------------|
+|--------|-------|-----|-----------|---------------------------------------------------|
 | Q629R5 | 11.48 | 1   | 0.0007036 | Pearson's Chi-squared test with Yates' correction |
@@
 |  |  |  |  |  |
-|---------------|---------------|---------------|---------------|---------------|
+|----|----|----|----|----|
 | Q629R6 | 20.113 | 1 | 7.301e-06 | Pearson's Chi-squared test with Yates' correction |
@@
 |        |        |     |        |                                                   |
-|---------------|---------------|---------------|---------------|---------------|
+|--------|--------|-----|--------|---------------------------------------------------|
 | Q629R7 | 2.6287 | 1   | 0.1049 | Pearson's Chi-squared test with Yates' correction |

### 2024-12-06 OldFiles/Naki_Main1.qmd
new files

--- 2024-12-06 Naki_Main.qmd
@@
 ---
+
+## Quarto
+
+Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.
+
+## Running Code
+
+When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:
+
+```{r}
+1 + 1
+```
+
+The `echo: false` option disables the printing of code (only output is displayed).
 
@@
 
-## Extracting relevant columns
+## Cleaning and factoring all columns in the df
+
 ```{r}
@@
 )
-```
 
-## Cleaning and factoring data for all columns
-```{r}
 # Create a new dataframe with selected variables
@@
 # Recode the outcome variable 'Q1R5' for readability
-dataset_new$Q1R5 <- factor(
-  ifelse(
+dataset_new$Q1R5 <- ifelse(
   dataset_new$Q1R5 == "(0) NO TO: Stopping discrimination against racial/ethnic minorities", 
@@
          "Not specified")
-  ),
-  levels = c("Not important", "Important", "Not specified") # Define factor levels
 )
@@
   "Q619_Q626R8",  # Perceptions of discrimination against Muslims
+
   "Q633"          # Discrimination impact on life satisfaction
@@
 
-# Recode 'Don' as 'Don't know' and ensure columns remain factors
+# Recode 'Don' as 'Don't know'
 dataset_new[perception_cols] <- lapply(dataset_new[perception_cols], function(x) {
-  x <- as.character(x)  # Temporarily convert to character for recoding
-  x <- ifelse(x == "(5) Don", "(5) Don't know", x)  # Recode 'Don' to 'Don't know'
-  factor(x)  # Convert back to factor
+  x <- ifelse(x == "Don", "Don't know", x)  # Recode 'Don' to 'Don't know'
+  return(x)
 })
 
-# Verify the changes in perception columns
+# Verify the changes
 head(dataset_new[perception_cols])
 
-# Define required columns to convert to factors (if not already defined)
-# Assuming `required_cols` is previously defined; otherwise, specify explicitly
-# e.g., required_cols <- c("Q1R5", "Q619_Q626R1", ...)
+# Convert all required columns to factors
 dataset_new[required_cols] <- lapply(dataset_new[required_cols], as.factor)
@@
 
+This code prepares the dataset for further analysis by selecting and cleaning up the relevant variables, ensuring they are in a suitable format (i.e., factors for categorical data).
+
+**Q633 & Q619_Q626R1 - Q619_Q626R8 Category Key**
+
+-   1 - A lot
+
+-   2 - Some
+
+-   3 - A little
+
+-   4 - None at all
+
+-   5 - Don't Know
+
+    ## Changing factor key
+
+    ```{r}
+    # Define the mapping from numeric values to the desired labels
+    perception_labels <- c("A lot", "Some", "A little", "None at all", "Don't know")
+
+    # Convert the factored perception columns to the correct labels
+    dataset_new[perception_cols] <- lapply(dataset_new[perception_cols], function(x) {
+      # Convert the numeric factors (1:5) to the corresponding perception labels
+      x <- factor(x, levels = 1:5, labels = perception_labels)
+      return(x)
+    })
+
+    # Verify the changes
+    head(dataset_new[perception_cols])
+
+    # Check the structure
+    str(dataset_new[perception_cols])
+
+    ```
+
     ## Consolidating Religion to one column
+
 ```{r}
@@
 head(dataset_clean)
+
 ```
@@
 ```{r}
-library(vcd)
-library(tidyverse)
-library(corrplot)
-```
-
-```{r}
 model <- glm(ImportanceOfStoppingDiscriminationAgainstMinorities ~ PrimaryRace +
@@
         data = dataset_clean, family = "binomial")
-```
 
-### Check VIF values
-```{r}
-# Check multicollinearity using VIF
-vif(model)
+# summary(model)
 ```
@@
 ```{r}
-## # Extract odds ratios
-## odds_ratios <- exp(coef(model))
-## 
-## # Get confidence intervals for coefficients
-## conf_intervals <- exp(confint(model))
-## 
-## # Extract p-values
-## p_values <- summary(model)$coefficients[, 4]
-## 
-## # Combine everything into a single data frame
-## results_model <- data.frame(
-##   OddsRatio = odds_ratios,
-##   LowerCI = conf_intervals[, 1],
-##   UpperCI = conf_intervals[, 2],
-##   PValue = p_values
-## )
-## 
-## # Display results
-## results_model
+# Extract odds ratios
+#odds_ratios <- exp(coef(model))
+
+# Get confidence intervals for coefficients
+#conf_intervals <- exp(confint(model))
+
+# Extract p-values
+#p_values <- summary(model)$coefficients[, 4]
+
+# Combine everything into a single data frame
+#results_model <- data.frame(
+#  OddsRatio = odds_ratios,
+#  LowerCI = conf_intervals[, 1],
+#  UpperCI = conf_intervals[, 2],
+#  PValue = p_values
+#)
+
+# Display results
+# results_model
+
+# Check multicollinearity using VIF
+# vif(model)
 ```
 
-## Perform stepwise selection to select best parameters
+## Model Stepwise
+
 ```{r}
@@
 
-### Check VIF Values
-```{r}
-# Check multicollinearity using VIF
-vif(model_stepwise)
-```
-
-## Extract Odds for each variable in the stepwise model
 ```{r}
@@
 results_model_stepwise
-```
 
-## Plot showing odds
-```{r}
-# Filter out only significant variables (p-value < 0.05)
-significant_data <- results_model_stepwise[results_model_stepwise$PValue < 0.051, ]
-
-# Create the plot with inverted axes and CI shown as error bars
-ggplot(significant_data, aes(y = rownames(significant_data), x = OddsRatio)) +
-  geom_point(size = 4, color = "blue") +  # Plot the odds ratio as points
-  geom_errorbarh(aes(xmin = LowerCI, xmax = UpperCI), height = 0.2, color = "black") +  # Add horizontal error bars
-  labs(title = "Significant Variables with Confidence Intervals (Odds Ratio)",
-       x = "Odds Ratio",
-       y = "Variables") +
-  theme(axis.text.y = element_text(angle = 0, hjust = 1))  # Keep y-axis labels horizontal
-  theme(plot.margin = margin(1, 2, 1, 1, "cm"))  # Adjust margins for larger size
-
-# Save the plot with larger dimensions
-ggsave("significant_variables_plot.png", width = 12, height = 8)
+# Check multicollinearity using VIF
+# vif(model_stepwise)
 ```
@@
 ```{r}
-# Filter for significant variables (p-value < 0.05)
-significant_data <- model_stepwise |> 
-  filter(PValue < 0.05)
+predicted_classes <- ifelse(predicted_probabilities > 0.7, "Not important", "Important")
+actual_classes <- dataset_clean$ImportanceOfStoppingDiscriminationAgainstMinorities
+```
 
-# Prepare data for plotting
-# You might need to adjust this section to ensure the data is in long format with appropriate variable names
-plot_data <- significant_data |>
-  select(Variable, OddsRatio) |>
-  mutate(Variable = factor(Variable, levels = unique(Variable)))
-
-# Create the boxplot
-ggplot(plot_data, aes(x = Variable, y = OddsRatio)) +
-  geom_boxplot() +
-  labs(title = "Boxplot of Significant Variables (Odds Ratio)",
-       x = "Variables",
-       y = "Odds Ratio") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```{r}
+#predicted_classes <- ifelse(predicted_probabilities > 0.5, 1, 0)
+confusion_matrix <- table(Predicted = predicted_classes, Actual = actual_classes)
+print(confusion_matrix)
 
 ```
+
+```{r}
+library(caret)
+confusion <- confusionMatrix(factor(predicted_classes), factor(actual_classes))
+print(confusion)
+```
+
+```{r}
+
+```

### 2024-12-06 OldFiles/Naki_Main_Old.qmd
new files

--- 2024-12-06 OldFiles/Naki_Main1.qmd
@@
----
-title: "Naki_Main"
-format: html
-editor: visual
----
-
-## Quarto
-
-Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.
-
-## Running Code
-
-When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:
-
 ```{r}
-1 + 1
-```
-
-The `echo: false` option disables the printing of code (only output is displayed).
-
-```{r}
-
 # Load the packages
@@
 
-# Introduction
-
-The Collaborative Multi-Racial Post-Election Survey (CMPS), United States, 2020 is an extensive dataset designed to capture diverse perspectives across racial and ethnic groups in the United States.
-Respondents were asked to address a wide range of issues, including political attitudes, voting behavior, social concerns, and policy preferences.
-
-# Preliminary Cleaning
-
-## Loading the Dataset
+# Loading dataset
 
 ```{r}
+# Set the file path
 file_path <- "/Users/nakiyahdhariwala/StatsProject/39096-0001-Data.rda"
 
-load(file_path)          # Load the RDA file
-ls()                     # List objects loaded into the environment
-dataset <- da39096.0001  # Access the dataset
+# Load the RDA file
+load(file_path)
 
+# List objects loaded into the environment
+ls()
+
+# Access the dataset (replace `da39096.0001` with the actual name found in `ls()`)
+dataset <- da39096.0001  # Use the correct object name
+```
+
+```{r}
 head(dataset)
@@
 ```
+
+The Collaborative Multi-Racial Post-Election Survey (CMPS), United States, 2020 is an extensive dataset designed to capture diverse perspectives across racial and ethnic groups in the United States.
+Respondents were asked to address a wide range of issues, including political attitudes, voting behavior, social concerns, and policy preferences.
 
@@
     col = c("lightblue", "salmon"))
+
 ```
 
-The outcome variable Q1R5 (Stopping discrimination against racial/ethnic minorities) for this Research Question is a binary variable which captures whether respondents view stopping racial and ethnic discrimination as an important issue (i.e., Important OR Not Important).
+The outcome variable Q1R5 (Stopping discrimination against racial/ethnic minorities) for this Research Question is a binary variable which captures whether respondents view stopping racial and ethnic discrimination as an important issue (i.e., 1 = Important, 0 = Not Important).
 Therefore, we will make use Logistic regression as it is suitable for modeling binary outcome variables and allows for the inclusion of both predictor and control variables.
@@
 
-## Cleaning and factoring all columns in the df
-
 ```{r}
@@
 required_cols <- c(
-  # Outcome Variable: Stopping discrimination against racial/ethnic minorities
+  # Outcome Variable: Stopping discrimination against racial/ethnic minorities (Binary: 0 = Not important, 1 = Important)
   "Q1R5",
@@
   "S2_RACE_PRIME",   # Primary Race
-  "S3",              # SexualOrientation
+  "S3",              # How you think of yourself
   "S3B",             # Gender
@@
   # Perceptions of discrimination
-  "Q619_Q626R1",     # Perceptions of discrimination against Whites
   "Q619_Q626R2",     # Perceptions of discrimination against Blacks
@@
   "Q619_Q626R5",     # Perceptions of discrimination against Immigrants
-  "Q619_Q626R6",     # Perceptions of discrimination against Latinos
-  "Q619_Q626R7",     # Perceptions of discrimination against GaysAndLesbians
-  "Q619_Q626R8"      # Perceptions of discrimination against Muslims
+  "Q619_Q626R6"      # Perceptions of discrimination against Latinos
 )
@@
 
-# Define the columns for perceptions of discrimination
-perception_cols <- c(
-  "Q619_Q626R1",  # Perceptions of discrimination against Whites
-  "Q619_Q626R2",  # Perceptions of discrimination against Blacks
-  "Q619_Q626R3",  # Perceptions of discrimination against Asian Americans
-  "Q619_Q626R4",  # Perceptions of discrimination against Native Americans
-  "Q619_Q626R5",  # Perceptions of discrimination against Immigrants
-  "Q619_Q626R6",  # Perceptions of discrimination against Latinos
-  "Q619_Q626R7",  # Perceptions of discrimination against GaysAndLesbians
-  "Q619_Q626R8",  # Perceptions of discrimination against Muslims
-
-  "Q633"          # Discrimination impact on life satisfaction
-)
-
-# Recode 'Don' as 'Don't know'
-dataset_new[perception_cols] <- lapply(dataset_new[perception_cols], function(x) {
-  x <- ifelse(x == "Don", "Don't know", x)  # Recode 'Don' to 'Don't know'
-  return(x)
-})
-
-# Verify the changes
-head(dataset_new[perception_cols])
-
 # Convert all required columns to factors
@@
 head(dataset_new)
-```
-
-This code prepares the dataset for further analysis by selecting and cleaning up the relevant variables, ensuring they are in a suitable format (i.e., factors for categorical data).
-
-**Q633 & Q619_Q626R1 - Q619_Q626R8 Category Key**
-
--   1 - A lot
-
--   2 - Some
-
--   3 - A little
-
--   4 - None at all
-
--   5 - Don't Know
-
-    ## Changing factor key
-
-    ```{r}
-    # Define the mapping from numeric values to the desired labels
-    perception_labels <- c("A lot", "Some", "A little", "None at all", "Don't know")
-
-    # Convert the factored perception columns to the correct labels
-    dataset_new[perception_cols] <- lapply(dataset_new[perception_cols], function(x) {
-      # Convert the numeric factors (1:5) to the corresponding perception labels
-      x <- factor(x, levels = 1:5, labels = perception_labels)
-      return(x)
-    })
-
-    # Verify the changes
-    head(dataset_new[perception_cols])
-
-    # Check the structure
-    str(dataset_new[perception_cols])
 
@@
 
-    ## Consolidating Religion to one column
+We first consolidated multiple columns related to respondents' religious affiliations into a single, categorical variable called Religion.
+We also add another option called 'Multiple Religions' where respondents identified with more than one religion.
+
+# Consolidating Religion into one column
 
@@
 dataset_new$Religion <- "None"
+dataset_new$Religion[dataset_new$Q58R1_numeric == 1] <- "Catholic"
+dataset_new$Religion[dataset_new$Q58R2_numeric == 1] <- "Protestant"
+dataset_new$Religion[dataset_new$Q58R3_numeric == 1] <- "Other Christian"
+dataset_new$Religion[dataset_new$Q58R4_numeric == 1] <- "Jewish"
+dataset_new$Religion[dataset_new$Q58R5_numeric == 1] <- "Muslim"
+dataset_new$Religion[dataset_new$Q58R6_numeric == 1] <- "Hindu"
+dataset_new$Religion[dataset_new$Q58R7_numeric == 1] <- "Buddhist"
+dataset_new$Religion[dataset_new$Q58R8_numeric == 1] <- "Mormon/LDS"
+dataset_new$Religion[dataset_new$Q58R9_numeric == 1] <- "Folk Religion"
+dataset_new$Religion[dataset_new$Q58R10_numeric == 1] <- "Daoist"
+dataset_new$Religion[dataset_new$Q58R11_numeric == 1] <- "Shamanism"
+dataset_new$Religion[dataset_new$Q58R12_numeric == 1] <- "Ancestral Religion"
+dataset_new$Religion[dataset_new$Q58R13_numeric == 1] <- "Spiritual but not religious"
+dataset_new$Religion[dataset_new$Q58R14_numeric == 1] <- "Atheist"
+dataset_new$Religion[dataset_new$Q58R15_numeric == 1] <- "Agnostic"
 
-# Assign religion based on the numeric values of each Q58R column
-religion_mapping <- c(
-  "Q58R1_numeric" = "Catholic", 
-  "Q58R2_numeric" = "Protestant", 
-  "Q58R3_numeric" = "Other Christian",
-  "Q58R4_numeric" = "Jewish", 
-  "Q58R5_numeric" = "Muslim", 
-  "Q58R6_numeric" = "Hindu", 
-  "Q58R7_numeric" = "Buddhist", 
-  "Q58R8_numeric" = "Mormon/LDS", 
-  "Q58R9_numeric" = "Folk Religion", 
-  "Q58R10_numeric" = "Daoist", 
-  "Q58R11_numeric" = "Shamanism", 
-  "Q58R12_numeric" = "Ancestral Religion", 
-  "Q58R13_numeric" = "Spiritual but not religious", 
-  "Q58R14_numeric" = "Atheist", 
-  "Q58R15_numeric" = "Agnostic"
-)
-
-# Loop through the mapping to assign values
-for (col in names(religion_mapping)) {
-  dataset_new$Religion[dataset_new[[col]] == 1] <- religion_mapping[[col]]
-}
+# Check if an individual reported multiple religious affiliations 
+#dataset_new$multiple_religions <- rowSums(dataset_new[, grep("_numeric", names(dataset_new))] == 1)
+#check <- table(dataset_new$multiple_religions)
 
@@
 
-# Convert Religion to a factor with specific levels
+# Convert Religion to a factor
 dataset_new$Religion <- factor(dataset_new$Religion,
@@
 redundant_cols <- c(
-  paste0("Q58R", 1:15), paste0("Q58R", 1:15, "_numeric")
+  "Q58R1", "Q58R2", "Q58R3", "Q58R4", "Q58R5", "Q58R6", "Q58R7", "Q58R8", 
+  "Q58R9", "Q58R10", "Q58R11", "Q58R12", "Q58R13", "Q58R14", "Q58R15",
+  "Q58R1_numeric", "Q58R2_numeric", "Q58R3_numeric", "Q58R4_numeric", 
+  "Q58R5_numeric", "Q58R6_numeric", "Q58R7_numeric", "Q58R8_numeric", 
+  "Q58R9_numeric", "Q58R10_numeric", "Q58R11_numeric", "Q58R12_numeric", 
+  "Q58R13_numeric", "Q58R14_numeric", "Q58R15_numeric"
 )
@@
 head(dataset_clean)
-
 ```
-
-We first consolidated multiple columns related to respondents' religious affiliations into a single, categorical variable called Religion.
-We also add another option called 'Multiple Religions' where respondents identified with more than one religion.
 
@@
 
-# Convert to a data frame
+# Convert to a data frame for better readability
 missing_summary <- data.frame(Variable = names(missing_proportions),
@@
 head(dataset_clean)
+#colnames(dataset_clean)
 ```
 
-## Rename the remaining column headers for ease of understanding
+# Rename the remaining column headers for ease of understanding
+
 ```{r}
@@
   "DiscriminationImpactOnLifeSatisfaction", 
-  "PerceptionsOfDiscriminationAgainstWhites",
   "PerceptionsOfDiscriminationAgainstBlacks", 
@@
   "PerceptionsOfDiscriminationAgainstLatinos", 
-  "PerceptionsOfDiscriminationAgainstGaysAndLesbians",
-  "PerceptionsOfDiscriminationAgainstMuslims",
   "Religion"
 )
-
+colnames(dataset_clean)
 ```
@@
 
-# Binomial Logistic Regression
+```{r}
+# Fit a logistic regression model
+model <- glm(Q1R5 ~ S2_RACE_PRIME + # Primary Race
+               S3 +                 # Orientation
+               S3B +                # Gender
+               S5_AGE +             # Age
+               S13 +                # Highest level of education
+               S14 +                # Community lived in
+               Q627 +               # Treated unfairly or discriminated against
+               Q633 +               # Discrimination impact on life satisfaction
+               Q619_Q626R2 +        # Perceptions of discrimination against Blacks
+               Q619_Q626R3 +        # Perceptions of discrimination against Asian Americans
+               Q619_Q626R4 +        # Perceptions of discrimination against Native Americans
+               Q619_Q626R5 +        # Perceptions of discrimination against Immigrants
+               Q619_Q626R6 +        # Perceptions of discrimination against Latinos
+               Religion +
+               S3B * Q627,
+             data = dataset_clean, family = binomial)
 
-## Model
-
-```{r}
-model <- glm(ImportanceOfStoppingDiscriminationAgainstMinorities ~ PrimaryRace +
-               Orientation + Gender + Age + 
-               HighestLevelOfEducation + CommunityLivedIn +
-               TreatedUnfairlyOrDiscriminatedAgainst + 
-               DiscriminationImpactOnLifeSatisfaction + 
-               PerceptionsOfDiscriminationAgainstWhites + 
-               PerceptionsOfDiscriminationAgainstBlacks +
-               PerceptionsOfDiscriminationAgainstAsianAmericans + 
-               PerceptionsOfDiscriminationAgainstNativeAmericans + 
-               PerceptionsOfDiscriminationAgainstImmigrants + 
-               PerceptionsOfDiscriminationAgainstLatinos + 
-               PerceptionsOfDiscriminationAgainstGaysAndLesbians + 
-               PerceptionsOfDiscriminationAgainstMuslims + 
-               Religion,
-
-        data = dataset_clean, family = "binomial")
-
-# summary(model)
+#summary(model)
 ```
@@
 ```{r}
-# Extract odds ratios
-#odds_ratios <- exp(coef(model))
-
-# Get confidence intervals for coefficients
-#conf_intervals <- exp(confint(model))
-
-# Extract p-values
-#p_values <- summary(model)$coefficients[, 4]
-
-# Combine everything into a single data frame
-#results_model <- data.frame(
-#  OddsRatio = odds_ratios,
-#  LowerCI = conf_intervals[, 1],
-#  UpperCI = conf_intervals[, 2],
-#  PValue = p_values
-#)
-
-# Display results
-# results_model
-
-# Check multicollinearity using VIF
-# vif(model)
-```
-
-## Model Stepwise
-
-```{r}
 # Perform stepwise selection
-model_stepwise <- step(model, 
-                       direction = "both",
-                       trace = 0)
+stepwise_model <- step(model, direction = "both", trace = 1)
 ```
@@
 ```{r}
-# Extract odds ratios
-odds_ratios <- exp(coef(model_stepwise))
+# New model based on stepwise regression
+model_final <- glm(Q1R5 ~ S2_RACE_PRIME + # Primary Race
+#REMOVED               S3 +                 # Orientation
+               S3B +                # Gender
+               S5_AGE +             # Age
+               S13 +                # Highest level of education
+#REMOVED               S14 +                # Community lived in
+               Q627 +               # Treated unfairly or discriminated against
+               Q633 +               # Discrimination impact on life satisfaction
+               Q619_Q626R2 +        # Perceptions of discrimination against Blacks
+               Q619_Q626R3 +        # Perceptions of discrimination against Asian Americans
+               Q619_Q626R4 +        # Perceptions of discrimination against Native Americans
+               Q619_Q626R5,        # Perceptions of discrimination against Immigrants
+#REMOVED               Q619_Q626R6 +        # Perceptions of discrimination against Latinos
+#REMOVED               Religion +
+#REMOVED              S2_RACE_PRIME * S3B,
 
-# Get confidence intervals for coefficients
-conf_intervals <- exp(confint(model_stepwise))
+            data = dataset_clean, family = binomial)
 
-# Extract p-values
-p_values <- summary(model_stepwise)$coefficients[, 4]
+summary(model_final)
+```
 
-# Combine everything into a single data frame
-results_model_stepwise <- data.frame(
-  OddsRatio = odds_ratios,
-  LowerCI = conf_intervals[, 1],
-  UpperCI = conf_intervals[, 2],
-  PValue = p_values
-)
+# Diagnostic Checks
 
-# Display results
-results_model_stepwise
+```{r}
+# Calculate VIFs
+vif_results <- vif(model)  # Uses type = "terms" automatically if predictors are weighted or non-linear
+vif_results_model_final <- vif(model_final)
 
-# Check multicollinearity using VIF
-# vif(model_stepwise)
+print(vif_results)
+print(vif_results_model_final)
 ```
@@
 ```{r}
-predicted_classes <- ifelse(predicted_probabilities > 0.7, "Not important", "Important")
-actual_classes <- dataset_clean$ImportanceOfStoppingDiscriminationAgainstMinorities
+plot(residuals(model, type = "deviance"))
+plot(residuals(model_final, type = "deviance"))
+
+influencePlot(model)
+influencePlot(model_final)
 ```
@@
 ```{r}
-#predicted_classes <- ifelse(predicted_probabilities > 0.5, 1, 0)
-confusion_matrix <- table(Predicted = predicted_classes, Actual = actual_classes)
-print(confusion_matrix)
-
+levels(as.factor(predictions))
+levels(as.factor(dataset_clean$Q1R5))
 ```
@@
 ```{r}
-library(caret)
-confusion <- confusionMatrix(factor(predicted_classes), factor(actual_classes))
-print(confusion)
-```
+levels(as.factor(testData$Q1R5))
+predictions <- predict(model_final, testData)
 
-```{r}
+set.seed(123)  # For reproducibility
+trainIndex <- createDataPartition(dataset_clean$Q1R5, p = 0.8, list = FALSE)
+trainData <- dataset_clean[trainIndex, ]
+testData <- dataset_clean[-trainIndex, ]
+
+levels(as.factor(predictions))
+levels(as.factor(testData$Q1R5))
+
+confusionMatrix(predictions, testData$Q1R5)
 

### 2024-12-06 peter_models.qmd
adding files
update chi-sq test, start modeling

--- 2024-12-02 peter_EDA.qmd
@@
 ---
-title: "peter_EDA"
+title: "peter_models"
 format: html
@@
 
-Conducting EDA on the 2020 CMPS dataset for Research Question 1
+##
+
+# Loading Packages
+
+```{r}
+library(tidyverse)
+library(ordinal)
+library(car)
+library(brant)
+library(MASS)
+```
 
@@
 
-# Selecting Variables
-
-# Cleaning Variables
-
-# Analysis of Missing Data
-
-# Save cleaned data
-
-# Creating summary table
+```{r}
+load("/Users/pdeguz01/Documents/git/Data/2020 CMPS/Output/2020cmps_clean.RData")
+```
 
@@
 
+```{r}
+# Create a summary table for all categorical variables
+variables <- c("Q29", "Q271", "S2_RACE_PRIME", "S3", "S3B", "S5_AGE", "S13", "S14", "S7", "Q21", "S2_M", "Q627", "Q628", "Q629R1", "Religion", "educ", "sex_orientation", "rural", "white", "race_recoded", "age_recoded")
+
+# Chi-square tests for each variable
+chi_square_results <- lapply(variables, function(var) {
+  chisq.test(table(dataset_clean[[var]], dataset_clean$Q29))
+})
+
+# Print chi-square results
+names(chi_square_results) <- variables
+print(chi_square_results)
+```
+
+| Variable | Chi-Sq Test Statistic | Degrees of Freedom (df) | p-value |
+|----|----|----|----|
+| Q271 | 291.49 | 12 | \< 2.2e-16 |
+| S2_RACE_PRIME | 291.12 | 21 | \< 2.2e-16 |
+| S3 | 101.7 | 15 | 6.196e-15 |
+| S3B | 296.35 | 9 | \< 2.2e-16 |
+| S5_AGE | NaN | 18 | NA |
+| S13 | 710.69 | 18 | \< 2.2e-16 |
+| S14 | 196.79 | 12 | \< 2.2e-16 |
+| S7 | 148.62 | 6 | \< 2.2e-16 |
+| Q21 | 1284.1 | 9 | \< 2.2e-16 |
+| S2_M | 6.0327 | 3 | 0.11 |
+| Q627 | 246.82 | 3 | \< 2.2e-16 |
+| Q628 | 9.0318 | 6 | 0.1718 |
+| Q629R1 | 6.9533 | 3 | 0.0734 |
+| Religion | 562.39 | 45 | \< 2.2e-16 |
+| educ | 608.65 | 6 | \< 2.2e-16 |
+| sex_orientation | 55.679 | 6 | 3.379e-10 |
+| rural/urban&suburban | 83.719 | 3 | \< 2.2e-16 |
+| white/non-white | 75.137 | 3 | 3.386e-16 |
+| race_recoded | 276.36 | 18 | \< 2.2e-16 |
+| age_recoded | 396.97 | 15 | \< 2.2e-16 |
+
+## Takeaways from Chi-Square Test
+
 # Fitting OLR Model
+```{r}
+
+```
+
+## Interpretation
 
 # Fitting MLR Model
+```{r}
+
+```
+
+## Interpretation
+
+# Fitting Proportional Odds Model
+```{r}
+
+```
+
+## Interpretation

### 2024-12-06 peter_EDA.qmd
updating recoding and plotting
updating data cleaning, plotting script, and plots
recoding vars into factors and running OLR model

@@
                                                                    S2_RACE_PRIME == "(7) Native Hawaiian" ~ "NHPI",
-                                                                   S2_RACE_PRIME == "(8) Not Hawaiian, but other Pacific Islander" ~ "NHPI"))
+                                                                   S2_RACE_PRIME == "(8) Not Hawaiian, but other Pacific Islander" ~ "NHPI"),
+                                          age_recoded = case_when(S5_AGE == "(1) 16-17" ~ "16-17",
+                                                                  S5_AGE == "(2) 18-29" ~ "18-29",
+                                                                  S5_AGE == "(3) 30-39" ~ "30-39",
+                                                                  S5_AGE == "(4) 40-49" ~ "40-49",
+                                                                  S5_AGE == "(5) 50-59" ~ "50-59",
+                                                                  S5_AGE == "(6) 60-69" ~ "60-69",
+                                                                  S5_AGE == "(7) 70 +" ~ "70+"))
+
+dataset_clean <- dataset_clean %>% mutate(race_recoded = factor(dataset_clean$race_recoded, levels = c("White", "Black", "Asian American", "Latino", "AIAN", "NHPI", "Arab/MENA")),
+                                         age_recoded = factor(dataset_clean$age_recoded, levels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+")),
+                                         educ = factor(dataset_clean$educ, levels = c("No college experience", "Some college experience", "College graduate")),
+                                         white = factor(dataset_clean$white, levels = c("White", "Non-White")),
+                                         rural = factor(dataset_clean$rural, levels= c("Rural", "Urban/suburban")),
+                                         sex_orientation = factor(dataset_clean$sex_orientation, levels = c("Straight", "Queer", "Refused/Don't Know")))
 

### 2024-12-06 peter_models.qmd
recoding vars into factors and running OLR model

@@
 
+Our outcome variable is Q29 (interest in politics).
+
+There appears to be a significant relationship between the following variables and Q29 (interest in politics):
+
+-   Age (Recoded)
+
+-   Race (Recoded)
+
+-   Gender
+
+-   White or Non-White status
+
+-   Rural or Urban & Suburban status
+
+-   Sexual Orientation
+
+-   Religion
+
+-   Educational Attainment
+
+-   Having experienced prior discrimination (Q627)
+
+-   Political Party ID
+
+-   Nativity (born in US or other)
+
 # Fitting OLR Model
+
 ```{r}
+dataset_clean$Q271_simple <- as.numeric(dataset_clean$Q271)
+dataset_clean$Q271 <- factor(dataset_clean$Q271,
+                            levels = c("(5) Not at all important",
+                                     "(4) Slightly important",
+                                     "(3) Moderately important", 
+                                     "(2) Very important",
+                                     "(1) Extremely important"),
+                            ordered = TRUE)
+
+m1 <- clm(Q29 ~ Q271 + race_recoded + age_recoded +
+            rural + # Rural or Urban/Suburban
+            sex_orientation + # Sexual Orientation
+            Religion + 
+            educ + # Educational Attainment
+            S3B + # Gender
+            S7 + # Nativity
+            Q21, # Party affiliation
+          data = dataset_clean)
+
+summary(m1)
+
+# Check for multicollinearity using a linear approximation
+vif_m1 <- vif(lm(as.numeric(Q29) ~ Q271 + race_recoded + age_recoded +
+            rural + # Rural or Urban/Suburban
+            sex_orientation + # Sexual Orientation
+            Religion + 
+            educ + # Educational Attainment
+            S3B + # Gender
+            S7 + # Nativity
+            Q21, # Party affiliation
+          data = dataset_clean))
+
+print("Variance Inflation Factors:")
+print(vif_m1)
+
+# Test proportional odds assumption
+print("Testing proportional odds assumption:")
+test1 <- nominal_test(m1)
+print("Nominal test for Model 1:")
+print(test1)
+
+# Alternative diagnostic approach using linear approximation
+lm_approx <- lm(as.numeric(Q29) ~ Q271 + race_recoded + age_recoded +
+            rural + # Rural or Urban/Suburban
+            sex_orientation + # Sexual Orientation
+            Religion + 
+            educ + # Educational Attainment
+            S3B + # Gender
+            S7 + # Nativity
+            Q21, # Party affiliation
+          data = dataset_clean)
+diagnostics_df <- data.frame(
+  fitted = fitted(lm_approx),
+  residuals = residuals(lm_approx),
+  index = 1:nrow(dataset_clean)
+)
+
+# Plot residuals vs fitted values using linear approximation
+p1 <- ggplot(diagnostics_df, aes(x = fitted, y = residuals)) +
+  geom_point(alpha = 0.1) +
+  geom_smooth(method = "loess") +
+  theme_minimal() +
+  labs(title = "Residuals vs Fitted Values (Linear Approximation)",
+       x = "Fitted Values",
+       y = "Residuals")
+print(p1)
+
+# Calculate and plot Cook's distance
+cooks_d <- cooks.distance(lm_approx)
+influence_df <- data.frame(
+  index = 1:length(cooks_d),
+  cooks_distance = cooks_d
+)
+
+p2 <- ggplot(influence_df, aes(x = index, y = cooks_distance)) +
+  geom_point() +
+  geom_hline(yintercept = 4/nrow(dataset_clean), color = "red", linetype = "dashed") +
+  theme_minimal() +
+  labs(title = "Cook's Distance Plot",
+       x = "Observation Number",
+       y = "Cook's Distance")
+print(p2)
+
+# Count influential observations
+influential_obs <- which(cooks_d > 4/nrow(dataset_clean))
+print("Number of influential observations:")
+print(length(influential_obs))
+
+# Create prediction visualization
+pred_data <- expand.grid(
+  Q271 = levels(dataset_clean$Q271),
+  race_recoded = levels(dataset_clean$race_recoded), 
+  age_recoded = levels(dataset_clean$age_recoded),
+  rural = levels(dataset_clean$rural),
+  sex_orientation = levels(dataset_clean$sex_orientation),
+  Religion = levels(dataset_clean$Religion),
+  educ = levels(dataset_clean$educ),
+  S3B = levels(dataset_clean$S3B),
+  S7 = levels(dataset_clean$S7),
+  Q21 = levels(dataset_clean$Q21)
+)
+
+# Get predicted probabilities
+preds <- predict(m1, newdata = pred_data, type = "prob")
+# Convert predictions to numeric scores (weighted average)
+pred_data$predicted_score <- apply(preds$fit, 1, function(x) sum(x * 1:4))
+
+# Create prediction plot
+p3 <- ggplot(pred_data, aes(x = Q271, y = predicted_score, color = race_recoded)) +
+  geom_point() +
+  geom_line(aes(group = race_recoded)) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  labs(title = "Predicted Political Interest by Racial Identity Importance and Race",
+       x = "Importance of Racial Identity",
+       y = "Predicted Political Interest Level (1-4)",
+       color = "Race/Ethnicity")
+print(p3)
+
+# Print the coefficients table
+print("Coefficients for Model 1:")
+coef_table <- summary(m1)$coefficients
+print(coef_table)
+
+# Calculate and print odds ratios for the coefficients
+odds_ratios <- exp(coef_table[, "Estimate"])
+print("Odds Ratios:")
+print(odds_ratios)
 
@@
 
+1.  Model Fit:
+    1.  AIC is 42342.54
+    2.  Log-likelihood is -21125.27
+2.  Proportional Odds Assumption
+
+-   The nominal test for Model 1 shows significant violations (p \< 0.05 for all predictor variables except for Religion
+
+-   This suggests the effect of racial identity importance varies across different levels of political interest
+
+1.  Multicollinearity
+
+-   Interpretation of VIF values
+
+1.  Residuals and Model Diagnostics
+
+-   The residuals plot shows:
+
+    -   A relatively flat line around zero
+
+    -   Straight downward slope patterns in the residuals
+
+1.  Influential Points:
+    -   There are 804 influential observations.
+
+    -   Cook's distance plot shows multiple observations above the threshold.
+
+    -   Given the large sample size (17,545), this number of influential points is not extremely concerning.
+2.  Key Findings from Model 1:
+    1.  The different levels of racial identity importance are all significant
+    2.  There are differences among racial/ethnic groups
+3.  Predicted Relationships (from plot)
+    1.  There appears to be a positive relationship between racial identity importance and political interest
+
 # Fitting MLR Model
+
 ```{r}
+
+# Load packages
+library(nnet)  # for multinomial logistic regression
+library(tidyverse)
+library(car)
+library(effects)
+
+# ensure outcome variable is a factor but not ordered
+dataset_clean$Q29_unordered <- factor(dataset_clean$Q29, ordered = FALSE)
+
+# Fit multinomial model
+m_mult <- multinom(Q29_unordered ~ Q271 + race_recoded + age_recoded +
+            rural + # Rural or Urban/Suburban
+            sex_orientation + # Sexual Orientation
+            Religion + 
+            educ + # Educational Attainment
+            S3B + # Gender
+            S7 + # Nativity
+            Q21, # Party affiliation
+          data = dataset_clean)
+
+# Calculate predicted probabilities of marginal effects
+# Create sequence of Q271_simple values
+x_vals <- seq(min(dataset_clean$Q271_simple), max(dataset_clean$Q271_simple), length.out = 100)
+
+# Function to calculate average predicted probabilities
+get_avg_probs <- function(model, newdata, var_name) {
+    preds <- predict(model, newdata = newdata, type = "probs")
+    colMeans(preds)
+}
+
+# Calculate marginal effects by race
+races <- unique(dataset_clean$race_recoded)
+marg_effects <- list()
+
+for(race in races) {
+    probs <- matrix(0, nrow = length(x_vals), ncol = length(levels(dataset_clean$Q29_unordered)))
+    for(i in seq_along(x_vals)) {
+        temp_data <- dataset_clean
+        temp_data$Q271_simple <- x_vals[i]
+        temp_data$race_major <- race
+        probs[i,] <- get_avg_probs(m_mult, temp_data, "Q271_simple")
+    }
+    marg_effects[[race]] <- data.frame(
+        x = x_vals,
+        race = race,
+        prob1 = probs[,1],
+        prob2 = probs[,2],
+        prob3 = probs[,3],
+        prob4 = probs[,4]
+    )
+}
+
+# Combine all marginal effects
+marg_effects_df <- bind_rows(marg_effects)
+
+# Create plots for marginal effects
+# Reshape data for plotting
+marg_effects_long <- marg_effects_df %>%
+    pivot_longer(cols = starts_with("prob"),
+                names_to = "outcome",
+                values_to = "probability") %>%
+    mutate(outcome = factor(outcome,
+                          levels = paste0("prob", 1:4),
+                          labels = levels(dataset_clean$Q29_unordered)))
+
+# Plot marginal effects
+p1 <- ggplot(marg_effects_long, 
+             aes(x = x, y = probability, color = outcome)) +
+    geom_line() +
+    facet_wrap(~race) +
+    theme_minimal() +
+    labs(title = "Marginal Effects of Racial Identity Importance by Race",
+         x = "Importance of Racial Identity",
+         y = "Predicted Probability",
+         color = "Political Interest Level") +
+    theme(legend.position = "bottom")
+print(p1)
+
+# Model fit statistics and comparisons
+print("Model summaries:")
+print(summary(m_mult))
+
+# Calculate McFadden's R-squared
+null_model <- multinom(Q29_unordered ~ 1, data = dataset_clean)
+mcfadden_r2 <- 1 - logLik(m_mult)/logLik(null_model)
+
+print("McFadden's R-squared:")
+print(Basic = mcfadden_r2)
+
+# Classification accuracy
+pred_class <- predict(m_mult)
+conf_matrix <- table(pred_class, dataset_clean$Q29_unordered)
+accuracy <- sum(diag(conf_matrix))/sum(conf_matrix)
+
+print("Classification accuracy:")
+print(accuracy)
+print("Confusion Matrix:")
+print(conf_matrix)
+
+# Print AIC
+print(Multinomial = AIC(m_mult))
+
+# Calculate predicted probabilities
+pred_data <- expand.grid(
+    Q271_simple = seq(1, 5, by = 0.1),
+    race_recoded = unique(dataset_clean$race_recoded)
+)
+
+# Get predictions
+preds <- predict(m_mult, newdata = pred_data, type = "probs")
+pred_data <- cbind(pred_data, as.data.frame(preds))
+
+# Create separate plots for each outcome level
+pred_long <- pred_data %>%
+    pivot_longer(cols = starts_with("("),
+                names_to = "interest_level",
+                values_to = "probability")
+
+# Plot for each outcome level
+p2 <- ggplot(pred_long, 
+             aes(x = Q271_simple, y = probability, color = race_recoded)) +
+    geom_line() +
+    facet_wrap(~interest_level) +
+    theme_minimal() +
+    labs(title = "Predicted Probabilities by Outcome Level",
+         x = "Importance of Racial Identity",
+         y = "Probability",
+         color = "Race/Ethnicity") +
+    theme(legend.position = "bottom")
+print(p2)
 
@@
 
-# Fitting Proportional Odds Model
+# Fitting Partial Proportional Odds Model
+
 ```{r}
+# Load required packages
+library(VGAM)
+library(ordinal)
+library(tidyverse)
+library(sandwich)
+
+# Create ordered 3-category version
+
+# Fit PPO model
+
+# create prediction data
+
+# get predictions for PPO model
+
+# Reshape for plotting
+
+# create plot
+
+# Model summaries
+
+# Cross-validation
+
+# calculate variable importance for PPO model
 
@@
 ## Interpretation
+
+# PPO Model with Interactions

### 2024-12-07 peter_models.qmd
update models

@@
 
-# Fitting OLR Model
+# Fitting OLR Model Using polr()
 
 ```{r}
-dataset_clean$Q271_simple <- as.numeric(dataset_clean$Q271)
-dataset_clean$Q271 <- factor(dataset_clean$Q271,
+library(MASS) # package for model
+library(caret) # package for confusion matrix
+
+dataset_clean$Q271_reordered <- factor(dataset_clean$Q271,
                             levels = c("(5) Not at all important",
@@
                                      "(2) Very important",
-                                     "(1) Extremely important"),
-                            ordered = TRUE)
+                                     "(1) Extremely important"))
+
+ord_mod <- polr(Q29 ~ Q271_reordered + race_recoded + age_recoded +
+            rural + # Rural or Urban/Suburban
+            sex_orientation + # Sexual Orientation
+            educ + # Educational Attainment
+            S3B + # Gender
+            S7 + # Nativity
+            Q21, # Party affiliation
+          data = dataset_clean, Hess=TRUE)
+
+summary(ord_mod)
+
+pvals <- pnorm(-abs(summary(ord_mod)$coef[,"t value"]))*2
+ctable <- cbind(summary(ord_mod)$coef,pvals)
+
+ctable
+
+(exp_coefs <- exp(cbind(OR=coef(ord_mod),confint(ord_mod))))
+
+head(ord_mod$fitted.values)
+head(predict(ord_mod))
+
+confusionMatrix(predict(ord_mod), dataset_clean$Q29, mode="everything")
+
+```
+
+```{r}
+# generate predictions for the ordinal model
+
+new_dataset <- data.frame
+
+predict(ord_mod, new_dataset)
+```
+
+# Fitting MLR Model
+
+```{r}
+
+# Load packages
+library(nnet)  # for multinomial logistic regression
+library(tidyverse)
+library(car)
+library(effects)
+
+# ensure outcome variable is a factor but not ordered
+dataset_clean$Q29_unordered <- factor(dataset_clean$Q29, ordered = FALSE)
+
+# Fit multinomial model
+m_mult <- multinom(Q29_unordered ~ Q271 + race_recoded + age_recoded +
+            rural + # Rural or Urban/Suburban
+            sex_orientation + # Sexual Orientation
+            educ + # Educational Attainment
+            S3B + # Gender
+            S7 + # Nativity
+            Q21, # Party affiliation
+          data = dataset_clean)
+
+# Calculate predicted probabilities of marginal effects
+# Create sequence of Q271_simple values
+x_vals <- seq(min(dataset_clean$Q271_simple), max(dataset_clean$Q271_simple), length.out = 100)
+
+# Function to calculate average predicted probabilities
+get_avg_probs <- function(model, newdata, var_name) {
+    preds <- predict(model, newdata = newdata, type = "probs")
+    colMeans(preds)
+}
+
+# Calculate marginal effects by race
+races <- unique(dataset_clean$race_recoded)
+marg_effects <- list()
+
+for(race in races) {
+    probs <- matrix(0, nrow = length(x_vals), ncol = length(levels(dataset_clean$Q29_unordered)))
+    for(i in seq_along(x_vals)) {
+        temp_data <- dataset_clean
+        temp_data$Q271_simple <- x_vals[i]
+        temp_data$race_major <- race
+        probs[i,] <- get_avg_probs(m_mult, temp_data, "Q271_simple")
+    }
+    marg_effects[[race]] <- data.frame(
+        x = x_vals,
+        race = race,
+        prob1 = probs[,1],
+        prob2 = probs[,2],
+        prob3 = probs[,3],
+        prob4 = probs[,4]
+    )
+}
+
+# Combine all marginal effects
+marg_effects_df <- bind_rows(marg_effects)
+
+# Create plots for marginal effects
+# Reshape data for plotting
+marg_effects_long <- marg_effects_df %>%
+    pivot_longer(cols = starts_with("prob"),
+                names_to = "outcome",
+                values_to = "probability") %>%
+    mutate(outcome = factor(outcome,
+                          levels = paste0("prob", 1:4),
+                          labels = levels(dataset_clean$Q29_unordered)))
+
+# Plot marginal effects
+p1 <- ggplot(marg_effects_long, 
+             aes(x = x, y = probability, color = outcome)) +
+    geom_line() +
+    facet_wrap(~race) +
+    theme_minimal() +
+    labs(title = "Marginal Effects of Racial Identity Importance by Race",
+         x = "Importance of Racial Identity",
+         y = "Predicted Probability",
+         color = "Political Interest Level") +
+    theme(legend.position = "bottom")
+print(p1)
+
+# Model fit statistics and comparisons
+print("Model summaries:")
+print(summary(m_mult))
+
+# Calculate McFadden's R-squared
+null_model <- multinom(Q29_unordered ~ 1, data = dataset_clean)
+mcfadden_r2 <- 1 - logLik(m_mult)/logLik(null_model)
+
+print("McFadden's R-squared:")
+print(mcfadden_r2)
+
+# Classification accuracy
+pred_class <- predict(m_mult)
+conf_matrix <- table(pred_class, dataset_clean$Q29_unordered)
+accuracy <- sum(diag(conf_matrix))/sum(conf_matrix)
+
+print("Classification accuracy:")
+print(accuracy)
+print("Confusion Matrix:")
+print(conf_matrix)
+
+# Print AIC
+print(AIC(m_mult))
+
+# Calculate predicted probabilities
+pred_data <- expand.grid(
+    Q271_simple = seq(1, 5, by = 0.1),
+    race_recoded = unique(dataset_clean$race_recoded)
+)
+
+# Get predictions
+preds <- predict(m_mult, newdata = pred_data, type = "probs")
+pred_data <- cbind(pred_data, as.data.frame(preds))
+
+# Create separate plots for each outcome level
+pred_long <- pred_data %>%
+    pivot_longer(cols = starts_with("("),
+                names_to = "interest_level",
+                values_to = "probability")
+
+# Plot for each outcome level
+p2 <- ggplot(pred_long, 
+             aes(x = Q271_simple, y = probability, color = race_recoded)) +
+    geom_line() +
+    facet_wrap(~interest_level) +
+    theme_minimal() +
+    labs(title = "Predicted Probabilities by Outcome Level",
+         x = "Importance of Racial Identity",
+         y = "Probability",
+         color = "Race/Ethnicity") +
+    theme(legend.position = "bottom")
+print(p2)
+
+```
+
+## Interpretation
+
+# Fitting OLR Model Using clm() (Following Jennifer's method)
+
+```{r}
+dataset_clean$Q271_simple <- as.numeric(dataset_clean$Q271)
+# dataset_clean$Q271 <- factor(dataset_clean$Q271,
+#                             levels = c("(5) Not at all important",
+#                                      "(4) Slightly important",
+#                                      "(3) Moderately important",
+#                                      "(2) Very important",
+#                                      "(1) Extremely important"),
+#                             ordered = TRUE)
 
@@
 
--   The nominal test for Model 1 shows significant violations (p \< 0.05 for all predictor variables except for Religion
+-   The nominal test for Model 1 shows significant violations (p \< 0.05 for all predictor variables except for Religion)
 
@@
 3.  Predicted Relationships (from plot)
-    1.  There appears to be a positive relationship between racial identity importance and political interest
-
-# Fitting MLR Model
-
-```{r}
-
-# Load packages
-library(nnet)  # for multinomial logistic regression
-library(tidyverse)
-library(car)
-library(effects)
-
-# ensure outcome variable is a factor but not ordered
-dataset_clean$Q29_unordered <- factor(dataset_clean$Q29, ordered = FALSE)
-
-# Fit multinomial model
-m_mult <- multinom(Q29_unordered ~ Q271 + race_recoded + age_recoded +
-            rural + # Rural or Urban/Suburban
-            sex_orientation + # Sexual Orientation
-            Religion + 
-            educ + # Educational Attainment
-            S3B + # Gender
-            S7 + # Nativity
-            Q21, # Party affiliation
-          data = dataset_clean)
-
-# Calculate predicted probabilities of marginal effects
-# Create sequence of Q271_simple values
-x_vals <- seq(min(dataset_clean$Q271_simple), max(dataset_clean$Q271_simple), length.out = 100)
-
-# Function to calculate average predicted probabilities
-get_avg_probs <- function(model, newdata, var_name) {
-    preds <- predict(model, newdata = newdata, type = "probs")
-    colMeans(preds)
-}
-
-# Calculate marginal effects by race
-races <- unique(dataset_clean$race_recoded)
-marg_effects <- list()
-
-for(race in races) {
-    probs <- matrix(0, nrow = length(x_vals), ncol = length(levels(dataset_clean$Q29_unordered)))
-    for(i in seq_along(x_vals)) {
-        temp_data <- dataset_clean
-        temp_data$Q271_simple <- x_vals[i]
-        temp_data$race_major <- race
-        probs[i,] <- get_avg_probs(m_mult, temp_data, "Q271_simple")
-    }
-    marg_effects[[race]] <- data.frame(
-        x = x_vals,
-        race = race,
-        prob1 = probs[,1],
-        prob2 = probs[,2],
-        prob3 = probs[,3],
-        prob4 = probs[,4]
-    )
-}
-
-# Combine all marginal effects
-marg_effects_df <- bind_rows(marg_effects)
-
-# Create plots for marginal effects
-# Reshape data for plotting
-marg_effects_long <- marg_effects_df %>%
-    pivot_longer(cols = starts_with("prob"),
-                names_to = "outcome",
-                values_to = "probability") %>%
-    mutate(outcome = factor(outcome,
-                          levels = paste0("prob", 1:4),
-                          labels = levels(dataset_clean$Q29_unordered)))
-
-# Plot marginal effects
-p1 <- ggplot(marg_effects_long, 
-             aes(x = x, y = probability, color = outcome)) +
-    geom_line() +
-    facet_wrap(~race) +
-    theme_minimal() +
-    labs(title = "Marginal Effects of Racial Identity Importance by Race",
-         x = "Importance of Racial Identity",
-         y = "Predicted Probability",
-         color = "Political Interest Level") +
-    theme(legend.position = "bottom")
-print(p1)
-
-# Model fit statistics and comparisons
-print("Model summaries:")
-print(summary(m_mult))
-
-# Calculate McFadden's R-squared
-null_model <- multinom(Q29_unordered ~ 1, data = dataset_clean)
-mcfadden_r2 <- 1 - logLik(m_mult)/logLik(null_model)
-
-print("McFadden's R-squared:")
-print(Basic = mcfadden_r2)
-
-# Classification accuracy
-pred_class <- predict(m_mult)
-conf_matrix <- table(pred_class, dataset_clean$Q29_unordered)
-accuracy <- sum(diag(conf_matrix))/sum(conf_matrix)
-
-print("Classification accuracy:")
-print(accuracy)
-print("Confusion Matrix:")
-print(conf_matrix)
-
-# Print AIC
-print(Multinomial = AIC(m_mult))
-
-# Calculate predicted probabilities
-pred_data <- expand.grid(
-    Q271_simple = seq(1, 5, by = 0.1),
-    race_recoded = unique(dataset_clean$race_recoded)
-)
-
-# Get predictions
-preds <- predict(m_mult, newdata = pred_data, type = "probs")
-pred_data <- cbind(pred_data, as.data.frame(preds))
-
-# Create separate plots for each outcome level
-pred_long <- pred_data %>%
-    pivot_longer(cols = starts_with("("),
-                names_to = "interest_level",
-                values_to = "probability")
-
-# Plot for each outcome level
-p2 <- ggplot(pred_long, 
-             aes(x = Q271_simple, y = probability, color = race_recoded)) +
-    geom_line() +
-    facet_wrap(~interest_level) +
-    theme_minimal() +
-    labs(title = "Predicted Probabilities by Outcome Level",
-         x = "Importance of Racial Identity",
-         y = "Probability",
-         color = "Race/Ethnicity") +
-    theme(legend.position = "bottom")
-print(p2)
-
-```
-
-## Interpretation
-
-# Fitting Partial Proportional Odds Model
-
-```{r}
-# Load required packages
-library(VGAM)
-library(ordinal)
-library(tidyverse)
-library(sandwich)
-
-# Create ordered 3-category version
-
-# Fit PPO model
-
-# create prediction data
-
-# get predictions for PPO model
-
-# Reshape for plotting
-
-# create plot
-
-# Model summaries
-
-# Cross-validation
-
-# calculate variable importance for PPO model
-
-```
-
-## Interpretation
-
-# PPO Model with Interactions
+    1.

### 2024-12-07 peter_models.qmd
update model

@@
 
+dataset_clean$Q29_reordered  <- factor(dataset_clean$Q29,
+                            levels = c("(4) Not at all interested in politics",
+                                     "(3) Not that interested in politics",
+                                     "(2) Somewhat interested",
+                                     "(1) Very interested in politics"))
+
 dataset_clean$Q271_reordered <- factor(dataset_clean$Q271,
@@
 
-ord_mod <- polr(Q29 ~ Q271_reordered + race_recoded + age_recoded +
-            rural + # Rural or Urban/Suburban
-            sex_orientation + # Sexual Orientation
+ord_mod <- polr(Q29 ~ Q271_reordered + race_recoded +
+                  age_recoded +
             educ +  # Educational Attainment
             S3B + # Gender
-            S7 + # Nativity
             Q21, # Party affiliation
@@
 # generate predictions for the ordinal model
+
+new_dataset <- data.frame
+
+predict(ord_mod, new_dataset)
+```
+
+# Fitting OLR with interaction terms
+
+```{r}
+library(MASS) # package for model
+library(caret) # package for confusion matrix
+
+# dataset_clean$Q271_reordered <- factor(dataset_clean$Q271,
+#                             levels = c("(5) Not at all important",
+#                                      "(4) Slightly important",
+#                                      "(3) Moderately important",
+#                                      "(2) Very important",
+#                                      "(1) Extremely important"))
+
+ord_mod_interactions <- polr(Q29 ~ Q271_reordered + race_recoded + age_recoded +
+            rural + # Rural or Urban/Suburban
+            sex_orientation + # Sexual Orientation
+            educ + # Educational Attainment
+            S3B + # Gender
+            S7 + # Nativity
+            Q21, # Party affiliation
+          data = dataset_clean, Hess=TRUE)
+
+summary(ord_mod_interactions)
+
+pvals_interact <- pnorm(-abs(summary(ord_mod_interactions)$coef[,"t value"]))*2
+ctable_interact <- cbind(summary(ord_mod_interactions)$coef,pvals)
+
+ctable_interact
+
+(exp_coefs_interact <- exp(cbind(OR=coef(ord_mod_interactions),confint(ord_mod_interactions))))
+
+head(ord_mod_interactions$fitted.values)
+head(predict(ord_mod_interactions))
+
+confusionMatrix(predict(ord_mod_interactions), dataset_clean$Q29, mode="everything")
+
+```
+
+```{r}
+# generate predictions for the ordinal model with interactions
 

### 2024-12-07 peter_models.qmd
updating OLR and MLR model

@@
 
-ord_mod_interactions <- polr(Q29 ~ Q271_reordered + race_recoded + age_recoded +
-            rural + # Rural or Urban/Suburban
-            sex_orientation + # Sexual Orientation
+ord_mod_interactions <- polr(Q29 ~ Q271_reordered + race_recoded +
+                  age_recoded +
             educ +  # Educational Attainment
             S3B + # Gender
-            S7 + # Nativity
             Q21, # Party affiliation
@@
 dataset_clean$Q29_unordered <- factor(dataset_clean$Q29, ordered = FALSE)
+dataset_clean$Q271_simple <- as.numeric(dataset_clean$Q271)
 
@@
 m_mult <- multinom(Q29_unordered ~ Q271 + race_recoded + age_recoded +
-            rural + # Rural or Urban/Suburban
-            sex_orientation + # Sexual Orientation
+            # rural + # Rural or Urban/Suburban
+            # sex_orientation + # Sexual Orientation
             educ + # Educational Attainment
             S3B + # Gender
-            S7 + # Nativity
+            # S7 + # Nativity
             Q21, # Party affiliation
@@
 pred_data <- expand.grid(
-    Q271_simple = seq(1, 5, by = 0.1),
-    race_recoded = unique(dataset_clean$race_recoded)
+    Q271 = unique(dataset_clean$Q271),
+    age_recoded = unique(dataset_clean$age_recoded),
+    race_recoded = unique(dataset_clean$race_recoded),
+    educ = unique(dataset_clean$educ),
+    S3B = unique(dataset_clean$S3B),
+    Q21 = unique(dataset_clean$Q21)
 )
@@
 p2 <- ggplot(pred_long, 
-             aes(x = Q271_simple, y = probability, color = race_recoded)) +
+             aes(x = Q271, y = probability, color = race_recoded)) +
     geom_line() +

### 2024-12-07 peter_models.qmd
updating olr models

@@
 
-# dataset_clean$Q271_reordered <- factor(dataset_clean$Q271,
-#                             levels = c("(5) Not at all important",
-#                                      "(4) Slightly important",
-#                                      "(3) Moderately important",
-#                                      "(2) Very important",
-#                                      "(1) Extremely important"))
+dataset_clean$Q271_reordered <- factor(dataset_clean$Q271,
+                            levels = c("(5) Not at all important",
+                                     "(4) Slightly important",
+                                     "(3) Moderately important",
+                                     "(2) Very important",
+                                     "(1) Extremely important"))
 
@@
             Q21, # Party affiliation
+            # Q271_reordered * race_recoded,  # Interaction term
           data = dataset_clean, Hess=TRUE)

### 2024-12-08 Naki_Main.qmd
done

@@
 ## Extracting relevant columns
+
 ```{r}
@@
 ## Cleaning and factoring data for all columns
+
 ```{r}
@@
 ## Consolidating Religion to one column
+
 ```{r}
@@
 ## Rename the remaining column headers for ease of understanding
+
 ```{r}
@@
 )
-
+dataset_clean.head()
 ```
@@
 ### Check VIF values
+
 ```{r}
@@
 ## Perform stepwise selection to select best parameters
+
 ```{r}
@@
 
+```{r}
+model_stepwise
+```
+
 ### Check VIF Values
+
 ```{r}
@@
 ## Extract Odds for each variable in the stepwise model
+
 ```{r}
@@
 ## Plot showing odds
+
 ```{r}
@@
 ```{r}
-# Filter for significant variables (p-value < 0.05)
-significant_data <- model_stepwise |> 
-  filter(PValue < 0.05)
+newdata <- dataset_clean[, c("ImportanceOfStoppingDiscriminationAgainstMinorities", 
+                             "PrimaryRace", "Gender", "HighestLevelOfEducation", "TreatedUnfairlyOrDiscriminatedAgainst",
+                             "DiscriminationImpactOnLifeSatisfaction", "PerceptionsOfDiscriminationAgainstWhites",
+                             "PerceptionsOfDiscriminationAgainstBlacks", "PerceptionsOfDiscriminationAgainstAsianAmericans",
+                             "PerceptionsOfDiscriminationAgainstNativeAmericans", "PerceptionsOfDiscriminationAgainstLatinos")]
+```
 
-# Prepare data for plotting
-# You might need to adjust this section to ensure the data is in long format with appropriate variable names
-plot_data <- significant_data |>
-  select(Variable, OddsRatio) |>
-  mutate(Variable = factor(Variable, levels = unique(Variable)))
+```{r}
+actual_data <- dataset_clean$ImportanceOfStoppingDiscriminationAgainstMinorities
+predicted_probs <- predict(model_stepwise, newdata = newdata, type = "response")
+predicted_probs
+```
 
-# Create the boxplot
-ggplot(plot_data, aes(x = Variable, y = OddsRatio)) +
-  geom_boxplot() +
-  labs(title = "Boxplot of Significant Variables (Odds Ratio)",
-       x = "Variables",
-       y = "Odds Ratio") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
+```{r}
+library(pROC)
+roc_curve <- roc(actual_data, predicted_probs)  # predicted_probs = probabilities for the positive class
+plot(roc_curve)
+auc(roc_curve)
 ```

### 2024-12-08 RQ2_main.qmd
done

--- 2024-12-08 Naki_Main.qmd
@@
 ---
-title: "Naki_Main"
-format: html
+title: "RQ2_main"
+format: 
+      pdf:
+       code-overflow: wrap
 editor: visual
@@
 
-```{r}
-
-# Load the packages
-library(ggplot2)
-library(broom)
-library(dplyr)
-library(car)
-library(caret)
+## Quarto
+The `echo: false` option disables the printing of code (only output is displayed).
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(
+  message = FALSE,   # Suppress messages
+  warning = FALSE    # Suppress warnings
+)
 ```
 
-# Introduction
+```{r echo = TRUE}
+# Load necessary libraries
+library(dplyr)   # Provides %>% and mutate
+library(tidyverse)  # Optionally load the entire tidyverse if needed
+library(pROC)
+```
 
-The Collaborative Multi-Racial Post-Election Survey (CMPS), United States, 2020 is an extensive dataset designed to capture diverse perspectives across racial and ethnic groups in the United States.
-Respondents were asked to address a wide range of issues, including political attitudes, voting behavior, social concerns, and policy preferences.
-
-# Preliminary Cleaning
-
-## Loading the Dataset
-
-```{r}
+```{r echo = TRUE}
 file_path <- "/Users/nakiyahdhariwala/StatsProject/39096-0001-Data.rda"
@@
 ```
-
-The dataset contains 17,545 rows, with each row representing an individual respondent's data, and 1,490 columns, each capturing a specific variable or question response.
-These variables include demographic information, opinions on political and social issues, and responses to survey questions.
-
-This particular analysis focused on a single topic within the larger dataset.
-Specifically, our research question is: **How do personal experiences with racism, in combination with demographic factors such as race, gender, and education, influence the importance respondents place on policies aimed at addressing racial and ethnic discrimination, controlling for other social and political factors?**
-
-```{r}
-# Calculate proportions
-proportions <- prop.table(table(dataset$Q1R5))
-proportions
-# Define custom labels
-custom_labels <- c("Not important", "Important")
-
-# Combine custom labels with percentages
-pie_labels <- paste(custom_labels, round(proportions * 100, 2), "%")
-
-# Plot the pie chart with custom labels
-pie(proportions,
-    labels = pie_labels,
-    main = "Q1R5: Stopping discrimination against racial/ethnic minorities",
-    col = c("lightblue", "salmon"))
-```
-
-The outcome variable Q1R5 (Stopping discrimination against racial/ethnic minorities) for this Research Question is a binary variable which captures whether respondents view stopping racial and ethnic discrimination as an important issue (i.e., Important OR Not Important).
-Therefore, we will make use Logistic regression as it is suitable for modeling binary outcome variables and allows for the inclusion of both predictor and control variables.
-While the distribution is skewed (74.25% say "No"), logistic regression is robust to imbalance in the outcome variable.
-However, extreme imbalance (e.g., 95%-5%) might require additional techniques like resampling or penalized logistic regression.
-
-We can expect the analysis to reveal whether and to what extent experiences with racism predict the importance of racial/ethnic discrimination policy, controlling for demographic factors.
-
-## Extracting relevant columns
 
@@
 
-## Cleaning and factoring data for all columns
-
 ```{r}
@@
     "Not important",
-    ifelse(dataset_new$Q1R5 == "(1) Stopping discrimination against racial/ethnic minorities", 
-           "Important", 
-           "Not specified")
+    "Important"  # Remaining cases are "Important"
   ),
-  levels = c("Not important", "Important", "Not specified") # Define factor levels
+  levels = c("Not important", "Important") # Define factor levels
 )
@@
 # Assuming `required_cols` is previously defined; otherwise, specify explicitly
-# e.g., required_cols <- c("Q1R5", "Q619_Q626R1", ...)
 dataset_new[required_cols] <- lapply(dataset_new[required_cols], as.factor)
@@
 head(dataset_new)
+
 ```
-
-## Consolidating Religion to one column
-
-```{r}
-# Convert individual religion columns to numeric
-for (i in 1:15) {
-  col_name <- paste0("Q58R", i, "_numeric")
-  dataset_new[[col_name]] <- as.numeric(sub("^\\((\\d+)\\).*", "\\1", dataset_new[[paste0("Q58R", i)]]))
-}
-
-# Create a new Religion variable
-dataset_new$Religion <- "None"
-
-# Assign religion based on the numeric values of each Q58R column
-religion_mapping <- c(
-  "Q58R1_numeric" = "Catholic", 
-  "Q58R2_numeric" = "Protestant", 
-  "Q58R3_numeric" = "Other Christian",
-  "Q58R4_numeric" = "Jewish", 
-  "Q58R5_numeric" = "Muslim", 
-  "Q58R6_numeric" = "Hindu", 
-  "Q58R7_numeric" = "Buddhist", 
-  "Q58R8_numeric" = "Mormon/LDS", 
-  "Q58R9_numeric" = "Folk Religion", 
-  "Q58R10_numeric" = "Daoist", 
-  "Q58R11_numeric" = "Shamanism", 
-  "Q58R12_numeric" = "Ancestral Religion", 
-  "Q58R13_numeric" = "Spiritual but not religious", 
-  "Q58R14_numeric" = "Atheist", 
-  "Q58R15_numeric" = "Agnostic"
-)
-
-# Loop through the mapping to assign values
-for (col in names(religion_mapping)) {
-  dataset_new$Religion[dataset_new[[col]] == 1] <- religion_mapping[[col]]
-}
-
-# Handle multiple religious affiliations
-dataset_new$Religion[rowSums(dataset_new[, grep("_numeric", names(dataset_new))] == 1) > 1] <- "Multiple Religions"
-
-# Convert Religion to a factor with specific levels
-dataset_new$Religion <- factor(dataset_new$Religion,
-                               levels = c("Catholic", "Protestant", "Other Christian", "Jewish", "Muslim",
-                                          "Hindu", "Buddhist", "Mormon/LDS", "Folk Religion",
-                                          "Daoist", "Shamanism", "Ancestral Religion",
-                                          "Spiritual but not religious", "Atheist", "Agnostic", 
-                                          "Multiple Religions", "None"))
-
-# Create a table of counts for the new Religion variable
-table(dataset_new$Religion)
-
-# Remove redundant religion columns (Q58R1 to Q58R15 and their numeric versions)
-redundant_cols <- c(
-  paste0("Q58R", 1:15), paste0("Q58R", 1:15, "_numeric")
-)
-
-# Remove redundant columns from the dataset
-dataset_clean <- dataset_new[, !(names(dataset_new) %in% redundant_cols)]
-
-# Take a look at the updated dataset
-head(dataset_clean)
-```
-
-We first consolidated multiple columns related to respondents' religious affiliations into a single, categorical variable called Religion.
-We also add another option called 'Multiple Religions' where respondents identified with more than one religion.
-
-# Checking for missing data
-
-```{r}
-# Check the number of missing values for each variable
-missing_summary <- sapply(dataset_clean, function(x) sum(is.na(x)))
-
-# Create a data frame to display the missing values clearly
-missing_df <- data.frame(Variable = names(missing_summary), Missing_Count = missing_summary)
-print(missing_df)
-
-# Extract the variables Q629R1 to Q629R7
-racism_vars <- dataset_clean[, grep("^Q629R[1-7]$", names(dataset_clean))]
-
-# Calculate the proportion of missing values for each variable
-missing_proportions <- colSums(is.na(racism_vars)) / nrow(racism_vars)
-
-# Convert to a data frame
-missing_summary <- data.frame(Variable = names(missing_proportions),
-                              Missing_Proportion = round(missing_proportions, 4))
-
-print(missing_summary)
-```
-
-Variables Q629R1 - Q629R7 each have 9,180 missing values, accounting for over 52% of their data. Due to this high proportion of missingness, we decided to remove these variables from the analysis.
-
-This decision is justified as these variables ("Q629R1" to "Q629R7") capture detailed experiences with unfair treatment but overlap significantly with the broader variable Q627 ("Treated unfairly or discriminated against"), which remains in the analysis.
-By retaining Q627, we ensure that the overall essence of experiences with racism is preserved, albeit in a more general and complete form.
-
-Removing these variables enhances data quality by addressing issues of excessive missingness, reduces the risk of bias, and strengthens the reliability and interpretability of the analysis.
-
-```{r}
-# Remove the variables Q629R1 to Q629R7
-dataset_clean <- dataset_clean[, !(names(dataset_clean) %in% c("Q629R1", "Q629R2", 
-                                                               "Q629R3", "Q629R4", 
-                                                               "Q629R5", "Q629R6", "Q629R7"))]
-
-head(dataset_clean)
-```
-
-## Rename the remaining column headers for ease of understanding
-
-```{r}
-colnames(dataset_clean)
-colnames(dataset_clean) <- c(
-  "ImportanceOfStoppingDiscriminationAgainstMinorities",
-  "PrimaryRace", 
-  "Orientation", 
-  "Gender", 
-  "Age", 
-  "HighestLevelOfEducation", 
-  "CommunityLivedIn", 
-  "TreatedUnfairlyOrDiscriminatedAgainst", 
-  "DiscriminationImpactOnLifeSatisfaction",
-  "PerceptionsOfDiscriminationAgainstWhites",
-  "PerceptionsOfDiscriminationAgainstBlacks", 
-  "PerceptionsOfDiscriminationAgainstAsianAmericans", 
-  "PerceptionsOfDiscriminationAgainstNativeAmericans", 
-  "PerceptionsOfDiscriminationAgainstImmigrants", 
-  "PerceptionsOfDiscriminationAgainstLatinos", 
-  "PerceptionsOfDiscriminationAgainstGaysAndLesbians",
-  "PerceptionsOfDiscriminationAgainstMuslims",
-  "Religion"
-)
-dataset_clean.head()
-```
-
-# Look for exploratory plots in the other documents
-
-# Binomial Logistic Regression
-
-## Model
-
-```{r}
-library(vcd)
-library(tidyverse)
-library(corrplot)
-```
-
-```{r}
-model <- glm(ImportanceOfStoppingDiscriminationAgainstMinorities ~ PrimaryRace +
-               Orientation + Gender + Age + 
-               HighestLevelOfEducation + CommunityLivedIn +
-               TreatedUnfairlyOrDiscriminatedAgainst + 
-               DiscriminationImpactOnLifeSatisfaction + 
-               PerceptionsOfDiscriminationAgainstWhites + 
-               PerceptionsOfDiscriminationAgainstBlacks +
-               PerceptionsOfDiscriminationAgainstAsianAmericans + 
-               PerceptionsOfDiscriminationAgainstNativeAmericans + 
-               PerceptionsOfDiscriminationAgainstImmigrants + 
-               PerceptionsOfDiscriminationAgainstLatinos + 
-               PerceptionsOfDiscriminationAgainstGaysAndLesbians + 
-               PerceptionsOfDiscriminationAgainstMuslims +
-               Religion,
-
-        data = dataset_clean, family = "binomial")
-```
-
-### Check VIF values
-
-```{r}
-# Check multicollinearity using VIF
-vif(model)
-```
-
-```{r}
-## # Extract odds ratios
-## odds_ratios <- exp(coef(model))
-## 
-## # Get confidence intervals for coefficients
-## conf_intervals <- exp(confint(model))
-## 
-## # Extract p-values
-## p_values <- summary(model)$coefficients[, 4]
-## 
-## # Combine everything into a single data frame
-## results_model <- data.frame(
-##   OddsRatio = odds_ratios,
-##   LowerCI = conf_intervals[, 1],
-##   UpperCI = conf_intervals[, 2],
-##   PValue = p_values
-## )
-## 
-## # Display results
-## results_model
-```
-
-## Perform stepwise selection to select best parameters
-
-```{r}
-# Perform stepwise selection
-model_stepwise <- step(model, 
-                       direction = "both",
-                       trace = 0)
-```
-
-```{r}
-model_stepwise
-```
-
-### Check VIF Values
-
-```{r}
-# Check multicollinearity using VIF
-vif(model_stepwise)
-```
-
-## Extract Odds for each variable in the stepwise model
-
-```{r}
-# Extract odds ratios
-odds_ratios <- exp(coef(model_stepwise))
-
-# Get confidence intervals for coefficients
-conf_intervals <- exp(confint(model_stepwise))
-
-# Extract p-values
-p_values <- summary(model_stepwise)$coefficients[, 4]
-
-# Combine everything into a single data frame
-results_model_stepwise <- data.frame(
-  OddsRatio = odds_ratios,
-  LowerCI = conf_intervals[, 1],
-  UpperCI = conf_intervals[, 2],
-  PValue = p_values
-)
-
-# Display results
-results_model_stepwise
-```
-
-## Plot showing odds
-
-```{r}
-# Filter out only significant variables (p-value < 0.05)
-significant_data <- results_model_stepwise[results_model_stepwise$PValue < 0.051, ]
-
-# Create the plot with inverted axes and CI shown as error bars
-ggplot(significant_data, aes(y = rownames(significant_data), x = OddsRatio)) +
-  geom_point(size = 4, color = "blue") +  # Plot the odds ratio as points
-  geom_errorbarh(aes(xmin = LowerCI, xmax = UpperCI), height = 0.2, color = "black") +  # Add horizontal error bars
-  labs(title = "Significant Variables with Confidence Intervals (Odds Ratio)",
-       x = "Odds Ratio",
-       y = "Variables") +
-  theme(axis.text.y = element_text(angle = 0, hjust = 1))  # Keep y-axis labels horizontal
-  theme(plot.margin = margin(1, 2, 1, 1, "cm"))  # Adjust margins for larger size
-
-# Save the plot with larger dimensions
-ggsave("significant_variables_plot.png", width = 12, height = 8)
-```
-
-```{r}
-newdata <- dataset_clean[, c("ImportanceOfStoppingDiscriminationAgainstMinorities", 
-                             "PrimaryRace", "Gender", "HighestLevelOfEducation", "TreatedUnfairlyOrDiscriminatedAgainst",
-                             "DiscriminationImpactOnLifeSatisfaction", "PerceptionsOfDiscriminationAgainstWhites",
-                             "PerceptionsOfDiscriminationAgainstBlacks", "PerceptionsOfDiscriminationAgainstAsianAmericans",
-                             "PerceptionsOfDiscriminationAgainstNativeAmericans", "PerceptionsOfDiscriminationAgainstLatinos")]
-```
-
-```{r}
-actual_data <- dataset_clean$ImportanceOfStoppingDiscriminationAgainstMinorities
-predicted_probs <- predict(model_stepwise, newdata = newdata, type = "response")
-predicted_probs
-```
-
-```{r}
-library(pROC)
-roc_curve <- roc(actual_data, predicted_probs)  # predicted_probs = probabilities for the positive class
-plot(roc_curve)
-auc(roc_curve)
-```

### 2024-12-08 peter_models.qmd
updating interactions model
updating modeling script

@@
 ```{r}
-# Create a summary table for all categorical variables
-variables <- c("Q29", "Q271", "S2_RACE_PRIME", "S3", "S3B", "S5_AGE", "S13", "S14", "S7", "Q21", "S2_M", "Q627", "Q628", "Q629R1", "Religion", "educ", "sex_orientation", "rural", "white", "race_recoded", "age_recoded")
-
-# Chi-square tests for each variable
-chi_square_results <- lapply(variables, function(var) {
-  chisq.test(table(dataset_clean[[var]], dataset_clean$Q29))
-})
-
-# Print chi-square results
-names(chi_square_results) <- variables
-print(chi_square_results)
+# # Create a summary table for all categorical variables
+# variables <- c("Q29", "Q271", "S2_RACE_PRIME", "S3", "S3B", "S5_AGE", "S13", "S14", "S7", "Q21", "S2_M", "Q627", "Q628", "Q629R1", "Religion", "educ", "sex_orientation", "rural", "white", "race_recoded", "age_recoded")
+# 
+# 
+# # Chi-square tests for each variable
+# chi_square_results <- lapply(variables, function(var) {
+#   chisq.test(table(dataset_clean[[var]], dataset_clean$Q29))
+# })
+# 
+# # Print chi-square results
+# names(chi_square_results) <- variables
+# print(chi_square_results)
 ```
@@
 # generate predictions for the ordinal model
-
+# 
 new_dataset <- data.frame
 
-predict(ord_mod, new_dataset)
+# predict(ord_mod, new_dataset)
 ```
@@
 
-ord_mod_interactions <- polr(Q29 ~ Q271_reordered + race_recoded +
+ord_mod_interactions <- polr(Q29 ~ Q271_reordered * S3B +
+                               race_recoded +
                   age_recoded +
             educ +  # Educational Attainment
-            S3B + # Gender
             Q21, # Party affiliation
-            # Q271_reordered * race_recoded,  # Interaction term
           data = dataset_clean, Hess=TRUE)
@@
 
-new_dataset <- data.frame
-
-predict(ord_mod, new_dataset)
+# new_dataset <- data.frame
+# 
+# predict(ord_mod, new_dataset)
 ```
@@
 ```{r}
-
-# Load packages
-library(nnet)  # for multinomial logistic regression
-library(tidyverse)
-library(car)
-library(effects)
-
-# ensure outcome variable is a factor but not ordered
-dataset_clean$Q29_unordered <- factor(dataset_clean$Q29, ordered = FALSE)
-dataset_clean$Q271_simple <- as.numeric(dataset_clean$Q271)
-
-# Fit multinomial model
-m_mult <- multinom(Q29_unordered ~ Q271 + race_recoded + age_recoded +
-            # rural + # Rural or Urban/Suburban
-            # sex_orientation + # Sexual Orientation
-            educ + # Educational Attainment
-            S3B + # Gender
-            # S7 + # Nativity
-            Q21, # Party affiliation
-          data = dataset_clean)
-
-# Calculate predicted probabilities of marginal effects
-# Create sequence of Q271_simple values
-x_vals <- seq(min(dataset_clean$Q271_simple), max(dataset_clean$Q271_simple), length.out = 100)
-
-# Function to calculate average predicted probabilities
-get_avg_probs <- function(model, newdata, var_name) {
-    preds <- predict(model, newdata = newdata, type = "probs")
-    colMeans(preds)
-}
-
-# Calculate marginal effects by race
-races <- unique(dataset_clean$race_recoded)
-marg_effects <- list()
-
-for(race in races) {
-    probs <- matrix(0, nrow = length(x_vals), ncol = length(levels(dataset_clean$Q29_unordered)))
-    for(i in seq_along(x_vals)) {
-        temp_data <- dataset_clean
-        temp_data$Q271_simple <- x_vals[i]
-        temp_data$race_major <- race
-        probs[i,] <- get_avg_probs(m_mult, temp_data, "Q271_simple")
-    }
-    marg_effects[[race]] <- data.frame(
-        x = x_vals,
-        race = race,
-        prob1 = probs[,1],
-        prob2 = probs[,2],
-        prob3 = probs[,3],
-        prob4 = probs[,4]
-    )
-}
-
-# Combine all marginal effects
-marg_effects_df <- bind_rows(marg_effects)
-
-# Create plots for marginal effects
-# Reshape data for plotting
-marg_effects_long <- marg_effects_df %>%
-    pivot_longer(cols = starts_with("prob"),
-                names_to = "outcome",
-                values_to = "probability") %>%
-    mutate(outcome = factor(outcome,
-                          levels = paste0("prob", 1:4),
-                          labels = levels(dataset_clean$Q29_unordered)))
-
-# Plot marginal effects
-p1 <- ggplot(marg_effects_long, 
-             aes(x = x, y = probability, color = outcome)) +
-    geom_line() +
-    facet_wrap(~race) +
-    theme_minimal() +
-    labs(title = "Marginal Effects of Racial Identity Importance by Race",
-         x = "Importance of Racial Identity",
-         y = "Predicted Probability",
-         color = "Political Interest Level") +
-    theme(legend.position = "bottom")
-print(p1)
-
-# Model fit statistics and comparisons
-print("Model summaries:")
-print(summary(m_mult))
-
-# Calculate McFadden's R-squared
-null_model <- multinom(Q29_unordered ~ 1, data = dataset_clean)
-mcfadden_r2 <- 1 - logLik(m_mult)/logLik(null_model)
-
-print("McFadden's R-squared:")
-print(mcfadden_r2)
-
-# Classification accuracy
-pred_class <- predict(m_mult)
-conf_matrix <- table(pred_class, dataset_clean$Q29_unordered)
-accuracy <- sum(diag(conf_matrix))/sum(conf_matrix)
-
-print("Classification accuracy:")
-print(accuracy)
-print("Confusion Matrix:")
-print(conf_matrix)
-
-# Print AIC
-print(AIC(m_mult))
-
-# Calculate predicted probabilities
-pred_data <- expand.grid(
-    Q271 = unique(dataset_clean$Q271),
-    age_recoded = unique(dataset_clean$age_recoded),
-    race_recoded = unique(dataset_clean$race_recoded),
-    educ = unique(dataset_clean$educ),
-    S3B = unique(dataset_clean$S3B),
-    Q21 = unique(dataset_clean$Q21)
-)
-
-# Get predictions
-preds <- predict(m_mult, newdata = pred_data, type = "probs")
-pred_data <- cbind(pred_data, as.data.frame(preds))
-
-# Create separate plots for each outcome level
-pred_long <- pred_data %>%
-    pivot_longer(cols = starts_with("("),
-                names_to = "interest_level",
-                values_to = "probability")
-
-# Plot for each outcome level
-p2 <- ggplot(pred_long, 
-             aes(x = Q271, y = probability, color = race_recoded)) +
-    geom_line() +
-    facet_wrap(~interest_level) +
-    theme_minimal() +
-    labs(title = "Predicted Probabilities by Outcome Level",
-         x = "Importance of Racial Identity",
-         y = "Probability",
-         color = "Race/Ethnicity") +
-    theme(legend.position = "bottom")
-print(p2)
+# 
+# # Load packages
+# library(nnet)  # for multinomial logistic regression
+# library(tidyverse)
+# library(car)
+# library(effects)
+# 
+# 
+# # ensure outcome variable is a factor but not ordered
+# dataset_clean$Q29_unordered <- factor(dataset_clean$Q29, ordered = FALSE)
+# dataset_clean$Q271_simple <- as.numeric(dataset_clean$Q271)
+# 
+# # Fit multinomial model
+# m_mult <- multinom(Q29_unordered ~ Q271 + race_recoded + age_recoded +
+#             # rural + # Rural or Urban/Suburban
+#             # sex_orientation + # Sexual Orientation
+#             educ + # Educational Attainment
+#             S3B + # Gender
+#             # S7 + # Nativity
+#             Q21, # Party affiliation
+#           data = dataset_clean)
+# 
+# 
+# 
+# # Calculate predicted probabilities of marginal effects
+# # Create sequence of Q271_simple values
+# x_vals <- seq(min(dataset_clean$Q271_simple), max(dataset_clean$Q271_simple), length.out = 100)
+# 
+# # Function to calculate average predicted probabilities
+# get_avg_probs <- function(model, newdata, var_name) {
+#     preds <- predict(model, newdata = newdata, type = "probs")
+#     colMeans(preds)
+# }
+# 
+# # Calculate marginal effects by race
+# races <- unique(dataset_clean$race_recoded)
+# marg_effects <- list()
+# 
+# for(race in races) {
+#     probs <- matrix(0, nrow = length(x_vals), ncol = length(levels(dataset_clean$Q29_unordered)))
+#     for(i in seq_along(x_vals)) {
+#         temp_data <- dataset_clean
+#         temp_data$Q271_simple <- x_vals[i]
+#         temp_data$race_major <- race
+#         probs[i,] <- get_avg_probs(m_mult, temp_data, "Q271_simple")
+#     }
+#     marg_effects[[race]] <- data.frame(
+#         x = x_vals,
+#         race = race,
+#         prob1 = probs[,1],
+#         prob2 = probs[,2],
+#         prob3 = probs[,3],
+#         prob4 = probs[,4]
+#     )
+# }
+# 
+# # Combine all marginal effects
+# marg_effects_df <- bind_rows(marg_effects)
+# 
+# # Create plots for marginal effects
+# # Reshape data for plotting
+# marg_effects_long <- marg_effects_df %>%
+#     pivot_longer(cols = starts_with("prob"),
+#                 names_to = "outcome",
+#                 values_to = "probability") %>%
+#     mutate(outcome = factor(outcome,
+#                           levels = paste0("prob", 1:4),
+#                           labels = levels(dataset_clean$Q29_unordered)))
+# 
+# # Plot marginal effects
+# p1 <- ggplot(marg_effects_long, 
+#              aes(x = x, y = probability, color = outcome)) +
+#     geom_line() +
+#     facet_wrap(~race) +
+#     theme_minimal() +
+#     labs(title = "Marginal Effects of Racial Identity Importance by Race",
+#          x = "Importance of Racial Identity",
+#          y = "Predicted Probability",
+#          color = "Political Interest Level") +
+#     theme(legend.position = "bottom")
+# print(p1)
+# 
+# # Model fit statistics and comparisons
+# print("Model summaries:")
+# print(summary(m_mult))
+# 
+# # Calculate McFadden's R-squared
+# null_model <- multinom(Q29_unordered ~ 1, data = dataset_clean)
+# mcfadden_r2 <- 1 - logLik(m_mult)/logLik(null_model)
+# 
+# print("McFadden's R-squared:")
+# print(mcfadden_r2)
+# 
+# # Classification accuracy
+# pred_class <- predict(m_mult)
+# conf_matrix <- table(pred_class, dataset_clean$Q29_unordered)
+# accuracy <- sum(diag(conf_matrix))/sum(conf_matrix)
+# 
+# print("Classification accuracy:")
+# print(accuracy)
+# print("Confusion Matrix:")
+# print(conf_matrix)
+# 
+# # Print AIC
+# print(AIC(m_mult))
+# 
+# # Calculate predicted probabilities
+# pred_data <- expand.grid(
+#     Q271 = unique(dataset_clean$Q271),
+#     age_recoded = unique(dataset_clean$age_recoded),
+#     race_recoded = unique(dataset_clean$race_recoded),
+#     educ = unique(dataset_clean$educ),
+#     S3B = unique(dataset_clean$S3B),
+#     Q21 = unique(dataset_clean$Q21)
+# )
+# 
+# # Get predictions
+# preds <- predict(m_mult, newdata = pred_data, type = "probs")
+# pred_data <- cbind(pred_data, as.data.frame(preds))
+# 
+# # Create separate plots for each outcome level
+# pred_long <- pred_data %>%
+#     pivot_longer(cols = starts_with("("),
+#                 names_to = "interest_level",
+#                 values_to = "probability")
+# 
+# 
+# # Plot for each outcome level
+# p2 <- ggplot(pred_long, 
+#              aes(x = Q271, y = probability, color = race_recoded)) +
+#     geom_line() +
+#     facet_wrap(~interest_level) +
+#     theme_minimal() +
+#     labs(title = "Predicted Probabilities by Outcome Level",
+#          x = "Importance of Racial Identity",
+#          y = "Probability",
+#          color = "Race/Ethnicity") +
+#     theme(legend.position = "bottom")
+# print(p2)
 
@@
 ```{r}
-dataset_clean$Q271_simple <- as.numeric(dataset_clean$Q271)
-# dataset_clean$Q271 <- factor(dataset_clean$Q271,
-#                             levels = c("(5) Not at all important",
-#                                      "(4) Slightly important",
-#                                      "(3) Moderately important",
-#                                      "(2) Very important",
-#                                      "(1) Extremely important"),
-#                             ordered = TRUE)
-
-m1 <- clm(Q29 ~ Q271 + race_recoded + age_recoded +
-            rural + # Rural or Urban/Suburban
-            sex_orientation + # Sexual Orientation
-            Religion + 
-            educ + # Educational Attainment
-            S3B + # Gender
-            S7 + # Nativity
-            Q21, # Party affiliation
-          data = dataset_clean)
-
-summary(m1)
-
-# Check for multicollinearity using a linear approximation
-vif_m1 <- vif(lm(as.numeric(Q29) ~ Q271 + race_recoded + age_recoded +
-            rural + # Rural or Urban/Suburban
-            sex_orientation + # Sexual Orientation
-            Religion + 
-            educ + # Educational Attainment
-            S3B + # Gender
-            S7 + # Nativity
-            Q21, # Party affiliation
-          data = dataset_clean))
-
-print("Variance Inflation Factors:")
-print(vif_m1)
-
-# Test proportional odds assumption
-print("Testing proportional odds assumption:")
-test1 <- nominal_test(m1)
-print("Nominal test for Model 1:")
-print(test1)
-
-# Alternative diagnostic approach using linear approximation
-lm_approx <- lm(as.numeric(Q29) ~ Q271 + race_recoded + age_recoded +
-            rural + # Rural or Urban/Suburban
-            sex_orientation + # Sexual Orientation
-            Religion + 
-            educ + # Educational Attainment
-            S3B + # Gender
-            S7 + # Nativity
-            Q21, # Party affiliation
-          data = dataset_clean)
-diagnostics_df <- data.frame(
-  fitted = fitted(lm_approx),
-  residuals = residuals(lm_approx),
-  index = 1:nrow(dataset_clean)
-)
-
-# Plot residuals vs fitted values using linear approximation
-p1 <- ggplot(diagnostics_df, aes(x = fitted, y = residuals)) +
-  geom_point(alpha = 0.1) +
-  geom_smooth(method = "loess") +
-  theme_minimal() +
-  labs(title = "Residuals vs Fitted Values (Linear Approximation)",
-       x = "Fitted Values",
-       y = "Residuals")
-print(p1)
-
-# Calculate and plot Cook's distance
-cooks_d <- cooks.distance(lm_approx)
-influence_df <- data.frame(
-  index = 1:length(cooks_d),
-  cooks_distance = cooks_d
-)
-
-p2 <- ggplot(influence_df, aes(x = index, y = cooks_distance)) +
-  geom_point() +
-  geom_hline(yintercept = 4/nrow(dataset_clean), color = "red", linetype = "dashed") +
-  theme_minimal() +
-  labs(title = "Cook's Distance Plot",
-       x = "Observation Number",
-       y = "Cook's Distance")
-print(p2)
-
-# Count influential observations
-influential_obs <- which(cooks_d > 4/nrow(dataset_clean))
-print("Number of influential observations:")
-print(length(influential_obs))
-
-# Create prediction visualization
-pred_data <- expand.grid(
-  Q271 = levels(dataset_clean$Q271),
-  race_recoded = levels(dataset_clean$race_recoded), 
-  age_recoded = levels(dataset_clean$age_recoded),
-  rural = levels(dataset_clean$rural),
-  sex_orientation = levels(dataset_clean$sex_orientation),
-  Religion = levels(dataset_clean$Religion),
-  educ = levels(dataset_clean$educ),
-  S3B = levels(dataset_clean$S3B),
-  S7 = levels(dataset_clean$S7),
-  Q21 = levels(dataset_clean$Q21)
-)
-
-# Get predicted probabilities
-preds <- predict(m1, newdata = pred_data, type = "prob")
-# Convert predictions to numeric scores (weighted average)
-pred_data$predicted_score <- apply(preds$fit, 1, function(x) sum(x * 1:4))
-
-# Create prediction plot
-p3 <- ggplot(pred_data, aes(x = Q271, y = predicted_score, color = race_recoded)) +
-  geom_point() +
-  geom_line(aes(group = race_recoded)) +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  labs(title = "Predicted Political Interest by Racial Identity Importance and Race",
-       x = "Importance of Racial Identity",
-       y = "Predicted Political Interest Level (1-4)",
-       color = "Race/Ethnicity")
-print(p3)
-
-# Print the coefficients table
-print("Coefficients for Model 1:")
-coef_table <- summary(m1)$coefficients
-print(coef_table)
-
-# Calculate and print odds ratios for the coefficients
-odds_ratios <- exp(coef_table[, "Estimate"])
-print("Odds Ratios:")
-print(odds_ratios)
+# dataset_clean$Q271_simple <- as.numeric(dataset_clean$Q271)
+# # dataset_clean$Q271 <- factor(dataset_clean$Q271,
+# #                             levels = c("(5) Not at all important",
+# #                                      "(4) Slightly important",
+# #                                      "(3) Moderately important",
+# #                                      "(2) Very important",
+# #                                      "(1) Extremely important"),
+# #                             ordered = TRUE)
+# 
+# m1 <- clm(Q29 ~ Q271 + race_recoded + age_recoded +
+#             rural + # Rural or Urban/Suburban
+#             sex_orientation + # Sexual Orientation
+#             Religion + 
+#             educ + # Educational Attainment
+#             S3B + # Gender
+#             S7 + # Nativity
+#             Q21, # Party affiliation
+#           data = dataset_clean)
+# 
+# summary(m1)
+# 
+# # Check for multicollinearity using a linear approximation
+# vif_m1 <- vif(lm(as.numeric(Q29) ~ Q271 + race_recoded + age_recoded +
+#             rural + # Rural or Urban/Suburban
+#             sex_orientation + # Sexual Orientation
+#             Religion + 
+#             educ + # Educational Attainment
+#             S3B + # Gender
+#             S7 + # Nativity
+#             Q21, # Party affiliation
+#           data = dataset_clean))
+# 
+# print("Variance Inflation Factors:")
+# print(vif_m1)
+# 
+# # Test proportional odds assumption
+# print("Testing proportional odds assumption:")
+# test1 <- nominal_test(m1)
+# print("Nominal test for Model 1:")
+# print(test1)
+# 
+# # Alternative diagnostic approach using linear approximation
+# lm_approx <- lm(as.numeric(Q29) ~ Q271 + race_recoded + age_recoded +
+#             rural + # Rural or Urban/Suburban
+#             sex_orientation + # Sexual Orientation
+#             Religion + 
+#             educ + # Educational Attainment
+#             S3B + # Gender
+#             S7 + # Nativity
+#             Q21, # Party affiliation
+#           data = dataset_clean)
+# diagnostics_df <- data.frame(
+#   fitted = fitted(lm_approx),
+#   residuals = residuals(lm_approx),
+#   index = 1:nrow(dataset_clean)
+# )
+# 
+# # Plot residuals vs fitted values using linear approximation
+# p1 <- ggplot(diagnostics_df, aes(x = fitted, y = residuals)) +
+#   geom_point(alpha = 0.1) +
+#   geom_smooth(method = "loess") +
+#   theme_minimal() +
+#   labs(title = "Residuals vs Fitted Values (Linear Approximation)",
+#        x = "Fitted Values",
+#        y = "Residuals")
+# print(p1)
+# 
+# # Calculate and plot Cook's distance
+# cooks_d <- cooks.distance(lm_approx)
+# influence_df <- data.frame(
+#   index = 1:length(cooks_d),
+#   cooks_distance = cooks_d
+# )
+# 
+# p2 <- ggplot(influence_df, aes(x = index, y = cooks_distance)) +
+#   geom_point() +
+#   geom_hline(yintercept = 4/nrow(dataset_clean), color = "red", linetype = "dashed") +
+#   theme_minimal() +
+#   labs(title = "Cook's Distance Plot",
+#        x = "Observation Number",
+#        y = "Cook's Distance")
+# print(p2)
+# 
+# # Count influential observations
+# influential_obs <- which(cooks_d > 4/nrow(dataset_clean))
+# print("Number of influential observations:")
+# print(length(influential_obs))
+# 
+# 
+# 
+# # Create prediction visualization
+# pred_data <- expand.grid(
+#   Q271 = levels(dataset_clean$Q271),
+#   race_recoded = levels(dataset_clean$race_recoded), 
+#   age_recoded = levels(dataset_clean$age_recoded),
+#   rural = levels(dataset_clean$rural),
+#   sex_orientation = levels(dataset_clean$sex_orientation),
+#   Religion = levels(dataset_clean$Religion),
+#   educ = levels(dataset_clean$educ),
+#   S3B = levels(dataset_clean$S3B),
+#   S7 = levels(dataset_clean$S7),
+#   Q21 = levels(dataset_clean$Q21)
+# )
+# 
+# # Get predicted probabilities
+# preds <- predict(m1, newdata = pred_data, type = "prob")
+# # Convert predictions to numeric scores (weighted average)
+# pred_data$predicted_score <- apply(preds$fit, 1, function(x) sum(x * 1:4))
+# 
+# # Create prediction plot
+# p3 <- ggplot(pred_data, aes(x = Q271, y = predicted_score, color = race_recoded)) +
+#   geom_point() +
+#   geom_line(aes(group = race_recoded)) +
+#   theme_minimal() +
+#   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+#   labs(title = "Predicted Political Interest by Racial Identity Importance and Race",
+#        x = "Importance of Racial Identity",
+#        y = "Predicted Political Interest Level (1-4)",
+#        color = "Race/Ethnicity")
+# print(p3)
+# 
+# # Print the coefficients table
+# print("Coefficients for Model 1:")
+# coef_table <- summary(m1)$coefficients
+# print(coef_table)
+# 
+# # Calculate and print odds ratios for the coefficients
+# odds_ratios <- exp(coef_table[, "Estimate"])
+# print("Odds Ratios:")
+# print(odds_ratios)
 

### 2024-12-08 peter_models.qmd
pushing MLR model update

--- 2024-12-07 peter_models.qmd
@@
 ```{r}
-# generate predictions for the ordinal model
+# generate new dataset to create predictions for the models
+# 
+new_dataset <- data.frame(race_recoded=c("White", "Black", "Asian American", "Latino", "AIAN", "NHPI", "Arab/MENA"),
+                          Q271_reordered=c("(5) Not at all important", "(4) Slightly important", "(3) Moderately important", "(2) Very important", "(1) Extremely important", "(4) Slightly important", "(5) Not at all important"),
+                          age_recoded = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+", "18-29"),
+                          S3B = c("(1) Man", "(2) Woman","(3) Non-binary","(4) Something else (Specify)", "(1) Man", "(2) Woman", "(1) Man"),
+                          educ = c("No college experience", "Some college experience", "College graduate", "No college experience", "Some college experience", "College graduate", "College graduate"),
+                          Q21 = c("(1) Republican",  "(2) Democrat","(3) Independent", "(4) Other party", "(1) Republican",  "(2) Democrat",  "(3) Independent"))
 
-new_dataset <- data.frame
-
-predict(ord_mod, new_dataset)
 ```
@@
 
-```{r}
-# generate predictions for the ordinal model with interactions
-
-new_dataset <- data.frame
-
-predict(ord_mod, new_dataset)
-```
-
 # Fitting MLR Model
@@
 # Fit multinomial model
-m_mult <- multinom(Q29_unordered ~ Q271 + race_recoded + age_recoded +
-            # rural + # Rural or Urban/Suburban
-            # sex_orientation + # Sexual Orientation
+m_mult <- multinom(Q29_unordered ~ Q271 * S3B + 
+                     race_recoded + 
+                     age_recoded +
             educ + # Educational Attainment
-            S3B + # Gender
-            # S7 + # Nativity
             Q21, # Party affiliation
@@
 marg_effects <- list()
-
+# 
 for(race in races) {
@@
 print(conf_matrix)
-
+# 
 # Print AIC
@@
 )
-
+# 
 # Get predictions
@@
 
-# Fitting OLR Model Using clm() (Following Jennifer's method)
+```{r}
+# generate predictions for the ordinal model and the ordinal model  with interactions
+predict(ord_mod, new_dataset, type="probs")
+predict(ord_mod_interactions, new_dataset, type="probs")
 
-```{r}
-dataset_clean$Q271_simple <- as.numeric(dataset_clean$Q271)
-# dataset_clean$Q271 <- factor(dataset_clean$Q271,
-#                             levels = c("(5) Not at all important",
-#                                      "(4) Slightly important",
-#                                      "(3) Moderately important",
-#                                      "(2) Very important",
-#                                      "(1) Extremely important"),
-#                             ordered = TRUE)
+new_dataset <- data.frame(race_recoded=c("White", "Black", "Asian American", "Latino", "AIAN", "NHPI", "Arab/MENA"),
+                          Q271=c("(5) Not at all important", "(4) Slightly important", "(3) Moderately important", "(2) Very important", "(1) Extremely important", "(4) Slightly important", "(5) Not at all important"),
+                          age_recoded = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+", "18-29"),
+                          S3B = c("(1) Man", "(2) Woman","(3) Non-binary","(4) Something else (Specify)", "(1) Man", "(2) Woman", "(1) Man"),
+                          educ = c("No college experience", "Some college experience", "College graduate", "No college experience", "Some college experience", "College graduate", "College graduate"),
+                          Q21 = c("(1) Republican",  "(2) Democrat","(3) Independent", "(4) Other party", "(1) Republican",  "(2) Democrat",  "(3) Independent"))
 
-m1 <- clm(Q29 ~ Q271 + race_recoded + age_recoded +
-            rural + # Rural or Urban/Suburban
-            sex_orientation + # Sexual Orientation
-            Religion + 
-            educ + # Educational Attainment
-            S3B + # Gender
-            S7 + # Nativity
-            Q21, # Party affiliation
-          data = dataset_clean)
-
-summary(m1)
-
-# Check for multicollinearity using a linear approximation
-vif_m1 <- vif(lm(as.numeric(Q29) ~ Q271 + race_recoded + age_recoded +
-            rural + # Rural or Urban/Suburban
-            sex_orientation + # Sexual Orientation
-            Religion + 
-            educ + # Educational Attainment
-            S3B + # Gender
-            S7 + # Nativity
-            Q21, # Party affiliation
-          data = dataset_clean))
-
-print("Variance Inflation Factors:")
-print(vif_m1)
-
-# Test proportional odds assumption
-print("Testing proportional odds assumption:")
-test1 <- nominal_test(m1)
-print("Nominal test for Model 1:")
-print(test1)
-
-# Alternative diagnostic approach using linear approximation
-lm_approx <- lm(as.numeric(Q29) ~ Q271 + race_recoded + age_recoded +
-            rural + # Rural or Urban/Suburban
-            sex_orientation + # Sexual Orientation
-            Religion + 
-            educ + # Educational Attainment
-            S3B + # Gender
-            S7 + # Nativity
-            Q21, # Party affiliation
-          data = dataset_clean)
-diagnostics_df <- data.frame(
-  fitted = fitted(lm_approx),
-  residuals = residuals(lm_approx),
-  index = 1:nrow(dataset_clean)
-)
-
-# Plot residuals vs fitted values using linear approximation
-p1 <- ggplot(diagnostics_df, aes(x = fitted, y = residuals)) +
-  geom_point(alpha = 0.1) +
-  geom_smooth(method = "loess") +
-  theme_minimal() +
-  labs(title = "Residuals vs Fitted Values (Linear Approximation)",
-       x = "Fitted Values",
-       y = "Residuals")
-print(p1)
-
-# Calculate and plot Cook's distance
-cooks_d <- cooks.distance(lm_approx)
-influence_df <- data.frame(
-  index = 1:length(cooks_d),
-  cooks_distance = cooks_d
-)
-
-p2 <- ggplot(influence_df, aes(x = index, y = cooks_distance)) +
-  geom_point() +
-  geom_hline(yintercept = 4/nrow(dataset_clean), color = "red", linetype = "dashed") +
-  theme_minimal() +
-  labs(title = "Cook's Distance Plot",
-       x = "Observation Number",
-       y = "Cook's Distance")
-print(p2)
-
-# Count influential observations
-influential_obs <- which(cooks_d > 4/nrow(dataset_clean))
-print("Number of influential observations:")
-print(length(influential_obs))
-
-# Create prediction visualization
-pred_data <- expand.grid(
-  Q271 = levels(dataset_clean$Q271),
-  race_recoded = levels(dataset_clean$race_recoded), 
-  age_recoded = levels(dataset_clean$age_recoded),
-  rural = levels(dataset_clean$rural),
-  sex_orientation = levels(dataset_clean$sex_orientation),
-  Religion = levels(dataset_clean$Religion),
-  educ = levels(dataset_clean$educ),
-  S3B = levels(dataset_clean$S3B),
-  S7 = levels(dataset_clean$S7),
-  Q21 = levels(dataset_clean$Q21)
-)
-
-# Get predicted probabilities
-preds <- predict(m1, newdata = pred_data, type = "prob")
-# Convert predictions to numeric scores (weighted average)
-pred_data$predicted_score <- apply(preds$fit, 1, function(x) sum(x * 1:4))
-
-# Create prediction plot
-p3 <- ggplot(pred_data, aes(x = Q271, y = predicted_score, color = race_recoded)) +
-  geom_point() +
-  geom_line(aes(group = race_recoded)) +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  labs(title = "Predicted Political Interest by Racial Identity Importance and Race",
-       x = "Importance of Racial Identity",
-       y = "Predicted Political Interest Level (1-4)",
-       color = "Race/Ethnicity")
-print(p3)
-
-# Print the coefficients table
-print("Coefficients for Model 1:")
-coef_table <- summary(m1)$coefficients
-print(coef_table)
-
-# Calculate and print odds ratios for the coefficients
-odds_ratios <- exp(coef_table[, "Estimate"])
-print("Odds Ratios:")
-print(odds_ratios)
-
+predict(m_mult, new_dataset, type = "probs")
 ```
-
-## Interpretation
-
-1.  Model Fit:
-    1.  AIC is 42342.54
-    2.  Log-likelihood is -21125.27
-2.  Proportional Odds Assumption
-
--   The nominal test for Model 1 shows significant violations (p \< 0.05 for all predictor variables except for Religion)
-
--   This suggests the effect of racial identity importance varies across different levels of political interest
-
-1.  Multicollinearity
-
--   Interpretation of VIF values
-
-1.  Residuals and Model Diagnostics
-
--   The residuals plot shows:
-
-    -   A relatively flat line around zero
-
-    -   Straight downward slope patterns in the residuals
-
-1.  Influential Points:
-    -   There are 804 influential observations.
-
-    -   Cook's distance plot shows multiple observations above the threshold.
-
-    -   Given the large sample size (17,545), this number of influential points is not extremely concerning.
-2.  Key Findings from Model 1:
-    1.  The different levels of racial identity importance are all significant
-    2.  There are differences among racial/ethnic groups
-3.  Predicted Relationships (from plot)
-    1.

### 2024-12-08 peter_models.qmd
updating with brant test

@@
 
+# Testing proportional odds assumption with Brant test
+```{r}
+library(brant)
+brant(ord_mod_interactions)
+```
+
 # Fitting MLR Model
@@
 ```
+# Comparing confusion matrix
+```{r}
+# Confusion matrix for ordinal model with interactions
+confusionMatrix(predict(ord_mod_interactions), dataset_clean$Q29, mode="everything")
+
+# Confusion matrix for multinomial model
+conf_matrix
+```

### 2024-12-08 OldMain/Jess_Main.Rmd
updated code

---
title: "Jess_Main"
output: pdf_document
date: "2024-12-08"
---

```{r}
# Load necessary libraries
library(dplyr)   # Provides %>% and mutate
library(tidyverse)  # Optionally load the entire tidyverse if needed
library(pROC)
```

```{r}
file_path <- "/Users/nakiyahdhariwala/StatsProject/39096-0001-Data.rda"

load(file_path)          # Load the RDA file
ls()                     # List objects loaded into the environment
dataset <- da39096.0001  # Access the dataset

head(dataset)
ncol(dataset)
nrow(dataset)
```

```{r}
# Calculate proportions
proportions <- prop.table(table(dataset$Q1R5))
proportions
# Define custom labels
custom_labels <- c("Not important", "Important", "Not ")

# Combine custom labels with percentages
pie_labels <- paste(custom_labels, round(proportions * 100, 2), "%")

# Plot the pie chart with custom labels
pie(proportions,
    labels = pie_labels,
    main = "Q1R5: Stopping discrimination against racial/ethnic minorities",
    col = c("lightblue", "salmon"))
```

```{r}
# Define the required columns
required_cols <- c(
  # Outcome Variable: Stopping discrimination against racial/ethnic minorities
  "Q1R5",

  # Demographic Variables
  "S2_RACE_PRIME",   # Primary Race
  "S3",              # SexualOrientation
  "S3B",             # Gender
  "S5_AGE",          # Age categories
  "S13",             # Highest level of education
  "S14",             # Community lived in

  # Variables capturing experiences with racism
  "Q629R1",          # Unfair treatment due to racial background/ethnicity
  "Q629R2",          # Unfair treatment due to skin color
  "Q629R3",          # Unfair treatment due to gender
  "Q629R4",          # Unfair treatment due to sexuality/orientation
  "Q629R5",          # Unfair treatment due to immigration status
  "Q629R6",          # Unfair treatment due to religion
  "Q629R7",          # Unfair treatment due to accent

  "Q627",            # Treated unfairly or discriminated against because of your race, ethnicity, gender, sexuality, being an immigrant, religious heritage or having an accent?
  "Q633",            # Discrimination impact on life satisfaction

  # Religion-related variables
  "Q58R1", "Q58R2", "Q58R3", "Q58R4", "Q58R5", "Q58R6", "Q58R7", 
  "Q58R8", "Q58R9", "Q58R10", "Q58R11", "Q58R12", "Q58R13", 
  "Q58R14", "Q58R15", # Various religious affiliations

  # Perceptions of discrimination
  "Q619_Q626R1",     # Perceptions of discrimination against Whites
  "Q619_Q626R2",     # Perceptions of discrimination against Blacks
  "Q619_Q626R3",     # Perceptions of discrimination against Asian Americans
  "Q619_Q626R4",     # Perceptions of discrimination against Native Americans
  "Q619_Q626R5",     # Perceptions of discrimination against Immigrants
  "Q619_Q626R6",     # Perceptions of discrimination against Latinos
  "Q619_Q626R7",     # Perceptions of discrimination against GaysAndLesbians
  "Q619_Q626R8"      # Perceptions of discrimination against Muslims
)
```

```{r}
# Create a new dataframe with selected variables
dataset_new <- dataset[, required_cols]

# Recode the outcome variable 'Q1R5' for readability
dataset_new$Q1R5 <- factor(
  ifelse(
    dataset_new$Q1R5 == "(0) NO TO: Stopping discrimination against racial/ethnic minorities", 
    "Not important",
    "Important"  # Remaining cases are "Important"
  ),
  levels = c("Not important", "Important") # Define factor levels
)

# Define the columns for perceptions of discrimination
perception_cols <- c(
  "Q619_Q626R1",  # Perceptions of discrimination against Whites
  "Q619_Q626R2",  # Perceptions of discrimination against Blacks
  "Q619_Q626R3",  # Perceptions of discrimination against Asian Americans
  "Q619_Q626R4",  # Perceptions of discrimination against Native Americans
  "Q619_Q626R5",  # Perceptions of discrimination against Immigrants
  "Q619_Q626R6",  # Perceptions of discrimination against Latinos
  "Q619_Q626R7",  # Perceptions of discrimination against GaysAndLesbians
  "Q619_Q626R8",  # Perceptions of discrimination against Muslims
  "Q633"          # Discrimination impact on life satisfaction
)

# Recode 'Don' as 'Don't know' and ensure columns remain factors
dataset_new[perception_cols] <- lapply(dataset_new[perception_cols], function(x) {
  x <- as.character(x)  # Temporarily convert to character for recoding
  x <- ifelse(x == "(5) Don", "(5) Don't know", x)  # Recode 'Don' to 'Don't know'
  factor(x)  # Convert back to factor
})

# Verify the changes in perception columns
head(dataset_new[perception_cols])

# Define required columns to convert to factors (if not already defined)
# Assuming `required_cols` is previously defined; otherwise, specify explicitly
dataset_new[required_cols] <- lapply(dataset_new[required_cols], as.factor)

# Verify the structure of the dataset
str(dataset_new[required_cols])

# Display the first few rows to check the changes
head(dataset_new)

```

```{r}
# Data Cleaning and Recoding for dataset_new
dataset_new <- dataset_new %>% 
  mutate(
    # Education recoding
    educ = case_when(
      S13 == "(1) Grades 1-8" ~ "No college experience",
      S13 == "(2) Some High School, but did not graduate" ~ "No college experience",
      S13 == "(3) High School graduate or GED" ~ "No college experience",
      S13 == "(4) Some college" ~ "Some college experience",
      S13 == "(5) Associates, 2-year degree" ~ "College graduate",
      S13 == "(6) Bachelors, 4-year degree" ~ "College graduate",
      S13 == "(7) Post-graduate degree" ~ "College graduate"
    ),
    # Sexual orientation recoding
    sex_orientation = case_when(
      S3 == "(01) Straight, that is, not gay or lesbian" ~ "Straight",
      S3 == "(02) Gay or lesbian" ~ "Queer",
      S3 == "(03) Bisexual" ~ "Queer",
      S3 == "(04) Something else (Specify)" ~ "Queer",
      S3 == "(88) Refused" ~ "Refused/Don't Know",
      S3 == "(99) I don" ~ "Refused/Don't Know"
    ),
    # Urban vs rural recoding
    rural = case_when(
      S14 == "(1) Large urban area" ~ "Urban/suburban",
      S14 == "(2) Large suburb near large city" ~ "Urban/suburban",
      S14 == "(3) Small suburb near small town or city" ~ "Urban/suburban",
      S14 == "(4) Small town or small city" ~ "Urban/suburban",
      S14 == "(5) Rural area" ~ "Rural"
    ),
    # Race recoding
    race_recoded = case_when(
      S2_RACE_PRIME == "(1) White" ~ "White",
      S2_RACE_PRIME == "(2) Hispanic or Latino" ~ "Latino",
      S2_RACE_PRIME == "(3) Black or African American" ~ "Black",
      S2_RACE_PRIME == "(4) Asian American" ~ "Asian American",
      S2_RACE_PRIME == "(5) American Indian/Native American" ~ "AIAN",
      S2_RACE_PRIME == "(6) Arab, Middle Eastern or North African" ~ "Arab/MENA",
      S2_RACE_PRIME == "(7) Native Hawaiian" ~ "NHPI",
      S2_RACE_PRIME == "(8) Not Hawaiian, but other Pacific Islander" ~ "NHPI"
    ),
    # Age recoding
    age_recoded = case_when(
      S5_AGE == "(1) 16-17" ~ "16-17",
      S5_AGE == "(2) 18-29" ~ "18-29",
      S5_AGE == "(3) 30-39" ~ "30-39",
      S5_AGE == "(4) 40-49" ~ "40-49",
      S5_AGE == "(5) 50-59" ~ "50-59",
      S5_AGE == "(6) 60-69" ~ "60-69",
      S5_AGE == "(7) 70 +" ~ "70+"
    )
  ) %>%
  mutate(
    # Converting variables to factors
    race_recoded = factor(race_recoded, levels = c("White", "Black", "Asian American", "Latino", "AIAN", "NHPI", "Arab/MENA")),
    age_recoded = factor(age_recoded, levels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+")),
    educ = factor(educ, levels = c("No college experience", "Some college experience", "College graduate")),
    rural = factor(rural, levels = c("Rural", "Urban/suburban")),
    sex_orientation = factor(sex_orientation, levels = c("Straight", "Queer", "Refused/Don't Know"))
  )

# Convert individual religion columns to numeric
for (i in 1:15) {
  col_name <- paste0("Q58R", i, "_numeric")
  dataset_new[[col_name]] <- as.numeric(sub("^\\((\\d+)\\).*", "\\1", dataset_new[[paste0("Q58R", i)]]))
}

# Create a new Religion variable
dataset_new$Religion <- "None"

# Assign religion based on the numeric values of each Q58R column
religion_mapping <- c(
  "Q58R1_numeric" = "Catholic", 
  "Q58R2_numeric" = "Protestant", 
  "Q58R3_numeric" = "Other Christian",
  "Q58R4_numeric" = "Jewish", 
  "Q58R5_numeric" = "Muslim", 
  "Q58R6_numeric" = "Hindu", 
  "Q58R7_numeric" = "Buddhist", 
  "Q58R8_numeric" = "Mormon/LDS", 
  "Q58R9_numeric" = "Folk Religion", 
  "Q58R10_numeric" = "Daoist", 
  "Q58R11_numeric" = "Shamanism", 
  "Q58R12_numeric" = "Ancestral Religion", 
  "Q58R13_numeric" = "Spiritual but not religious", 
  "Q58R14_numeric" = "Atheist", 
  "Q58R15_numeric" = "Agnostic"
)

# Loop through the mapping to assign religion values
for (col in names(religion_mapping)) {
  dataset_new$Religion[dataset_new[[col]] == 1] <- religion_mapping[[col]]
}

# Handle multiple religious affiliations
dataset_new$Religion[rowSums(dataset_new[, grep("_numeric", names(dataset_new))] == 1) > 1] <- "Multiple Religions"

# Convert Religion to a factor with specific levels
dataset_new$Religion <- factor(dataset_new$Religion,
                               levels = c("Catholic", "Protestant", "Other Christian", "Jewish", "Muslim",
                                          "Hindu", "Buddhist", "Mormon/LDS", "Folk Religion",
                                          "Daoist", "Shamanism", "Ancestral Religion",
                                          "Spiritual but not religious", "Atheist", "Agnostic", 
                                          "Multiple Religions", "None"))

# Remove redundant religion columns
redundant_cols <- c(
  paste0("Q58R", 1:15), paste0("Q58R", 1:15, "_numeric")
)

# Remove redundant columns from the dataset
dataset_clean <- dataset_new[, !(names(dataset_new) %in% redundant_cols)]

# Take a look at the updated dataset
head(dataset_clean)

```

```{r}
# Check the number of missing values for each variable
missing_summary <- sapply(dataset_clean, function(x) sum(is.na(x)))

# Create a data frame to display the missing values clearly
missing_df <- data.frame(Variable = names(missing_summary), Missing_Count = missing_summary)
print(missing_df)

# Extract the variables Q629R1 to Q629R7
racism_vars <- dataset_clean[, grep("^Q629R[1-7]$", names(dataset_clean))]

# Calculate the proportion of missing values for each variable
missing_proportions <- colSums(is.na(racism_vars)) / nrow(racism_vars)

# Convert to a data frame
missing_summary <- data.frame(Variable = names(missing_proportions),
                              Missing_Proportion = round(missing_proportions, 4))

print(missing_summary)
```

```{r}
# Remove the variables Q629R1 to Q629R7
dataset_clean <- dataset_clean[, !(names(dataset_clean) %in% c("Q629R1", "Q629R2", 
                                                               "Q629R3", "Q629R4", 
                                                               "Q629R5", "Q629R6", "Q629R7"))]

head(dataset_clean)
```

```{r}
# List of redundant columns to remove
redundant_cols <- c("S2_RACE_PRIME", "S13", "S3", "S5_AGE","S14")

# Remove the redundant columns
dataset_clean <- dataset_clean %>% select(-all_of(redundant_cols))

# Verify the resulting dataset
colnames(dataset_clean)

```

```{r}
  # Rename columns in dataset_clean
colnames(dataset_clean) <- c(
  "ImportanceOfStoppingDiscriminationAgainstMinorities",  # Q1R5
  "Gender",                                              # S3B
  "CommunityLivedIn",                                    # S14
  "DiscriminationImpactOnLifeSatisfaction",              # Q633
  "PerceptionsOfDiscriminationAgainstWhites",            # Q619_Q626R1
  "PerceptionsOfDiscriminationAgainstBlacks",            # Q619_Q626R2
  "PerceptionsOfDiscriminationAgainstAsianAmericans",    # Q619_Q626R3
  "PerceptionsOfDiscriminationAgainstNativeAmericans",   # Q619_Q626R4
  "PerceptionsOfDiscriminationAgainstImmigrants",        # Q619_Q626R5
  "PerceptionsOfDiscriminationAgainstLatinos",           # Q619_Q626R6
  "PerceptionsOfDiscriminationAgainstGaysAndLesbians",   # Q619_Q626R7
  "PerceptionsOfDiscriminationAgainstMuslims",           # Q619_Q626R8
  "Educ",                                                # educ
  "SexOrientation",                                      # sex_orientation
  "Rural",                                               # rural
  "RaceRecoded",                                         # race_recoded
  "AgeRecoded",                                          # age_recoded
  "Religion"                                             # Religion
)

# Verify the renamed columns
colnames(dataset_clean)

```

# model

```{r}
library(vcd)

library(car)
```

```{r}
# Adjusted model without `TreatedUnfairlyOrDiscriminatedAgainst`
model <- glm(ImportanceOfStoppingDiscriminationAgainstMinorities ~ RaceRecoded +
               SexOrientation + Gender + AgeRecoded + 
               Educ + CommunityLivedIn + 
               DiscriminationImpactOnLifeSatisfaction + 
               PerceptionsOfDiscriminationAgainstWhites + 
               PerceptionsOfDiscriminationAgainstBlacks +
               PerceptionsOfDiscriminationAgainstAsianAmericans + 
               PerceptionsOfDiscriminationAgainstNativeAmericans + 
               PerceptionsOfDiscriminationAgainstImmigrants + 
               PerceptionsOfDiscriminationAgainstLatinos + 
               PerceptionsOfDiscriminationAgainstGaysAndLesbians + 
               PerceptionsOfDiscriminationAgainstMuslims +
               Religion,

             data = dataset_clean, family = "binomial")

```

```{r}
# Check multicollinearity using VIF
vif(model)
```

```{r}
# Perform stepwise selection
model_stepwise <- step(model, 
                       direction = "both",
                       trace = 0)
```

```{r}
# Check multicollinearity using VIF
vif(model_stepwise)
```

```{r}
# Extract odds ratios
odds_ratios <- exp(coef(model_stepwise))

# Get confidence intervals for coefficients
conf_intervals <- exp(confint(model_stepwise))

# Extract p-values
p_values <- summary(model_stepwise)$coefficients[, 4]

# Combine everything into a single data frame
results_model_stepwise <- data.frame(
  OddsRatio = odds_ratios,
  LowerCI = conf_intervals[, 1],
  UpperCI = conf_intervals[, 2],
  PValue = p_values
)

# Display results
results_model_stepwise
```

```{r}
# Filter for significant variables (p-value < 0.05)
significant_data <- results_model_stepwise[results_model_stepwise$PValue < 0.05, ]

# Add a column for variable names to use in the plot
significant_data <- significant_data %>% 
  mutate(Variable = rownames(significant_data))

# Create the plot
ggplot(significant_data, aes(x = OddsRatio, y = reorder(Variable, OddsRatio))) +
  geom_point(size = 4, color = "blue") +  # Add points for odds ratios
  geom_errorbarh(aes(xmin = LowerCI, xmax = UpperCI), height = 0.3, color = "black") +  # Add horizontal error bars
  geom_vline(xintercept = 1, linetype = "dashed", color = "red", size = 0.7) +  # Add a vertical reference line at Odds Ratio = 1
  labs(
    title = "Significant Variables with Confidence Intervals (Odds Ratio)",
    x = "Odds Ratio",
    y = "Variables"
  ) +
  theme_minimal() +  # Use a minimal theme for better aesthetics
  theme(
    axis.text.y = element_text(size = 10, color = "black"),  # Improve y-axis text readability
    axis.title.x = element_text(size = 12, face = "bold"),  # Highlight x-axis title
    axis.title.y = element_text(size = 12, face = "bold"),  # Highlight y-axis title
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),  # Center the title
    plot.margin = margin(1, 1, 1, 1, "cm")  # Adjust margins
  )

# Save the plot with larger dimensions
ggsave("Significant_Variables_Plot_Enhanced.png", width = 20, height = 10)

```

```{r}
# Set seed for reproducibility
set.seed(123)

# Split data into training (70%) and testing (30%)
library(caret)
train_index <- createDataPartition(dataset_clean$ImportanceOfStoppingDiscriminationAgainstMinorities, p = 0.7, list = FALSE)
train_data <- dataset_clean[train_index, ]
test_data <- dataset_clean[-train_index, ]

# Train stepwise logistic regression model on training data
library(MASS)  # Required for stepAIC if used in stepwise selection

# Fit the initial model (full model)
full_model <- glm(ImportanceOfStoppingDiscriminationAgainstMinorities ~ ., 
                  data = train_data, family = "binomial")

# Perform stepwise selection
final_model_stepwise <- stepAIC(full_model, direction = "both", trace = FALSE)

# Generate predictions on the test data
predicted_probs_stepwise <- predict(final_model_stepwise, newdata = test_data, type = "response")

# Convert probabilities to binary outcomes using a threshold (e.g., 0.5)
predicted_classes_stepwise <- ifelse(predicted_probs_stepwise > 0.5, "Important", "Not important")

# Ensure levels match
predicted_classes_stepwise <- factor(predicted_classes_stepwise, levels = c("Not important", "Important"))
actual_classes_stepwise <- factor(test_data$ImportanceOfStoppingDiscriminationAgainstMinorities, 
                                  levels = c("Not important", "Important"))

# Create confusion matrix
conf_matrix_stepwise <- confusionMatrix(predicted_classes_stepwise, actual_classes_stepwise)

# Print the confusion matrix
print(conf_matrix_stepwise)

```

-   The model is heavily biased toward predicting the majority class ("Not important"), as indicated by the high sensitivity but very low specificity.

-   While the overall accuracy is decent, it is largely influenced by the majority class prevalence.

-   The specificity (16.83%) suggests poor performance in identifying minority class instances ("Important").

```{r}
# Calculate the ROC curve
roc_curve <- roc(test_data$ImportanceOfStoppingDiscriminationAgainstMinorities, 
                 predicted_probs_stepwise, 
                 levels = c("Not important", "Important"))

# Print the AUC
auc_value <- auc(roc_curve)
cat("AUC:", auc_value, "\n")

# Plot the ROC curve
plot(roc_curve, col = "blue", main = "ROC Curve")
abline(a = 0, b = 1, col = "red", lty = 2)  # Add diagonal line for random guess

# Save the ROC plot
ggsave("roc_curve_plot.png", width = 8, height = 6)

```

1.  **AUC Value (0.7187)**:

    -   This is above the threshold of 0.7, which indicates acceptable discrimination ability.

    -   However, it’s not in the "excellent" range (\>0.9). The model still struggles with distinguishing between the classes, likely due to the class imbalance and overlap in predictors.

2.  **ROC Curve**:

    -   The curve is above the diagonal (random guess), which reflects the model's ability to predict better than chance.

    -   The closer the curve gets to the top-left corner, the better the performance.

###

### 2024-12-08 RQ2_main.qmd
updated code

@@
 ## Quarto
+
 The `echo: false` option disables the printing of code (only output is displayed).
+
 ```{r setup, include=FALSE}
@@
 
-```{r echo = TRUE}
+```{r echo = FALSE}
 # Load necessary libraries
@@
 library(pROC)
+library(vcd)
+library(car)
 ```
 
-```{r echo = TRUE}
+```{r echo = FALSE, results = FALSE}
 file_path <- "/Users/nakiyahdhariwala/StatsProject/39096-0001-Data.rda"
@@
 
-```{r}
+```{r echo = FALSE, results=FALSE}
 # Define the required columns
@@
 
-```{r}
+```{r echo=FALSE, results = FALSE}
 # Create a new dataframe with selected variables
@@
 
-# Verify the changes in perception columns
-head(dataset_new[perception_cols])
-
 # Define required columns to convert to factors (if not already defined)
@@
 head(dataset_new)
+```
+
+```{r echo=FALSE, results=FALSE}
+# Data Cleaning and Recoding for dataset_new
+dataset_new <- dataset_new |> 
+  mutate(
+    # Education recoding
+    educ = case_when(
+      S13 == "(1) Grades 1-8" ~ "No college experience",
+      S13 == "(2) Some High School, but did not graduate" ~ "No college experience",
+      S13 == "(3) High School graduate or GED" ~ "No college experience",
+      S13 == "(4) Some college" ~ "Some college experience",
+      S13 == "(5) Associates, 2-year degree" ~ "College graduate",
+      S13 == "(6) Bachelors, 4-year degree" ~ "College graduate",
+      S13 == "(7) Post-graduate degree" ~ "College graduate"
+    ),
+    # Sexual orientation recoding
+    sex_orientation = case_when(
+      S3 == "(01) Straight, that is, not gay or lesbian" ~ "Straight",
+      S3 == "(02) Gay or lesbian" ~ "Queer",
+      S3 == "(03) Bisexual" ~ "Queer",
+      S3 == "(04) Something else (Specify)" ~ "Queer",
+      S3 == "(88) Refused" ~ "Refused/Don't Know",
+      S3 == "(99) I don" ~ "Refused/Don't Know"
+    ),
+    # Urban vs rural recoding
+    rural = case_when(
+      S14 == "(1) Large urban area" ~ "Urban/suburban",
+      S14 == "(2) Large suburb near large city" ~ "Urban/suburban",
+      S14 == "(3) Small suburb near small town or city" ~ "Urban/suburban",
+      S14 == "(4) Small town or small city" ~ "Urban/suburban",
+      S14 == "(5) Rural area" ~ "Rural"
+    ),
+    # Race recoding
+    race_recoded = case_when(
+      S2_RACE_PRIME == "(1) White" ~ "White",
+      S2_RACE_PRIME == "(2) Hispanic or Latino" ~ "Latino",
+      S2_RACE_PRIME == "(3) Black or African American" ~ "Black",
+      S2_RACE_PRIME == "(4) Asian American" ~ "Asian American",
+      S2_RACE_PRIME == "(5) American Indian/Native American" ~ "AIAN",
+      S2_RACE_PRIME == "(6) Arab, Middle Eastern or North African" ~ "Arab/MENA",
+      S2_RACE_PRIME == "(7) Native Hawaiian" ~ "NHPI",
+      S2_RACE_PRIME == "(8) Not Hawaiian, but other Pacific Islander" ~ "NHPI"
+    ),
+    # Age recoding
+    age_recoded = case_when(
+      S5_AGE == "(1) 16-17" ~ "16-17",
+      S5_AGE == "(2) 18-29" ~ "18-29",
+      S5_AGE == "(3) 30-39" ~ "30-39",
+      S5_AGE == "(4) 40-49" ~ "40-49",
+      S5_AGE == "(5) 50-59" ~ "50-59",
+      S5_AGE == "(6) 60-69" ~ "60-69",
+      S5_AGE == "(7) 70 +" ~ "70+"
+    )
+  ) %>%
+  mutate(
+    # Converting variables to factors
+    race_recoded = factor(race_recoded, levels = c("White", "Black", "Asian American", "Latino", "AIAN", "NHPI", "Arab/MENA")),
+    age_recoded = factor(age_recoded, levels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+")),
+    educ = factor(educ, levels = c("No college experience", "Some college experience", "College graduate")),
+    rural = factor(rural, levels = c("Rural", "Urban/suburban")),
+    sex_orientation = factor(sex_orientation, levels = c("Straight", "Queer", "Refused/Don't Know"))
+  )
+
+# Convert individual religion columns to numeric
+for (i in 1:15) {
+  col_name <- paste0("Q58R", i, "_numeric")
+  dataset_new[[col_name]] <- as.numeric(sub("^\\((\\d+)\\).*", "\\1", dataset_new[[paste0("Q58R", i)]]))
+}
+
+# Create a new Religion variable
+dataset_new$Religion <- "None"
+
+# Assign religion based on the numeric values of each Q58R column
+religion_mapping <- c(
+  "Q58R1_numeric" = "Catholic", 
+  "Q58R2_numeric" = "Protestant", 
+  "Q58R3_numeric" = "Other Christian",
+  "Q58R4_numeric" = "Jewish", 
+  "Q58R5_numeric" = "Muslim", 
+  "Q58R6_numeric" = "Hindu", 
+  "Q58R7_numeric" = "Buddhist", 
+  "Q58R8_numeric" = "Mormon/LDS", 
+  "Q58R9_numeric" = "Folk Religion", 
+  "Q58R10_numeric" = "Daoist", 
+  "Q58R11_numeric" = "Shamanism", 
+  "Q58R12_numeric" = "Ancestral Religion", 
+  "Q58R13_numeric" = "Spiritual but not religious", 
+  "Q58R14_numeric" = "Atheist", 
+  "Q58R15_numeric" = "Agnostic"
+)
+
+# Loop through the mapping to assign religion values
+for (col in names(religion_mapping)) {
+  dataset_new$Religion[dataset_new[[col]] == 1] <- religion_mapping[[col]]
+}
+
+# Handle multiple religious affiliations
+dataset_new$Religion[rowSums(dataset_new[, grep("_numeric", names(dataset_new))] == 1) > 1] <- "Multiple Religions"
+
+# Convert Religion to a factor with specific levels
+dataset_new$Religion <- factor(dataset_new$Religion,
+                               levels = c("Catholic", "Protestant", "Other Christian", "Jewish", "Muslim",
+                                          "Hindu", "Buddhist", "Mormon/LDS", "Folk Religion",
+                                          "Daoist", "Shamanism", "Ancestral Religion",
+                                          "Spiritual but not religious", "Atheist", "Agnostic", 
+                                          "Multiple Religions", "None"))
+
+# Remove redundant religion columns
+redundant_cols <- c(
+  paste0("Q58R", 1:15), paste0("Q58R", 1:15, "_numeric")
+)
+
+# Remove redundant columns from the dataset
+dataset_clean <- dataset_new[, !(names(dataset_new) %in% redundant_cols)]
+
+# Take a look at the updated dataset
+head(dataset_clean)
+```
+
+```{r echo=FALSE, results=FALSE}
+# Check the number of missing values for each variable
+missing_summary <- sapply(dataset_clean, function(x) sum(is.na(x)))
+
+# Create a data frame to display the missing values clearly
+missing_df <- data.frame(Variable = names(missing_summary), Missing_Count = missing_summary)
+print(missing_df)
+
+# Extract the variables Q629R1 to Q629R7
+racism_vars <- dataset_clean[, grep("^Q629R[1-7]$", names(dataset_clean))]
+
+# Calculate the proportion of missing values for each variable
+missing_proportions <- colSums(is.na(racism_vars)) / nrow(racism_vars)
+
+# Convert to a data frame
+missing_summary <- data.frame(Variable = names(missing_proportions),
+                              Missing_Proportion = round(missing_proportions, 4))
+
+print(missing_summary)
+```
+
+```{r echo=FALSE, results=FALSE}
+# Remove all redundant columns
+redundant_columns <- c("Q629R1", "Q629R2", "Q629R3", "Q629R4", 
+                       "Q629R5", "Q629R6", "Q629R7",
+                       "S2_RACE_PRIME", "S13", "S3", "S5_AGE","S14")
+
+dataset_clean <- dataset_clean[, !(names(dataset_clean) %in% redundant_columns)]
+
+head(dataset_clean)
+```
+
+```{r echo=FALSE, results=FALSE}
+# Rename remaining columns in dataset_clean
+colnames(dataset_clean) <- c(
+  "ImportanceOfStoppingDiscriminationAgainstMinorities",  # Q1R5
+  "Gender",                                               # S3B
+  "TreatedUnfairlyOrDiscriminatedAgainst",                # Q627
+  "DiscriminationImpactOnLifeSatisfaction",              # Q633
+  "PerceptionsOfDiscriminationAgainstWhites",            # Q619_Q626R1
+  "PerceptionsOfDiscriminationAgainstBlacks",            # Q619_Q626R2
+  "PerceptionsOfDiscriminationAgainstAsianAmericans",    # Q619_Q626R3
+  "PerceptionsOfDiscriminationAgainstNativeAmericans",   # Q619_Q626R4
+  "PerceptionsOfDiscriminationAgainstImmigrants",        # Q619_Q626R5
+  "PerceptionsOfDiscriminationAgainstLatinos",           # Q619_Q626R6
+  "PerceptionsOfDiscriminationAgainstGaysAndLesbians",   # Q619_Q626R7
+  "PerceptionsOfDiscriminationAgainstMuslims",           # Q619_Q626R8
+  "Education",                                           # educ
+  "SexOrientation",                                      # sex_orientation
+  "Rural",                                               # rural
+  "Race",                                                # race_recoded
+  "Age",                                                 # age_recoded
+  "Religion"                                             # Religion
+)
+
+# Verify the renamed columns
+colnames(dataset_clean)
+```
+
+# Model Building
+
+```{r echo=FALSE, results=FALSE}
+# Regression model
+model <- glm(ImportanceOfStoppingDiscriminationAgainstMinorities ~ Gender +
+               TreatedUnfairlyOrDiscriminatedAgainst + 
+               DiscriminationImpactOnLifeSatisfaction +
+               PerceptionsOfDiscriminationAgainstWhites + 
+               PerceptionsOfDiscriminationAgainstBlacks +
+               PerceptionsOfDiscriminationAgainstAsianAmericans + 
+               PerceptionsOfDiscriminationAgainstNativeAmericans + 
+               PerceptionsOfDiscriminationAgainstImmigrants + 
+               PerceptionsOfDiscriminationAgainstLatinos + 
+               PerceptionsOfDiscriminationAgainstGaysAndLesbians + 
+               PerceptionsOfDiscriminationAgainstMuslims +
+               Education + 
+                 SexOrientation +
+                 Rural +
+                 Race +
+                 Age +
+               Religion,
+
+             data = dataset_clean, family = "binomial")
+
+# Check multicollinearity using VIF
+vif(model)
+```
+
+## Model 2: stepwise regression
+```{r echo=FALSE, results=FALSE}
+# Perform stepwise selection
+model_stepwise <- step(model, 
+                       direction = "both",
+                       trace = 0)
+```
+
+```{r echo=FALSE, results=FALSE}
+# print(model_stepwise)
+vif(model_stepwise)
+```
+
+```{r echo=FALSE, results=FALSE}
+
+# Extract odds ratios
+odds_ratios <- exp(coef(model_stepwise))
+
+# Get confidence intervals for coefficients
+conf_intervals <- exp(confint(model_stepwise))
+
+# Extract p-values
+p_values <- summary(model_stepwise)$coefficients[, 4]
+
+# Combine everything into a single data frame
+results_model_stepwise <- data.frame(
+  OddsRatio = odds_ratios,
+  LowerCI = conf_intervals[, 1],
+  UpperCI = conf_intervals[, 2],
+  PValue = p_values
+)
+
+# Display results
+results_model_stepwise
+```
+
+## Model Diagnostics
+```{r echo=FALSE, results=FALSE}
+# Set seed for reproducibility
+set.seed(123)
+
+# Split data into training (70%) and testing (30%)
+library(caret)
+train_index <- createDataPartition(dataset_clean$ImportanceOfStoppingDiscriminationAgainstMinorities, p = 0.7, list = FALSE)
+train_data <- dataset_clean[train_index, ]
+test_data <- dataset_clean[-train_index, ]
+
+# Train stepwise logistic regression model on training data
+
+# Fit the initial model (full model)
+full_model <- glm(ImportanceOfStoppingDiscriminationAgainstMinorities ~ Gender +
+               TreatedUnfairlyOrDiscriminatedAgainst + 
+               DiscriminationImpactOnLifeSatisfaction +
+               PerceptionsOfDiscriminationAgainstWhites + 
+               PerceptionsOfDiscriminationAgainstBlacks +
+               PerceptionsOfDiscriminationAgainstAsianAmericans + 
+               PerceptionsOfDiscriminationAgainstNativeAmericans + 
+               PerceptionsOfDiscriminationAgainstImmigrants + 
+               PerceptionsOfDiscriminationAgainstLatinos + 
+               PerceptionsOfDiscriminationAgainstGaysAndLesbians + 
+               PerceptionsOfDiscriminationAgainstMuslims +
+               Education + 
+                 SexOrientation +
+                 Rural +
+                 Race +
+                 Age +
+               Religion, 
+                  data = train_data, family = "binomial")
+
+# Perform stepwise selection
+final_model <- step(full_model, 
+                       direction = "both",
+                       trace = 0)
+
+# Generate predictions on the test data
+predicted_probs <- predict(final_model, newdata = test_data, type = "response")
+
+# Convert probabilities to binary outcomes using a threshold (e.g., 0.5)
+predicted_classes <- ifelse(predicted_probs > 0.5, "Important", "Not important")
+
+# Convert predicted classes to factor and align levels with actual classes
+predicted_classes <- factor(predicted_classes, levels = c("Not important", "Important"))
+actual_classes <- factor(test_data$ImportanceOfStoppingDiscriminationAgainstMinorities, 
+                                  levels = c("Not important", "Important"))
+
+# Create confusion matrix
+conf_matrix <- confusionMatrix(predicted_classes, actual_classes)
+
+# Print the confusion matrix
+print(conf_matrix)
+```
+
+```{r echo=FALSE, results=FALSE}
+# Calculate the ROC curve
+roc_curve <- roc(test_data$ImportanceOfStoppingDiscriminationAgainstMinorities, 
+                 predicted_probs_stepwise, 
+                 levels = c("Not important", "Important"))
+
+# Print the AUC
+auc_value <- auc(roc_curve)
+cat("AUC:", auc_value, "\n")
+
+# Plot the ROC curve
+plot(roc_curve, col = "blue", main = "ROC Curve")
+abline(a = 0, b = 1, col = "red", lty = 2)  # Add diagonal line for random guess
+
+# Save the ROC plot
+ggsave("roc_curve_plot.png", width = 8, height = 6)
 
 ```
+
+```{r echo=FALSE}
+## Figure 2: 
+# Filter for significant variables (p-value < 0.05)
+significant_data <- results_model_stepwise[results_model_stepwise$PValue < 0.05, ]
+
+# Add a column for variable names to use in the plot
+significant_data <- significant_data %>% 
+  mutate(Variable = rownames(significant_data))
+
+# Create the plot
+ggplot(significant_data, aes(x = OddsRatio, y = reorder(Variable, OddsRatio))) +
+  geom_point(size = 4, color = "blue") +  # Add points for odds ratios
+  geom_errorbarh(aes(xmin = LowerCI, xmax = UpperCI), height = 0.3, color = "black") +  # Add horizontal error bars
+  geom_vline(xintercept = 1, linetype = "dashed", color = "red", size = 0.7) +  # Add a vertical reference line at Odds Ratio = 1
+  labs(
+    title = "Significant Variables with Confidence Intervals (Odds Ratio)",
+    x = "Odds Ratio",
+    y = "Variables"
+  ) +
+  theme_minimal() +  # Use a minimal theme for better aesthetics
+  theme(
+    axis.text.y = element_text(size = 6, color = "black"),  # Improve y-axis text readability
+    axis.title.x = element_text(size = 12, face = "bold"),  # Highlight x-axis title
+    axis.title.y = element_text(size = 12, face = "bold"),  # Highlight y-axis title
+    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),  # Center the title
+    plot.margin = margin(1, 1, 1, 1, "cm"),  # Adjust margins
+    axis.text.x = element_text(angle = 45, hjust = 1, size = 8),  # Reduce x-axis text size
+    legend.text = element_text(size = 8),  # Reduce legend text size
+    legend.title = element_text(size = 9)  # Reduce legend title size
+  )
+
+# Save the plot with larger dimensions
+ggsave("Significant_Variables_Plot_Enhanced.png", width = 20, height = 10)
+```
+
+## Figure 1: Proportion of outcome variable for RQ 1
+```{r echo=FALSE}
+# Calculate proportions
+proportions <- prop.table(table(dataset$Q1R5))
+proportions
+# Define custom labels
+custom_labels <- c("Not important", "Important")
+
+# Combine custom labels with percentages
+pie_labels <- paste(custom_labels, round(proportions * 100, 2), "%")
+
+# Plot the pie chart with custom labels
+pie(proportions,
+    labels = pie_labels,
+    main = "Q1R5: Stopping discrimination against racial/ethnic minorities",
+    col = c("lightblue", "salmon"))
+
+ggsave("Proportion_Outcome_variable_RQ2.png", width = 20, height = 10)
+```

### 2024-12-09 peter_models.qmd
updating model qmd
updating summary table

--- 2024-12-08 peter_models.qmd
@@
 
+# Prop Tables
+
+```{r}
+library(janitor)
+tabyl(dataset_clean$Q29)
+tabyl(dataset_clean$Q271)
+
+```
+
 # Fitting OLR Model Using polr()
@@
 ```
+
+# Testing proportional odds assumption with Brant test
+
+```{r}
+library(brant)
+brant(ord_mod_interactions)
+```
+
+# Summary Table - OLR with Interactions
+
+| Variable | Odds Ratio | 2.5% (Confidence Interval) | 97.5% (Confidence Interval) |
+|------------------|------------------|------------------|------------------|
+| Q271 (4) Slightly important | 1.0338781 | 0.86761871 | 1.2317529 |
+| Q271 (3) Moderately important | 1.0228231 | 0.88466618 | 1.1826555 |
+| Q271 (2) Very important | 0.8366761 | 0.72314779 | 0.9680337 |
+| Q271 (1) Extremely important | 0.6069588 | 0.52449164 | 0.7024084 |
+| S3B(2) Woman | 1.5887610 | 1.36312914 | 1.8518245 |
+| S3B(3) Non-binary | 0.5380781 | 0.28888679 | 0.9978406 |
+| S3B(4) Something else (Specify) | 0.4648621 | 0.07401426 | 2.6422344 |
+| race Black | 1.1525119 | 1.04988995 | 1.2652177 |
+| race Asian American | 1.6819184 | 1.53325820 | 1.8451528 |
+| race Latino | 1.2583888 | 1.14546256 | 1.3825307 |
+| race AIAN | 1.4882734 | 1.27404185 | 1.7384045 |
+| race NHPI | 1.7838764 | 1.45756175 | 2.1825886 |
+| race Arab/MENA | 0.8628526 | 0.66889425 | 1.1122887 |
+| age 30-39 | 0.9698580 | 0.89048308 | 1.0563119 |
+| age 40-49 | 1.0813779 | 0.98549367 | 1.1865407 |
+| age 50-59 | 0.9107924 | 0.82952560 | 0.9996833 |
+| age 60-69 | 0.7553567 | 0.68762045 | 0.8297158 |
+| age 70+ | 0.5918432 | 0.52889988 | 0.6621363 |
+| educSome college experience | 0.6551005 | 0.60189454 | 0.7129797 |
+| educCollege graduate | 0.4476726 | 0.41528212 | 0.4825867 |
+| Q21(2) Democrat | 0.8823547 | 0.81416780 | 0.9562372 |
+| Q21(3) Independent | 1.7222132 | 1.58330889 | 1.8733830 |
+| Q21(4) Other party | 4.7003458 | 4.03685844 | 5.4748427 |
+| Q271 4) Slightly important:S3B(2) Woman | 1.0044380 | 0.78966197 | 1.2778491 |
+| Q271 (3) Moderately important:S3B(2) Woman | 1.0569625 | 0.87144637 | 1.2819704 |
+| Q271 (2) Very important:S3B(2) Woman | 1.0212671 | 0.84557349 | 1.2335590 |
+| Q271 (1) Extremely important:S3B(2) Woman | 1.2324558 | 1.02421823 | 1.4830676 |
+| Q271 (4) Slightly important:S3B(3) Non-binary | 1.5393710 | 0.46529911 | 5.0344630 |
+| Q271 (3) Moderately important:S3B(3) Non-binary | 1.5246008 | 0.57882727 | 4.0084234 |
+| Q271 (2) Very important:S3B(3) Non-binary | 1.9464262 | 0.77448045 | 4.8746428 |
+| Q271 (1) Extremely important:S3B(3) Non-binary | 1.1408623 | 0.45797295 | 2.8216144 |
+| Q271 (4) Slightly important:S3B(4) Something else (Specify) | 4.4125437 | 0.49596064 | 41.9119559 |
+| Q271 (3) Moderately important:S3B(4) Something else (Specify) | 0.5879278 | 0.05055911 | 4.9095993 |
+| Q271 2) Very important:S3B(4) Something else (Specify) | 9.6145175 | 0.70357378 | 160.9916657 |
+| Q271 1) Extremely important:S3B(4) Something else (Specify) | 1.5327973 | 0.20059858 | 12.5345043 |
+
+# Producing odds ratio plot for ordinal model with interactions
 
@@
 ```
+
+# Comparing confusion matrix
+
+```{r}
+# Confusion matrix for ordinal model with interactions
+confusionMatrix(predict(ord_mod_interactions), dataset_clean$Q29, mode="everything")
+
+# Confusion matrix for multinomial model
+conf_matrix
+```

### 2024-12-09 draft_finalreport.qmd
adding draft report qmd

---
title: "The Effects of Racial Identity and Discrimination on Political Participation and Political Attitudes"
author: "Peter de Guzman, Jennifer Li, Nakiyah Dhariwala, Sizhe Chen"
format: html
date: 2024-12-08
editor: visual
---

# Abstract

Following the Immigration and Nationality Act of 1965, which allowed for increased immigration into the United States, the United States has become increasingly racially and ethnically diverse. However, studies of political participation and public opinion have often failed to include large samples of racial and ethnic minority Americans, limiting researchers from drawing conclusions on the attitudes of these groups. Using survey data from the 2020 Collaborative Multi-Racial Post-Election Survey, which includes oversamples of racial and ethnic minority populations in the U.S., we investigate the impact of racial identity and racism on political participation and public opinion. First, we analyze whether valuing belonging in a racial group can influence a respondent’s interest in politics. Then, we analyze how personal experiences with racism may influence how respondents value policies aimed at addressing racial and ethnic discrimination. Our findings suggest that racial group belonging is not a predictor of political interest, while gender identity, racial identity, and party affiliation may predict higher levels of political interest. Additionally, gender identity, racial identity, and personal experiences with discrimination may predict increased levels of support for anti-discrimination policies.

# Introduction

The United States is becoming increasingly racially and ethnically diverse. In the decades following the 1965 Immigration and Nationality Act, the population of Americans born outside of the U.S. increased by more than 400% \[1\]. This increased diversity was also visible in the results from the 2020 decennial U.S. Census. According to data from the 2020 U.S. Census, the share of Americans that are White alone (and do not identify as any other race or ethnicity in addition to White) has declined by approximately 9% since 2010. In addition, the population of multiracial Americans has experienced a 276% increase since the 2010 decennial Census. The percentage of people who reported multiple races increased more than all of the single-race groups, increasing from approximately 3% of the population in 2010 to 10% of the population in 2020 \[2\].

As a result of these trends that have led to increased shares of non-white groups in the U.S. the electoral power of racial and ethnic minority groups has become more salient in American politics. Researchers, policymakers, and advocacy groups must understand the factors that drive political participation and issue preferences among racial and ethnic minority groups, as well as differences between groups.

Researchers commonly understand the drivers for participation in American politics through the “resource model of political participation”, popularized by Henry E. Brady, Sidney Verba and Kay Lehman Schlozman \[3\]. This theory asserts that socioeconomic status predicts political participation, with a specific focus on resources such as time, money, and the organizational skills that are used to navigate civic life (“civic skills”). However, other scholars have pushed back on the generalizability of this theory for non-white groups, as these analyses were based on studies of white Americans and their conclusions may not hold true for members of minority groups \[4\]. In this study, we analyze how racial and ethnic minority groups navigate American politics and the factors that influence their political participation and issue preference.

In a similar vein, researchers have studied whether racial and ethnic discrimination can have a causal effect on the political participation of minority communities. Recent studies on Latino Americans and Indian Americans \[5\] \[6\], have found mixed effects. We address how racial and ethnic discrimination may potentially shape the issue preferences of American voters in this study.

Our analysis focuses on two topics within the field of racial and ethnic politics.

1.  Are Americans who believe that their race is important to their identity more likely to be interested in politics than those that are not?

2.  How do personal experiences with racism, in combination with demographic factors such as race, gender, and education, influence the importance respondents place on policies aimed at addressing racial and ethnic discrimination, controlling for other social and political factors?

To investigate these two research questions, we used survey data from the 2020 Collaborative Multi-Racial Post-Election Survey (CMPS) \[7\]. The CMPS is an extensive dataset designed to capture perspectives across racial and ethnic groups in the United States. The survey was offered in multiple languages, including English, Spanish, Chinese, Korean, Vietnamese, Arabic, Urdu, Farsi, and Haitian Creole.

# Methods

## Data

The 2020 Collaborative Multi-Racial Post-Election Survey (CMPS) dataset contains 17,545 rows, with each row representing an individual respondent's data, and 1,490 columns, each capturing a specific variable or question response. These variables include demographic information and responses to survey questions on political and social issues.

## Variable Selection

Prior to modelling, we conducted exploratory data analysis to examine relationships between our predictor and outcome variables. First, we created summary tables for multiple survey questions. Then, we conducted chi-square tests to understand the relationship between the categorical variables. Finally, we plotted the relationships between the outcome variables and demographic variables to account for potential differences for subgroups in our dataset.

We streamlined the dataset by consolidating religious affiliations into a single categorical variable, Religion, with a new category, Multiple Religions, for respondents reporting more than one affiliation. Variables with over 50% missing values were excluded. Key demographic variables, including educational attainment, sexual orientation, and rurality were re-coded into simplified categories. Age and race were structured into meaningful groups. Extraneous variables were removed, and categorical variables were converted to ordered factors to ensure consistency and clarity for analysis.

## Model Fitting and Evaluation

### Research Question 1: How Racial Group Belonging Impacts Interest in Politics

To model the outcome variable Q29 (interest in participating in politics), we used an ordinal logistic regression with predictors including Q271 (importance of belonging to a racial group to one’s identity), race, age, educational attainment, and political party affiliation. We also included an interaction term between gender and Q271 to assess whether the effect of the importance of belonging to a racial group on one’s identity differs based on one’s gender identity. To assess the model we used a confusion matrix to compare predicted and observed values. Finally, we used the Brant test to check for violations of the proportional odds assumption.

### Research Question 2: How Experiences with Discrimination Affect the Prioritization of Anti-Discrimination Policies

We used a binomial logistic regression to model the outcome variable, Q1R5 (importance placed on anti-discrimination policies), with predictors including race, gender identity, sexual orientation, age, education, community type, experiences with racism, and perception of racism for different groups. An interaction term between gender and race was excluded due to multicollinearity. Stepwise regression and model diagnostics, including VIF and AIC, ensured a stable and interpretable model.

# Results

### Research Question 1:

For the first research question, the outcome variable Q29 (interest in participating in politics) was modeled using ordinal logistic regression. Q29, an ordinal variable captured responses to the question: “Some people are very interested in politics while other people can’t stand politics, how about you? Are you…” with response options “Very interested in politics”, “Somewhat interested”, “Not that interested in politics”, and “Not at all interested in politics.” The overall sample reported high levels of interest in politics, with 43% of respondents reporting that they were somewhat interested and one in four (25%) saying they were very interested in politics.

The model also included the predictor variable Q271, a categorical variable that captured responses to the question: “How important is being \[RACIAL GROUP\] to your identity?”, with response options “Extremely important”, “Very important”, “Moderately important”, “Slightly important”, and “Not at all important”. The respondent’s identified racial group was brought into the survey question based on their prior survey responses. Over half (56%) of respondents reported that being a member of their racial group was very or extremely important to their identity. Only 23% of respondents said that being a member of their racial group was slightly or not at all important to their identity.

The model also included multiple predictor variables that control for the impact of other aspects of a respondent’s identity, socioeconomic status, and political preferences These included demographic variables such as the primary race reported by a respondent, their age, and gender, as well as a respondent’s reported level of educational attainment and political party affiliation. Finally, the model included an interaction term between gender and Q271 (the importance a respondent placed in belonging to their racial group) to investigate whether there are combined effects of gender and racial belonging.

(Odds Ratio Plot)

Controlling for age, race, gender identity, educational attainment, and party affiliation, respondents who considered belonging in their racial group to be slightly or moderately important to their identity were marginally more likely to report a higher level of interest in politics than those that considered belonging in their racial group not at all important to their identity, but these effects were not statistically significant. Surprisingly, respondents who considered belonging in their racial group to be very or extremely important to their identity were less likely to report higher levels of interest in politics than the reference category (those who did not place a high importance on belonging to their racial group), holding the other demographic variables constant. These results were statistically significant.

Holding other variables constant, all racial groups in our sample reported increased odds of reporting a higher level of interest in politics compared to white respondents, except for Arab, Middle East & North African (MENA) respondents. However, the effects for respondents of Arab/MENA heritage was not statistically significant. Black respondents were 1.15 times more likely to report a higher level of interest in politics than White respondents. Latino individuals were 1.26 times more likely to report a higher level of interest in politics compared to white individuals. These effects were even more pronounced for smaller racial/ethnic minority groups. Asian Americans were 1.68 times more likely to report a higher level of interest in politics compared to white respondents. Native American (1.49 times more likely) and Native Hawaiian and Pacific Islander respondents (1.78 times more likely) were also more likely to report a higher level of interest in politics compared to white respondents.

For women, the odds of reporting a higher level of interest in politics were 1.58 times higher than men, holding the other demographic variables constant. Non-binary respondents and respondents who specified a different gender identity were less likely to report higher levels of interest in politics than men.  For women and non-binary respondents, reporting that belonging to their racial group was very or extremely important to their racial identity appeared to increase their odds of reporting higher levels of interest in politics, holding other variables constant. However, none of the effects of the interaction term between gender and the importance a respondent placed in belonging to their racial group were statistically significant.

Contrary to the resource model of participation, increases in educational attainment did not result in increased odds of reporting a higher level of interest in politics. Similarly, older respondents reported decreased odds of reporting a higher level of interest in politics compared to the reference category of individuals aged 18-29 years old.

For respondents that identified as Democrats, holding other factors constant, their odds of reporting a higher level of interest in politics decreased 12% compared to Republican respondents. However, these results differed for respondents that identified as Independent or belonging to another party. For Independent respondents, their odds of reporting a higher level of interest in politics were 1.72 times higher than Republican respondents,  holding other variables constant. For respondents that identified with another party, the odds of reporting a higher level of interest in politics were 4.7 times higher than Republican respondents, holding other variables constant. This may indicate that individuals who identify with parties outside of the dominant two party system in the U.S. may already be more interested in politics than typical American respondents.

To determine if the ordinal logistic regression model violated the proportional odds assumption, we produced a confusion matrix. The confusion matrix revealed that the model had a low level of precision, with an accuracy value of 0.4503. The sensitivity of the model was low for each class except for the response option of being “Somewhat interested” in politics, which had an associated sensitivity of (0.8973). These results indicate that the model struggled to correctly identify respondents who were “Very interested”, “Not that interested”, or “Not at all interested” in politics. The specificity of the model was high for each class, except for the response option of being “Somewhat interested” in politics, indicating that the model classified a high rate of false positives for this class, and it was potentially overly biased to predicting that respondents were “Somewhat interested” in politics.

We also produced a confusion matrix for a multinomial model with the same predictor variables. The confusion matrices were very similar, providing evidence that the proportional odds assumption was not violated. In addition, we conducted a Brant test. The Brant test result indicated that the proportional odds assumption was upheld.

### Research Question 2:

For the second research question, the variable outcome variable Q1R5 (i.e. importance placed on stopping discrimination against racial/ethnic minorities) was modeled using binomial logistic regression, an appropriate method for binary outcomes. Despite the skewed distribution of the dependent variable—74.25% of respondents considered it "Not Important" as shown in fig. 1—logistic regression is robust to such imbalances.

(INSERT PIE CHART)

The model incorporated predictors to examine factors influencing the importance respondents placed on anti-discrimination policies. These included personal experiences with racism, captured by **Q627** (unfair treatment or discrimination) and **Q633** (impact of discrimination on life satisfaction), alongside demographic variables such as **S2_RACE_PRIME** (primary race), **S3** (sexual orientation), **S3B** (gender identity), **S5_AGE** (age), **S13** (education), and **S14** (community type: rural, suburban, urban). Variables **Q629R1–Q629R7** were excluded due to substantial missing data (\>52%). Nonetheless, because it overlapped significantly with **Q627 (see fig. 2)**, the overall essence of experiences with racism was retained in a more general and complete form.

(INSERT ODDS RATIO CHART)

To investigate the combined effects of gender and race and engage with intersectionality theory, an interaction term between Gender and Primary Race was initially included in the model. However, this introduced multicollinearity, as indicated by aliased coefficients and elevated Variance Inflation Factor (VIF) values, likely due to data sparsity and the underrepresentation of certain gender-race combinations, resulting in near-perfect linear dependencies. To resolve this, the interaction term was removed, and additional interactions were avoided to mitigate similar issues. Simplifying the model enhanced interpretability and reduced the risk of overfitting.

To further address high VIF values observed in predictors such as **Perceptions of Discrimination Against Latinos** (GVIF = 12.45) and **Perceptions of Discrimination Against Immigrants** (GVIF = 11.53), forward and backward stepwise regression was employed. This approach optimized model fit using the Akaike Information Criterion (AIC) while minimizing multicollinearity. The iterative process ensured that only the most significant predictors were retained, reducing VIF values and improving the model's stability.

Gender and race emerged as critical determinants in prioritizing anti-discrimination policies. Women were 1.25 times more likely than men to consider these policies important, while non-binary individuals exhibited 1.40 times higher odds, though this result was not statistically significant. Race had an even more pronounced impact; Black respondents had 3.55 times higher odds, Asian Americans had 2.12 times higher odds, Latinos had 1.78 times higher odds, Native Hawaiians/Pacific Islanders had 2.17 times higher odds, and Arab/MENA individuals had 2.67 times higher odds of prioritizing these policies compared to White respondents.

Personal experiences with racism further shaped support for these policies. Respondents who reported experiencing unfair treatment or discrimination were 1.12 times more likely to prioritize anti-discrimination efforts. The impact of discrimination on life satisfaction exhibited a notable gradient: individuals who reported no impact on their life satisfaction were 0.73 times less likely to consider these policies important, while those who experienced greater impacts demonstrated stronger support.

Perceptions of discrimination also played a significant role. Respondents perceiving lower levels of discrimination against Black individuals had 0.48–0.65 times lower odds of prioritizing anti-discrimination policies, whereas perceptions of minimal or no discrimination against Whites were associated with 1.47–2.01 times higher odds of supporting such policies. Sexual orientation also influenced prioritization, with queer respondents having 1.15 times higher odds of supporting anti-discrimination policies compared to non-queer respondents.

The confusion matrix reveals that the model has high **sensitivity** (0.9488), thus accurately identifying respondents who consider policies "Not important." However, the low **specificity** (0.1683) indicates that it struggles to correctly classify those who view the policies as "Important." This is expected given the dataset's skew toward "Not important" (74.25%). The **AUC** of 0.7186 suggests moderate predictive ability, but the model's low specificity and imbalance in class distribution highlight areas for improvement.

# Conclusion

In this analysis, we investigated whether valuing belonging in a racial group can influence a respondent’s interest in politics. We found that the importance an individual placed in belonging to their specific racial group did not have a significant effect on their interest in politics. Contrary to the resource model of participation, age and increases in educational attainment did not appear to result in increased odds of reporting a higher level of interest in politics. The most significant predictors of increased levels of interest in politics were an individual’s gender identity, racial identity, and political party affiliation.

Then, we analyzed how personal experiences with racism may influence how respondents value policies aimed at addressing racial and ethnic discrimination. Our findings underscore the intersectional influences of race, gender, and personal experiences with discrimination on public attitudes toward anti-discrimination policies. Furthermore, perceptions of discrimination against various groups highlight the complex social dynamics that shape policy prioritization. The results provide essential insights for tailoring advocacy and policy efforts to address racial and ethnic discrimination effectively.

Despite the value of this survey’s oversamples of racial and ethnic minority communities, some limitations were present in our study. For the ordinal logistic regression, we used educational attainment as a proxy for socioeconomic status. This is a commonly used proxy measure in behavioral and social sciences \[8\]. Further research on the influence of socioeconomic status of political participation for racial/ethnic minorities and the use of other measures such as household income or wealth would help us understand if the aforementioned resource model of participation is indeed generalizable to other groups in the American electorate.

The effect of race on political interest appeared to be more pronounced for members of very small racial/ethnic minority groups in our analysis such as Asian Americans, Native Americans, Native Hawaiians and Pacific Islanders. Both models had low specificity, suggesting that they struggled to correctly identify cases where respondents had low levels of political interest or respondents who considered anti-discrimination policies to be important.

These challenges and opportunities highlight the importance of disaggregated data and further research in partnership with small racial and ethnic minority communities within the United States. Future work on political participation and public opinion among racial and ethnic minorities could support researchers, policymakers, and advocates efforts to better understand and support these communities.

# Appendix

# References

1.  Noe-Bustamante, A. B., Christine Tamir, Lauren Mora and Luis. (2020, August 20). Facts on U.S. immigrants, 2018. Pew Research Center. https://www.pewresearch.org/race-and-ethnicity/2020/08/20/facts-on-u-s-immigrants/

2.  Jones, N., Marks, R., Ramirez, R., & Rios-Vargas, M. (2021, August 12). Improved Race and Ethnicity Measures Reveal U.S. Population Is Much More Multiracial. The United States Census Bureau. https://www.census.gov/library/stories/2021/08/improved-race-ethnicity-measures-reveal-united-states-population-much-more-multiracial.html

3.  Brady, H. E., Verba, S., & Schlozman, K. L. (1995). Beyond Ses: A Resource Model of Political Participation. The American Political Science Review, 89(2), 271–294. https://doi.org/10.2307/2082425

4.  Leighley, J. E. (2021). Strength in Numbers? Princeton University Press.

5.  Besco, R., Garcia-Rios, S., Lagodny, J., Lajevardi, N., Oskooii, K., & Tolley, E. (2022). Fight not flight: The effects of explicit racism on minority political engagement. Electoral Studies, 80, 102515. https://doi.org/10.1016/j.electstud.2022.102515

6.  Wood, Z. J., & John-Henderson, N. A. (2024). Perceived discrimination, political efficacy, and political participation in American Indian adults. *Frontiers in Political Science*, *6*. https://doi.org/10.3389/fpos.2024.1328521

7.  Frasure, Lorrie, Wong, Janelle, Vargas, Edward, and Barreto, Matt. Collaborative Multi-racial Post-election Survey (CMPS), United States, 2020. Inter-university Consortium for Political and Social Research \[distributor\], 2024-06-11. https://doi.org/10.3886/ICPSR39096.v1

8.  Oakes, M. (n.d.). Measuring Socioeconomic Status 1. Learning Objectives. https://obssr.od.nih.gov/sites/obssr/files/Measuring-Socioeconomic-Status.pdf

### 2024-12-09 draft_finalreport.qmd
adding nakiyah analysis to draft report

@@
 ---
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(
+  message = FALSE,   # Suppress messages
+  warning = FALSE    # Suppress warnings
+)
+```
+
+```{r echo = FALSE}
+# Libraries
+library(tidyverse)  
+library(pROC)
+library(vcd)
+library(car)
+```
+
+```{r echo = FALSE, results = FALSE}
+# identifying requireed columns
+required_cols <- c(
+  # Outcome Variable: Stopping discrimination against racial/ethnic minorities
+  "Q1R5",
+
+  # Demographic Variables
+  "S2_RACE_PRIME",   # Primary Race
+  "S3",              # SexualOrientation
+  "S3B",             # Gender
+  "S5_AGE",          # Age categories
+  "S13",             # Highest level of education
+  "S14",             # Community lived in
+
+  # Variables capturing experiences with racism
+  "Q629R1",          # Unfair treatment due to racial background/ethnicity
+  "Q629R2",          # Unfair treatment due to skin color
+  "Q629R3",          # Unfair treatment due to gender
+  "Q629R4",          # Unfair treatment due to sexuality/orientation
+  "Q629R5",          # Unfair treatment due to immigration status
+  "Q629R6",          # Unfair treatment due to religion
+  "Q629R7",          # Unfair treatment due to accent
+
+  "Q627",            # Treated unfairly or discriminated against because of your race, ethnicity, gender, sexuality, being an immigrant, religious heritage or having an accent?
+  "Q633",            # Discrimination impact on life satisfaction
+
+  # Religion-related variables
+  "Q58R1", "Q58R2", "Q58R3", "Q58R4", "Q58R5", "Q58R6", "Q58R7", 
+  "Q58R8", "Q58R9", "Q58R10", "Q58R11", "Q58R12", "Q58R13", 
+  "Q58R14", "Q58R15", # Various religious affiliations
+
+  # Perceptions of discrimination
+  "Q619_Q626R1",     # Perceptions of discrimination against Whites
+  "Q619_Q626R2",     # Perceptions of discrimination against Blacks
+  "Q619_Q626R3",     # Perceptions of discrimination against Asian Americans
+  "Q619_Q626R4",     # Perceptions of discrimination against Native Americans
+  "Q619_Q626R5",     # Perceptions of discrimination against Immigrants
+  "Q619_Q626R6",     # Perceptions of discrimination against Latinos
+  "Q619_Q626R7",     # Perceptions of discrimination against GaysAndLesbians
+  "Q619_Q626R8"      # Perceptions of discrimination against Muslims
+)
+
+```
+
+```{r echo=FALSE, results=FALSE}
+# Create a new dataframe with selected variables
+dataset_new <- dataset[, required_cols]
+
+# Recode the outcome variable 'Q1R5' for readability
+dataset_new$Q1R5 <- factor(
+  ifelse(
+    dataset_new$Q1R5 == "(0) NO TO: Stopping discrimination against racial/ethnic minorities", 
+    "Not important",
+    "Important"  # Remaining cases are "Important"
+  ),
+  levels = c("Not important", "Important") # Define factor levels
+)
+
+# Define the columns for perceptions of discrimination
+perception_cols <- c(
+  "Q619_Q626R1",  # Perceptions of discrimination against Whites
+  "Q619_Q626R2",  # Perceptions of discrimination against Blacks
+  "Q619_Q626R3",  # Perceptions of discrimination against Asian Americans
+  "Q619_Q626R4",  # Perceptions of discrimination against Native Americans
+  "Q619_Q626R5",  # Perceptions of discrimination against Immigrants
+  "Q619_Q626R6",  # Perceptions of discrimination against Latinos
+  "Q619_Q626R7",  # Perceptions of discrimination against GaysAndLesbians
+  "Q619_Q626R8",  # Perceptions of discrimination against Muslims
+  "Q633"          # Discrimination impact on life satisfaction
+)
+
+# Recode 'Don' as 'Don't know' and ensure columns remain factors
+dataset_new[perception_cols] <- lapply(dataset_new[perception_cols], function(x) {
+  x <- as.character(x)  # Temporarily convert to character for recoding
+  x <- ifelse(x == "(5) Don", "(5) Don't know", x)  # Recode 'Don' to 'Don't know'
+  factor(x)  # Convert back to factor
+})
+
+# Define required columns to convert to factors (if not already defined)
+# Assuming `required_cols` is previously defined; otherwise, specify explicitly
+dataset_new[required_cols] <- lapply(dataset_new[required_cols], as.factor)
+
+# Verify the structure of the dataset
+str(dataset_new[required_cols])
+```
+
+```{r}
+# Cleaning and recoding dataset_new
+dataset_new <- dataset_new |> 
+  mutate(
+    # Education recoding
+    educ = case_when(
+      S13 == "(1) Grades 1-8" ~ "No college experience",
+      S13 == "(2) Some High School, but did not graduate" ~ "No college experience",
+      S13 == "(3) High School graduate or GED" ~ "No college experience",
+      S13 == "(4) Some college" ~ "Some college experience",
+      S13 == "(5) Associates, 2-year degree" ~ "College graduate",
+      S13 == "(6) Bachelors, 4-year degree" ~ "College graduate",
+      S13 == "(7) Post-graduate degree" ~ "College graduate"
+    ),
+    # Sexual orientation recoding
+    sex_orientation = case_when(
+      S3 == "(01) Straight, that is, not gay or lesbian" ~ "Straight",
+      S3 == "(02) Gay or lesbian" ~ "Queer",
+      S3 == "(03) Bisexual" ~ "Queer",
+      S3 == "(04) Something else (Specify)" ~ "Queer",
+      S3 == "(88) Refused" ~ "Refused/Don't Know",
+      S3 == "(99) I don" ~ "Refused/Don't Know"
+    ),
+    # Urban vs rural recoding
+    rural = case_when(
+      S14 == "(1) Large urban area" ~ "Urban/suburban",
+      S14 == "(2) Large suburb near large city" ~ "Urban/suburban",
+      S14 == "(3) Small suburb near small town or city" ~ "Urban/suburban",
+      S14 == "(4) Small town or small city" ~ "Urban/suburban",
+      S14 == "(5) Rural area" ~ "Rural"
+    ),
+    # Race recoding
+    race_recoded = case_when(
+      S2_RACE_PRIME == "(1) White" ~ "White",
+      S2_RACE_PRIME == "(2) Hispanic or Latino" ~ "Latino",
+      S2_RACE_PRIME == "(3) Black or African American" ~ "Black",
+      S2_RACE_PRIME == "(4) Asian American" ~ "Asian American",
+      S2_RACE_PRIME == "(5) American Indian/Native American" ~ "AIAN",
+      S2_RACE_PRIME == "(6) Arab, Middle Eastern or North African" ~ "Arab/MENA",
+      S2_RACE_PRIME == "(7) Native Hawaiian" ~ "NHPI",
+      S2_RACE_PRIME == "(8) Not Hawaiian, but other Pacific Islander" ~ "NHPI"
+    ),
+    # Age recoding
+    age_recoded = case_when(
+      S5_AGE == "(1) 16-17" ~ "16-17",
+      S5_AGE == "(2) 18-29" ~ "18-29",
+      S5_AGE == "(3) 30-39" ~ "30-39",
+      S5_AGE == "(4) 40-49" ~ "40-49",
+      S5_AGE == "(5) 50-59" ~ "50-59",
+      S5_AGE == "(6) 60-69" ~ "60-69",
+      S5_AGE == "(7) 70 +" ~ "70+"
+    )
+  ) %>%
+  mutate(
+    # Converting variables to factors
+    race_recoded = factor(race_recoded, levels = c("White", "Black", "Asian American", "Latino", "AIAN", "NHPI", "Arab/MENA")),
+    age_recoded = factor(age_recoded, levels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+")),
+    educ = factor(educ, levels = c("No college experience", "Some college experience", "College graduate")),
+    rural = factor(rural, levels = c("Rural", "Urban/suburban")),
+    sex_orientation = factor(sex_orientation, levels = c("Straight", "Queer", "Refused/Don't Know"))
+  )
+
+# Convert individual religion columns to numeric
+for (i in 1:15) {
+  col_name <- paste0("Q58R", i, "_numeric")
+  dataset_new[[col_name]] <- as.numeric(sub("^\\((\\d+)\\).*", "\\1", dataset_new[[paste0("Q58R", i)]]))
+}
+
+# Create a new Religion variable
+dataset_new$Religion <- "None"
+
+# Assign religion based on the numeric values of each Q58R column
+religion_mapping <- c(
+  "Q58R1_numeric" = "Catholic", 
+  "Q58R2_numeric" = "Protestant", 
+  "Q58R3_numeric" = "Other Christian",
+  "Q58R4_numeric" = "Jewish", 
+  "Q58R5_numeric" = "Muslim", 
+  "Q58R6_numeric" = "Hindu", 
+  "Q58R7_numeric" = "Buddhist", 
+  "Q58R8_numeric" = "Mormon/LDS", 
+  "Q58R9_numeric" = "Folk Religion", 
+  "Q58R10_numeric" = "Daoist", 
+  "Q58R11_numeric" = "Shamanism", 
+  "Q58R12_numeric" = "Ancestral Religion", 
+  "Q58R13_numeric" = "Spiritual but not religious", 
+  "Q58R14_numeric" = "Atheist", 
+  "Q58R15_numeric" = "Agnostic"
+)
+
+# Loop through the mapping to assign religion values
+for (col in names(religion_mapping)) {
+  dataset_new$Religion[dataset_new[[col]] == 1] <- religion_mapping[[col]]
+}
+
+# Handle multiple religious affiliations
+dataset_new$Religion[rowSums(dataset_new[, grep("_numeric", names(dataset_new))] == 1) > 1] <- "Multiple Religions"
+
+# Convert Religion to a factor with specific levels
+dataset_new$Religion <- factor(dataset_new$Religion,
+                               levels = c("Catholic", "Protestant", "Other Christian", "Jewish", "Muslim",
+                                          "Hindu", "Buddhist", "Mormon/LDS", "Folk Religion",
+                                          "Daoist", "Shamanism", "Ancestral Religion",
+                                          "Spiritual but not religious", "Atheist", "Agnostic", 
+                                          "Multiple Religions", "None"))
+
+# Remove redundant religion columns
+redundant_cols <- c(
+  paste0("Q58R", 1:15), paste0("Q58R", 1:15, "_numeric")
+)
+
+# Remove redundant columns from the dataset
+dataset_clean <- dataset_new[, !(names(dataset_new) %in% redundant_cols)]
+
+```
+
+```{r echo=FALSE, results=FALSE}
+# Check the number of missing values for each variable
+missing_summary <- sapply(dataset_clean, function(x) sum(is.na(x)))
+
+# Create a data frame to display the missing values clearly
+missing_df <- data.frame(Variable = names(missing_summary), Missing_Count = missing_summary)
+print(missing_df)
+
+# Extract the variables Q629R1 to Q629R7
+racism_vars <- dataset_clean[, grep("^Q629R[1-7]$", names(dataset_clean))]
+
+# Calculate the proportion of missing values for each variable
+missing_proportions <- colSums(is.na(racism_vars)) / nrow(racism_vars)
+
+# Convert to a data frame
+missing_summary <- data.frame(Variable = names(missing_proportions),
+                              Missing_Proportion = round(missing_proportions, 4))
+
+print(missing_summary)
+```
+
+```{r echo=FALSE, results=FALSE}
+# Remove all redundant columns
+redundant_columns <- c("Q629R1", "Q629R2", "Q629R3", "Q629R4", 
+                       "Q629R5", "Q629R6", "Q629R7",
+                       "S2_RACE_PRIME", "S13", "S3", "S5_AGE","S14")
+
+dataset_clean <- dataset_clean[, !(names(dataset_clean) %in% redundant_columns)]
+
+head(dataset_clean)
+```
+
+```{r echo=FALSE, results=FALSE}
+# Rename remaining columns in dataset_clean
+colnames(dataset_clean) <- c(
+  "ImportanceOfStoppingDiscriminationAgainstMinorities",  # Q1R5
+  "Gender",                                               # S3B
+  "TreatedUnfairlyOrDiscriminatedAgainst",                # Q627
+  "DiscriminationImpactOnLifeSatisfaction",              # Q633
+  "PerceptionsOfDiscriminationAgainstWhites",            # Q619_Q626R1
+  "PerceptionsOfDiscriminationAgainstBlacks",            # Q619_Q626R2
+  "PerceptionsOfDiscriminationAgainstAsianAmericans",    # Q619_Q626R3
+  "PerceptionsOfDiscriminationAgainstNativeAmericans",   # Q619_Q626R4
+  "PerceptionsOfDiscriminationAgainstImmigrants",        # Q619_Q626R5
+  "PerceptionsOfDiscriminationAgainstLatinos",           # Q619_Q626R6
+  "PerceptionsOfDiscriminationAgainstGaysAndLesbians",   # Q619_Q626R7
+  "PerceptionsOfDiscriminationAgainstMuslims",           # Q619_Q626R8
+  "Education",                                           # educ
+  "SexOrientation",                                      # sex_orientation
+  "Rural",                                               # rural
+  "Race",                                                # race_recoded
+  "Age",                                                 # age_recoded
+  "Religion"                                             # Religion
+)
+
+# Verify the renamed columns
+colnames(dataset_clean)
+```
+
+# Model Building
+
+```{r echo=FALSE, results=FALSE}
+# Regression model
+model <- glm(ImportanceOfStoppingDiscriminationAgainstMinorities ~ Gender +
+               TreatedUnfairlyOrDiscriminatedAgainst + 
+               DiscriminationImpactOnLifeSatisfaction +
+               PerceptionsOfDiscriminationAgainstWhites + 
+               PerceptionsOfDiscriminationAgainstBlacks +
+               PerceptionsOfDiscriminationAgainstAsianAmericans + 
+               PerceptionsOfDiscriminationAgainstNativeAmericans + 
+               PerceptionsOfDiscriminationAgainstImmigrants + 
+               PerceptionsOfDiscriminationAgainstLatinos + 
+               PerceptionsOfDiscriminationAgainstGaysAndLesbians + 
+               PerceptionsOfDiscriminationAgainstMuslims +
+               Education + 
+                 SexOrientation +
+                 Rural +
+                 Race +
+                 Age +
+               Religion,
+
+             data = dataset_clean, family = "binomial")
+
+# Check multicollinearity using VIF
+vif(model)
+```
+
+## Model 2: stepwise regression
+```{r echo=FALSE, results=FALSE}
+# Perform stepwise selection
+model_stepwise <- step(model, 
+                       direction = "both",
+                       trace = 0)
+```
+
+```{r echo=FALSE, results=FALSE}
+# print(model_stepwise)
+vif(model_stepwise)
+```
+
+```{r echo=FALSE, results=FALSE}
+
+# Extract odds ratios
+odds_ratios <- exp(coef(model_stepwise))
+
+# Get confidence intervals for coefficients
+conf_intervals <- exp(confint(model_stepwise))
+
+# Extract p-values
+p_values <- summary(model_stepwise)$coefficients[, 4]
+
+# Combine everything into a single data frame
+results_model_stepwise <- data.frame(
+  OddsRatio = odds_ratios,
+  LowerCI = conf_intervals[, 1],
+  UpperCI = conf_intervals[, 2],
+  PValue = p_values
+)
+
+# Display results
+results_model_stepwise
+```
+
+## Model Diagnostics
+```{r echo=FALSE, results=FALSE}
+# Set seed for reproducibility
+set.seed(123)
+
+# Split data into training (70%) and testing (30%)
+library(caret)
+train_index <- createDataPartition(dataset_clean$ImportanceOfStoppingDiscriminationAgainstMinorities, p = 0.7, list = FALSE)
+train_data <- dataset_clean[train_index, ]
+test_data <- dataset_clean[-train_index, ]
+
+# Train stepwise logistic regression model on training data
+
+# Fit the initial model (full model)
+full_model <- glm(ImportanceOfStoppingDiscriminationAgainstMinorities ~ Gender +
+               TreatedUnfairlyOrDiscriminatedAgainst + 
+               DiscriminationImpactOnLifeSatisfaction +
+               PerceptionsOfDiscriminationAgainstWhites + 
+               PerceptionsOfDiscriminationAgainstBlacks +
+               PerceptionsOfDiscriminationAgainstAsianAmericans + 
+               PerceptionsOfDiscriminationAgainstNativeAmericans + 
+               PerceptionsOfDiscriminationAgainstImmigrants + 
+               PerceptionsOfDiscriminationAgainstLatinos + 
+               PerceptionsOfDiscriminationAgainstGaysAndLesbians + 
+               PerceptionsOfDiscriminationAgainstMuslims +
+               Education + 
+                 SexOrientation +
+                 Rural +
+                 Race +
+                 Age +
+               Religion, 
+                  data = train_data, family = "binomial")
+
+# Perform stepwise selection
+final_model <- step(full_model, 
+                       direction = "both",
+                       trace = 0)
+
+# Generate predictions on the test data
+predicted_probs <- predict(final_model, newdata = test_data, type = "response")
+
+# Convert probabilities to binary outcomes using a threshold (e.g., 0.5)
+predicted_classes <- ifelse(predicted_probs > 0.5, "Important", "Not important")
+
+# Convert predicted classes to factor and align levels with actual classes
+predicted_classes <- factor(predicted_classes, levels = c("Not important", "Important"))
+actual_classes <- factor(test_data$ImportanceOfStoppingDiscriminationAgainstMinorities, 
+                                  levels = c("Not important", "Important"))
+
+# Create confusion matrix
+conf_matrix <- confusionMatrix(predicted_classes, actual_classes)
+
+# Print the confusion matrix
+print(conf_matrix)
+```
+
+```{r echo=FALSE, results=FALSE}
+# Calculate the ROC curve
+roc_curve <- roc(test_data$ImportanceOfStoppingDiscriminationAgainstMinorities, 
+                 predicted_probs_stepwise, 
+                 levels = c("Not important", "Important"))
+
+# Print the AUC
+auc_value <- auc(roc_curve)
+cat("AUC:", auc_value, "\n")
+
+# Plot the ROC curve
+plot(roc_curve, col = "blue", main = "ROC Curve")
+abline(a = 0, b = 1, col = "red", lty = 2)  # Add diagonal line for random guess
+
+# Save the ROC plot
+#ggsave("roc_curve_plot.png", width = 8, height = 6)
+
+```
 
@@
 
+| Response to Q29                   | Count  | Share of Total Respondents |
+|-----------------------------------|--------|----------------------------|
+| Very interested in politics       | 4,306  | 25%                        |
+| Somewhat interested               | 7,612  | 43%                        |
+| Not that interested in politics   | 3,643  | 21%                        |
+| Not at all interested in politics | 1,984  | 11%                        |
+| Total                             | 17,545 | 100%                       |
+
 The model also included the predictor variable Q271, a categorical variable that captured responses to the question: “How important is being \[RACIAL GROUP\] to your identity?”, with response options “Extremely important”, “Very important”, “Moderately important”, “Slightly important”, and “Not at all important”.
@@
 
-(INSERT PIE CHART)
+```{r echo=FALSE}
+# Calculate proportions
+proportions <- prop.table(table(dataset$Q1R5))
+proportions
+# Define custom labels
+custom_labels <- c("Not important", "Important")
+
+# Combine custom labels with percentages
+pie_labels <- paste(custom_labels, round(proportions * 100, 2), "%")
+
+# Plot the pie chart with custom labels
+pie(proportions,
+    labels = pie_labels,
+    main = "Q1R5: Stopping discrimination against racial/ethnic minorities",
+    col = c("lightblue", "salmon"))
+
+ggsave("Proportion_Outcome_variable_RQ2.png", width = 20, height = 10)
+```
 
 The model incorporated predictors to examine factors influencing the importance respondents placed on anti-discrimination policies.
-These included personal experiences with racism, captured by **Q627** (unfair treatment or discrimination) and **Q633** (impact of discrimination on life satisfaction), alongside demographic variables such as **S2_RACE_PRIME** (primary race), **S3** (sexual orientation), **S3B** (gender identity), **S5_AGE** (age), **S13** (education), and **S14** (community type: rural, suburban, urban).
-Variables **Q629R1–Q629R7** were excluded due to substantial missing data (\>52%).
-Nonetheless, because it overlapped significantly with **Q627 (see fig.
-2)**, the overall essence of experiences with racism was retained in a more general and complete form.
+These included personal experiences with racism, captured by Q627 (unfair treatment or discrimination) and Q633 (impact of discrimination on life satisfaction), alongside demographic variables such as S2_RACE_PRIME (primary race), S3 (sexual orientation), S3B (gender identity), S5_AGE (age), S13 (education), and S14 (community type: rural, suburban, urban).
+Variables Q629R1–Q629R7 were excluded due to substantial missing data (\>52%).
+Nonetheless, because it overlapped significantly with Q627 (see fig.
+2), the overall essence of experiences with racism was retained in a more general and complete form.
 
-(INSERT ODDS RATIO CHART)
+```{r echo = FALSE}
+# Filter for significant variables (p-value < 0.05)
+significant_data <- results_model_stepwise[results_model_stepwise$PValue < 0.05, ]
+
+# Add a column for variable names to use in the plot
+significant_data <- significant_data %>% 
+  mutate(Variable = rownames(significant_data))
+
+# Create the plot
+ggplot(significant_data, aes(x = OddsRatio, y = reorder(Variable, OddsRatio))) +
+  geom_point(size = 4, color = "blue") +  # Add points for odds ratios
+  geom_errorbarh(aes(xmin = LowerCI, xmax = UpperCI), height = 0.3, color = "black") +  # Add horizontal error bars
+  geom_vline(xintercept = 1, linetype = "dashed", color = "red", size = 0.7) +  # Add a vertical reference line at Odds Ratio = 1
+  labs(
+    title = "Significant Variables with Confidence Intervals (Odds Ratio)",
+    x = "Odds Ratio",
+    y = "Variables"
+  ) +
+  theme_minimal() +  # Use a minimal theme for better aesthetics
+  theme(
+    axis.text.y = element_text(size = 6, color = "black"),  # Improve y-axis text readability
+    axis.title.x = element_text(size = 12, face = "bold"),  # Highlight x-axis title
+    axis.title.y = element_text(size = 12, face = "bold"),  # Highlight y-axis title
+    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),  # Center the title
+    plot.margin = margin(1, 1, 1, 1, "cm"),  # Adjust margins
+    axis.text.x = element_text(angle = 45, hjust = 1, size = 8),  # Reduce x-axis text size
+    legend.text = element_text(size = 8),  # Reduce legend text size
+    legend.title = element_text(size = 9)  # Reduce legend title size
+  )
+
+# Save the plot with larger dimensions
+ggsave("Significant_Variables_Plot_Enhanced.png", width = 20, height = 10)
+```
 
@@
 
-To further address high VIF values observed in predictors such as **Perceptions of Discrimination Against Latinos** (GVIF = 12.45) and **Perceptions of Discrimination Against Immigrants** (GVIF = 11.53), forward and backward stepwise regression was employed.
+To further address high VIF values observed in predictors such as Perceptions of Discrimination Against Latinos (GVIF = 12.45) and Perceptions of Discrimination Against Immigrants (GVIF = 11.53), forward and backward stepwise regression was employed.
 This approach optimized model fit using the Akaike Information Criterion (AIC) while minimizing multicollinearity.
@@
 
-The confusion matrix reveals that the model has high **sensitivity** (0.9488), thus accurately identifying respondents who consider policies "Not important." However, the low **specificity** (0.1683) indicates that it struggles to correctly classify those who view the policies as "Important." This is expected given the dataset's skew toward "Not important" (74.25%).
-The **AUC** of 0.7186 suggests moderate predictive ability, but the model's low specificity and imbalance in class distribution highlight areas for improvement.
+The confusion matrix reveals that the model has high sensitivity (0.9488), thus accurately identifying respondents who consider policies "Not important." However, the low specificity (0.1683) indicates that it struggles to correctly classify those who view the policies as "Important." This is expected given the dataset's skew toward "Not important" (74.25%).
+The AUC of 0.7186 suggests moderate predictive ability, but the model's low specificity and imbalance in class distribution highlight areas for improvement.
 
@@
 Future work on political participation and public opinion among racial and ethnic minorities could support researchers, policymakers, and advocates efforts to better understand and support these communities.
-
-# Appendix
 

### 2024-12-09 draft_finalreport.qmd
update draft and mdoels

@@
 library(car)
+library(MASS) # package for model
+library(caret) # package for confusion matrix
+library(janitor)
+library(brant)
+
+```
+
+```{r echo = FALSE, results = FALSE}
+file_path <- "//Users/pdeguz01/Documents/git/Data/2020 CMPS/DS0001/39096-0001-Data.rda"
+
+load(file_path)          # Load the RDA file
+ls()                     # List objects loaded into the environment
+dataset <- da39096.0001  # Access the dataset
+
 ```
@@
 
-```{r}
+```{r, echo=FALSE, results=FALSE}
 # Cleaning and recoding dataset_new
@@
 
-# Model Building
-
 ```{r echo=FALSE, results=FALSE}
@@
 
-## Model 2: stepwise regression
 ```{r echo=FALSE, results=FALSE}
@@
 
-## Model Diagnostics
 ```{r echo=FALSE, results=FALSE}
@@
 # Calculate the ROC curve
-roc_curve <- roc(test_data$ImportanceOfStoppingDiscriminationAgainstMinorities, 
-                 predicted_probs_stepwise, 
-                 levels = c("Not important", "Important"))
-
-# Print the AUC
-auc_value <- auc(roc_curve)
-cat("AUC:", auc_value, "\n")
-
-# Plot the ROC curve
-plot(roc_curve, col = "blue", main = "ROC Curve")
-abline(a = 0, b = 1, col = "red", lty = 2)  # Add diagonal line for random guess
+# roc_curve <- roc(test_data$ImportanceOfStoppingDiscriminationAgainstMinorities, 
+#                  predicted_probs_stepwise, 
+#                  levels = c("Not important", "Important"))
+# 
+# # Print the AUC
+# auc_value <- auc(roc_curve)
+# cat("AUC:", auc_value, "\n")
+# 
+# # Plot the ROC curve
+# plot(roc_curve, col = "blue", main = "ROC Curve")
+# abline(a = 0, b = 1, col = "red", lty = 2)  # Add diagonal line for random guess
 
@@
 
+```
+
+```{r, echo=FALSE, results=FALSE}
+# data load and cleaning for ordinal regression model
+required_cols_ord <- c("Q29", # Outcome Variable; interest in politics
+                   "Q271", # Predictor Variable; importance of race to identity
+                   "S2_RACE_PRIME", # Race
+                   "S3", # Orientation
+                   "S3B", # Gender
+                   "S5_AGE", # Age Category
+                   "S13", # Educational Attainment
+                   "S14", # Community Type
+                   "S7", # Nativity
+                   "Q21", # Party ID
+                   "S2_M", # Multiracial
+
+                   # Religion-related variables
+                   "Q58R1",            # Catholic
+                   "Q58R2",            # Protestant
+                   "Q58R3",            # Other Christian
+                   "Q58R4",            # Jewish
+                   "Q58R5",            # Muslim
+                   "Q58R6",            # Hindu
+                   "Q58R7",            # Buddhist
+                   "Q58R8",            # Mormon/LDS
+                   "Q58R9",            # Folk Religion
+                   "Q58R10",           # Daoist
+                   "Q58R11",           # Shamanism
+                   "Q58R12",           # Ancestral Religion
+                   "Q58R13",           # Spiritual but not religious
+                   "Q58R14",           # Atheist
+                   "Q58R15",            # Agnostic
+
+                   # Variables capturing experiences with racism
+                   "Q627", # Treated unfairly or discriminated against ,
+                   "Q628", # Location of discrimination
+                   "Q629R1" # Unfair treatment due to racial background/ethnicity
+                   )
+
+# Create new df
+data_filtered <- dataset[, required_cols_ord]
+
+# Convert the required columns to factors
+data_filtered[required_cols] <- lapply(data_filtered[required_cols_ord], as.factor)
+
+# Convert individual religion columns to numeric
+for (i in 1:15) {
+  col_name <- paste0("Q58R", i, "_numeric")
+  data_filtered[[col_name]] <- as.numeric(sub("^\\((\\d+)\\).*", "\\1", data_filtered[[paste0("Q58R", i)]]))
+}
+
+# Create a new Religion variable
+data_filtered$Religion <- "None"
+data_filtered$Religion[data_filtered$Q58R1_numeric == 1] <- "Catholic"
+data_filtered$Religion[data_filtered$Q58R2_numeric == 1] <- "Protestant"
+data_filtered$Religion[data_filtered$Q58R3_numeric == 1] <- "Other Christian"
+data_filtered$Religion[data_filtered$Q58R4_numeric == 1] <- "Jewish"
+data_filtered$Religion[data_filtered$Q58R5_numeric == 1] <- "Muslim"
+data_filtered$Religion[data_filtered$Q58R6_numeric == 1] <- "Hindu"
+data_filtered$Religion[data_filtered$Q58R7_numeric == 1] <- "Buddhist"
+data_filtered$Religion[data_filtered$Q58R8_numeric == 1] <- "Mormon/LDS"
+data_filtered$Religion[data_filtered$Q58R9_numeric == 1] <- "Folk Religion"
+data_filtered$Religion[data_filtered$Q58R10_numeric == 1] <- "Daoist"
+data_filtered$Religion[data_filtered$Q58R11_numeric == 1] <- "Shamanism"
+data_filtered$Religion[data_filtered$Q58R12_numeric == 1] <- "Ancestral Religion"
+data_filtered$Religion[data_filtered$Q58R13_numeric == 1] <- "Spiritual but not religious"
+data_filtered$Religion[data_filtered$Q58R14_numeric == 1] <- "Atheist"
+data_filtered$Religion[data_filtered$Q58R15_numeric == 1] <- "Agnostic"
+
+# Convert Religion to a factor
+data_filtered$Religion <- factor(data_filtered$Religion,
+                               levels = c("Catholic", "Protestant", "Other Christian", "Jewish", "Muslim",
+                                          "Hindu", "Buddhist", "Mormon/LDS", "Folk Religion",
+                                          "Daoist", "Shamanism", "Ancestral Religion",
+                                          "Spiritual but not religious", "Atheist", "Agnostic", "None"))
+
+# Remove redundant religion columns (Q58R1 to Q58R15 and their numeric versions)
+redundant_cols <- c(
+  "Q58R1", "Q58R2", "Q58R3", "Q58R4", "Q58R5", "Q58R6", "Q58R7", "Q58R8", 
+  "Q58R9", "Q58R10", "Q58R11", "Q58R12", "Q58R13", "Q58R14", "Q58R15",
+  "Q58R1_numeric", "Q58R2_numeric", "Q58R3_numeric", "Q58R4_numeric", 
+  "Q58R5_numeric", "Q58R6_numeric", "Q58R7_numeric", "Q58R8_numeric", 
+  "Q58R9_numeric", "Q58R10_numeric", "Q58R11_numeric", "Q58R12_numeric", 
+  "Q58R13_numeric", "Q58R14_numeric", "Q58R15_numeric"
+)
+
+# Remove redundant columns from the dataset
+dataset_clean_ord <- data_filtered[, !(names(data_filtered) %in% redundant_cols)]
+
+dataset_clean_ord <- dataset_clean_ord %>% mutate(educ = case_when(S13 == "(1) Grades 1-8" ~ "No college experience",
+                                                           S13 == "(2) Some High School, but did not graduate" ~ "No college experience",
+                                                           S13 == "(3) High School graduate or GED" ~ "No college experience",
+                                                           S13 == "(4) Some college" ~ "Some college experience",
+                                                           S13 == "(5) Associates, 2-year degree" ~ "College graduate",
+                                                           S13 == "(6) Bachelors, 4-year degree" ~ "College graduate",
+                                                           S13 == "(7) Post-graduate degree" ~ "College graduate"),
+                                          sex_orientation = case_when(S3 == "(01) Straight, that is, not gay or lesbian" ~ "Straight",
+                                                                     S3 == "(02) Gay or lesbian" ~ "Queer",
+                                                                     S3 == "(03) Bisexual" ~ "Queer",
+                                                                     S3 == "(04) Something else (Specify)" ~ "Queer",
+                                                                     S3 == "(88) Refused" ~ "Refused/Don't Know",
+                                                                     S3 == "(99) I don" ~ "Refused/Don't Know"),
+                                          rural = case_when(S14 == "(1) Large urban area" ~ "Urban/suburban",
+                                                            S14 == "(2) Large suburb near large city" ~ "Urban/suburban",
+                                                            S14 == "(3) Small suburb near small town or city" ~ "Urban/suburban",
+                                                            S14 == "(4) Small town or small city" ~ "Urban/suburban",
+                                                            S14 == "(5) Rural area" ~ "Rural"),
+                                          white = case_when(S2_RACE_PRIME == "(1) White" ~ "White", 
+                                                            S2_RACE_PRIME == "(2) Hispanic or Latino" ~ "Non-White",
+                                                            S2_RACE_PRIME == "(3) Black or African American" ~ "Non-White",
+                                                            S2_RACE_PRIME == "(4) Asian American" ~ "Non-White",
+                                                            S2_RACE_PRIME == "(5) American Indian/Native American" ~ "Non-White",
+                                                            S2_RACE_PRIME == "(6) Arab, Middle Eastern or North African" ~ "Non-White",
+                                                            S2_RACE_PRIME == "(7) Native Hawaiian" ~ "Non-White",
+                                                            S2_RACE_PRIME == "(8) Not Hawaiian, but other Pacific Islander" ~ "Non-White"),
+                                          race_recoded = case_when(S2_RACE_PRIME == "(1) White" ~ "White",
+                                                                   S2_RACE_PRIME == "(2) Hispanic or Latino" ~ "Latino",
+                                                                   S2_RACE_PRIME == "(3) Black or African American" ~ "Black",
+                                                                   S2_RACE_PRIME == "(4) Asian American" ~ "Asian American",
+                                                                   S2_RACE_PRIME == "(5) American Indian/Native American" ~ "AIAN",
+                                                                   S2_RACE_PRIME == "(6) Arab, Middle Eastern or North African" ~ "Arab/MENA",
+                                                                   S2_RACE_PRIME == "(7) Native Hawaiian" ~ "NHPI",
+                                                                   S2_RACE_PRIME == "(8) Not Hawaiian, but other Pacific Islander" ~ "NHPI"),
+                                          age_recoded = case_when(S5_AGE == "(1) 16-17" ~ "16-17",
+                                                                  S5_AGE == "(2) 18-29" ~ "18-29",
+                                                                  S5_AGE == "(3) 30-39" ~ "30-39",
+                                                                  S5_AGE == "(4) 40-49" ~ "40-49",
+                                                                  S5_AGE == "(5) 50-59" ~ "50-59",
+                                                                  S5_AGE == "(6) 60-69" ~ "60-69",
+                                                                  S5_AGE == "(7) 70 +" ~ "70+"))
+
+dataset_clean_ord <- dataset_clean_ord %>% mutate(race_recoded = factor(dataset_clean_ord$race_recoded, levels = c("White", "Black", "Asian American", "Latino", "AIAN", "NHPI", "Arab/MENA")),
+                                         age_recoded = factor(dataset_clean_ord$age_recoded, levels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+")),
+                                         educ = factor(dataset_clean_ord$educ, levels = c("No college experience", "Some college experience", "College graduate")),
+                                         white = factor(dataset_clean_ord$white, levels = c("White", "Non-White")),
+                                         rural = factor(dataset_clean_ord$rural, levels= c("Rural", "Urban/suburban")),
+                                         sex_orientation = factor(dataset_clean_ord$sex_orientation, levels = c("Straight", "Queer", "Refused/Don't Know")))
+
+```
+
+```{r, echo=FALSE, results=FALSE}
+# Exploratory data analysis
+# # Create a summary table for all categorical variables
+variables <- c("Q29", "Q271", "S2_RACE_PRIME", "S3", "S3B", "S5_AGE", "S13", "S14", "S7", "Q21", "S2_M", "Q627", "Q628", "Q629R1", "Religion", "educ", "sex_orientation", "rural", "white", "race_recoded", "age_recoded")
+
+# # Chi-square tests for each variable
+# chi_square_results <- lapply(variables, function(var) {
+#   chisq.test(table(dataset_clean_ord[[var]], dataset_clean_ord$Q29))
+# })
+# 
+# # Print chi-square results
+# names(chi_square_results) <- variables
+# print(chi_square_results)
+# 
+# tabyl(dataset_clean_ord$Q29)
+# tabyl(dataset_clean_ord$Q271)
+
+```
+
+```{r, echo=FALSE, results=FALSE}
+# ordinal regression model and assessment
+# loading cleaned data
+load("/Users/pdeguz01/Documents/git/Data/2020 CMPS/Output/2020cmps_clean.RData")
+dataset_clean_ord <- dataset_clean
+
+dataset_clean_ord$Q271_reordered <- factor(dataset_clean_ord$Q271,
+                            levels = c("(5) Not at all important",
+                                     "(4) Slightly important",
+                                     "(3) Moderately important",
+                                     "(2) Very important",
+                                     "(1) Extremely important"))
+
+ord_mod_interactions <- polr(Q29 ~ Q271_reordered * S3B +
+                               race_recoded +
+                  age_recoded +
+            educ +  # Educational Attainment
+            Q21, # Party affiliation
+          data = dataset_clean_ord, Hess=TRUE)
+
+summary(ord_mod_interactions)
+# 
+# 
+# pvals_interact <- pnorm(-abs(summary(ord_mod_interactions)$coef[,"t value"]))*2
+# ctable_interact <- cbind(summary(ord_mod_interactions)$coef,pvals)
+# 
+# ctable_interact
+# 
+# (exp_coefs_interact <- exp(cbind(OR=coef(ord_mod_interactions),confint(ord_mod_interactions))))
+# 
+# head(ord_mod_interactions$fitted.values)
+# head(predict(ord_mod_interactions))
+# 
+# confusionMatrix(predict(ord_mod_interactions), dataset_clean_ord$Q29, mode="everything")
+# 
+# # brant test
+# brant(ord_mod_interactions)
+# 
+# # generating predictions
+# new_dataset <- data.frame(race_recoded=c("White", "Black", "Asian American", "Latino", "AIAN", "NHPI", "Arab/MENA"),
+#                           Q271_reordered=c("(5) Not at all important", "(4) Slightly important", "(3) Moderately important", "(2) Very important", "(1) Extremely important", "(4) Slightly important", "(5) Not at all important"),
+#                           age_recoded = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+", "18-29"),
+#                           S3B = c("(1) Man", "(2) Woman","(3) Non-binary","(4) Something else (Specify)", "(1) Man", "(2) Woman", "(1) Man"),
+#                           educ = c("No college experience", "Some college experience", "College graduate", "No college experience", "Some college experience", "College graduate", "College graduate"),
+#                           Q21 = c("(1) Republican",  "(2) Democrat","(3) Independent", "(4) Other party", "(1) Republican",  "(2) Democrat",  "(3) Independent"))
+# predict(ord_mod_interactions, new_dataset, type="probs")
 ```
@@
 
-(Odds Ratio Plot)
+```{r, echo=FALSE, results=FALSE}
+# odds_ratio_plot <- function(model) {
+#   # Get odds ratios and CIs
+#   or_data <- exp(cbind(OR=coef(model),confint(model)))
+#   
+#   # Calculate p-values
+#   p_values <- pnorm(-abs(summary(model)$coefficients[,"t value"]))*2
+#   
+#   # Keep only significant variables
+#   sig_indices <- p_values[names(coef(model))] < 0.05
+#   or_data_sig <- or_data[sig_indices, ]
+#   
+#   ggplot() +
+#     # Add vertical line at OR = 1
+#     geom_vline(xintercept = 1, linetype = "dashed", color = "red", size = 0.7) +
+#     # Add points and error bars
+#     geom_point(aes(x = or_data_sig[,1], 
+#                    y = reorder(rownames(or_data_sig), or_data_sig[,1])), 
+#               size = 4, color = "blue") +
+#     geom_errorbarh(aes(xmin = or_data_sig[,2], 
+#                        xmax = or_data_sig[,3],
+#                        y = reorder(rownames(or_data_sig), or_data_sig[,1])),
+#                    height = 0.3, color = "black") +
+#     # Customize theme
+#     theme_minimal() +
+#     theme(
+#       axis.text.y = element_text(size = 6, color = "black"),
+#       axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
+#       axis.title.x = element_text(size = 12, face = "bold"),
+#       axis.title.y = element_text(size = 12, face = "bold"),
+#       plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
+#       plot.margin = margin(1, 1, 1, 1, "cm"),
+#       legend.text = element_text(size = 8),
+#       legend.title = element_text(size = 9)
+#     ) +
+#     # Add labels
+#     labs(
+#       title = "Significant Variables with Confidence Intervals (Odds Ratio)",
+#       x = "Odds Ratio",
+#       y = "Variables"
+#     )
+# }
+# 
+# # Generate the plot
+# odds_ratio_plot(ord_mod_interactions)
+# ggsave("RQ1_oddsratioplot.png", width = 20, height = 10)
+
+```
+
+![](RQ1_oddsratio.png)
 
@@
 
-```{r echo=FALSE}
+```{r echo=FALSE, results=FALSE}
 # Calculate proportions
@@
 
-ggsave("Proportion_Outcome_variable_RQ2.png", width = 20, height = 10)
+#ggsave("Proportion_Outcome_variable_RQ2.png", width = 20, height = 10)
 ```
@@
 
-```{r echo = FALSE}
+```{r echo = FALSE, results=FALSE}
 # Filter for significant variables (p-value < 0.05)
@@
 # Save the plot with larger dimensions
-ggsave("Significant_Variables_Plot_Enhanced.png", width = 20, height = 10)
+#ggsave("Significant_Variables_Plot_Enhanced.png", width = 20, height = 10)
 ```
+
+![](Significant_Variables_Plot_Enhanced.png){width="20"}
 

### 2024-12-09 peter_models.qmd
update draft and mdoels

@@
 
+```{r}
+odds_ratio_plot <- function(model) {
+  # Get odds ratios and CIs
+  or_data <- exp(cbind(OR=coef(model),confint(model)))
+
+  # Calculate p-values
+  p_values <- pnorm(-abs(summary(model)$coefficients[,"t value"]))*2
+
+  # Keep only significant variables
+  sig_indices <- p_values[names(coef(model))] < 0.05
+  or_data_sig <- or_data[sig_indices, ]
+
+  ggplot() +
+    # Add vertical line at OR = 1
+    geom_vline(xintercept = 1, linetype = "dashed", color = "red", size = 0.7) +
+    # Add points and error bars
+    geom_point(aes(x = or_data_sig[,1], 
+                   y = reorder(rownames(or_data_sig), or_data_sig[,1])), 
+              size = 4, color = "blue") +
+    geom_errorbarh(aes(xmin = or_data_sig[,2], 
+                       xmax = or_data_sig[,3],
+                       y = reorder(rownames(or_data_sig), or_data_sig[,1])),
+                   height = 0.3, color = "black") +
+    # Customize theme
+    theme_minimal() +
+    theme(
+      axis.text.y = element_text(size = 6, color = "black"),
+      axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
+      axis.title.x = element_text(size = 12, face = "bold"),
+      axis.title.y = element_text(size = 12, face = "bold"),
+      plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
+      plot.margin = margin(1, 1, 1, 1, "cm"),
+      legend.text = element_text(size = 8),
+      legend.title = element_text(size = 9)
+    ) +
+    # Add labels
+    labs(
+      title = "Significant Variables with Confidence Intervals (Odds Ratio)",
+      x = "Odds Ratio",
+      y = "Variables"
+    )
+}
+
+# Generate the plot
+odds_ratio_plot(ord_mod)
+
+```
+
 # Fitting MLR Model

### 2024-12-09 copy_nakiyah.qmd
adding a copy of nakiyah script for testing

--- 2024-12-08 RQ2_main.qmd
@@
 ```{r echo = FALSE, results = FALSE}
-file_path <- "/Users/nakiyahdhariwala/StatsProject/39096-0001-Data.rda"
+file_path <- "//Users/pdeguz01/Documents/git/Data/2020 CMPS/DS0001/39096-0001-Data.rda"
 
@@
 ```{r echo=FALSE, results=FALSE}
-# Calculate the ROC curve
-roc_curve <- roc(test_data$ImportanceOfStoppingDiscriminationAgainstMinorities, 
-                 predicted_probs_stepwise, 
-                 levels = c("Not important", "Important"))
-
-# Print the AUC
-auc_value <- auc(roc_curve)
-cat("AUC:", auc_value, "\n")
-
-# Plot the ROC curve
-plot(roc_curve, col = "blue", main = "ROC Curve")
-abline(a = 0, b = 1, col = "red", lty = 2)  # Add diagonal line for random guess
-
-# Save the ROC plot
-ggsave("roc_curve_plot.png", width = 8, height = 6)
+# # Calculate the ROC curve
+# roc_curve <- roc(test_data$ImportanceOfStoppingDiscriminationAgainstMinorities, 
+#                  predicted_probs_stepwise, 
+#                  levels = c("Not important", "Important"))
+# 
+# # Print the AUC
+# auc_value <- auc(roc_curve)
+# cat("AUC:", auc_value, "\n")
+# 
+# # Plot the ROC curve
+# plot(roc_curve, col = "blue", main = "ROC Curve")
+# abline(a = 0, b = 1, col = "red", lty = 2)  # Add diagonal line for random guess
+# 
+# # Save the ROC plot
+# ggsave("roc_curve_plot.png", width = 8, height = 6)
 

### 2024-12-13 peter_models.qmd
updating models

@@
 # Fitting OLR with interaction terms 
+## This is the model we include in the final report.
 
@@
                                      "(1) Extremely important"))
+dataset_clean$Q29_reordered <- factor(dataset_clean$Q29,
+                            levels = c("(4) Not at all interested in politics",
+                                     "(3) Not that interested in politics",
+                                     "(2) Somewhat interested",
+                                     "(1) Very interested in politics"))
 
-ord_mod_interactions <- polr(Q29 ~ Q271_reordered * S3B +
+ord_mod_interactions <- polr(Q29_reordered ~ Q271_reordered * S3B +
                                race_recoded +
@@
 pvals_interact <- pnorm(-abs(summary(ord_mod_interactions)$coef[,"t value"]))*2
-ctable_interact <- cbind(summary(ord_mod_interactions)$coef,pvals)
+ctable_interact <- cbind(summary(ord_mod_interactions)$coef,pvals_interact)
 
@@
 # Generate the plot
-odds_ratio_plot(ord_mod)
+odds_ratio_plot(ord_mod_interactions)
+
+```
+
+# Fitting OLR Model with Interaction Term (Q271 * Q21 Party ID)
+
+```{r}
+
+library(MASS) # package for model
+library(caret) # package for confusion matrix
+
+# dataset_clean$Q271_reordered <- factor(dataset_clean$Q271,
+#                             levels = c("(5) Not at all important",
+#                                      "(4) Slightly important",
+#                                      "(3) Moderately important",
+#                                      "(2) Very important",
+#                                      "(1) Extremely important"))
+
+ord_mod_interactpoliticalID <- polr(Q29 ~ Q271_reordered * Q21 +
+                               race_recoded +
+                  age_recoded +
+            educ +  # Educational Attainment
+            S3B, # Gender
+          data = dataset_clean, Hess=TRUE)
+
+summary(ord_mod_interactpoliticalID)
+
+pvals_interactpoliticalID <- pnorm(-abs(summary(ord_mod_interactpoliticalID)$coef[,"t value"]))*2
+ctable_interactpoliticalID <- cbind(summary(ord_mod_interactpoliticalID)$coef,pvals_interactpoliticalID)
+
+ctable_interactpoliticalID
+
+(exp_coefs_interact <- exp(cbind(OR=coef(ord_mod_interactpoliticalID),confint(ord_mod_interactpoliticalID))))
+
+head(ord_mod_interactpoliticalID$fitted.values)
+head(predict(ord_mod_interactpoliticalID))
+
+confusionMatrix(predict(ord_mod_interactpoliticalID), dataset_clean$Q29, mode="everything")
 

### 2024-12-13 peter_models.qmd
updating interaction terms

@@
                                      "(2) Somewhat interested",
-                                     "(1) Very interested in politics"))
+                                     "(1) Very interested in politics"), ordered=TRUE)
 
@@
 
-ord_mod_interactpoliticalID <- polr(Q29 ~ Q271_reordered * Q21 +
+ord_mod_interactpoliticalID <- polr(Q29_reordered ~ Q271_reordered * Q21 +
                                race_recoded +

### 2024-12-13 peter_models.qmd
updating models, commenting out deprecated models

@@
 ```{r}
-library(MASS) # package for model
-library(caret) # package for confusion matrix
-
-dataset_clean$Q29_reordered  <- factor(dataset_clean$Q29,
-                            levels = c("(4) Not at all interested in politics",
-                                     "(3) Not that interested in politics",
-                                     "(2) Somewhat interested",
-                                     "(1) Very interested in politics"))
-
-dataset_clean$Q271_reordered <- factor(dataset_clean$Q271,
-                            levels = c("(5) Not at all important",
-                                     "(4) Slightly important",
-                                     "(3) Moderately important",
-                                     "(2) Very important",
-                                     "(1) Extremely important"))
-
-ord_mod <- polr(Q29 ~ Q271_reordered + race_recoded +
-                  age_recoded +
-            educ +  # Educational Attainment
-            S3B + # Gender
-            Q21, # Party affiliation
-          data = dataset_clean, Hess=TRUE)
-
-summary(ord_mod)
-
-pvals <- pnorm(-abs(summary(ord_mod)$coef[,"t value"]))*2
-ctable <- cbind(summary(ord_mod)$coef,pvals)
-
-ctable
-
-(exp_coefs <- exp(cbind(OR=coef(ord_mod),confint(ord_mod))))
-
-head(ord_mod$fitted.values)
-head(predict(ord_mod))
-
-confusionMatrix(predict(ord_mod), dataset_clean$Q29, mode="everything")
+# library(MASS) # package for model
+# library(caret) # package for confusion matrix
+# 
+# dataset_clean$Q29_reordered  <- factor(dataset_clean$Q29,
+#                             levels = c("(4) Not at all interested in politics",
+#                                      "(3) Not that interested in politics",
+#                                      "(2) Somewhat interested",
+#                                      "(1) Very interested in politics"))
+# 
+# 
+# dataset_clean$Q271_reordered <- factor(dataset_clean$Q271,
+#                             levels = c("(5) Not at all important",
+#                                      "(4) Slightly important",
+#                                      "(3) Moderately important",
+#                                      "(2) Very important",
+#                                      "(1) Extremely important"))
+# 
+# ord_mod <- polr(Q29 ~ Q271_reordered + race_recoded +
+#                   age_recoded +
+#             educ +  # Educational Attainment
+#             S3B + # Gender
+#             Q21, # Party affiliation
+#           data = dataset_clean, Hess=TRUE)
+# 
+# summary(ord_mod)
+# 
+# 
+# pvals <- pnorm(-abs(summary(ord_mod)$coef[,"t value"]))*2
+# ctable <- cbind(summary(ord_mod)$coef,pvals)
+# 
+# ctable
+# 
+# (exp_coefs <- exp(cbind(OR=coef(ord_mod),confint(ord_mod))))
+# 
+# head(ord_mod$fitted.values)
+# head(predict(ord_mod))
+# 
+# confusionMatrix(predict(ord_mod), dataset_clean$Q29, mode="everything")
 
@@
 
-library(MASS) # package for model
-library(caret) # package for confusion matrix
-
-# dataset_clean$Q271_reordered <- factor(dataset_clean$Q271,
-#                             levels = c("(5) Not at all important",
-#                                      "(4) Slightly important",
-#                                      "(3) Moderately important",
-#                                      "(2) Very important",
-#                                      "(1) Extremely important"))
-
-ord_mod_interactpoliticalID <- polr(Q29_reordered ~ Q271_reordered * Q21 +
-                               race_recoded +
-                  age_recoded +
-            educ +  # Educational Attainment
-            S3B, # Gender
-          data = dataset_clean, Hess=TRUE)
-
-summary(ord_mod_interactpoliticalID)
-
-pvals_interactpoliticalID <- pnorm(-abs(summary(ord_mod_interactpoliticalID)$coef[,"t value"]))*2
-ctable_interactpoliticalID <- cbind(summary(ord_mod_interactpoliticalID)$coef,pvals_interactpoliticalID)
-
-ctable_interactpoliticalID
-
-(exp_coefs_interact <- exp(cbind(OR=coef(ord_mod_interactpoliticalID),confint(ord_mod_interactpoliticalID))))
-
-head(ord_mod_interactpoliticalID$fitted.values)
-head(predict(ord_mod_interactpoliticalID))
-
-confusionMatrix(predict(ord_mod_interactpoliticalID), dataset_clean$Q29, mode="everything")
+# library(MASS) # package for model
+# library(caret) # package for confusion matrix
+# 
+# # dataset_clean$Q271_reordered <- factor(dataset_clean$Q271,
+# #                             levels = c("(5) Not at all important",
+# #                                      "(4) Slightly important",
+# #                                      "(3) Moderately important",
+# #                                      "(2) Very important",
+# #                                      "(1) Extremely important"))
+# 
+# ord_mod_interactpoliticalID <- polr(Q29_reordered ~ Q271_reordered * Q21 +
+#                                race_recoded +
+#                   age_recoded +
+#             educ +  # Educational Attainment
+#             S3B, # Gender
+#           data = dataset_clean, Hess=TRUE)
+# 
+# 
+# 
+# summary(ord_mod_interactpoliticalID)
+# 
+# 
+# pvals_interactpoliticalID <- pnorm(-abs(summary(ord_mod_interactpoliticalID)$coef[,"t value"]))*2
+# ctable_interactpoliticalID <- cbind(summary(ord_mod_interactpoliticalID)$coef,pvals_interactpoliticalID)
+# 
+# ctable_interactpoliticalID
+# 
+# (exp_coefs_interact <- exp(cbind(OR=coef(ord_mod_interactpoliticalID),confint(ord_mod_interactpoliticalID))))
+# 
+# head(ord_mod_interactpoliticalID$fitted.values)
+# head(predict(ord_mod_interactpoliticalID))
+# 
+# confusionMatrix(predict(ord_mod_interactpoliticalID), dataset_clean$Q29, mode="everything")
 
@@
 ```{r}
-
-# Load packages
-library(nnet)  # for multinomial logistic regression
-library(tidyverse)
-library(car)
-library(effects)
-
-# ensure outcome variable is a factor but not ordered
-dataset_clean$Q29_unordered <- factor(dataset_clean$Q29, ordered = FALSE)
-dataset_clean$Q271_simple <- as.numeric(dataset_clean$Q271)
-
-# Fit multinomial model
-m_mult <- multinom(Q29_unordered ~ Q271 * S3B + 
-                     race_recoded + 
-                     age_recoded +
-            educ + # Educational Attainment
-            Q21, # Party affiliation
-          data = dataset_clean)
-
-# Calculate predicted probabilities of marginal effects
-# Create sequence of Q271_simple values
-x_vals <- seq(min(dataset_clean$Q271_simple), max(dataset_clean$Q271_simple), length.out = 100)
-
-# Function to calculate average predicted probabilities
-get_avg_probs <- function(model, newdata, var_name) {
-    preds <- predict(model, newdata = newdata, type = "probs")
-    colMeans(preds)
-}
-
-# Calculate marginal effects by race
-races <- unique(dataset_clean$race_recoded)
-marg_effects <- list()
 # 
-for(race in races) {
-    probs <- matrix(0, nrow = length(x_vals), ncol = length(levels(dataset_clean$Q29_unordered)))
-    for(i in seq_along(x_vals)) {
-        temp_data <- dataset_clean
-        temp_data$Q271_simple <- x_vals[i]
-        temp_data$race_major <- race
-        probs[i,] <- get_avg_probs(m_mult, temp_data, "Q271_simple")
-    }
-    marg_effects[[race]] <- data.frame(
-        x = x_vals,
-        race = race,
-        prob1 = probs[,1],
-        prob2 = probs[,2],
-        prob3 = probs[,3],
-        prob4 = probs[,4]
-    )
-}
-
-# Combine all marginal effects
-marg_effects_df <- bind_rows(marg_effects)
-
-# Create plots for marginal effects
-# Reshape data for plotting
-marg_effects_long <- marg_effects_df %>%
-    pivot_longer(cols = starts_with("prob"),
-                names_to = "outcome",
-                values_to = "probability") %>%
-    mutate(outcome = factor(outcome,
-                          levels = paste0("prob", 1:4),
-                          labels = levels(dataset_clean$Q29_unordered)))
-
-# Plot marginal effects
-p1 <- ggplot(marg_effects_long,
-             aes(x = x, y = probability, color = outcome)) +
-    geom_line() +
-    facet_wrap(~race) +
-    theme_minimal() +
-    labs(title = "Marginal Effects of Racial Identity Importance by Race",
-         x = "Importance of Racial Identity",
-         y = "Predicted Probability",
-         color = "Political Interest Level") +
-    theme(legend.position = "bottom")
-print(p1)
-
-# Model fit statistics and comparisons
-print("Model summaries:")
-print(summary(m_mult))
-
-# Calculate McFadden's R-squared
-null_model <- multinom(Q29_unordered ~ 1, data = dataset_clean)
-mcfadden_r2 <- 1 - logLik(m_mult)/logLik(null_model)
-
-print("McFadden's R-squared:")
-print(mcfadden_r2)
-
-# Classification accuracy
-pred_class <- predict(m_mult)
-conf_matrix <- table(pred_class, dataset_clean$Q29_unordered)
-accuracy <- sum(diag(conf_matrix))/sum(conf_matrix)
-
-print("Classification accuracy:")
-print(accuracy)
-print("Confusion Matrix:")
-print(conf_matrix)
+# # Load packages
+# library(nnet)  # for multinomial logistic regression
+# library(tidyverse)
+# library(car)
+# library(effects)
 # 
-# Print AIC
-print(AIC(m_mult))
-
-# Calculate predicted probabilities
-pred_data <- expand.grid(
-    Q271 = unique(dataset_clean$Q271),
-    age_recoded = unique(dataset_clean$age_recoded),
-    race_recoded = unique(dataset_clean$race_recoded),
-    educ = unique(dataset_clean$educ),
-    S3B = unique(dataset_clean$S3B),
-    Q21 = unique(dataset_clean$Q21)
-)
 # 
-# Get predictions
-preds <- predict(m_mult, newdata = pred_data, type = "probs")
-pred_data <- cbind(pred_data, as.data.frame(preds))
-
-# Create separate plots for each outcome level
-pred_long <- pred_data %>%
-    pivot_longer(cols = starts_with("("),
-                names_to = "interest_level",
-                values_to = "probability")
-
-# Plot for each outcome level
-p2 <- ggplot(pred_long,
-             aes(x = Q271, y = probability, color = race_recoded)) +
-    geom_line() +
-    facet_wrap(~interest_level) +
-    theme_minimal() +
-    labs(title = "Predicted Probabilities by Outcome Level",
-         x = "Importance of Racial Identity",
-         y = "Probability",
-         color = "Race/Ethnicity") +
-    theme(legend.position = "bottom")
-print(p2)
+# # ensure outcome variable is a factor but not ordered
+# dataset_clean$Q29_unordered <- factor(dataset_clean$Q29, ordered = FALSE)
+# dataset_clean$Q271_simple <- as.numeric(dataset_clean$Q271)
+# 
+# # Fit multinomial model
+# m_mult <- multinom(Q29_unordered ~ Q271 * S3B + 
+#                      race_recoded + 
+#                      age_recoded +
+#             educ + # Educational Attainment
+#             Q21, # Party affiliation
+#           data = dataset_clean)
+# 
+# 
+# 
+# # Calculate predicted probabilities of marginal effects
+# # Create sequence of Q271_simple values
+# x_vals <- seq(min(dataset_clean$Q271_simple), max(dataset_clean$Q271_simple), length.out = 100)
+# 
+# # Function to calculate average predicted probabilities
+# get_avg_probs <- function(model, newdata, var_name) {
+#     preds <- predict(model, newdata = newdata, type = "probs")
+#     colMeans(preds)
+# }
+# 
+# # Calculate marginal effects by race
+# races <- unique(dataset_clean$race_recoded)
+# marg_effects <- list()
+# # 
+# for(race in races) {
+#     probs <- matrix(0, nrow = length(x_vals), ncol = length(levels(dataset_clean$Q29_unordered)))
+#     for(i in seq_along(x_vals)) {
+#         temp_data <- dataset_clean
+#         temp_data$Q271_simple <- x_vals[i]
+#         temp_data$race_major <- race
+#         probs[i,] <- get_avg_probs(m_mult, temp_data, "Q271_simple")
+#     }
+#     marg_effects[[race]] <- data.frame(
+#         x = x_vals,
+#         race = race,
+#         prob1 = probs[,1],
+#         prob2 = probs[,2],
+#         prob3 = probs[,3],
+#         prob4 = probs[,4]
+#     )
+# }
+# 
+# # Combine all marginal effects
+# marg_effects_df <- bind_rows(marg_effects)
+# 
+# # Create plots for marginal effects
+# # Reshape data for plotting
+# marg_effects_long <- marg_effects_df %>%
+#     pivot_longer(cols = starts_with("prob"),
+#                 names_to = "outcome",
+#                 values_to = "probability") %>%
+#     mutate(outcome = factor(outcome,
+#                           levels = paste0("prob", 1:4),
+#                           labels = levels(dataset_clean$Q29_unordered)))
+# 
+# # Plot marginal effects
+# p1 <- ggplot(marg_effects_long,
+#              aes(x = x, y = probability, color = outcome)) +
+#     geom_line() +
+#     facet_wrap(~race) +
+#     theme_minimal() +
+#     labs(title = "Marginal Effects of Racial Identity Importance by Race",
+#          x = "Importance of Racial Identity",
+#          y = "Predicted Probability",
+#          color = "Political Interest Level") +
+#     theme(legend.position = "bottom")
+# print(p1)
+# 
+# # Model fit statistics and comparisons
+# print("Model summaries:")
+# print(summary(m_mult))
+# 
+# # Calculate McFadden's R-squared
+# null_model <- multinom(Q29_unordered ~ 1, data = dataset_clean)
+# mcfadden_r2 <- 1 - logLik(m_mult)/logLik(null_model)
+# 
+# print("McFadden's R-squared:")
+# print(mcfadden_r2)
+# 
+# # Classification accuracy
+# pred_class <- predict(m_mult)
+# conf_matrix <- table(pred_class, dataset_clean$Q29_unordered)
+# accuracy <- sum(diag(conf_matrix))/sum(conf_matrix)
+# 
+# print("Classification accuracy:")
+# print(accuracy)
+# print("Confusion Matrix:")
+# print(conf_matrix)
+# # 
+# # Print AIC
+# print(AIC(m_mult))
+# 
+# # Calculate predicted probabilities
+# pred_data <- expand.grid(
+#     Q271 = unique(dataset_clean$Q271),
+#     age_recoded = unique(dataset_clean$age_recoded),
+#     race_recoded = unique(dataset_clean$race_recoded),
+#     educ = unique(dataset_clean$educ),
+#     S3B = unique(dataset_clean$S3B),
+#     Q21 = unique(dataset_clean$Q21)
+# )
+# # 
+# # Get predictions
+# preds <- predict(m_mult, newdata = pred_data, type = "probs")
+# pred_data <- cbind(pred_data, as.data.frame(preds))
+# 
+# # Create separate plots for each outcome level
+# pred_long <- pred_data %>%
+#     pivot_longer(cols = starts_with("("),
+#                 names_to = "interest_level",
+#                 values_to = "probability")
+# 
+# 
+# # Plot for each outcome level
+# p2 <- ggplot(pred_long,
+#              aes(x = Q271, y = probability, color = race_recoded)) +
+#     geom_line() +
+#     facet_wrap(~interest_level) +
+#     theme_minimal() +
+#     labs(title = "Predicted Probabilities by Outcome Level",
+#          x = "Importance of Racial Identity",
+#          y = "Probability",
+#          color = "Race/Ethnicity") +
+#     theme(legend.position = "bottom")
+# print(p2)
 
@@
 # generate predictions for the ordinal model and the ordinal model  with interactions
-predict(ord_mod, new_dataset, type="probs")
+# predict(ord_mod, new_dataset, type="probs")
 predict(ord_mod_interactions, new_dataset, type="probs")
 
-new_dataset <- data.frame(race_recoded=c("White", "Black", "Asian American", "Latino", "AIAN", "NHPI", "Arab/MENA"),
-                          Q271=c("(5) Not at all important", "(4) Slightly important", "(3) Moderately important", "(2) Very important", "(1) Extremely important", "(4) Slightly important", "(5) Not at all important"),
-                          age_recoded = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+", "18-29"),
-                          S3B = c("(1) Man", "(2) Woman","(3) Non-binary","(4) Something else (Specify)", "(1) Man", "(2) Woman", "(1) Man"),
-                          educ = c("No college experience", "Some college experience", "College graduate", "No college experience", "Some college experience", "College graduate", "College graduate"),
-                          Q21 = c("(1) Republican",  "(2) Democrat","(3) Independent", "(4) Other party", "(1) Republican",  "(2) Democrat",  "(3) Independent"))
+# new_dataset <- data.frame(race_recoded=c("White", "Black", "Asian American", "Latino", "AIAN", "NHPI", "Arab/MENA"),
+#                           Q271=c("(5) Not at all important", "(4) Slightly important", "(3) Moderately important", "(2) Very important", "(1) Extremely important", "(4) Slightly important", "(5) Not at all important"),
+#                           age_recoded = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+", "18-29"),
+#                           S3B = c("(1) Man", "(2) Woman","(3) Non-binary","(4) Something else (Specify)", "(1) Man", "(2) Woman", "(1) Man"),
+#                           educ = c("No college experience", "Some college experience", "College graduate", "No college experience", "Some college experience", "College graduate", "College graduate"),
+#                           Q21 = c("(1) Republican",  "(2) Democrat","(3) Independent", "(4) Other party", "(1) Republican",  "(2) Democrat",  "(3) Independent"))
 
-predict(m_mult, new_dataset, type = "probs")
+# predict(m_mult, new_dataset, type = "probs")
 ```
@@
 # Confusion matrix for multinomial model
-conf_matrix
+
 ```

### 2024-12-14 peter_models.qmd
updating plot to label y aaxis variables

@@
 
+# Testing plotting function
+```{r}
+or_data_test <- exp(cbind(OR=coef(ord_mod_interactions),confint(ord_mod_interactions)))
+
+  # Calculate p-values
+p_values_test <- pnorm(-abs(summary(ord_mod_interactions)$coefficients[,"t value"]))*2
+
+  # Keep only significant variables
+sig_indices_test <- p_values_test[names(coef(ord_mod_interactions))] < 0.05
+or_data_sig_test <- or_data_test[sig_indices_test, ]
+
+rename_map <- c(
+    "Q271_reordered(2) Very important" = "Racial Belonging - Very Important",
+    "Q271_reordered(1) Extremely important" = "Racial Belonging - Extremely Important",
+    "S3B(2) Woman" = "Gender - Woman",
+    "S3B(3) Non-binary" = "Gender - Non-binary",
+    "race_recodedBlack" = "Race - Black",
+    "race_recodedAsian American" = "Race - Asian",
+    "race_recodedLatino" = "Race - Latino",
+    "race_recodedAIAN" = "Race - Native",
+    "race_recodedNHPI" = "Race - NHPI",
+    "age_recoded50-59" = "Age - 50-59",
+    "age_recoded60-69" = "Age - 60-69",
+    "age_recoded70+" = "Age - 70+",
+    "educSome college experience" = "Education - Some College",
+    "educCollege graduate" = "Education - College Graduate",
+    "Q21(2) Democrat" = "Party ID - Democrat",
+    "Q21(3) Independent" = "Party ID - Independent",
+    "Q21(4) Other party" = "Party ID - Other",
+    "Q271_reordered(1) Extremely important:S3B(2) Woman" = "Interaction: Racial Belonging & Gender: Woman"
+  )
+
+rownames(or_data_sig_test) <- rename_map[rownames(or_data_sig_test)]
+
+or_data_sig_test
+
+```
+
+# Trying to run the plot again but with explicit y-axis Variable labels
+
+```{r}
+odds_ratio_plot_labelled <- function(model) {
+  # Get odds ratios and CIs
+  or_data <- exp(cbind(OR=coef(model),confint(model)))
+
+  # Calculate p-values
+  p_values <- pnorm(-abs(summary(model)$coefficients[,"t value"]))*2
+
+  # Keep only significant variables
+  sig_indices <- p_values[names(coef(model))] < 0.05
+  or_data_sig <- or_data[sig_indices, ]
+
+  rename_map <- c(
+    "Q271_reordered(2) Very important" = "Racial Belonging - Very Important",
+    "Q271_reordered(1) Extremely important" = "Racial Belonging - Extremely Important",
+    "S3B(2) Woman" = "Gender - Woman",
+    "S3B(3) Non-binary" = "Gender - Non-binary",
+    "race_recodedBlack" = "Race - Black",
+    "race_recodedAsian American" = "Race - Asian",
+    "race_recodedLatino" = "Race - Latino",
+    "race_recodedAIAN" = "Race - Native",
+    "race_recodedNHPI" = "Race - NHPI",
+    "age_recoded50-59" = "Age - 50-59",
+    "age_recoded60-69" = "Age - 60-69",
+    "age_recoded70+" = "Age - 70+",
+    "educSome college experience" = "Education - Some College",
+    "educCollege graduate" = "Education - College Graduate",
+    "Q21(2) Democrat" = "Party ID - Democrat",
+    "Q21(3) Independent" = "Party ID - Independent",
+    "Q21(4) Other party" = "Party ID - Other",
+    "Q271_reordered(1) Extremely important:S3B(2) Woman" = "Interaction: Racial Belonging & Gender: Woman"
+  )
+  or_data_sig_renamed <- or_data_sig
+  rownames(or_data_sig_renamed) <- rename_map[rownames(or_data_sig_renamed)]
+
+  ggplot() +
+    # Add vertical line at OR = 1
+    geom_vline(xintercept = 1, linetype = "dashed", color = "red", size = 0.7) +
+    # Add points and error bars
+    geom_point(aes(x = or_data_sig_renamed[,1], 
+                   y = reorder(rownames(or_data_sig_renamed), or_data_sig_renamed[,1])), 
+              size = 4, color = "blue") +
+    geom_errorbarh(aes(xmin = or_data_sig[,2], 
+                       xmax = or_data_sig[,3],
+                       y = reorder(rownames(or_data_sig_renamed), or_data_sig_renamed[,1])),
+                   height = 0.3, color = "black") +
+    # Customize theme
+    theme_minimal() +
+    theme(
+      axis.text.y = element_text(size = 6, color = "black"),
+      axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
+      axis.title.x = element_text(size = 12, face = "bold"),
+      axis.title.y = element_text(size = 12, face = "bold"),
+      plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
+      plot.margin = margin(1, 1, 1, 1, "cm"),
+      legend.text = element_text(size = 8),
+      legend.title = element_text(size = 9)
+    ) +
+    # Add labels
+    labs(
+      title = "Significant Variables with Confidence Intervals",
+      x = "Odds Ratio",
+      y = "Variables"
+    )
+
+}
+
+# Generate the plot
+odds_ratio_plot_labelled(ord_mod_interactions)
+
+```
+
 # Fitting OLR Model with Interaction Term (Q271 * Q21 Party ID)

### 2024-12-14 peter_models.qmd
updating model

@@
 # Fitting OLR with interaction terms
+
 ## This is the model we include in the final report.
@@
 |------------------|------------------|------------------|------------------|
-| Q271 (4) Slightly important | 1.0338781 | 0.86761871 | 1.2317529 |
-| Q271 (3) Moderately important | 1.0228231 | 0.88466618 | 1.1826555 |
-| Q271 (2) Very important | 0.8366761 | 0.72314779 | 0.9680337 |
-| Q271 (1) Extremely important | 0.6069588 | 0.52449164 | 0.7024084 |
-| S3B(2) Woman | 1.5887610 | 1.36312914 | 1.8518245 |
-| S3B(3) Non-binary | 0.5380781 | 0.28888679 | 0.9978406 |
-| S3B(4) Something else (Specify) | 0.4648621 | 0.07401426 | 2.6422344 |
-| race Black | 1.1525119 | 1.04988995 | 1.2652177 |
-| race Asian American | 1.6819184 | 1.53325820 | 1.8451528 |
-| race Latino | 1.2583888 | 1.14546256 | 1.3825307 |
-| race AIAN | 1.4882734 | 1.27404185 | 1.7384045 |
-| race NHPI | 1.7838764 | 1.45756175 | 2.1825886 |
-| race Arab/MENA | 0.8628526 | 0.66889425 | 1.1122887 |
-| age 30-39 | 0.9698580 | 0.89048308 | 1.0563119 |
-| age 40-49 | 1.0813779 | 0.98549367 | 1.1865407 |
-| age 50-59 | 0.9107924 | 0.82952560 | 0.9996833 |
-| age 60-69 | 0.7553567 | 0.68762045 | 0.8297158 |
-| age 70+ | 0.5918432 | 0.52889988 | 0.6621363 |
-| educSome college experience | 0.6551005 | 0.60189454 | 0.7129797 |
-| educCollege graduate | 0.4476726 | 0.41528212 | 0.4825867 |
-| Q21(2) Democrat | 0.8823547 | 0.81416780 | 0.9562372 |
-| Q21(3) Independent | 1.7222132 | 1.58330889 | 1.8733830 |
-| Q21(4) Other party | 4.7003458 | 4.03685844 | 5.4748427 |
-| Q271 4) Slightly important:S3B(2) Woman | 1.0044380 | 0.78966197 | 1.2778491 |
-| Q271 (3) Moderately important:S3B(2) Woman | 1.0569625 | 0.87144637 | 1.2819704 |
-| Q271 (2) Very important:S3B(2) Woman | 1.0212671 | 0.84557349 | 1.2335590 |
-| Q271 (1) Extremely important:S3B(2) Woman | 1.2324558 | 1.02421823 | 1.4830676 |
-| Q271 (4) Slightly important:S3B(3) Non-binary | 1.5393710 | 0.46529911 | 5.0344630 |
-| Q271 (3) Moderately important:S3B(3) Non-binary | 1.5246008 | 0.57882727 | 4.0084234 |
-| Q271 (2) Very important:S3B(3) Non-binary | 1.9464262 | 0.77448045 | 4.8746428 |
-| Q271 (1) Extremely important:S3B(3) Non-binary | 1.1408623 | 0.45797295 | 2.8216144 |
-| Q271 (4) Slightly important:S3B(4) Something else (Specify) | 4.4125437 | 0.49596064 | 41.9119559 |
-| Q271 (3) Moderately important:S3B(4) Something else (Specify) | 0.5879278 | 0.05055911 | 4.9095993 |
-| Q271 2) Very important:S3B(4) Something else (Specify) | 9.6145175 | 0.70357378 | 160.9916657 |
-| Q271 1) Extremely important:S3B(4) Something else (Specify) | 1.5327973 | 0.20059858 | 12.5345043 |
+| Q271 (4) Slightly important | 0.9672589 | 0.811845068 | 1.1525850 |
+| Q271 (3) Moderately important | 0.9776608 | 0.845556718 | 1.1303745 |
+| Q271 (2) Very important | 1.1952004 | 1.033019394 | 1.3828451 |
+| Q271 (1) Extremely important | 1.6474969 | 1.423670208 | 1.9066091 |
+| S3B(2) Woman | 0.6294256 | 0.540015144 | 0.7336072 |
+| S3B(3) Non-binary | 1.8576658 | 1.002509318 | 3.4617409 |
+| S3B(4) Something else (Specify) | 2.1359493 | 0.377740111 | 13.5124525 |
+| race Black | 0.8676620 | 0.790376010 | 0.9524806 |
+| race Asian American | 0.5945609 | 0.541964870 | 0.6522143 |
+| race Latino | 0.7946622 | 0.575235415 | 0.7848896 |
+| race AIAN | 0.6719204 | 1.27404185 | 1.7384045 |
+| race NHPI | 0.5606006 | 0.458205702 | 0.6860792 |
+| race Arab/MENA | 1.1589043 | 0.899447122 | 1.4956228 |
+| age 30-39 | 1.0310801 | 0.946696738 | 1.0563119 |
+| age 40-49 | 0.9247623 | 0.98549367 | 1.1229868 |
+| age 50-59 | 1.0979487 | 1.000027913 | 1.2054892 |
+| age 60-69 | 1.3238874 | 1.205240800 | 1.4542874 |
+| age 70+ | 1.6896464 | 1.510226543 | 1.8907050 |
+| educSome college experience | 1.5264694 | 1.402573135 | 1.6614197 |
+| educCollege graduate | 2.2337286 | 2.072162224 | 2.4081069 |
+| Q21(2) Democrat | 1.1333513 | 1.045765624 | 1.2282275 |
+| Q21(3) Independent | 0.5806653 | 0.533801211 | 0.6315870 |
+| Q21(4) Other party | 0.2127592 | 0.182654104 | 0.2477198 |
+| Q271 4) Slightly important:S3B(2) Woman | 0.9955364 | 0.782564309 | 1.2663323 |
+| Q271 (3) Moderately important:S3B(2) Woman | 0.9461281 | 0.780040309 | 1.1475378 |
+| Q271 (2) Very important:S3B(2) Woman | 0.9791675 | 0.810674379 | 1.1826356 |
+| Q271 (1) Extremely important:S3B(2) Woman | 0.8114102 | 0.674275440 | 0.9763646 |
+| Q271 (4) Slightly important:S3B(3) Non-binary | 0.6505015 | 0.198565161 | 2.1505278 |
+| Q271 (3) Moderately important:S3B(3) Non-binary | 0.6563254 | 0.249594245 | 1.7277048 |
+| Q271 (2) Very important:S3B(3) Non-binary | 0.5140828 | 0.205132143 | 1.2905898 |
+| Q271 (1) Extremely important:S3B(3) Non-binary | 0.8770903 | 0.354410521 | 2.1835545 |
+| Q271 (4) Slightly important:S3B(4) Something else (Specify) | 0.2279084 | 0.023857021 | 2.0164000 |
+| Q271 (3) Moderately important:S3B(4) Something else (Specify) | 1.7158304 | 0.146865519 | 19.7516328 |
+| Q271 2) Very important:S3B(4) Something else (Specify) | 0.1049752 | 0.006213617 | 1.4210951 |
+| Q271 1) Extremely important:S3B(4) Something else (Specify) | 0.6584933 | 0.079780177 | 4.9850479 |
 
@@
 # Testing plotting function
+
 ```{r}
@@
 
-# Fitting OLR Model with Interaction Term (Q271 * Q21 Party ID)
+# Fitting OLR Model with Interaction Term (Q271 \* Q21 Party ID)
 

### 2024-12-14 RQ1_Plot.qmd
add OR plot code

--- 2024-12-14 peter_models.qmd
@@
 ```{r}
-load("/Users/pdeguz01/Documents/git/Data/2020 CMPS/Output/2020cmps_clean.RData")
+load("/Users/jenniferli/Desktop/2020cmps_clean2.RData")
+```
+
+```{r}
+```
+
+```{r}
+summary(dataset_clean)
 ```
@@
     "race_recodedLatino" = "Race - Latino",
-    "race_recodedAIAN" = "Race - Native",
+    "race_recodedAIAN" = "Race - Native", # why use Native instead of AIAN?
     "race_recodedNHPI" = "Race - NHPI",
@@
 ```
+
+# Plot with Andrea's following comments all considered:
+
+1.  make sure to use appropriate variable labels.
+2.
+I see that you ordered them based on the OR values, but I think it’s confusing to see the responses to the questions in different places.
+I would still group the questions together, and maybe put them in descending order within each question?
+3.  Also make sure the reference levels are clear so that the reader can appropriately interpret the plot
+
+```{r}
+odds_ratio_plot_labelled <- function(model) {
+  # Get odds ratios and CIs
+  or_data <- exp(cbind(OR=coef(model), confint(model)))
+
+  # Calculate p-values
+  p_values <- pnorm(-abs(summary(model)$coefficients[,"t value"]))*2
+
+  # Keep only significant variables
+  sig_indices <- p_values[names(coef(model))] < 0.05
+  or_data_sig <- or_data[sig_indices, ]
+
+  # Create a mapping for variable names with grouping prefixes
+  rename_map <- c(
+    # Education Group
+    "educCollege graduate" = "1. Education: College Graduate",
+    "educSome college experience" = "1. Education: Some College",
+
+    # Gender Group
+    "S3B(2) Woman" = "2. Gender: Woman",
+    "S3B(3) Non-binary" = "2. Gender: Non-binary",
+
+    # Age Group
+    "age_recoded70+" = "3. Age: 70+",
+    "age_recoded60-69" = "3. Age: 60-69",
+    "age_recoded50-59" = "3. Age: 50-59",
+
+    # Race Group
+    "race_recodedBlack" = "4. Race: Black",
+    "race_recodedAsian American" = "4. Race: Asian",
+    "race_recodedLatino" = "4. Race: Latino",
+    "race_recodedAIAN" = "4. Race: Native",
+    "race_recodedNHPI" = "4. Race: NHPI",
+
+    # Racial Belonging Group
+    "Q271_reordered(2) Very important" = "5. Racial Belonging: Very Important",
+    "Q271_reordered(1) Extremely important" = "5. Racial Belonging: Extremely Important",
+
+    # Party ID Group
+    "Q21(2) Democrat" = "6. Party ID: Democrat",
+    "Q21(3) Independent" = "6. Party ID: Independent",
+    "Q21(4) Other party" = "6. Party ID: Other",
+
+    # Interactions
+    "Q271_reordered(1) Extremely important:S3B(2) Woman" = "7. Interaction: Racial Belonging & Gender: Woman"
+  )
+
+  or_data_sig_renamed <- or_data_sig
+  rownames(or_data_sig_renamed) <- rename_map[rownames(or_data_sig_renamed)]
+
+  # Custom sorting function
+  sort_by_group_and_name <- function(x) {
+    # Split the strings to get group numbers and names
+    groups <- as.numeric(sub("^(\\d+)\\..+$", "\\1", x))
+    # Get the part after the colon for sorting within groups
+    names <- sub("^\\d+\\.[^:]+: (.+) \\(ref.+$", "\\1", x)
+    # Create a data frame for sorting
+    sort_df <- data.frame(
+      original = x,
+      group = groups,
+      name = names,
+      stringsAsFactors = FALSE
+    )
+    # Sort by group (ascending) and then by name (descending)
+    sort_df <- sort_df[order(sort_df$group, -xtfrm(sort_df$name)), ]
+    return(sort_df$original)
+  }
+
+  # Apply the custom sorting
+  sorted_names <- sort_by_group_and_name(rownames(or_data_sig_renamed))
+  or_data_sig_renamed <- or_data_sig_renamed[sorted_names, ]
+
+  ggplot() +
+    geom_vline(xintercept = 1, linetype = "dashed", color = "red", size = 0.7) +
+    geom_point(aes(x = or_data_sig_renamed[,1], 
+                   y = factor(rownames(or_data_sig_renamed), 
+                          levels = rev(rownames(or_data_sig_renamed)))), 
+              size = 4, color = "blue") +
+    geom_errorbarh(aes(xmin = or_data_sig[,2], 
+                       xmax = or_data_sig[,3],
+                       y = factor(rownames(or_data_sig_renamed), 
+                              levels = rev(rownames(or_data_sig_renamed)))),
+                   height = 0.3, color = "black") +
+    theme_minimal() +
+    theme(
+      axis.text.y = element_text(size = 6, color = "black"),
+      axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
+      axis.title.x = element_text(size = 12, face = "bold"),
+      axis.title.y = element_text(size = 12, face = "bold"),
+      plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
+      plot.margin = margin(1, 1, 1, 1, "cm"),
+      panel.grid.major.y = element_line(color = "gray90"),
+      panel.grid.minor.y = element_blank()
+    ) +
+    labs(
+      title = "Significant Variables with Confidence Intervals",
+      x = "Odds Ratio",
+      y = "Variables"
+    )
+}
+
+# Generate the plot
+odds_ratio_plot_labelled(ord_mod_interactions)
+
+```
+
+Note: Some variable levels are excluded due to filtering for significant variables. To view the reference level for each variable group, please see the below table.
+
+## Reference Levels Table
+
+| Variable Group | Reference Level |
+|----|----|
+| 1\. Education | No college experience |
+| 2\. Gender | Man |
+| 3\. Age | 18-29 years old |
+| 4\. Race | White |
+| 5\. Racial Belonging | Not At All Important |
+| 6\. Party ID | Republican |
+| 7\. Interaction: Racial Belonging & Gender: Woman | Man, Not At All Important |
 

### 2024-12-14 RQ2_final.qmd
Look at RQ2

---
title: "RQ2_final"
format: html
editor: visual

execute: 
  warning: false
---

### Loading Libraries

```{r echo = FALSE}
# Load necessary libraries
library(dplyr)   # Provides %>% and mutate
library(tidyverse)  # Optionally load the entire tidyverse if needed
library(pROC)
library(vcd)
library(car)
library(knitr)
library(gt)
library(caret)
```

```{r echo = FALSE, results = FALSE}
# Read the dataset
file_path <- "39096-0001-Data.rda"

load(file_path)          # Load the RDA file
ls()                     # List objects loaded into the environment
dataset <- da39096.0001  # Access the dataset

head(dataset)
ncol(dataset)
nrow(dataset)
```

## Pre-processing Data

```{r echo = FALSE, results=FALSE}
# Define the required columns
required_cols <- c(
  # Outcome Variable: Stopping discrimination against racial/ethnic minorities
  "Q1R5",

  # Demographic Variables
  "S2_RACE_PRIME",   # Primary Race
  "S3",              # SexualOrientation
  "S3B",             # Gender
  "S5_AGE",          # Age categories
  "S13",             # Highest level of education
  "S14",             # Community lived in

  # Variables capturing experiences with racism
  "Q629R1",          # Unfair treatment due to racial background/ethnicity
  "Q629R2",          # Unfair treatment due to skin color
  "Q629R3",          # Unfair treatment due to gender
  "Q629R4",          # Unfair treatment due to sexuality/orientation
  "Q629R5",          # Unfair treatment due to immigration status
  "Q629R6",          # Unfair treatment due to religion
  "Q629R7",          # Unfair treatment due to accent

  "Q627",            # Treated unfairly or discriminated against because of your race, ethnicity, gender, sexuality, being an immigrant, religious heritage or having an accent?
  "Q633",            # Discrimination impact on life satisfaction

  # Perceptions of discrimination
  "Q619_Q626R1",     # Perceptions of discrimination against Whites
  "Q619_Q626R2",     # Perceptions of discrimination against Blacks
  "Q619_Q626R3",     # Perceptions of discrimination against Asian Americans
  "Q619_Q626R4",     # Perceptions of discrimination against Native Americans
  "Q619_Q626R5",     # Perceptions of discrimination against Immigrants
  "Q619_Q626R6",     # Perceptions of discrimination against Latinos
  "Q619_Q626R7",     # Perceptions of discrimination against GaysAndLesbians
  "Q619_Q626R8"      # Perceptions of discrimination against Muslims
)
```

```{r echo=FALSE, results = FALSE}
# Create a new dataframe with selected variables
dataset_new <- dataset[, required_cols]

# Recode the outcome variable 'Q1R5' for readability
dataset_new$Q1R5 <- factor(
  ifelse(
    dataset_new$Q1R5 == "(0) NO TO: Stopping discrimination against racial/ethnic minorities", 
    "Not important",
    "Important"  # Remaining cases are "Important"
  ),
  levels = c("Not important", "Important") # Define factor levels
)

# Define the columns for perceptions of discrimination
perception_cols <- c(
  "Q619_Q626R1",  # Perceptions of discrimination against Whites
  "Q619_Q626R2",  # Perceptions of discrimination against Blacks
  "Q619_Q626R3",  # Perceptions of discrimination against Asian Americans
  "Q619_Q626R4",  # Perceptions of discrimination against Native Americans
  "Q619_Q626R5",  # Perceptions of discrimination against Immigrants
  "Q619_Q626R6",  # Perceptions of discrimination against Latinos
  "Q619_Q626R7",  # Perceptions of discrimination against GaysAndLesbians
  "Q619_Q626R8",  # Perceptions of discrimination against Muslims
  "Q633"          # Discrimination impact on life satisfaction
)

# Recode 'Don' as 'Don't know' and ensure columns remain factors
dataset_new[perception_cols] <- lapply(dataset_new[perception_cols], function(x) {
  x <- as.character(x)  # Temporarily convert to character for recoding
  x <- ifelse(x == "(5) Don", "(5) Don't know", x)  # Recode 'Don' to 'Don't know'
  factor(x)  # Convert back to factor
})

# Define required columns to convert to factors (if not already defined)
# Assuming `required_cols` is previously defined; otherwise, specify explicitly
dataset_new[required_cols] <- lapply(dataset_new[required_cols], as.factor)

# Verify the structure of the dataset
str(dataset_new[required_cols])

# Display the first few rows to check the changes
head(dataset_new)
```

```{r echo=FALSE, results=FALSE}
# Data Cleaning and Recoding for dataset_new
dataset_clean <- dataset_new |> 
  mutate(
    # Education recoding
    educ = case_when(
      S13 == "(1) Grades 1-8" ~ "No college experience",
      S13 == "(2) Some High School, but did not graduate" ~ "No college experience",
      S13 == "(3) High School graduate or GED" ~ "No college experience",
      S13 == "(4) Some college" ~ "Some college experience",
      S13 == "(5) Associates, 2-year degree" ~ "College graduate",
      S13 == "(6) Bachelors, 4-year degree" ~ "College graduate",
      S13 == "(7) Post-graduate degree" ~ "College graduate"
    ),
    # Sexual orientation recoding
    sex_orientation = case_when(
      S3 == "(01) Straight, that is, not gay or lesbian" ~ "Straight",
      S3 == "(02) Gay or lesbian" ~ "Queer",
      S3 == "(03) Bisexual" ~ "Queer",
      S3 == "(04) Something else (Specify)" ~ "Queer",
      S3 == "(88) Refused" ~ "Refused/Don't Know",
      S3 == "(99) I don" ~ "Refused/Don't Know"
    ),
    # Urban vs rural recoding
    rural = case_when(
      S14 == "(1) Large urban area" ~ "Urban/suburban",
      S14 == "(2) Large suburb near large city" ~ "Urban/suburban",
      S14 == "(3) Small suburb near small town or city" ~ "Urban/suburban",
      S14 == "(4) Small town or small city" ~ "Urban/suburban",
      S14 == "(5) Rural area" ~ "Rural"
    ),
    # Race recoding
    race_recoded = case_when(
      S2_RACE_PRIME == "(1) White" ~ "White",
      S2_RACE_PRIME == "(2) Hispanic or Latino" ~ "Latino",
      S2_RACE_PRIME == "(3) Black or African American" ~ "Black",
      S2_RACE_PRIME == "(4) Asian American" ~ "Asian American",
      S2_RACE_PRIME == "(5) American Indian/Native American" ~ "AIAN",
      S2_RACE_PRIME == "(6) Arab, Middle Eastern or North African" ~ "Arab/MENA",
      S2_RACE_PRIME == "(7) Native Hawaiian" ~ "NHPI",
      S2_RACE_PRIME == "(8) Not Hawaiian, but other Pacific Islander" ~ "NHPI"
    ),
    # Age recoding
    age_recoded = case_when(
      S5_AGE == "(1) 16-17" ~ "16-17",
      S5_AGE == "(2) 18-29" ~ "18-29",
      S5_AGE == "(3) 30-39" ~ "30-39",
      S5_AGE == "(4) 40-49" ~ "40-49",
      S5_AGE == "(5) 50-59" ~ "50-59",
      S5_AGE == "(6) 60-69" ~ "60-69",
      S5_AGE == "(7) 70 +" ~ "70+"
    )
  ) |>
  mutate(
    # Converting variables to factors
    race_recoded = factor(race_recoded, levels = c("White", "Black", "Asian American", "Latino", "AIAN", "NHPI", "Arab/MENA")),
    age_recoded = factor(age_recoded, levels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+")),
    educ = factor(educ, levels = c("No college experience", "Some college experience", "College graduate")),
    rural = factor(rural, levels = c("Rural", "Urban/suburban")),
    sex_orientation = factor(sex_orientation, levels = c("Straight", "Queer", "Refused/Don't Know"))
  )

# Take a look at the updated dataset
head(dataset_clean)
```

```{r echo=FALSE, results=FALSE}
# Remove all redundant columns
redundant_columns <- c("Q629R1", "Q629R2", "Q629R3", "Q629R4", 
                       "Q629R5", "Q629R6", "Q629R7",
                       "S2_RACE_PRIME", "S13", "S3", "S5_AGE","S14")

dataset_clean <- dataset_clean[, !(names(dataset_clean) %in% redundant_columns)]

```

```{r echo=FALSE, results=FALSE}
# Rename remaining columns in dataset_clean
colnames(dataset_clean) <- c(
  "ImportanceOfStoppingDiscriminationAgainstMinorities",  # Q1R5
  "Gender",                                               # S3B
  "TreatedUnfairlyOrDiscriminatedAgainst",                # Q627
  "DiscriminationImpactOnLifeSatisfaction",              # Q633
  "PerceptionsOfDiscriminationAgainstWhites",            # Q619_Q626R1
  "PerceptionsOfDiscriminationAgainstBlacks",            # Q619_Q626R2
  "PerceptionsOfDiscriminationAgainstAsianAmericans",    # Q619_Q626R3
  "PerceptionsOfDiscriminationAgainstNativeAmericans",   # Q619_Q626R4
  "PerceptionsOfDiscriminationAgainstImmigrants",        # Q619_Q626R5
  "PerceptionsOfDiscriminationAgainstLatinos",           # Q619_Q626R6
  "PerceptionsOfDiscriminationAgainstGaysAndLesbians",   # Q619_Q626R7
  "PerceptionsOfDiscriminationAgainstMuslims",           # Q619_Q626R8
  "Education",                                           # educ
  "SexOrientation",                                      # sex_orientation
  "Rural",                                               # rural
  "Race",                                                # race_recoded
  "Age"                                                  # age_recoded
)

```

# Model Building

## Model 1: Main

```{r echo=FALSE, results=FALSE}
# Regression model
base_model <- glm(ImportanceOfStoppingDiscriminationAgainstMinorities ~ Gender +
               TreatedUnfairlyOrDiscriminatedAgainst + 
               DiscriminationImpactOnLifeSatisfaction +
               PerceptionsOfDiscriminationAgainstWhites + 
               PerceptionsOfDiscriminationAgainstBlacks +
               PerceptionsOfDiscriminationAgainstAsianAmericans + 
               PerceptionsOfDiscriminationAgainstNativeAmericans + 
               PerceptionsOfDiscriminationAgainstImmigrants + 
               PerceptionsOfDiscriminationAgainstLatinos + 
               PerceptionsOfDiscriminationAgainstGaysAndLesbians + 
               PerceptionsOfDiscriminationAgainstMuslims +
               Education + 
                 SexOrientation +
                 Rural +
                 Race +
                 Age,

             data = dataset_clean, family = "binomial")
```

## Model 2: Model with interaction term

```{r echo=FALSE, results=FALSE}
interaction_model <- glm(ImportanceOfStoppingDiscriminationAgainstMinorities ~ 
               TreatedUnfairlyOrDiscriminatedAgainst + 
               DiscriminationImpactOnLifeSatisfaction +
               PerceptionsOfDiscriminationAgainstWhites + 
               PerceptionsOfDiscriminationAgainstBlacks +
               PerceptionsOfDiscriminationAgainstAsianAmericans + 
               PerceptionsOfDiscriminationAgainstNativeAmericans + 
               PerceptionsOfDiscriminationAgainstImmigrants + 
               PerceptionsOfDiscriminationAgainstLatinos + 
               PerceptionsOfDiscriminationAgainstGaysAndLesbians + 
               PerceptionsOfDiscriminationAgainstMuslims +
               Education + 
                 SexOrientation +
                 Rural +
                 Age + 
                 Gender * Race,

             data = dataset_clean, family = "binomial")
```

#VIF Values

```{r}
# Check multicollinearity using VIF
vif_base <- vif(base_model)
vif_interaction <- vif(interaction_model)
```

|   | Normalized VIF: Base Model | Normalized VIF: Interaction Model |
|------------------------|------------------------|------------------------|
| Gender | 1.022 | 1.978 |
| Treated Unfairly or Discriminated Against | 1.125 | 1.126 |
| Discrimination Impact on Life Satisfaction | 1.036 | 1.037 |
| Perceptions Of Discrimination Against Whites | 1.105 | 1.106 |
| Perceptions Of Discrimination Against Blacks | 1.329 | 1.329 |
| Perceptions Of Discrimination Against Asian Americans | 1.308 | 1.308 |
| Perceptions Of Discrimination Against Native Americans | 1.224 | 1.224 |
| Perceptions Of Discrimination Against Immigrants | 1.357 | 1.359 |
| Perceptions Of Discrimination Against Latinos | 1.370 | 1.371 |
| Perceptions Of Discrimination Against Gays and Lesbians | 1.262 | 1.262 |
| Perceptions Of Discrimination Against Muslims | 1.341 | 1.342 |
| Education | 1.058 | 1.060 |
| Orientation | 1.042 | 1.044 |
| Community Type | 1.017 | 1.018 |
| Age | 1.028 | 1.030 |
| Race | 1.027 | 1.725 |
| Gender \* Race |  | 1.308 |

: Table 1: VIF Values

# Selecting the better model

## Calculating model's goodness-of-fit through deviance and chi-square test

```{r}
anova(base_model, interaction_model, test="Chisq")
```

## Calculating Odds Ratio, Confidence Interval, and p-values (base model)

```{r}
# Extract odds ratios
odds_ratios <- exp(coef(base_model))

# Get confidence intervals for coefficients
conf_intervals <- exp(confint(base_model))

# Extract p-values
p_values <- summary(base_model)$coefficients[, 4]

# Combine everything into a single data frame
results_base_model <- data.frame(
  OddsRatio = odds_ratios,
  LowerCI = conf_intervals[, 1],
  UpperCI = conf_intervals[, 2],
  PValue = p_values
)

# Display results
results_base_model
```

```{r}
# Verify the structure of the dataset
str(dataset_clean)
```

## Creating the analysis table

```{r}
names <- c("Intercept",
           "Gender: Women",
           "Gender: Non-Binary",
           "Gender: Something Else",
           "Treated Unfairly Or Discriminated Against: No", # Yes
           "Discrimination Impact On Life Satisfaction: Some", # A lot 
           "Discrimination Impact On Life Satisfaction: A Little",
           "Discrimination Impact On Life Satisfaction: Not At All",
           "Discrimination Impact On Life Satisfaction: Don't Know",
           "Discrimination Impact On Life Satisfaction: Refused",
           "Perceptions Of Discrimination Against Whites: Some", # A lot 
           "Perceptions Of Discrimination Against Whites: A Little",
           "Perceptions Of Discrimination Against Whites: None At All",
           "Perceptions Of Discrimination Against Whites: Don't Know",
           "Perceptions Of Discrimination Against Blacks: Some", # A lot 
           "Perceptions Of Discrimination Against Blacks: A Little",
           "Perceptions Of Discrimination Against Blacks: None At All",
           "Perceptions Of Discrimination Against Blacks: Don't Know",
           "Perceptions Of Discrimination Against Asian Americans: Some", # A lot 
           "Perceptions Of Discrimination Against Asian Americans: A Little",
           "Perceptions Of Discrimination Against Asian Americans: None At All",
           "Perceptions Of Discrimination Against Asian Americans: Don't Know",
           "Perceptions Of Discrimination Against Native Americans: Some", # A lot 
           "Perceptions Of Discrimination Against Native Americans: A Little",
           "Perceptions Of Discrimination Against Native Americans: None At All",
           "Perceptions Of Discrimination Against Native Americans: Don't Know",
           "Perceptions Of Discrimination Against Immigrants: Some", # A lot 
           "Perceptions Of Discrimination Against Immigrants: A Little",
           "Perceptions Of Discrimination Against Immigrants: None At All",
           "Perceptions Of Discrimination Against Immigrants: Don't Know",
           "Perceptions Of Discrimination Against Latinos: Some", # A lot 
           "Perceptions Of Discrimination Against Latinos: A Little",
           "Perceptions Of Discrimination Against Latinos: None At All",
           "Perceptions Of Discrimination Against Latinos: Don't Know",
           "Perceptions Of Discrimination Against Gays And Lesbians: Some", # A lot 
           "Perceptions Of Discrimination Against Gays And Lesbians: A Little",
           "Perceptions Of Discrimination Against Gays And Lesbians: None At All",
           "Perceptions Of Discrimination Against Gays And Lesbians: Don't Know",
           "Perceptions Of Discrimination Against Muslims: Some",# A lot 
           "Perceptions Of Discrimination Against Muslims: A Little",
           "Perceptions Of Discrimination Against Muslims: None At All",
           "Perceptions Of Discrimination Against Muslims: Don't Know",
           "Education: Some College Experience", # No College Experience
           "Education: College Graduate",
           "Orientation: Queer", #Straight
           "Orientation: Refused",
           "Community Type: Urban/Suburban", # Rural
           "Race: Black", # White
           "Race: Asian American",
           "Race: Latino",
           "Race: AIAN",
           "Race: NHPI",
           "Race: Arab/MENA",
           "Age: 30-39 Years", # 18-29
           "Age: 40-49 Years",
           "Age: 50-59 Years",
           "Age: 60-69 Years",
           "Age: 70+ Years"
           )
```

```         
```

```{r}
# Add row names as a new column
analysis_table <- cbind(RowName = names, results_base_model)

# Render the table
table <- analysis_table |>
  gt() |>
  tab_header(
    title = "Odds Ratios with Confidence Intervals and P-Values for base model"
  ) |>
  cols_label(
    RowName = "Description",  # Change the column header for the row names
    OddsRatio = "Odds Ratio",
    LowerCI = "Lower Confidence Interval",
    UpperCI = "Upper Confidence Interval",
    PValue = "P-Value"
  ) |>
  fmt_number(
    columns = c(OddsRatio, LowerCI, UpperCI),
    decimals = 4
  ) |>
  fmt_scientific(
    columns = c(PValue),
    decimals = 2  # Show p-values in scientific notation
  )

# View the table
table
```

```{r}
# Update the column names in the dataframe
colnames(analysis_table) <- c("Description", "Odds Ratio", "Lower Confidence Interval", 
                              "Upper Confidence Interval", "P-Value")

# View analysis table
analysis_table

# View the updated dataframe
head(analysis_table)
```

## Figure 1: Odds values

```{r}
# Filter for significant variables (p-value < 0.05)
significant_data <- analysis_table[analysis_table$`P-Value` < 0.05, ]

# Add a column for variable names to use in the plot (from the "Description" column)
significant_data <- significant_data |>
  mutate(Variable = Description)

# Create the plot
ggplot(significant_data, aes(x = `Odds Ratio`, y = Variable)) +
  geom_point(size = 4, color = "blue") +  # Add points for odds ratios
  geom_errorbarh(aes(xmin = `Lower Confidence Interval`, 
                     xmax = `Upper Confidence Interval`), 
                 height = 0.3, color = "black") +  # Add horizontal error bars
  geom_vline(xintercept = 1, linetype = "dashed", color = "red", size = 0.7) +  # Add a vertical reference line at Odds Ratio = 1
  labs(
    title = "Odds Ratio of Significant Variables",
    x = "Odds Ratio",
    y = "Variable"
  ) +
  theme_minimal() +  # Use a minimal theme for better aesthetics
  theme(
    axis.text.y = element_text(size = 6, color = "black"),  # Improve y-axis text readability
    axis.title.x = element_text(size = 12, face = "bold"),  # Highlight x-axis title
    axis.title.y = element_text(size = 12, face = "bold"),  # Highlight y-axis title
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),  # Center the title
    plot.margin = margin(1, 1, 1, 1, "cm"),  # Adjust margins
    axis.text.x = element_text(angle = 45, hjust = 1, size = 8),  # Reduce x-axis text size
    legend.text = element_text(size = 8),  # Reduce legend text size
    legend.title = element_text(size = 9)  # Reduce legend title size
  )

# Save the plot with larger dimensions
ggsave("Significant_Variables_Plot_Enhanced.png", width = 40, height = 20)

```

# Model Diagnostics and Evaluation

```{r}
# Generating a prediction as binary values (0 or 1), where:
## 1 indicates a predicted probability greater than 0.5 (positive class, "Important").
## 0 indicates a predicted probability less than or equal to 0.5 (negative class, "Not important").
model_prediction <- ifelse(predict(base_model, 
                                   newdata = dataset_clean, 
                                   type = "response") > 0.5, 1, 0)

# Converting the predictions factors
## 0 Not important, 1 Important
model_prediction_factor <- factor(model_prediction, 
                                  levels = c(0, 1), 
                                  labels = c("Not important", "Important"))  # Adjusting labels

# Convert the response variable to a factor with matching levels
response_factor <- factor(dataset_clean$ImportanceOfStoppingDiscriminationAgainstMinorities,
                          levels = c("Not important", "Important"))  # Ensuring matching order

# Checking if factors are identical (you can uncomment this to check the levels)
# levels(model_prediction_factor)
# levels(response_factor)

# Generate confusion matrix (if needed)
confusionMatrix(table(model_prediction_factor, response_factor),
                positive = "Important",
                mode = "everything")

```

|                                      |                   |               |
|--------------------------------------|-------------------|---------------|
| [**Predicted / Actual**]{.underline} | **Not Important** | **Important** |
| **Not Important**                    | 12361             | 3812          |
| **Important**                        | 666               | 706           |

: Table CONFUSION: Confusion Matrix for Base model

|                           | Metric  |
|---------------------------|---------|
| Sensitivity (Recall)      | 0.1563  |
| Specificity               | 0.9489  |
| Precision                 | 0.5146  |
| F1                        | 0.2397  |
| Balanced Accuracy         | 0.5526  |
| No Information Rate (NIR) | 0.7425  |
| Kappa                     | 0.1361  |
| Mcnemar's Test P-Value    | \<2e-16 |

: Table MODEL: Base Model Statistics

## ROC Curve (fig2)

```{r}
# Generate predicted probabilities (not just binary predictions) from the model
model_probabilities <- predict(base_model, newdata = dataset_clean, type = "response")

# Calculate ROC curve
roc_curve <- roc(response_factor, model_probabilities)

# Calculate AUC (Area Under the Curve)
auc_value <- auc(roc_curve)
print(paste("AUC: ", auc_value))

# Plot ROC curve
plot(roc_curve, main = "ROC Curve", col = "blue", lwd = 2)
legend("bottomright", legend = paste("AUC =", round(auc_value, 3)), col = "blue", lwd = 2)
```

### 2024-12-14 RQ2_final_Sizhe.qmd
Look at RQ2

--- 2024-12-08 RQ2_main.qmd
@@
 ---
-title: "RQ2_main"
-format: 
-      pdf:
-       code-overflow: wrap
+title: "RQ2_final"
+format: html
 editor: visual
+
+execute: 
+  warning: false
 ---
 
-## Quarto
-The `echo: false` option disables the printing of code (only output is displayed).
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(
-  message = FALSE,   # Suppress messages
-  warning = FALSE    # Suppress warnings
-)
-```
+### Loading Libraries
 
-```{r echo = TRUE}
+```{r echo = FALSE}
 # Load necessary libraries
@@
 library(pROC)
+library(vcd)
+library(car)
 ```
 
-```{r echo = TRUE}
-file_path <- "/Users/nakiyahdhariwala/StatsProject/39096-0001-Data.rda"
+```{r echo = FALSE, results = FALSE}
+# Read the dataset
+file_path <- "/Users/chensi/Desktop/MIDS/Fall 2024/IDS 702/ICPSR_39096/DS0001/39096-0001-Data.rda"
 
@@
 
-```{r}
+## Pre-processing Data
+
+```{r echo = FALSE, results=FALSE}
 # Define the required columns
@@
 head(dataset_new)
+```
+
+```{r echo=FALSE, results=FALSE}
+# Data Cleaning and Recoding for dataset_new
+# Adding religion column
+dataset_new <- dataset_new |> 
+  mutate(
+    # Education recoding
+    educ = case_when(
+      S13 == "(1) Grades 1-8" ~ "No college experience",
+      S13 == "(2) Some High School, but did not graduate" ~ "No college experience",
+      S13 == "(3) High School graduate or GED" ~ "No college experience",
+      S13 == "(4) Some college" ~ "Some college experience",
+      S13 == "(5) Associates, 2-year degree" ~ "College graduate",
+      S13 == "(6) Bachelors, 4-year degree" ~ "College graduate",
+      S13 == "(7) Post-graduate degree" ~ "College graduate"
+    ),
+    # Sexual orientation recoding
+    sex_orientation = case_when(
+      S3 == "(01) Straight, that is, not gay or lesbian" ~ "Straight",
+      S3 == "(02) Gay or lesbian" ~ "Queer",
+      S3 == "(03) Bisexual" ~ "Queer",
+      S3 == "(04) Something else (Specify)" ~ "Queer",
+      S3 == "(88) Refused" ~ "Refused/Don't Know",
+      S3 == "(99) I don" ~ "Refused/Don't Know"
+    ),
+    # Urban vs rural recoding
+    rural = case_when(
+      S14 == "(1) Large urban area" ~ "Urban/suburban",
+      S14 == "(2) Large suburb near large city" ~ "Urban/suburban",
+      S14 == "(3) Small suburb near small town or city" ~ "Urban/suburban",
+      S14 == "(4) Small town or small city" ~ "Urban/suburban",
+      S14 == "(5) Rural area" ~ "Rural"
+    ),
+    # Race recoding
+    race_recoded = case_when(
+      S2_RACE_PRIME == "(1) White" ~ "White",
+      S2_RACE_PRIME == "(2) Hispanic or Latino" ~ "Latino",
+      S2_RACE_PRIME == "(3) Black or African American" ~ "Black",
+      S2_RACE_PRIME == "(4) Asian American" ~ "Asian American",
+      S2_RACE_PRIME == "(5) American Indian/Native American" ~ "AIAN",
+      S2_RACE_PRIME == "(6) Arab, Middle Eastern or North African" ~ "Arab/MENA",
+      S2_RACE_PRIME == "(7) Native Hawaiian" ~ "NHPI",
+      S2_RACE_PRIME == "(8) Not Hawaiian, but other Pacific Islander" ~ "NHPI"
+    ),
+    # Age recoding
+    age_recoded = case_when(
+      S5_AGE == "(1) 16-17" ~ "16-17",
+      S5_AGE == "(2) 18-29" ~ "18-29",
+      S5_AGE == "(3) 30-39" ~ "30-39",
+      S5_AGE == "(4) 40-49" ~ "40-49",
+      S5_AGE == "(5) 50-59" ~ "50-59",
+      S5_AGE == "(6) 60-69" ~ "60-69",
+      S5_AGE == "(7) 70 +" ~ "70+"
+    )
+  ) |>
+  mutate(
+    # Converting variables to factors
+    race_recoded = factor(race_recoded, levels = c("White", "Black", "Asian American", "Latino", "AIAN", "NHPI", "Arab/MENA")),
+    age_recoded = factor(age_recoded, levels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+")),
+    educ = factor(educ, levels = c("No college experience", "Some college experience", "College graduate")),
+    rural = factor(rural, levels = c("Rural", "Urban/suburban")),
+    sex_orientation = factor(sex_orientation, levels = c("Straight", "Queer", "Refused/Don't Know"))
+  )
+
+# Convert individual religion columns to numeric
+for (i in 1:15) {
+  col_name <- paste0("Q58R", i, "_numeric")
+  dataset_new[[col_name]] <- as.numeric(sub("^\\((\\d+)\\).*", "\\1", dataset_new[[paste0("Q58R", i)]]))
+}
+
+# Create a new Religion variable
+dataset_new$Religion <- "None"
+
+# Assign religion based on the numeric values of each Q58R column
+religion_mapping <- c(
+  "Q58R1_numeric" = "Catholic", 
+  "Q58R2_numeric" = "Protestant", 
+  "Q58R3_numeric" = "Other Christian",
+  "Q58R4_numeric" = "Jewish", 
+  "Q58R5_numeric" = "Muslim", 
+  "Q58R6_numeric" = "Hindu", 
+  "Q58R7_numeric" = "Buddhist", 
+  "Q58R8_numeric" = "Mormon/LDS", 
+  "Q58R9_numeric" = "Folk Religion", 
+  "Q58R10_numeric" = "Daoist", 
+  "Q58R11_numeric" = "Shamanism", 
+  "Q58R12_numeric" = "Ancestral Religion", 
+  "Q58R13_numeric" = "Spiritual but not religious", 
+  "Q58R14_numeric" = "Atheist", 
+  "Q58R15_numeric" = "Agnostic"
+)
+
+# Loop through the mapping to assign religion values
+for (col in names(religion_mapping)) {
+  dataset_new$Religion[dataset_new[[col]] == 1] <- religion_mapping[[col]]
+}
+
+# Handle multiple religious affiliations
+dataset_new$Religion[rowSums(dataset_new[, grep("_numeric", names(dataset_new))] == 1) > 1] <- "Multiple Religions"
+
+# Convert Religion to a factor with specific levels
+dataset_new$Religion <- factor(dataset_new$Religion,
+                               levels = c("Catholic", "Protestant", "Other Christian", "Jewish", "Muslim",
+                                          "Hindu", "Buddhist", "Mormon/LDS", "Folk Religion",
+                                          "Daoist", "Shamanism", "Ancestral Religion",
+                                          "Spiritual but not religious", "Atheist", "Agnostic", 
+                                          "Multiple Religions", "None"))
+
+# Remove redundant religion columns
+redundant_cols <- c(
+  paste0("Q58R", 1:15), paste0("Q58R", 1:15, "_numeric")
+)
+
+# Remove redundant columns from the dataset
+dataset_clean <- dataset_new[, !(names(dataset_new) %in% redundant_cols)]
+
+# Take a look at the updated dataset
+head(dataset_clean)
+```
+
+```{r echo=FALSE, results=FALSE}
+# Remove all redundant columns
+redundant_columns <- c("Q629R1", "Q629R2", "Q629R3", "Q629R4", 
+                       "Q629R5", "Q629R6", "Q629R7",
+                       "S2_RACE_PRIME", "S13", "S3", "S5_AGE","S14")
+
+dataset_clean <- dataset_clean[, !(names(dataset_clean) %in% redundant_columns)]
+
+head(dataset_clean)
+```
+
+```{r echo=FALSE, results=FALSE}
+# Rename remaining columns in dataset_clean
+colnames(dataset_clean) <- c(
+  "ImportanceOfStoppingDiscriminationAgainstMinorities",  # Q1R5
+  "Gender",                                               # S3B
+  "TreatedUnfairlyOrDiscriminatedAgainst",                # Q627
+  "DiscriminationImpactOnLifeSatisfaction",              # Q633
+  "PerceptionsOfDiscriminationAgainstWhites",            # Q619_Q626R1
+  "PerceptionsOfDiscriminationAgainstBlacks",            # Q619_Q626R2
+  "PerceptionsOfDiscriminationAgainstAsianAmericans",    # Q619_Q626R3
+  "PerceptionsOfDiscriminationAgainstNativeAmericans",   # Q619_Q626R4
+  "PerceptionsOfDiscriminationAgainstImmigrants",        # Q619_Q626R5
+  "PerceptionsOfDiscriminationAgainstLatinos",           # Q619_Q626R6
+  "PerceptionsOfDiscriminationAgainstGaysAndLesbians",   # Q619_Q626R7
+  "PerceptionsOfDiscriminationAgainstMuslims",           # Q619_Q626R8
+  "Education",                                           # educ
+  "SexOrientation",                                      # sex_orientation
+  "Rural",                                               # rural
+  "Race",                                                # race_recoded
+  "Age",                                                 # age_recoded
+  "Religion"                                             # Religion
+)
+
+# Verify the renamed columns
+colnames(dataset_clean)
+head(dataset_clean)
+```
+
+# Model Building
+
+## Model 1: Main
+
+```{r echo=FALSE, results=FALSE}
+# Regression model
+base_model <- glm(ImportanceOfStoppingDiscriminationAgainstMinorities ~ Gender +
+               TreatedUnfairlyOrDiscriminatedAgainst + 
+               DiscriminationImpactOnLifeSatisfaction +
+               PerceptionsOfDiscriminationAgainstWhites + 
+               PerceptionsOfDiscriminationAgainstBlacks +
+               PerceptionsOfDiscriminationAgainstAsianAmericans + 
+               PerceptionsOfDiscriminationAgainstNativeAmericans + 
+               PerceptionsOfDiscriminationAgainstImmigrants + 
+               PerceptionsOfDiscriminationAgainstLatinos + 
+               PerceptionsOfDiscriminationAgainstGaysAndLesbians + 
+               PerceptionsOfDiscriminationAgainstMuslims +
+               Education + 
+                 SexOrientation +
+                 Rural +
+                 Race +
+                 Age,
+
+             data = dataset_clean, family = "binomial")
+
+# Check multicollinearity using VIF
+vif(base_model)
+```
+
+## Model 2: Model with interaction term
+
+```{r echo=FALSE, results=FALSE}
+interaction_model <- glm(ImportanceOfStoppingDiscriminationAgainstMinorities ~ 
+               TreatedUnfairlyOrDiscriminatedAgainst + 
+               DiscriminationImpactOnLifeSatisfaction +
+               PerceptionsOfDiscriminationAgainstWhites + 
+               PerceptionsOfDiscriminationAgainstBlacks +
+               PerceptionsOfDiscriminationAgainstAsianAmericans + 
+               PerceptionsOfDiscriminationAgainstNativeAmericans + 
+               PerceptionsOfDiscriminationAgainstImmigrants + 
+               PerceptionsOfDiscriminationAgainstLatinos + 
+               PerceptionsOfDiscriminationAgainstGaysAndLesbians + 
+               PerceptionsOfDiscriminationAgainstMuslims +
+               Education + 
+                 SexOrientation +
+                 Rural +
+                 Age + 
+                 Gender * Race,
+
+             data = dataset_clean, family = "binomial")
+
+# Check multicollinearity using VIF
+vif(interaction_model)
+```
+
+# Selecting the better model
+
+## Calculating model's goodness-of-fit through deviance and chi-square test
+
+```{r}
+anova(base_model, interaction_model, test="Chisq")
+```
+
+# confusion matrix
+
+```{r}
+# Generate predicted probabilities and binary predictions
+predicted_probs <- predict(base_model, type = "response")
+
+# Generate binary predictions with a threshold of 0.5
+predicted_classes <- ifelse(predicted_probs > 0.5, 1, 0)
+
+# Create a confusion matrix
+confusion_matrix <- table(
+  Actual = dataset_clean$ImportanceOfStoppingDiscriminationAgainstMinorities,
+  Predicted = predicted_classes
+)
+print("Confusion Matrix:")
+print(confusion_matrix)
+
+# Calculate metrics from the confusion matrix
+TP <- confusion_matrix[2, 2]  # True Positives
+TN <- confusion_matrix[1, 1]  # True Negatives
+FP <- confusion_matrix[1, 2]  # False Positives
+FN <- confusion_matrix[2, 1]  # False Negatives
+
+# Compute performance metrics
+accuracy <- (TP + TN) / sum(confusion_matrix)
+precision <- TP / (TP + FP)
+recall <- TP / (TP + FN)
+specificity <- TN / (TN + FP)
+f1_score <- 2 * (precision * recall) / (precision + recall)
+
+cat("Metrics:\n")
+cat(sprintf("Accuracy: %.4f\n", accuracy))
+cat(sprintf("Precision: %.4f\n", precision))
+cat(sprintf("Recall (Sensitivity): %.4f\n", recall))
+cat(sprintf("Specificity: %.4f\n", specificity))
+cat(sprintf("F1-Score: %.4f\n", f1_score))
 
 ```
+
+The model performs well for the majority class ("Not important") but struggles significantly with the minority class ("Important"), as shown by the low recall and F1-Score.
+
+-   **Imbalance in Recognizing Importance**:
+
+    -   The model has high specificity (94.86%), meaning it effectively identifies respondents who do **not** place high importance on anti-discrimination policies.
+
+    -   However, the low sensitivity (15.63%) shows the model struggles to identify respondents who do place high importance on such policies.
+
+    -   This could suggest a possible imbalance in how different factors (e.g., experiences with racism, demographics) are weighed in the model, potentially requiring better feature selection or adjustments.
+
+-   **Predictive Power of the Model**:
+
+    -   The moderate accuracy (74.48%) suggests the model provides a reasonable prediction of respondents' views, but the low F1-Score (23.97%) indicates that it has difficulty capturing nuanced relationships, especially for respondents who prioritize such policies.
+
+    -   This highlights a need for deeper exploration into how experiences of racism interact with demographic factors to influence policy importance.
+
+# ROC
+
+```{r}
+# ROC Curve
+library(pROC)
+roc_curve <- roc(dataset_clean$ImportanceOfStoppingDiscriminationAgainstMinorities, predicted_probs)
+plot(roc_curve, main = "ROC Curve", col = "blue", lwd = 2)
+auc_value <- auc(roc_curve)
+cat(sprintf("AUC: %.4f\n", auc_value))
+```
+
+The ROC curve confirms that the model has discriminatory power but could be improved, particularly in recognizing minority class cases.
+
+-   **Moderate Discriminatory Power**:
+
+    -   The ROC curve and AUC suggest the model moderately distinguishes between respondents who place high vs. low importance on policies addressing racial discrimination.
+
+    -   This could mean that personal experiences with racism and demographic factors do play a role but may not fully account for the variance in the outcome.
+
+-   **Potential for Model Improvement**:
+
+    -   The curve indicates room for improvement in how the model uses predictors.
+For example, interactions between demographic factors (e.g., education and race) and personal experiences with racism might enhance predictive accuracy.

### 2024-12-14 RQ2_main copy.qmd
Look at RQ2

--- 2024-12-08 RQ2_main.qmd
@@
 ```{r echo = FALSE, results = FALSE}
-file_path <- "/Users/nakiyahdhariwala/StatsProject/39096-0001-Data.rda"
+file_path <- "39096-0001-Data.rda"
 
@@
 
+```{r}
+head(dataset_clean)
+```
+
 # Model Building
 
+## Model 1: Main
 ```{r echo=FALSE, results=FALSE}
@@
 
-## Model 2: stepwise regression
-```{r echo=FALSE, results=FALSE}
-# Perform stepwise selection
-model_stepwise <- step(model, 
-                       direction = "both",
-                       trace = 0)
-```
-
-```{r echo=FALSE, results=FALSE}
-# print(model_stepwise)
-vif(model_stepwise)
-```
-
-```{r echo=FALSE, results=FALSE}
-
+```{r}
 # Extract odds ratios
-odds_ratios <- exp(coef(model_stepwise))
+odds_ratios <- exp(coef(model))
 
 # Get confidence intervals for coefficients
-conf_intervals <- exp(confint(model_stepwise))
+conf_intervals <- exp(confint(model))
 
 # Extract p-values
-p_values <- summary(model_stepwise)$coefficients[, 4]
+p_values <- summary(model)$coefficients[, 4]
 
 # Combine everything into a single data frame
-results_model_stepwise <- data.frame(
+results_model <- data.frame(
   OddsRatio = odds_ratios,
@@
 # Display results
-results_model_stepwise
+results_model
 ```
 
-## Model Diagnostics
+## Model 2: Without Religion
+```{r echo=FALSE, results=FALSE}
+model_without_religion <- glm(ImportanceOfStoppingDiscriminationAgainstMinorities ~ Gender +
+               TreatedUnfairlyOrDiscriminatedAgainst + 
+               DiscriminationImpactOnLifeSatisfaction +
+               PerceptionsOfDiscriminationAgainstWhites + 
+               PerceptionsOfDiscriminationAgainstBlacks +
+               PerceptionsOfDiscriminationAgainstAsianAmericans + 
+               PerceptionsOfDiscriminationAgainstNativeAmericans + 
+               PerceptionsOfDiscriminationAgainstImmigrants + 
+               PerceptionsOfDiscriminationAgainstLatinos + 
+               PerceptionsOfDiscriminationAgainstGaysAndLesbians + 
+               PerceptionsOfDiscriminationAgainstMuslims +
+               Education + 
+                 SexOrientation +
+                 Rural +
+                 Race +
+                 Age,
+
+             data = dataset_clean, family = "binomial")
+
+# Check multicollinearity using VIF
+vif(model_without_religion)
+```
+
+```{r echo=FALSE, results=FALSE}
+
+# Extract odds ratios
+odds_ratios <- exp(coef(model_without_religion))
+
+# Get confidence intervals for coefficients
+conf_intervals <- exp(confint(model_without_religion))
+
+# Extract p-values
+p_values <- summary(model_without_religion)$coefficients[, 4]
+
+# Combine everything into a single data frame
+results_model_without_religion <- data.frame(
+  OddsRatio = odds_ratios,
+  LowerCI = conf_intervals[, 1],
+  UpperCI = conf_intervals[, 2],
+  PValue = p_values
+)
+
+# Display results
+results_model_without_religion
+```
+
+## Model 3: Model without religion + interaction term
+```{r echo=FALSE, results=FALSE}
+model_noreligion_Interaction <- glm(ImportanceOfStoppingDiscriminationAgainstMinorities ~ Gender +
+               TreatedUnfairlyOrDiscriminatedAgainst + 
+               DiscriminationImpactOnLifeSatisfaction +
+               PerceptionsOfDiscriminationAgainstWhites + 
+               PerceptionsOfDiscriminationAgainstBlacks +
+               PerceptionsOfDiscriminationAgainstAsianAmericans + 
+               PerceptionsOfDiscriminationAgainstNativeAmericans + 
+               PerceptionsOfDiscriminationAgainstImmigrants + 
+               PerceptionsOfDiscriminationAgainstLatinos + 
+               PerceptionsOfDiscriminationAgainstGaysAndLesbians + 
+               PerceptionsOfDiscriminationAgainstMuslims +
+               Education + 
+                 SexOrientation +
+                 Rural +
+                 Race +
+                 Age + 
+                 Gender * Race,
+
+             data = dataset_clean, family = "binomial")
+
+# Check multicollinearity using VIF
+vif(model_noreligion_Interaction)
+```
+
+```{r echo=FALSE, results=FALSE}
+
+# Extract odds ratios
+odds_ratios <- exp(coef(model_noreligion_Interaction))
+
+# Get confidence intervals for coefficients
+conf_intervals <- exp(confint(model_noreligion_Interaction))
+
+# Extract p-values
+p_values <- summary(model_noreligion_Interaction)$coefficients[, 4]
+
+# Combine everything into a single data frame
+results_model_noreligion_Interaction <- data.frame(
+  OddsRatio = odds_ratios,
+  LowerCI = conf_intervals[, 1],
+  UpperCI = conf_intervals[, 2],
+  PValue = p_values
+)
+
+# Display results
+model_noreligion_Interaction
+```
+
+```{r}
+results_model_noreligion_Interaction
+```
+
+```{r}
+model_noreligion_Interaction
+model_without_religion
+
+anova(model_without_religion, model_noreligion_Interaction, test="Chisq")
+
+```
+
+```{r}
+results_model_noreligion_Interaction
+```
+
+# Model Diagnostics
+
+```{r}
+# Convert the prediction to a factor with correct levels
+model_prediction_factor <- factor(model_prediction, 
+                                  levels = c(0, 1), 
+                                  labels = c("Important", "Not important"))
+
+# Convert the response variable to a factor with matching levels
+response_factor <- factor(dataset_clean$ImportanceOfStoppingDiscriminationAgainstMinorities,
+                          levels = c("Important", "Not important"))
+
+# Generate the confusion matrix
+confusionMatrix(table(model_prediction_factor, response_factor),
+                positive = "Important",
+                mode = "everything")
+```
+
+## OLD DATA
+
 ```{r echo=FALSE, results=FALSE}
@@
 ## Figure 1: Proportion of outcome variable for RQ 1
+
 ```{r echo=FALSE}

### 2024-12-14 RQ2_main.qmd
Look at RQ2

@@
 ## Model 2: stepwise regression
+
 ```{r echo=FALSE, results=FALSE}
@@
 ## Model Diagnostics
+
 ```{r echo=FALSE, results=FALSE}

### 2024-12-15 draft_finalreport.qmd
updating notebooks

--- 2024-12-09 draft_finalreport.qmd
@@
 
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(
+  message = FALSE,   # Suppress messages
+  warning = FALSE    # Suppress warnings
+)
+```
+
+```{r echo = FALSE}
+# Libraries
+library(tidyverse)  
+library(pROC)
+library(vcd)
+library(car)
+library(MASS) # package for model
+library(caret) # package for confusion matrix
+library(janitor)
+library(brant)
+library(knitr)
+library(gt)
+```
+
+```{r echo = FALSE, results = FALSE}
+file_path <- "//Users/pdeguz01/Documents/git/Data/2020 CMPS/DS0001/39096-0001-Data.rda"
+
+load(file_path)          # Load the RDA file
+ls()                     # List objects loaded into the environment
+dataset <- da39096.0001  # Access the dataset
+
+```
+
+```{r echo = FALSE, results = FALSE}
+# Define the required columns
+required_cols <- c(
+  # Outcome Variable: Stopping discrimination against racial/ethnic minorities
+  "Q1R5",
+
+  # Demographic Variables
+  "S2_RACE_PRIME",   # Primary Race
+  "S3",              # SexualOrientation
+  "S3B",             # Gender
+  "S5_AGE",          # Age categories
+  "S13",             # Highest level of education
+  "S14",             # Community lived in
+
+  # Variables capturing experiences with racism
+  "Q629R1",          # Unfair treatment due to racial background/ethnicity
+  "Q629R2",          # Unfair treatment due to skin color
+  "Q629R3",          # Unfair treatment due to gender
+  "Q629R4",          # Unfair treatment due to sexuality/orientation
+  "Q629R5",          # Unfair treatment due to immigration status
+  "Q629R6",          # Unfair treatment due to religion
+  "Q629R7",          # Unfair treatment due to accent
+
+  "Q627",            # Treated unfairly or discriminated against because of your race, ethnicity, gender, sexuality, being an immigrant, religious heritage or having an accent?
+  "Q633",            # Discrimination impact on life satisfaction
+
+  # Perceptions of discrimination
+  "Q619_Q626R1",     # Perceptions of discrimination against Whites
+  "Q619_Q626R2",     # Perceptions of discrimination against Blacks
+  "Q619_Q626R3",     # Perceptions of discrimination against Asian Americans
+  "Q619_Q626R4",     # Perceptions of discrimination against Native Americans
+  "Q619_Q626R5",     # Perceptions of discrimination against Immigrants
+  "Q619_Q626R6",     # Perceptions of discrimination against Latinos
+  "Q619_Q626R7",     # Perceptions of discrimination against GaysAndLesbians
+  "Q619_Q626R8"      # Perceptions of discrimination against Muslims
+)
+
+```
+
+```{r echo=FALSE, results=FALSE}
+# Create a new dataframe with selected variables
+dataset_new <- dataset[, required_cols]
+
+# Recode the outcome variable 'Q1R5' for readability
+dataset_new$Q1R5 <- factor(
+  ifelse(
+    dataset_new$Q1R5 == "(0) NO TO: Stopping discrimination against racial/ethnic minorities", 
+    "Not important",
+    "Important"  # Remaining cases are "Important"
+  ),
+  levels = c("Not important", "Important") # Define factor levels
+)
+
+# Define the columns for perceptions of discrimination
+perception_cols <- c(
+  "Q619_Q626R1",  # Perceptions of discrimination against Whites
+  "Q619_Q626R2",  # Perceptions of discrimination against Blacks
+  "Q619_Q626R3",  # Perceptions of discrimination against Asian Americans
+  "Q619_Q626R4",  # Perceptions of discrimination against Native Americans
+  "Q619_Q626R5",  # Perceptions of discrimination against Immigrants
+  "Q619_Q626R6",  # Perceptions of discrimination against Latinos
+  "Q619_Q626R7",  # Perceptions of discrimination against GaysAndLesbians
+  "Q619_Q626R8",  # Perceptions of discrimination against Muslims
+  "Q633"          # Discrimination impact on life satisfaction
+)
+
+# Recode 'Don' as 'Don't know' and ensure columns remain factors
+dataset_new[perception_cols] <- lapply(dataset_new[perception_cols], function(x) {
+  x <- as.character(x)  # Temporarily convert to character for recoding
+  x <- ifelse(x == "(5) Don", "(5) Don't know", x)  # Recode 'Don' to 'Don't know'
+  factor(x)  # Convert back to factor
+})
+
+# Define required columns to convert to factors (if not already defined)
+# Assuming `required_cols` is previously defined; otherwise, specify explicitly
+dataset_new[required_cols] <- lapply(dataset_new[required_cols], as.factor)
+
+# Verify the structure of the dataset
+str(dataset_new[required_cols])
+
+# Display the first few rows to check the changes
+head(dataset_new)
+```
+
+```{r, echo=FALSE, results=FALSE}
+# Data Cleaning and Recoding for dataset_new
+dataset_clean <- dataset_new |> 
+  mutate(
+    # Education recoding
+    educ = case_when(
+      S13 == "(1) Grades 1-8" ~ "No college experience",
+      S13 == "(2) Some High School, but did not graduate" ~ "No college experience",
+      S13 == "(3) High School graduate or GED" ~ "No college experience",
+      S13 == "(4) Some college" ~ "Some college experience",
+      S13 == "(5) Associates, 2-year degree" ~ "College graduate",
+      S13 == "(6) Bachelors, 4-year degree" ~ "College graduate",
+      S13 == "(7) Post-graduate degree" ~ "College graduate"
+    ),
+    # Sexual orientation recoding
+    sex_orientation = case_when(
+      S3 == "(01) Straight, that is, not gay or lesbian" ~ "Straight",
+      S3 == "(02) Gay or lesbian" ~ "Queer",
+      S3 == "(03) Bisexual" ~ "Queer",
+      S3 == "(04) Something else (Specify)" ~ "Queer",
+      S3 == "(88) Refused" ~ "Refused/Don't Know",
+      S3 == "(99) I don" ~ "Refused/Don't Know"
+    ),
+    # Urban vs rural recoding
+    rural = case_when(
+      S14 == "(1) Large urban area" ~ "Urban/suburban",
+      S14 == "(2) Large suburb near large city" ~ "Urban/suburban",
+      S14 == "(3) Small suburb near small town or city" ~ "Urban/suburban",
+      S14 == "(4) Small town or small city" ~ "Urban/suburban",
+      S14 == "(5) Rural area" ~ "Rural"
+    ),
+    # Race recoding
+    race_recoded = case_when(
+      S2_RACE_PRIME == "(1) White" ~ "White",
+      S2_RACE_PRIME == "(2) Hispanic or Latino" ~ "Latino",
+      S2_RACE_PRIME == "(3) Black or African American" ~ "Black",
+      S2_RACE_PRIME == "(4) Asian American" ~ "Asian American",
+      S2_RACE_PRIME == "(5) American Indian/Native American" ~ "AIAN",
+      S2_RACE_PRIME == "(6) Arab, Middle Eastern or North African" ~ "Arab/MENA",
+      S2_RACE_PRIME == "(7) Native Hawaiian" ~ "NHPI",
+      S2_RACE_PRIME == "(8) Not Hawaiian, but other Pacific Islander" ~ "NHPI"
+    ),
+    # Age recoding
+    age_recoded = case_when(
+      S5_AGE == "(1) 16-17" ~ "16-17",
+      S5_AGE == "(2) 18-29" ~ "18-29",
+      S5_AGE == "(3) 30-39" ~ "30-39",
+      S5_AGE == "(4) 40-49" ~ "40-49",
+      S5_AGE == "(5) 50-59" ~ "50-59",
+      S5_AGE == "(6) 60-69" ~ "60-69",
+      S5_AGE == "(7) 70 +" ~ "70+"
+    )
+  ) |>
+  mutate(
+    # Converting variables to factors
+    race_recoded = factor(race_recoded, levels = c("White", "Black", "Asian American", "Latino", "AIAN", "NHPI", "Arab/MENA")),
+    age_recoded = factor(age_recoded, levels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+")),
+    educ = factor(educ, levels = c("No college experience", "Some college experience", "College graduate")),
+    rural = factor(rural, levels = c("Rural", "Urban/suburban")),
+    sex_orientation = factor(sex_orientation, levels = c("Straight", "Queer", "Refused/Don't Know"))
+  )
+
+# Take a look at the updated dataset
+head(dataset_clean)
+
+```
+
+```{r echo=FALSE, results=FALSE}
+# Remove all redundant columns
+redundant_columns <- c("Q629R1", "Q629R2", "Q629R3", "Q629R4", 
+                       "Q629R5", "Q629R6", "Q629R7",
+                       "S2_RACE_PRIME", "S13", "S3", "S5_AGE","S14")
+
+dataset_clean <- dataset_clean[, !(names(dataset_clean) %in% redundant_columns)]
+```
+
+```{r echo=FALSE, results=FALSE}
+# creating the analysis table for RQ2
+
+names <- c("Intercept",
+           "Gender: Women",
+           "Gender: Non-Binary",
+           "Gender: Something Else",
+           "Treated Unfairly Or Discriminated Against: No", # Yes
+           "Discrimination Impact On Life Satisfaction: Some", # A lot 
+           "Discrimination Impact On Life Satisfaction: A Little",
+           "Discrimination Impact On Life Satisfaction: Not At All",
+           "Discrimination Impact On Life Satisfaction: Don't Know",
+           "Discrimination Impact On Life Satisfaction: Refused",
+           "Perceptions Of Discrimination Against Whites: Some", # A lot 
+           "Perceptions Of Discrimination Against Whites: A Little",
+           "Perceptions Of Discrimination Against Whites: None At All",
+           "Perceptions Of Discrimination Against Whites: Don't Know",
+           "Perceptions Of Discrimination Against Blacks: Some", # A lot 
+           "Perceptions Of Discrimination Against Blacks: A Little",
+           "Perceptions Of Discrimination Against Blacks: None At All",
+           "Perceptions Of Discrimination Against Blacks: Don't Know",
+           "Perceptions Of Discrimination Against Asian Americans: Some", # A lot 
+           "Perceptions Of Discrimination Against Asian Americans: A Little",
+           "Perceptions Of Discrimination Against Asian Americans: None At All",
+           "Perceptions Of Discrimination Against Asian Americans: Don't Know",
+           "Perceptions Of Discrimination Against Native Americans: Some", # A lot 
+           "Perceptions Of Discrimination Against Native Americans: A Little",
+           "Perceptions Of Discrimination Against Native Americans: None At All",
+           "Perceptions Of Discrimination Against Native Americans: Don't Know",
+           "Perceptions Of Discrimination Against Immigrants: Some", # A lot 
+           "Perceptions Of Discrimination Against Immigrants: A Little",
+           "Perceptions Of Discrimination Against Immigrants: None At All",
+           "Perceptions Of Discrimination Against Immigrants: Don't Know",
+           "Perceptions Of Discrimination Against Latinos: Some", # A lot 
+           "Perceptions Of Discrimination Against Latinos: A Little",
+           "Perceptions Of Discrimination Against Latinos: None At All",
+           "Perceptions Of Discrimination Against Latinos: Don't Know",
+           "Perceptions Of Discrimination Against Gays And Lesbians: Some", # A lot 
+           "Perceptions Of Discrimination Against Gays And Lesbians: A Little",
+           "Perceptions Of Discrimination Against Gays And Lesbians: None At All",
+           "Perceptions Of Discrimination Against Gays And Lesbians: Don't Know",
+           "Perceptions Of Discrimination Against Muslims: Some",# A lot 
+           "Perceptions Of Discrimination Against Muslims: A Little",
+           "Perceptions Of Discrimination Against Muslims: None At All",
+           "Perceptions Of Discrimination Against Muslims: Don't Know",
+           "Education: Some College Experience", # No College Experience
+           "Education: College Graduate",
+           "Orientation: Queer", #Straight
+           "Orientation: Refused",
+           "Community Type: Urban/Suburban", # Rural
+           "Race: Black", # White
+           "Race: Asian American",
+           "Race: Latino",
+           "Race: AIAN",
+           "Race: NHPI",
+           "Race: Arab/MENA",
+           "Age: 30-39 Years", # 18-29
+           "Age: 40-49 Years",
+           "Age: 50-59 Years",
+           "Age: 60-69 Years",
+           "Age: 70+ Years"
+           )
+
+# Add row names as a new column
+analysis_table <- cbind(RowName = names, results_base_model)
+
+# Render the table
+table <- analysis_table |>
+  gt() |>
+  tab_header(
+    title = "Odds Ratios with Confidence Intervals and P-Values for base model"
+  ) |>
+  cols_label(
+    RowName = "Description",  # Change the column header for the row names
+    OddsRatio = "Odds Ratio",
+    LowerCI = "Lower Confidence Interval",
+    UpperCI = "Upper Confidence Interval",
+    PValue = "P-Value"
+  ) |>
+  fmt_number(
+    columns = c(OddsRatio, LowerCI, UpperCI),
+    decimals = 4
+  ) |>
+  fmt_scientific(
+    columns = c(PValue),
+    decimals = 2  # Show p-values in scientific notation
+  )
+
+# View the table
+#table
+
+# Update the column names in the dataframe
+colnames(analysis_table) <- c("Description", "Odds Ratio", "Lower Confidence Interval", 
+                              "Upper Confidence Interval", "P-Value")
+
+# View analysis table
+#analysis_table
+
+# View the updated dataframe
+#head(analysis_table)
+
+```
+
+```{r echo=FALSE, results=FALSE}
+# Regression model - RQ2
+base_model <- glm(ImportanceOfStoppingDiscriminationAgainstMinorities ~ Gender +
+               TreatedUnfairlyOrDiscriminatedAgainst + 
+               DiscriminationImpactOnLifeSatisfaction +
+               PerceptionsOfDiscriminationAgainstWhites + 
+               PerceptionsOfDiscriminationAgainstBlacks +
+               PerceptionsOfDiscriminationAgainstAsianAmericans + 
+               PerceptionsOfDiscriminationAgainstNativeAmericans + 
+               PerceptionsOfDiscriminationAgainstImmigrants + 
+               PerceptionsOfDiscriminationAgainstLatinos + 
+               PerceptionsOfDiscriminationAgainstGaysAndLesbians + 
+               PerceptionsOfDiscriminationAgainstMuslims +
+               Education + 
+                 SexOrientation +
+                 Rural +
+                 Race +
+                 Age,
+
+             data = dataset_clean, family = "binomial")
+```
+
+```{r echo=FALSE, results=FALSE}
+# Regression Model with interaction term - RQ2
+interaction_model <- glm(ImportanceOfStoppingDiscriminationAgainstMinorities ~ 
+               TreatedUnfairlyOrDiscriminatedAgainst + 
+               DiscriminationImpactOnLifeSatisfaction +
+               PerceptionsOfDiscriminationAgainstWhites + 
+               PerceptionsOfDiscriminationAgainstBlacks +
+               PerceptionsOfDiscriminationAgainstAsianAmericans + 
+               PerceptionsOfDiscriminationAgainstNativeAmericans + 
+               PerceptionsOfDiscriminationAgainstImmigrants + 
+               PerceptionsOfDiscriminationAgainstLatinos + 
+               PerceptionsOfDiscriminationAgainstGaysAndLesbians + 
+               PerceptionsOfDiscriminationAgainstMuslims +
+               Education + 
+                 SexOrientation +
+                 Rural +
+                 Age + 
+                 Gender * Race,
+
+             data = dataset_clean, family = "binomial")
+```
+
+# ordinal model
+
+# ```{r, echo=FALSE, results=FALSE}
+# # data load and cleaning for ordinal regression model
+# required_cols_ord <- c("Q29", # Outcome Variable; interest in politics
+#                    "Q271", # Predictor Variable; importance of race to identity
+#                    "S2_RACE_PRIME", # Race
+#                    "S3", # Orientation
+#                    "S3B", # Gender
+#                    "S5_AGE", # Age Category
+#                    "S13", # Educational Attainment
+#                    "S14", # Community Type
+#                    "S7", # Nativity
+#                    "Q21", # Party ID
+#                    "S2_M", # Multiracial
+#                    
+#                    # Religion-related variables
+#                    "Q58R1",            # Catholic
+#                    "Q58R2",            # Protestant
+#                    "Q58R3",            # Other Christian
+#                    "Q58R4",            # Jewish
+#                    "Q58R5",            # Muslim
+#                    "Q58R6",            # Hindu
+#                    "Q58R7",            # Buddhist
+#                    "Q58R8",            # Mormon/LDS
+#                    "Q58R9",            # Folk Religion
+#                    "Q58R10",           # Daoist
+#                    "Q58R11",           # Shamanism
+#                    "Q58R12",           # Ancestral Religion
+#                    "Q58R13",           # Spiritual but not religious
+#                    "Q58R14",           # Atheist
+#                    "Q58R15",            # Agnostic
+#                    
+#                    # Variables capturing experiences with racism
+#                    "Q627", # Treated unfairly or discriminated against ,
+#                    "Q628", # Location of discrimination
+#                    "Q629R1" # Unfair treatment due to racial background/ethnicity
+#                    )
+#                  
+#                    
+#                    
+# # Create new df
+# data_filtered <- dataset[, required_cols_ord]
+# 
+# # Convert the required columns to factors
+# data_filtered[required_cols] <- lapply(data_filtered[required_cols_ord], as.factor)
+# 
+# # Convert individual religion columns to numeric
+# for (i in 1:15) {
+#   col_name <- paste0("Q58R", i, "_numeric")
+#   data_filtered[[col_name]] <- as.numeric(sub("^\\((\\d+)\\).*", "\\1", data_filtered[[paste0("Q58R", i)]]))
+# }
+# 
+# # Create a new Religion variable
+# data_filtered$Religion <- "None"
+# data_filtered$Religion[data_filtered$Q58R1_numeric == 1] <- "Catholic"
+# data_filtered$Religion[data_filtered$Q58R2_numeric == 1] <- "Protestant"
+# data_filtered$Religion[data_filtered$Q58R3_numeric == 1] <- "Other Christian"
+# data_filtered$Religion[data_filtered$Q58R4_numeric == 1] <- "Jewish"
+# data_filtered$Religion[data_filtered$Q58R5_numeric == 1] <- "Muslim"
+# data_filtered$Religion[data_filtered$Q58R6_numeric == 1] <- "Hindu"
+# data_filtered$Religion[data_filtered$Q58R7_numeric == 1] <- "Buddhist"
+# data_filtered$Religion[data_filtered$Q58R8_numeric == 1] <- "Mormon/LDS"
+# data_filtered$Religion[data_filtered$Q58R9_numeric == 1] <- "Folk Religion"
+# data_filtered$Religion[data_filtered$Q58R10_numeric == 1] <- "Daoist"
+# data_filtered$Religion[data_filtered$Q58R11_numeric == 1] <- "Shamanism"
+# data_filtered$Religion[data_filtered$Q58R12_numeric == 1] <- "Ancestral Religion"
+# data_filtered$Religion[data_filtered$Q58R13_numeric == 1] <- "Spiritual but not religious"
+# data_filtered$Religion[data_filtered$Q58R14_numeric == 1] <- "Atheist"
+# data_filtered$Religion[data_filtered$Q58R15_numeric == 1] <- "Agnostic"
+# 
+# # Convert Religion to a factor
+# data_filtered$Religion <- factor(data_filtered$Religion,
+#                                levels = c("Catholic", "Protestant", "Other Christian", "Jewish", "Muslim",
+#                                           "Hindu", "Buddhist", "Mormon/LDS", "Folk Religion",
+#                                           "Daoist", "Shamanism", "Ancestral Religion",
+#                                           "Spiritual but not religious", "Atheist", "Agnostic", "None"))
+# 
+# # Remove redundant religion columns (Q58R1 to Q58R15 and their numeric versions)
+# redundant_cols <- c(
+#   "Q58R1", "Q58R2", "Q58R3", "Q58R4", "Q58R5", "Q58R6", "Q58R7", "Q58R8", 
+#   "Q58R9", "Q58R10", "Q58R11", "Q58R12", "Q58R13", "Q58R14", "Q58R15",
+#   "Q58R1_numeric", "Q58R2_numeric", "Q58R3_numeric", "Q58R4_numeric", 
+#   "Q58R5_numeric", "Q58R6_numeric", "Q58R7_numeric", "Q58R8_numeric", 
+#   "Q58R9_numeric", "Q58R10_numeric", "Q58R11_numeric", "Q58R12_numeric", 
+#   "Q58R13_numeric", "Q58R14_numeric", "Q58R15_numeric"
+# )
+# 
+# # Remove redundant columns from the dataset
+# dataset_clean_ord <- data_filtered[, !(names(data_filtered) %in% redundant_cols)]
+# 
+# dataset_clean_ord <- dataset_clean_ord %>% mutate(educ = case_when(S13 == "(1) Grades 1-8" ~ "No college experience",
+#                                                            S13 == "(2) Some High School, but did not graduate" ~ "No college experience",
+#                                                            S13 == "(3) High School graduate or GED" ~ "No college experience",
+#                                                            S13 == "(4) Some college" ~ "Some college experience",
+#                                                            S13 == "(5) Associates, 2-year degree" ~ "College graduate",
+#                                                            S13 == "(6) Bachelors, 4-year degree" ~ "College graduate",
+#                                                            S13 == "(7) Post-graduate degree" ~ "College graduate"),
+#                                           sex_orientation = case_when(S3 == "(01) Straight, that is, not gay or lesbian" ~ "Straight",
+#                                                                      S3 == "(02) Gay or lesbian" ~ "Queer",
+#                                                                      S3 == "(03) Bisexual" ~ "Queer",
+#                                                                      S3 == "(04) Something else (Specify)" ~ "Queer",
+#                                                                      S3 == "(88) Refused" ~ "Refused/Don't Know",
+#                                                                      S3 == "(99) I don" ~ "Refused/Don't Know"),
+#                                           rural = case_when(S14 == "(1) Large urban area" ~ "Urban/suburban",
+#                                                             S14 == "(2) Large suburb near large city" ~ "Urban/suburban",
+#                                                             S14 == "(3) Small suburb near small town or city" ~ "Urban/suburban",
+#                                                             S14 == "(4) Small town or small city" ~ "Urban/suburban",
+#                                                             S14 == "(5) Rural area" ~ "Rural"),
+#                                           white = case_when(S2_RACE_PRIME == "(1) White" ~ "White", 
+#                                                             S2_RACE_PRIME == "(2) Hispanic or Latino" ~ "Non-White",
+#                                                             S2_RACE_PRIME == "(3) Black or African American" ~ "Non-White",
+#                                                             S2_RACE_PRIME == "(4) Asian American" ~ "Non-White",
+#                                                             S2_RACE_PRIME == "(5) American Indian/Native American" ~ "Non-White",
+#                                                             S2_RACE_PRIME == "(6) Arab, Middle Eastern or North African" ~ "Non-White",
+#                                                             S2_RACE_PRIME == "(7) Native Hawaiian" ~ "Non-White",
+#                                                             S2_RACE_PRIME == "(8) Not Hawaiian, but other Pacific Islander" ~ "Non-White"),
+#                                           race_recoded = case_when(S2_RACE_PRIME == "(1) White" ~ "White",
+#                                                                    S2_RACE_PRIME == "(2) Hispanic or Latino" ~ "Latino",
+#                                                                    S2_RACE_PRIME == "(3) Black or African American" ~ "Black",
+#                                                                    S2_RACE_PRIME == "(4) Asian American" ~ "Asian American",
+#                                                                    S2_RACE_PRIME == "(5) American Indian/Native American" ~ "AIAN",
+#                                                                    S2_RACE_PRIME == "(6) Arab, Middle Eastern or North African" ~ "Arab/MENA",
+#                                                                    S2_RACE_PRIME == "(7) Native Hawaiian" ~ "NHPI",
+#                                                                    S2_RACE_PRIME == "(8) Not Hawaiian, but other Pacific Islander" ~ "NHPI"),
+#                                           age_recoded = case_when(S5_AGE == "(1) 16-17" ~ "16-17",
+#                                                                   S5_AGE == "(2) 18-29" ~ "18-29",
+#                                                                   S5_AGE == "(3) 30-39" ~ "30-39",
+#                                                                   S5_AGE == "(4) 40-49" ~ "40-49",
+#                                                                   S5_AGE == "(5) 50-59" ~ "50-59",
+#                                                                   S5_AGE == "(6) 60-69" ~ "60-69",
+#                                                                   S5_AGE == "(7) 70 +" ~ "70+"))
+# 
+# dataset_clean_ord <- dataset_clean_ord %>% mutate(race_recoded = factor(dataset_clean_ord$race_recoded, levels = c("White", "Black", "Asian American", "Latino", "AIAN", "NHPI", "Arab/MENA")),
+#                                          age_recoded = factor(dataset_clean_ord$age_recoded, levels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+")),
+#                                          educ = factor(dataset_clean_ord$educ, levels = c("No college experience", "Some college experience", "College graduate")),
+#                                          white = factor(dataset_clean_ord$white, levels = c("White", "Non-White")),
+#                                          rural = factor(dataset_clean_ord$rural, levels= c("Rural", "Urban/suburban")),
+#                                          sex_orientation = factor(dataset_clean_ord$sex_orientation, levels = c("Straight", "Queer", "Refused/Don't Know")))
+# 
+# 
+# ```
+
 # Abstract
@@
 Then, we analyze how personal experiences with racism may influence how respondents value policies aimed at addressing racial and ethnic discrimination.
-Our findings suggest that racial group belonging is not a predictor of political interest, while gender identity, racial identity, and party affiliation may predict higher levels of political interest.
-Additionally, gender identity, racial identity, and personal experiences with discrimination may predict increased levels of support for anti-discrimination policies.
+Our findings suggest that racial group belonging is not associated with higher odds  of increased political interest, while gender identity, racial identity, and party affiliation may be associated with higher levels of political interest.
+Additionally, gender identity, racial identity, and personal experiences with discrimination may be associated with higher odds of increased levels of support for anti-discrimination policies.
 
@@
 
-## Data
-
 The 2020 Collaborative Multi-Racial Post-Election Survey (CMPS) dataset contains 17,545 rows, with each row representing an individual respondent's data, and 1,490 columns, each capturing a specific variable or question response.
 These variables include demographic information and responses to survey questions on political and social issues.
-
-## Variable Selection
 
@@
 
-## Model Fitting and Evaluation
+To model the outcome variable, interest in participating in politics, in the first research question, we will use an ordinal logistic regression with predictors including the importance of belonging to a racial group to one’s identity, race, age, educational attainment, and political party affiliation.
+The outcome variable has levels “Not at all interested in politics”, “Not that interested in politics”, “Somewhat interested”, and “Very interested in politics”.
+We will also include an interaction term between gender and the importance of belonging to a racial group to one’s identity to assess whether the effect of the importance of belonging to a racial group on one’s identity differs based on one’s gender identity.
+To assess the model we will use a confusion matrix to compare predicted and observed values, and compare the model’s accuracy and no information rate.
+Finally, we will conduct a Brant test to provide additional evidence regarding whether the proportional odds assumption was violated.
 
-### Research Question 1: How Racial Group Belonging Impacts Interest in Politics
-
-To model the outcome variable Q29 (interest in participating in politics), we used an ordinal logistic regression with predictors including Q271 (importance of belonging to a racial group to one’s identity), race, age, educational attainment, and political party affiliation.
-We also included an interaction term between gender and Q271 to assess whether the effect of the importance of belonging to a racial group on one’s identity differs based on one’s gender identity.
-To assess the model we used a confusion matrix to compare predicted and observed values.
-Finally, we used the Brant test to check for violations of the proportional odds assumption.
-
-### Research Question 2: How Experiences with Discrimination Affect the Prioritization of Anti-Discrimination Policies
-
-We used a binomial logistic regression to model the outcome variable, Q1R5 (importance placed on anti-discrimination policies), with predictors including race, gender identity, sexual orientation, age, education, community type, experiences with racism, and perception of racism for different groups.
-An interaction term between gender and race was excluded due to multicollinearity.
-Stepwise regression and model diagnostics, including VIF and AIC, ensured a stable and interpretable model.
+For the second research question, we will make use of a binomial logistic regression to model the outcome variable ’Importance placed on anti-discrimination policies’ with predictors including race, gender identity, sexual orientation, age, education, community type, experiences with racism, and perception of racism for different groups.
+Additionally, we will include an interaction term between gender and race to explore intersectionality theory and determine which model performs better.
+After selecting the better model, we will evaluate and assess its performance using various tests such as the confusion matrix, ROC, AUC, accuracy, and other relevant metrics.
 
@@
 
-For the first research question, the outcome variable Q29 (interest in participating in politics) was modeled using ordinal logistic regression.
-Q29, an ordinal variable captured responses to the question: “Some people are very interested in politics while other people can’t stand politics, how about you?
+For the first research question, the outcome variable, interest in participating in politics, was modeled using ordinal logistic regression.
+Interest in participating in politics, an ordinal variable, captured responses to the question: “Some people are very interested in politics while other people can’t stand politics, how about you?
 Are you…” with response options “Very interested in politics”, “Somewhat interested”, “Not that interested in politics”, and “Not at all interested in politics.” The overall sample reported high levels of interest in politics, with 43% of respondents reporting that they were somewhat interested and one in four (25%) saying they were very interested in politics.
 
-The model also included the predictor variable Q271, a categorical variable that captured responses to the question: “How important is being \[RACIAL GROUP\] to your identity?”, with response options “Extremely important”, “Very important”, “Moderately important”, “Slightly important”, and “Not at all important”.
+| Response to Q29                   | Count  | Share of Total Respondents |
+|-----------------------------------|--------|----------------------------|
+| Very interested in politics       | 4,306  | 25%                        |
+| Somewhat interested               | 7,612  | 43%                        |
+| Not that interested in politics   | 3,643  | 21%                        |
+| Not at all interested in politics | 1,984  | 11%                        |
+| Total                             | 17,545 | 100%                       |
+
+The model also included the predictor variable regarding the importance of belonging to a racial group to one’s identity, a categorical variable that captured responses to the question: “How important is being [RACIAL GROUP] to your identity?”, with response options “Extremely important”, “Very important”, “Moderately important”, “Slightly important”, and “Not at all important”.
 The respondent’s identified racial group was brought into the survey question based on their prior survey responses.
@@
 The model also included multiple predictor variables that control for the impact of other aspects of a respondent’s identity, socioeconomic status, and political preferences These included demographic variables such as the primary race reported by a respondent, their age, and gender, as well as a respondent’s reported level of educational attainment and political party affiliation.
-Finally, the model included an interaction term between gender and Q271 (the importance a respondent placed in belonging to their racial group) to investigate whether there are combined effects of gender and racial belonging.
+Finally, the model included an interaction term between gender and the importance a respondent placed in belonging to their racial group to investigate whether the relationship between racial belonging and political interest differs depending on gender.
+This interaction term allows us to evaluate if the effect of racial belonging on political interest is different for males, females, non-binary respondents, or respondents who specified a different gender identity.
 
-(Odds Ratio Plot)
+(Insert Odds Ratio Plot for RQ1)
 
-Controlling for age, race, gender identity, educational attainment, and party affiliation, respondents who considered belonging in their racial group to be slightly or moderately important to their identity were marginally more likely to report a higher level of interest in politics than those that considered belonging in their racial group not at all important to their identity, but these effects were not statistically significant.
-Surprisingly, respondents who considered belonging in their racial group to be very or extremely important to their identity were less likely to report higher levels of interest in politics than the reference category (those who did not place a high importance on belonging to their racial group), holding the other demographic variables constant.
-These results were statistically significant.
+Note: Some variable levels are excluded due to filtering for significant variables. To view the reference level for each variable group, please see the below table.
 
-Holding other variables constant, all racial groups in our sample reported increased odds of reporting a higher level of interest in politics compared to white respondents, except for Arab, Middle East & North African (MENA) respondents.
+INSERT TABLE
+
+Controlling for age, gender identity, educational attainment, and party affiliation, all racial groups in our sample reported lower odds of reporting a higher level of interest in politics compared to white respondents, except for Arab, Middle East & North African (MENA) respondents.
 However, the effects for respondents of Arab/MENA heritage was not statistically significant.
-Black respondents were 1.15 times more likely to report a higher level of interest in politics than White respondents.
-Latino individuals were 1.26 times more likely to report a higher level of interest in politics compared to white individuals.
+The odds of Black respondents being in a higher level of interest in politics were 0.87 times as high as White respondents (CI: 0.79-0.95).
+The odds of Latino respondents being in a higher level of interest in politics were 0.79 times as high as White respondents (CI: 0.72-0.87).
 These effects were even more pronounced for smaller racial/ethnic minority groups.
-Asian Americans were 1.68 times more likely to report a higher level of interest in politics compared to white respondents.
-Native American (1.49 times more likely) and Native Hawaiian and Pacific Islander respondents (1.78 times more likely) were also more likely to report a higher level of interest in politics compared to white respondents.
+Asian Americans’ odds of reporting a higher level of interest in politics were 0.60 times as high as that of White respondents (CI: 0.54-0.65).
 
-For women, the odds of reporting a higher level of interest in politics were 1.58 times higher than men, holding the other demographic variables constant.
-Non-binary respondents and respondents who specified a different gender identity were less likely to report higher levels of interest in politics than men.  For women and non-binary respondents, reporting that belonging to their racial group was very or extremely important to their racial identity appeared to increase their odds of reporting higher levels of interest in politics, holding other variables constant.
-However, none of the effects of the interaction term between gender and the importance a respondent placed in belonging to their racial group were statistically significant.
+Increases in educational attainment were associated with higher odds of reporting a higher level of interest in politics.
+Respondents with some college experience had 1.53 times higher odds of being in a higher level of interest in politics than respondents with no college experience (CI: 1.40-1.66).
+Respondents who were college graduates had 2.23 times higher odds of being in a higher level of interest in politics than respondents with no college experience (CI: 2.07-2.40).
+This represents the strongest positive effect in our model.
 
-Contrary to the resource model of participation, increases in educational attainment did not result in increased odds of reporting a higher level of interest in politics.
-Similarly, older respondents reported decreased odds of reporting a higher level of interest in politics compared to the reference category of individuals aged 18-29 years old.
+Respondents in older age groups had odds of being in a higher level of interest in politics that were higher than individuals ages 18-29 years old.
+Respondents ages 60-69 years old had 1.32 times higher odds of being in a higher level of interest in politics than respondents ages 18-29 years old (CI: 1.21-1.45).
+Respondents that were 70+ had 1.69 times higher odds of being in a higher level of interest in politics than respondents ages 18-29 years old (CI: 1.51-1.89).
 
-For respondents that identified as Democrats, holding other factors constant, their odds of reporting a higher level of interest in politics decreased 12% compared to Republican respondents.
-However, these results differed for respondents that identified as Independent or belonging to another party.
-For Independent respondents, their odds of reporting a higher level of interest in politics were 1.72 times higher than Republican respondents,  holding other variables constant.
-For respondents that identified with another party, the odds of reporting a higher level of interest in politics were 4.7 times higher than Republican respondents, holding other variables constant.
-This may indicate that individuals who identify with parties outside of the dominant two party system in the U.S.
-may already be more interested in politics than typical American respondents.
+Regarding gender differences, women showed significantly lower odds of political interest compared to men (OR: 0.63, CI: 0.54-0.73).
+The reference group for the interaction term of the importance of racial belonging and gender identity is male respondents who reported that their belonging to a specific racial group is not at all important to their identity.
+Respondents who said that their belonging to a specific racial group was “Very important” to their racial identity had 1.19 times higher odds of reporting a higher level of interest in politics than this reference group (CI: 1.03-1.38).
+Respondents who said that their belonging to a specific racial group was “Extremely important” to their racial identity had 1.64 times higher odds than this reference group (CI: 1.42-1.90).
 
@@
 The confusion matrix revealed that the model had a low level of precision, with an accuracy value of 0.4503.
-The sensitivity of the model was low for each class except for the response option of being “Somewhat interested” in politics, which had an associated sensitivity of (0.8973).
+The model’s accuracy is slightly higher than the No Information Rate (0.4339), indicating that the model does not perform much better than a random guess based on the distribution of respondents in various levels of the outcome variable, interest in politics.
+
+INSERT TABLE
+
+The values for the sensitivity and specificity of the model for each class are present in the above table.
+The sensitivity of the model was low (below 0.20) for the levels “Very interested in politics”, “Not that interested in politics”, and “Not at all interested in politics.”.
+The only level with a high sensitivity value was the response option of being “Somewhat interested” in politics, which had an associated sensitivity of 0.90.
 These results indicate that the model struggled to correctly identify respondents who were “Very interested”, “Not that interested”, or “Not at all interested” in politics.
-The specificity of the model was high for each class, except for the response option of being “Somewhat interested” in politics, indicating that the model classified a high rate of false positives for this class, and it was potentially overly biased to predicting that respondents were “Somewhat interested” in politics.
-
+The specificity of the model was high for each level, except for the response option of being “Somewhat interested” in politics (Specificity: 0.16), indicating that the model classified a high rate of false positives for this level, and it was potentially overly biased to predicting that respondents were “Somewhat interested” in politics.
 We also produced a confusion matrix for a multinomial model with the same predictor variables.
-The confusion matrices were very similar, providing evidence that the proportional odds assumption was not violated.
+The confusion matrices were very similar, providing evidence that there is not a sizable difference in performance between the ordinal model and the multinomial model.
 In addition, we conducted a Brant test.
-The Brant test result indicated that the proportional odds assumption was upheld.
+The Brant test result indicated that there is not strong evidence that the proportional odds assumption was violated.
 
@@
 
-For the second research question, the variable outcome variable Q1R5 (i.e.
-importance placed on stopping discrimination against racial/ethnic minorities) was modeled using binomial logistic regression, an appropriate method for binary outcomes.
-Despite the skewed distribution of the dependent variable—74.25% of respondents considered it "Not Important" as shown in fig.
-1—logistic regression is robust to such imbalances.
+For the second research question, the outcome variable ‘Importance placed on stopping discrimination against racial/ethnic minorities’ was modeled using binomial logistic regression.
+Despite the skewed distribution of the dependent variable—74.25% of respondents considered it "Not Important"—logistic regression is robust to such imbalances.
 
-(INSERT PIE CHART)
+To examine factors influencing the importance of anti-discrimination policies, the model included predictors like personal experiences with racism (measured via ‘Unfair treatment or discrimination’ and its impact on life satisfaction), alongside demographic variables such as primary race, sexual orientation, gender, age, education, and community type.
+Variables Q629R1–Q629R7, which captured specific experiences of unfair treatment (e.g., due to race or gender), were excluded due to substantial missing data (>52%).
+However, this was justified since the broader question on ‘unfair treatment or discrimination’ retained the key essence of experiences with racism.
 
-The model incorporated predictors to examine factors influencing the importance respondents placed on anti-discrimination policies.
-These included personal experiences with racism, captured by **Q627** (unfair treatment or discrimination) and **Q633** (impact of discrimination on life satisfaction), alongside demographic variables such as **S2_RACE_PRIME** (primary race), **S3** (sexual orientation), **S3B** (gender identity), **S5_AGE** (age), **S13** (education), and **S14** (community type: rural, suburban, urban).
-Variables **Q629R1–Q629R7** were excluded due to substantial missing data (\>52%).
-Nonetheless, because it overlapped significantly with **Q627 (see fig.
-2)**, the overall essence of experiences with racism was retained in a more general and complete form.
+To investigate the factors influencing the importance respondents placed on anti-discrimination policies, we developed two regression models.
+The first, termed the "base model," excluded interaction terms.
+The second, the "interaction model," included an interaction term between ‘race’ and ‘gender’ to explore the principles of intersectionality theory.
 
-(INSERT ODDS RATIO CHART)
+We assessed multicollinearity in both models by calculating the normalized Variance Inflation Factor (VIF) values.
+In all cases, the VIF values (APPENDIX TABLE have made a vif at the end of the R code) were consistently below 2, indicating minimal correlation among predictors, suggesting that both models are robust and not significantly affected by multicollinearity.
 
-To investigate the combined effects of gender and race and engage with intersectionality theory, an interaction term between Gender and Primary Race was initially included in the model.
-However, this introduced multicollinearity, as indicated by aliased coefficients and elevated Variance Inflation Factor (VIF) values, likely due to data sparsity and the underrepresentation of certain gender-race combinations, resulting in near-perfect linear dependencies.
-To resolve this, the interaction term was removed, and additional interactions were avoided to mitigate similar issues.
-Simplifying the model enhanced interpretability and reduced the risk of overfitting.
+To further assess the goodness-of-fit for the two models, a chi-square test was conducted with their residual degrees of freedom and deviance for were as follows:
+-   Base Model: Residual Degrees of Freedom = 17,487, Residual Deviance = 17,934
+-   Interaction Model: Residual Degrees of Freedom = 17,469, Residual Deviance = 17,913
 
-To further address high VIF values observed in predictors such as **Perceptions of Discrimination Against Latinos** (GVIF = 12.45) and **Perceptions of Discrimination Against Immigrants** (GVIF = 11.53), forward and backward stepwise regression was employed.
-This approach optimized model fit using the Akaike Information Criterion (AIC) while minimizing multicollinearity.
-The iterative process ensured that only the most significant predictors were retained, reducing VIF values and improving the model's stability.
+The difference in deviance between the two models was 21.145 with 18 degrees of freedom, indicating the interaction model to be slightly superior.
+However, the p-value of 0.2722 exceeded the 0.05 significance level, suggesting that including the interaction term did not significantly improve the model fit.
+Consequently, the base model was selected as it was more parsimonious with comparable explanatory power.
 
-Gender and race emerged as critical determinants in prioritizing anti-discrimination policies.
-Women were 1.25 times more likely than men to consider these policies important, while non-binary individuals exhibited 1.40 times higher odds, though this result was not statistically significant.
-Race had an even more pronounced impact; Black respondents had 3.55 times higher odds, Asian Americans had 2.12 times higher odds, Latinos had 1.78 times higher odds, Native Hawaiians/Pacific Islanders had 2.17 times higher odds, and Arab/MENA individuals had 2.67 times higher odds of prioritizing these policies compared to White respondents.
+The findings from the base model revealed that personal experiences with discrimination and certain demographic factors significantly influence the importance respondents place on policies aimed at addressing racial and ethnic discrimination.
 
-Personal experiences with racism further shaped support for these policies.
-Respondents who reported experiencing unfair treatment or discrimination were 1.12 times more likely to prioritize anti-discrimination efforts.
-The impact of discrimination on life satisfaction exhibited a notable gradient: individuals who reported no impact on their life satisfaction were 0.73 times less likely to consider these policies important, while those who experienced greater impacts demonstrated stronger support.
+```{r}
 
-Perceptions of discrimination also played a significant role.
-Respondents perceiving lower levels of discrimination against Black individuals had 0.48–0.65 times lower odds of prioritizing anti-discrimination policies, whereas perceptions of minimal or no discrimination against Whites were associated with 1.47–2.01 times higher odds of supporting such policies.
-Sexual orientation also influenced prioritization, with queer respondents having 1.15 times higher odds of supporting anti-discrimination policies compared to non-queer respondents.
+# Filter for significant variables (p-value < 0.05)
+significant_data <- analysis_table[analysis_table$`P-Value` < 0.05, ]
 
-The confusion matrix reveals that the model has high **sensitivity** (0.9488), thus accurately identifying respondents who consider policies "Not important." However, the low **specificity** (0.1683) indicates that it struggles to correctly classify those who view the policies as "Important." This is expected given the dataset's skew toward "Not important" (74.25%).
-The **AUC** of 0.7186 suggests moderate predictive ability, but the model's low specificity and imbalance in class distribution highlight areas for improvement.
+# Add a column for variable names to use in the plot (from the "Description" column)
+significant_data <- significant_data |>
+  mutate(Variable = Description)
+
+# Create the plot
+ggplot(significant_data, aes(x = `Odds Ratio`, y = Variable)) +
+  geom_point(size = 4, color = "blue") +  # Add points for odds ratios
+  geom_errorbarh(aes(xmin = `Lower Confidence Interval`, 
+                     xmax = `Upper Confidence Interval`), 
+                 height = 0.3, color = "black") +  # Add horizontal error bars
+  geom_vline(xintercept = 1, linetype = "dashed", color = "red", size = 0.7) +  # Add a vertical reference line at Odds Ratio = 1
+  labs(
+    title = "Odds Ratio of Significant Variables",
+    x = "Odds Ratio",
+    y = "Variable"
+  ) +
+  theme_minimal() +  # Use a minimal theme for better aesthetics
+  theme(
+    axis.text.y = element_text(size = 6, color = "black"),  # Improve y-axis text readability
+    axis.title.x = element_text(size = 12, face = "bold"),  # Highlight x-axis title
+    axis.title.y = element_text(size = 12, face = "bold"),  # Highlight y-axis title
+    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),  # Center the title
+    plot.margin = margin(1, 1, 1, 1, "cm"),  # Adjust margins
+    axis.text.x = element_text(angle = 45, hjust = 1, size = 8),  # Reduce x-axis text size
+    legend.text = element_text(size = 8),  # Reduce legend text size
+    legend.title = element_text(size = 9)  # Reduce legend title size
+  )
+
+# Save the plot with larger dimensions
+#ggsave("Significant_Variables_Plot_Enhanced.png", width = 40, height = 20)
+
+```
+
+As shown in the figure above, gender and race were the most consistent and influential predictors.
+Women had 1.25 times higher odds of prioritizing anti-discrimination policies than men, while Black respondents had 3.53 times higher odds than White respondents.
+Latino and Asian American respondents had 1.74 and 2.15 times higher odds, respectively, to emphasize such policies compared to their White counterparts.
+These results underscore the significance of addressing the unique experiences of marginalized groups, whose direct experiences with racism may shape their policy attitudes.
+
+Additionally, individuals who perceive higher levels of discrimination against racial or ethnic groups tend to prioritize anti-discrimination policies.
+For example, those who believed there was "a little" discrimination against Blacks were 45% less likely to prioritize these policies compared to those who saw "a lot" of discrimination, demonstrating that personal perceptions of racial injustice are key in shaping views on anti-discrimination measures.
+In contrast, education and sexual orientation did not have a statistically significant impact on prioritizing anti-discrimination policies.
+This reflects that, in this model, personal experiences of discrimination and perceptions of racial inequality have a more direct effect on attitudes toward such policies.
+While education and sexual orientation may influence political views in other contexts, they seem less relevant to attitudes on racial discrimination.
+The lack of significance for these variables may indicate that perceptions of racial or ethnic discrimination are stronger motivators for supporting anti-discrimination policies than other demographic factors.
+
+The base model's evaluation (as shown in Table MODEL) provides insights into how personal experiences with discrimination and demographic factors influence the prioritization of anti-discrimination policies.
+The model achieved an overall accuracy of 74.48%, which closely aligns with the No Information Rate (NIR) of 74.25%, indicating that its performance is not significantly better than predicting the majority class.
+
+From the confusion matrix in Table CONFUSION, it is evident that the model performs well in identifying respondents who did not prioritize anti-discrimination policies ("Not Important"), as reflected by its high specificity of 94.89%., However, it struggled with the sensitivity (recall) for the "Important" class, which was only 15.63%.
+This low sensitivity indicates that the model missed a large proportion of respondents who do prioritize these policies.
+Precision for the "Important" class was moderate at 51.46%, but the F1 score of 0.2397 highlights poor overall performance in identifying the minority class.
+The model's AUC (Area Under the Curve) value (see Fig.
+2) of 0.72 suggests that it has a moderate ability to discriminate between the two classes, though it still falls short of ideal performance.
+
+The balanced accuracy was 55.26%, reflecting the model’s challenges in handling class imbalance.
+Despite achieving statistical significance with McNemar’s test (p-value < 2e-16), the model’s predictions deviate considerably from the true data.
+In conclusion, while the model effectively identifies respondents who do not prioritize anti-discrimination policies, it performs poorly in predicting those who do.
+This imbalance signals the need for improvement in predicting the minority class.
 
@@
 In this analysis, we investigated whether valuing belonging in a racial group can influence a respondent’s interest in politics.
-We found that the importance an individual placed in belonging to their specific racial group did not have a significant effect on their interest in politics.
-Contrary to the resource model of participation, age and increases in educational attainment did not appear to result in increased odds of reporting a higher level of interest in politics.
-The most significant predictors of increased levels of interest in politics were an individual’s gender identity, racial identity, and political party affiliation.
+We found that the higher levels of importance an individual placed in belonging to their specific racial group (reporting that belonging to their group was “Very important” or “Extremely important”) had a significant positive effect on their interest in politics when compared to male respondents who did not place a high level of importance on belonging to their specific racial group.
+Further research is needed to understand how an individual’s racial belonging can impact their political participation while controlling for other demographic characteristics.
 
-Then, we analyzed how personal experiences with racism may influence how respondents value policies aimed at addressing racial and ethnic discrimination.
-Our findings underscore the intersectional influences of race, gender, and personal experiences with discrimination on public attitudes toward anti-discrimination policies.
-Furthermore, perceptions of discrimination against various groups highlight the complex social dynamics that shape policy prioritization.
-The results provide essential insights for tailoring advocacy and policy efforts to address racial and ethnic discrimination effectively.
+Following the resource model of participation, age and increases in educational attainment appeared to be associated with higher odds of reporting a higher level of interest in politics.
+The most significant positive predictors of increased levels of interest in politics were an individual’s age and level of educational attainment.
+In particular, individuals who graduated from college had much higher odds of reporting a higher level of interest in politics than individuals with no college experience.
+This follows popular findings that educational attainment can lead to increases in political participation [8], and that older Americans vote at higher rates than Americans ages 18-29 years old [9].
+
+For our second question, we analyzed how personal experiences with racism influence respondents' prioritization of policies addressing racial and ethnic discrimination.
+Our findings highlight the significant impact of race, gender, and lived experiences with discrimination, as well as the critical role of perceptions of racism against various groups in shaping public attitudes.
+These results underscore the complex social dynamics that drive policy preferences and provide valuable insights for designing targeted advocacy and policy efforts to address the unique experiences and needs of marginalized communities.
 
@@
 For the ordinal logistic regression, we used educational attainment as a proxy for socioeconomic status.
-This is a commonly used proxy measure in behavioral and social sciences \[8\].
+This is a commonly used proxy measure in behavioral and social sciences [10].
 Further research on the influence of socioeconomic status of political participation for racial/ethnic minorities and the use of other measures such as household income or wealth would help us understand if the aforementioned resource model of participation is indeed generalizable to other groups in the American electorate.
-
 The effect of race on political interest appeared to be more pronounced for members of very small racial/ethnic minority groups in our analysis such as Asian Americans, Native Americans, Native Hawaiians and Pacific Islanders.
 Both models had low specificity, suggesting that they struggled to correctly identify cases where respondents had low levels of political interest or respondents who considered anti-discrimination policies to be important.
+Similarly, in our second research question, while the model effectively identified respondents who do not prioritize anti-discrimination policies, it struggled to predict those who do.
+This imbalance underscores the challenges of accurately modeling the minority class—respondents who view anti-discrimination policies as important—and highlights the need for more nuanced approaches to capture these perspectives.
 
-These challenges and opportunities highlight the importance of disaggregated data and further research in partnership with small racial and ethnic minority communities within the United States.
-Future work on political participation and public opinion among racial and ethnic minorities could support researchers, policymakers, and advocates efforts to better understand and support these communities.
-
-# Appendix
+Overall, these challenges and opportunities highlight the importance of disaggregated data and further research in partnership with small racial and ethnic minority communities within the United States.
+Future work on political participation and public opinion among racial and ethnic minorities could support researchers, policymakers, and advocates in designing more inclusive policies and interventions.
+Additionally, incorporating qualitative methodologies, such as interviews or community-based participatory research, could provide richer context and amplify the voices of underrepresented communities.
+Longitudinal studies could also complement these efforts, capturing changes over time and offering deeper insights into the lived experiences of marginalized communities.
+By addressing these gaps, future research can better capture the diverse and intersectional realities of minority groups and contribute to more effective advocacy and policy efforts.
 
@@
 https://www.pewresearch.org/race-and-ethnicity/2020/08/20/facts-on-u-s-immigrants/
-
 2.
@@
 https://www.census.gov/library/stories/2021/08/improved-race-ethnicity-measures-reveal-united-states-population-much-more-multiracial.html
-
-3.  Brady, H. E., Verba, S., & Schlozman, K. L. (1995). Beyond Ses: A Resource Model of Political Participation. The American Political Science Review, 89(2), 271–294. https://doi.org/10.2307/2082425
-
+3.
+Brady, H.
+E., Verba, S., & Schlozman, K.
+L.
+(1995).
+Beyond Ses: A Resource Model of Political Participation.
+The American Political Science Review, 89(2), 271–294.
+https://doi.org/10.2307/2082425
 4.    Leighley, J. E. (2021). Strength in Numbers? Princeton University Press.
-
 5.
@@
 https://doi.org/10.1016/j.electstud.2022.102515
-
 6.
@@
 Perceived discrimination, political efficacy, and political participation in American Indian adults.
-*Frontiers in Political Science*, *6*.
+Frontiers in Political Science, 6.
 https://doi.org/10.3389/fpos.2024.1328521
-
 7.
@@
 Collaborative Multi-racial Post-election Survey (CMPS), United States, 2020.
-Inter-university Consortium for Political and Social Research \[distributor\], 2024-06-11.
+Inter-university Consortium for Political and Social Research [distributor], 2024-06-11.
 https://doi.org/10.3886/ICPSR39096.v1
+8.    Kim, Y. (2023). Absolutely Relative: How Education Shapes Voter Turnout in the United States. Social Indicators Research, 168(1-3), 447–469. https://doi.org/10.1007/s11205-023-03146-1
+9.
+McDonald, M.
+(n.d.).
+Voter Turnout Demographics [Review of Voter Turnout Demographics].
+U.S.
+Elections Project.
+Retrieved December 13, 2024, from https://www.electproject.org/election-data/voter-turnout-demographics
+10.   Oakes, M. (n.d.). Measuring Socioeconomic Status 1. Learning Objectives. https://obssr.od.nih.gov/sites/obssr/files/Measuring-Socioeconomic-Status.pdf
 
-8.  Oakes, M. (n.d.). Measuring Socioeconomic Status 1. Learning Objectives. https://obssr.od.nih.gov/sites/obssr/files/Measuring-Socioeconomic-Status.pdf
+# Appendix
+
+## VIF Table - Research Question 2
+
+```{r}
+# Check multicollinearity using VIF
+vif_base <- vif(base_model)
+vif_interaction <- vif(interaction_model)
+```
+
+## Table CONFUSION: Confusion Matrix for Base model
+|                                      |                   |               |
+|--------------------------------------|-------------------|---------------|
+| [**Predicted / Actual**]{.underline} | **Not Important** | **Important** |
+| **Not Important**                    | 12361             | 3812          |
+| **Important**                        | 666               | 706           |
+
+## Table MODEL: Base Model Statistics
+|                           | Metric  |
+|---------------------------|---------|
+| Sensitivity (Recall)      | 0.1563  |
+| Specificity               | 0.9489  |
+| Precision                 | 0.5146  |
+| F1                        | 0.2397  |
+| Balanced Accuracy         | 0.5526  |
+| No Information Rate (NIR) | 0.7425  |
+| Kappa                     | 0.1361  |
+| Mcnemar's Test P-Value    | \<2e-16 |

### 2024-12-15 peter_models.qmd
updating notebooks

@@
       x = "Odds Ratio",
-      y = "Variables"
+      y = "Variables",
+      caption = "Note: Some variable levels are excluded due to filtering for significant variables. \n To view the reference level for each variable group, please see the below table."
     )
@@
 ```
+
+Table of Reference Levels
+
+| Variable Group                              | Reference Level            |
+|---------------------------------------------|----------------------------|
+| Education                                   | No college experience      |
+| Gender                                      | Man                        |
+| Age                                         | 18-29 years old            |
+| Race                                        | White                      |
+| Importance of Racial Belonging              | Not At All Important       |
+| Interaction Term: Racial Belonging & Gender | Male, Not At All Important |
 

### 2024-12-15 draft_finalreport.qmd
updating final report

@@
   "Q1R5",
+
+  # Outcome Variable: Level of Political Interest
+  "Q29",
+
+  # Predictor: Importance of Racial Belonging
+  "Q271",
+
+  # Predictor: Party ID
+  "Q21",
 
@@
 # Data Cleaning and Recoding for dataset_new
-dataset_clean <- dataset_new |> 
+dataset_clean_allvars <- dataset_new |> 
   mutate(
@@
 
-# Take a look at the updated dataset
-head(dataset_clean)
-
 ```
@@
 
-dataset_clean <- dataset_clean[, !(names(dataset_clean) %in% redundant_columns)]
+dataset_clean <- dataset_clean_allvars[, !(names(dataset_clean_allvars) %in% redundant_columns)]
+```
+
+```{r echo=FALSE, results=FALSE}
+# Rename remaining columns in dataset_clean
+colnames(dataset_clean) <- c(
+  "ImportanceOfStoppingDiscriminationAgainstMinorities",  # Q1R5
+  "Gender",                                               # S3B
+  "TreatedUnfairlyOrDiscriminatedAgainst",                # Q627
+  "DiscriminationImpactOnLifeSatisfaction",              # Q633
+  "PerceptionsOfDiscriminationAgainstWhites",            # Q619_Q626R1
+  "PerceptionsOfDiscriminationAgainstBlacks",            # Q619_Q626R2
+  "PerceptionsOfDiscriminationAgainstAsianAmericans",    # Q619_Q626R3
+  "PerceptionsOfDiscriminationAgainstNativeAmericans",   # Q619_Q626R4
+  "PerceptionsOfDiscriminationAgainstImmigrants",        # Q619_Q626R5
+  "PerceptionsOfDiscriminationAgainstLatinos",           # Q619_Q626R6
+  "PerceptionsOfDiscriminationAgainstGaysAndLesbians",   # Q619_Q626R7
+  "PerceptionsOfDiscriminationAgainstMuslims",           # Q619_Q626R8
+  "Education",                                           # educ
+  "SexOrientation",                                      # sex_orientation
+  "Rural",                                               # rural
+  "Race",                                                # race_recoded
+  "Age"                                                  # age_recoded
+)
+
+```
+
+```{r echo=FALSE, results=FALSE}
+# Regression model - RQ2
+base_model <- glm(ImportanceOfStoppingDiscriminationAgainstMinorities ~ Gender +
+               TreatedUnfairlyOrDiscriminatedAgainst + 
+               DiscriminationImpactOnLifeSatisfaction +
+               PerceptionsOfDiscriminationAgainstWhites + 
+               PerceptionsOfDiscriminationAgainstBlacks +
+               PerceptionsOfDiscriminationAgainstAsianAmericans + 
+               PerceptionsOfDiscriminationAgainstNativeAmericans + 
+               PerceptionsOfDiscriminationAgainstImmigrants + 
+               PerceptionsOfDiscriminationAgainstLatinos + 
+               PerceptionsOfDiscriminationAgainstGaysAndLesbians + 
+               PerceptionsOfDiscriminationAgainstMuslims +
+               Education + 
+                 SexOrientation +
+                 Rural +
+                 Race +
+                 Age,
+
+             data = dataset_clean, family = "binomial")
+```
+
+```{r echo=FALSE, results=FALSE}
+# Regression Model with interaction term - RQ2
+interaction_model <- glm(ImportanceOfStoppingDiscriminationAgainstMinorities ~ 
+               TreatedUnfairlyOrDiscriminatedAgainst + 
+               DiscriminationImpactOnLifeSatisfaction +
+               PerceptionsOfDiscriminationAgainstWhites + 
+               PerceptionsOfDiscriminationAgainstBlacks +
+               PerceptionsOfDiscriminationAgainstAsianAmericans + 
+               PerceptionsOfDiscriminationAgainstNativeAmericans + 
+               PerceptionsOfDiscriminationAgainstImmigrants + 
+               PerceptionsOfDiscriminationAgainstLatinos + 
+               PerceptionsOfDiscriminationAgainstGaysAndLesbians + 
+               PerceptionsOfDiscriminationAgainstMuslims +
+               Education + 
+                 SexOrientation +
+                 Rural +
+                 Age + 
+                 Gender * Race,
+
+             data = dataset_clean, family = "binomial")
+```
+
+```{r echo=FALSE, results=FALSE}
+
+## Calculating Odds Ratio, Confidence Interval, and p-values (base model)
+
+# Extract odds ratios
+odds_ratios <- exp(coef(base_model))
+
+# Get confidence intervals for coefficients
+conf_intervals <- exp(confint(base_model))
+
+# Extract p-values
+p_values <- summary(base_model)$coefficients[, 4]
+
+# Combine everything into a single data frame
+results_base_model <- data.frame(
+  OddsRatio = odds_ratios,
+  LowerCI = conf_intervals[, 1],
+  UpperCI = conf_intervals[, 2],
+  PValue = p_values
+)
+
+# Display results
+results_base_model
 ```
@@
 
-```
-
-```{r echo=FALSE, results=FALSE}
-# Regression model - RQ2
-base_model <- glm(ImportanceOfStoppingDiscriminationAgainstMinorities ~ Gender +
-               TreatedUnfairlyOrDiscriminatedAgainst + 
-               DiscriminationImpactOnLifeSatisfaction +
-               PerceptionsOfDiscriminationAgainstWhites + 
-               PerceptionsOfDiscriminationAgainstBlacks +
-               PerceptionsOfDiscriminationAgainstAsianAmericans + 
-               PerceptionsOfDiscriminationAgainstNativeAmericans + 
-               PerceptionsOfDiscriminationAgainstImmigrants + 
-               PerceptionsOfDiscriminationAgainstLatinos + 
-               PerceptionsOfDiscriminationAgainstGaysAndLesbians + 
-               PerceptionsOfDiscriminationAgainstMuslims +
-               Education + 
-                 SexOrientation +
-                 Rural +
-                 Race +
-                 Age,
-
-             data = dataset_clean, family = "binomial")
-```
-
-```{r echo=FALSE, results=FALSE}
-# Regression Model with interaction term - RQ2
-interaction_model <- glm(ImportanceOfStoppingDiscriminationAgainstMinorities ~ 
-               TreatedUnfairlyOrDiscriminatedAgainst + 
-               DiscriminationImpactOnLifeSatisfaction +
-               PerceptionsOfDiscriminationAgainstWhites + 
-               PerceptionsOfDiscriminationAgainstBlacks +
-               PerceptionsOfDiscriminationAgainstAsianAmericans + 
-               PerceptionsOfDiscriminationAgainstNativeAmericans + 
-               PerceptionsOfDiscriminationAgainstImmigrants + 
-               PerceptionsOfDiscriminationAgainstLatinos + 
-               PerceptionsOfDiscriminationAgainstGaysAndLesbians + 
-               PerceptionsOfDiscriminationAgainstMuslims +
-               Education + 
-                 SexOrientation +
-                 Rural +
-                 Age + 
-                 Gender * Race,
-
-             data = dataset_clean, family = "binomial")
 ```
@@
 
-## VIF Table - Research Question 2
+## Research Question 1
+
+### Model Summary - OLR Model with Interactions
+```{r}
+dataset_clean_allvars$Q271_reordered <- factor(dataset_clean_allvars$Q271,
+                            levels = c("(5) Not at all important",
+                                     "(4) Slightly important",
+                                     "(3) Moderately important",
+                                     "(2) Very important",
+                                     "(1) Extremely important"))
+dataset_clean_allvars$Q29_reordered <- factor(dataset_clean_allvars$Q29,
+                            levels = c("(4) Not at all interested in politics",
+                                     "(3) Not that interested in politics",
+                                     "(2) Somewhat interested",
+                                     "(1) Very interested in politics"), ordered=TRUE)
+
+ord_mod_interactions <- polr(Q29_reordered ~ Q271_reordered * S3B +
+                               race_recoded +
+                  age_recoded +
+            educ +  # Educational Attainment
+            Q21, # Party affiliation
+          data = dataset_clean_allvars, Hess=TRUE)
+
+summary(ord_mod_interactions)
+
+pvals_interact <- pnorm(-abs(summary(ord_mod_interactions)$coef[,"t value"]))*2
+ctable_interact <- cbind(summary(ord_mod_interactions)$coef,pvals_interact)
+
+ctable_interact
+
+(exp_coefs_interact <- exp(cbind(OR=coef(ord_mod_interactions),confint(ord_mod_interactions))))
+
+head(ord_mod_interactions$fitted.values)
+head(predict(ord_mod_interactions))
+
+confusionMatrix(predict(ord_mod_interactions), dataset_clean$Q29, mode="everything")
+
+```
+
+### Odds Ratio Table 
+| Variable | Odds Ratio | 2.5% (Confidence Interval) | 97.5% (Confidence Interval) |
+|------------------|------------------|------------------|------------------|
+| Q271 (4) Slightly important | 0.9672589 | 0.811845068 | 1.1525850 |
+| Q271 (3) Moderately important | 0.9776608 | 0.845556718 | 1.1303745 |
+| Q271 (2) Very important | 1.1952004 | 1.033019394 | 1.3828451 |
+| Q271 (1) Extremely important | 1.6474969 | 1.423670208 | 1.9066091 |
+| S3B(2) Woman | 0.6294256 | 0.540015144 | 0.7336072 |
+| S3B(3) Non-binary | 1.8576658 | 1.002509318 | 3.4617409 |
+| S3B(4) Something else (Specify) | 2.1359493 | 0.377740111 | 13.5124525 |
+| race Black | 0.8676620 | 0.790376010 | 0.9524806 |
+| race Asian American | 0.5945609 | 0.541964870 | 0.6522143 |
+| race Latino | 0.7946622 | 0.575235415 | 0.7848896 |
+| race AIAN | 0.6719204 | 1.27404185 | 1.7384045 |
+| race NHPI | 0.5606006 | 0.458205702 | 0.6860792 |
+| race Arab/MENA | 1.1589043 | 0.899447122 | 1.4956228 |
+| age 30-39 | 1.0310801 | 0.946696738 | 1.0563119 |
+| age 40-49 | 0.9247623 | 0.98549367 | 1.1229868 |
+| age 50-59 | 1.0979487 | 1.000027913 | 1.2054892 |
+| age 60-69 | 1.3238874 | 1.205240800 | 1.4542874 |
+| age 70+ | 1.6896464 | 1.510226543 | 1.8907050 |
+| educSome college experience | 1.5264694 | 1.402573135 | 1.6614197 |
+| educCollege graduate | 2.2337286 | 2.072162224 | 2.4081069 |
+| Q21(2) Democrat | 1.1333513 | 1.045765624 | 1.2282275 |
+| Q21(3) Independent | 0.5806653 | 0.533801211 | 0.6315870 |
+| Q21(4) Other party | 0.2127592 | 0.182654104 | 0.2477198 |
+| Q271 4) Slightly important:S3B(2) Woman | 0.9955364 | 0.782564309 | 1.2663323 |
+| Q271 (3) Moderately important:S3B(2) Woman | 0.9461281 | 0.780040309 | 1.1475378 |
+| Q271 (2) Very important:S3B(2) Woman | 0.9791675 | 0.810674379 | 1.1826356 |
+| Q271 (1) Extremely important:S3B(2) Woman | 0.8114102 | 0.674275440 | 0.9763646 |
+| Q271 (4) Slightly important:S3B(3) Non-binary | 0.6505015 | 0.198565161 | 2.1505278 |
+| Q271 (3) Moderately important:S3B(3) Non-binary | 0.6563254 | 0.249594245 | 1.7277048 |
+| Q271 (2) Very important:S3B(3) Non-binary | 0.5140828 | 0.205132143 | 1.2905898 |
+| Q271 (1) Extremely important:S3B(3) Non-binary | 0.8770903 | 0.354410521 | 2.1835545 |
+| Q271 (4) Slightly important:S3B(4) Something else (Specify) | 0.2279084 | 0.023857021 | 2.0164000 |
+| Q271 (3) Moderately important:S3B(4) Something else (Specify) | 1.7158304 | 0.146865519 | 19.7516328 |
+| Q271 2) Very important:S3B(4) Something else (Specify) | 0.1049752 | 0.006213617 | 1.4210951 |
+| Q271 1) Extremely important:S3B(4) Something else (Specify) | 0.6584933 | 0.079780177 | 4.9850479 |
+
+## Research Question 2
+
+### VIF Table
 
@@
 ```
+
+### Odds Ratio Table
 
@@
 
-## Table MODEL: Base Model Statistics
+### Table MODEL: Base Model Statistics
 |                           | Metric  |

### 2024-12-15 draft_finalreport.qmd
updating draft of final

@@
 author: "Peter de Guzman, Jennifer Li, Nakiyah Dhariwala, Sizhe Chen"
-format: html
+format: pdf
 date: 2024-12-08
@@
                        "Q629R5", "Q629R6", "Q629R7",
-                       "S2_RACE_PRIME", "S13", "S3", "S5_AGE","S14")
+                       "S2_RACE_PRIME", "S13", "S3", "S5_AGE","S14", "Q29", "Q21")
 
@@
   "ImportanceOfStoppingDiscriminationAgainstMinorities",  # Q1R5
+  "RacialBelonging",                                      #Q271
   "Gender",                                               # S3B
@@
 
-# ordinal model
-
-# ```{r, echo=FALSE, results=FALSE}
-# # data load and cleaning for ordinal regression model
-# required_cols_ord <- c("Q29", # Outcome Variable; interest in politics
-#                    "Q271", # Predictor Variable; importance of race to identity
-#                    "S2_RACE_PRIME", # Race
-#                    "S3", # Orientation
-#                    "S3B", # Gender
-#                    "S5_AGE", # Age Category
-#                    "S13", # Educational Attainment
-#                    "S14", # Community Type
-#                    "S7", # Nativity
-#                    "Q21", # Party ID
-#                    "S2_M", # Multiracial
-#                    
-#                    # Religion-related variables
-#                    "Q58R1",            # Catholic
-#                    "Q58R2",            # Protestant
-#                    "Q58R3",            # Other Christian
-#                    "Q58R4",            # Jewish
-#                    "Q58R5",            # Muslim
-#                    "Q58R6",            # Hindu
-#                    "Q58R7",            # Buddhist
-#                    "Q58R8",            # Mormon/LDS
-#                    "Q58R9",            # Folk Religion
-#                    "Q58R10",           # Daoist
-#                    "Q58R11",           # Shamanism
-#                    "Q58R12",           # Ancestral Religion
-#                    "Q58R13",           # Spiritual but not religious
-#                    "Q58R14",           # Atheist
-#                    "Q58R15",            # Agnostic
-#                    
-#                    # Variables capturing experiences with racism
-#                    "Q627", # Treated unfairly or discriminated against ,
-#                    "Q628", # Location of discrimination
-#                    "Q629R1" # Unfair treatment due to racial background/ethnicity
-#                    )
-#                  
-#                    
-#                    
-# # Create new df
-# data_filtered <- dataset[, required_cols_ord]
-# 
-# # Convert the required columns to factors
-# data_filtered[required_cols] <- lapply(data_filtered[required_cols_ord], as.factor)
-# 
-# # Convert individual religion columns to numeric
-# for (i in 1:15) {
-#   col_name <- paste0("Q58R", i, "_numeric")
-#   data_filtered[[col_name]] <- as.numeric(sub("^\\((\\d+)\\).*", "\\1", data_filtered[[paste0("Q58R", i)]]))
-# }
-# 
-# # Create a new Religion variable
-# data_filtered$Religion <- "None"
-# data_filtered$Religion[data_filtered$Q58R1_numeric == 1] <- "Catholic"
-# data_filtered$Religion[data_filtered$Q58R2_numeric == 1] <- "Protestant"
-# data_filtered$Religion[data_filtered$Q58R3_numeric == 1] <- "Other Christian"
-# data_filtered$Religion[data_filtered$Q58R4_numeric == 1] <- "Jewish"
-# data_filtered$Religion[data_filtered$Q58R5_numeric == 1] <- "Muslim"
-# data_filtered$Religion[data_filtered$Q58R6_numeric == 1] <- "Hindu"
-# data_filtered$Religion[data_filtered$Q58R7_numeric == 1] <- "Buddhist"
-# data_filtered$Religion[data_filtered$Q58R8_numeric == 1] <- "Mormon/LDS"
-# data_filtered$Religion[data_filtered$Q58R9_numeric == 1] <- "Folk Religion"
-# data_filtered$Religion[data_filtered$Q58R10_numeric == 1] <- "Daoist"
-# data_filtered$Religion[data_filtered$Q58R11_numeric == 1] <- "Shamanism"
-# data_filtered$Religion[data_filtered$Q58R12_numeric == 1] <- "Ancestral Religion"
-# data_filtered$Religion[data_filtered$Q58R13_numeric == 1] <- "Spiritual but not religious"
-# data_filtered$Religion[data_filtered$Q58R14_numeric == 1] <- "Atheist"
-# data_filtered$Religion[data_filtered$Q58R15_numeric == 1] <- "Agnostic"
-# 
-# # Convert Religion to a factor
-# data_filtered$Religion <- factor(data_filtered$Religion,
-#                                levels = c("Catholic", "Protestant", "Other Christian", "Jewish", "Muslim",
-#                                           "Hindu", "Buddhist", "Mormon/LDS", "Folk Religion",
-#                                           "Daoist", "Shamanism", "Ancestral Religion",
-#                                           "Spiritual but not religious", "Atheist", "Agnostic", "None"))
-# 
-# # Remove redundant religion columns (Q58R1 to Q58R15 and their numeric versions)
-# redundant_cols <- c(
-#   "Q58R1", "Q58R2", "Q58R3", "Q58R4", "Q58R5", "Q58R6", "Q58R7", "Q58R8", 
-#   "Q58R9", "Q58R10", "Q58R11", "Q58R12", "Q58R13", "Q58R14", "Q58R15",
-#   "Q58R1_numeric", "Q58R2_numeric", "Q58R3_numeric", "Q58R4_numeric", 
-#   "Q58R5_numeric", "Q58R6_numeric", "Q58R7_numeric", "Q58R8_numeric", 
-#   "Q58R9_numeric", "Q58R10_numeric", "Q58R11_numeric", "Q58R12_numeric", 
-#   "Q58R13_numeric", "Q58R14_numeric", "Q58R15_numeric"
-# )
-# 
-# # Remove redundant columns from the dataset
-# dataset_clean_ord <- data_filtered[, !(names(data_filtered) %in% redundant_cols)]
-# 
-# dataset_clean_ord <- dataset_clean_ord %>% mutate(educ = case_when(S13 == "(1) Grades 1-8" ~ "No college experience",
-#                                                            S13 == "(2) Some High School, but did not graduate" ~ "No college experience",
-#                                                            S13 == "(3) High School graduate or GED" ~ "No college experience",
-#                                                            S13 == "(4) Some college" ~ "Some college experience",
-#                                                            S13 == "(5) Associates, 2-year degree" ~ "College graduate",
-#                                                            S13 == "(6) Bachelors, 4-year degree" ~ "College graduate",
-#                                                            S13 == "(7) Post-graduate degree" ~ "College graduate"),
-#                                           sex_orientation = case_when(S3 == "(01) Straight, that is, not gay or lesbian" ~ "Straight",
-#                                                                      S3 == "(02) Gay or lesbian" ~ "Queer",
-#                                                                      S3 == "(03) Bisexual" ~ "Queer",
-#                                                                      S3 == "(04) Something else (Specify)" ~ "Queer",
-#                                                                      S3 == "(88) Refused" ~ "Refused/Don't Know",
-#                                                                      S3 == "(99) I don" ~ "Refused/Don't Know"),
-#                                           rural = case_when(S14 == "(1) Large urban area" ~ "Urban/suburban",
-#                                                             S14 == "(2) Large suburb near large city" ~ "Urban/suburban",
-#                                                             S14 == "(3) Small suburb near small town or city" ~ "Urban/suburban",
-#                                                             S14 == "(4) Small town or small city" ~ "Urban/suburban",
-#                                                             S14 == "(5) Rural area" ~ "Rural"),
-#                                           white = case_when(S2_RACE_PRIME == "(1) White" ~ "White", 
-#                                                             S2_RACE_PRIME == "(2) Hispanic or Latino" ~ "Non-White",
-#                                                             S2_RACE_PRIME == "(3) Black or African American" ~ "Non-White",
-#                                                             S2_RACE_PRIME == "(4) Asian American" ~ "Non-White",
-#                                                             S2_RACE_PRIME == "(5) American Indian/Native American" ~ "Non-White",
-#                                                             S2_RACE_PRIME == "(6) Arab, Middle Eastern or North African" ~ "Non-White",
-#                                                             S2_RACE_PRIME == "(7) Native Hawaiian" ~ "Non-White",
-#                                                             S2_RACE_PRIME == "(8) Not Hawaiian, but other Pacific Islander" ~ "Non-White"),
-#                                           race_recoded = case_when(S2_RACE_PRIME == "(1) White" ~ "White",
-#                                                                    S2_RACE_PRIME == "(2) Hispanic or Latino" ~ "Latino",
-#                                                                    S2_RACE_PRIME == "(3) Black or African American" ~ "Black",
-#                                                                    S2_RACE_PRIME == "(4) Asian American" ~ "Asian American",
-#                                                                    S2_RACE_PRIME == "(5) American Indian/Native American" ~ "AIAN",
-#                                                                    S2_RACE_PRIME == "(6) Arab, Middle Eastern or North African" ~ "Arab/MENA",
-#                                                                    S2_RACE_PRIME == "(7) Native Hawaiian" ~ "NHPI",
-#                                                                    S2_RACE_PRIME == "(8) Not Hawaiian, but other Pacific Islander" ~ "NHPI"),
-#                                           age_recoded = case_when(S5_AGE == "(1) 16-17" ~ "16-17",
-#                                                                   S5_AGE == "(2) 18-29" ~ "18-29",
-#                                                                   S5_AGE == "(3) 30-39" ~ "30-39",
-#                                                                   S5_AGE == "(4) 40-49" ~ "40-49",
-#                                                                   S5_AGE == "(5) 50-59" ~ "50-59",
-#                                                                   S5_AGE == "(6) 60-69" ~ "60-69",
-#                                                                   S5_AGE == "(7) 70 +" ~ "70+"))
-# 
-# dataset_clean_ord <- dataset_clean_ord %>% mutate(race_recoded = factor(dataset_clean_ord$race_recoded, levels = c("White", "Black", "Asian American", "Latino", "AIAN", "NHPI", "Arab/MENA")),
-#                                          age_recoded = factor(dataset_clean_ord$age_recoded, levels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+")),
-#                                          educ = factor(dataset_clean_ord$educ, levels = c("No college experience", "Some college experience", "College graduate")),
-#                                          white = factor(dataset_clean_ord$white, levels = c("White", "Non-White")),
-#                                          rural = factor(dataset_clean_ord$rural, levels= c("Rural", "Urban/suburban")),
-#                                          sex_orientation = factor(dataset_clean_ord$sex_orientation, levels = c("Straight", "Queer", "Refused/Don't Know")))
-# 
-# 
-# ```
-
 # Abstract
@@
 
-The model also included the predictor variable regarding the importance of belonging to a racial group to one’s identity, a categorical variable that captured responses to the question: “How important is being [RACIAL GROUP] to your identity?”, with response options “Extremely important”, “Very important”, “Moderately important”, “Slightly important”, and “Not at all important”.
+The model also included the predictor variable regarding the importance of belonging to a racial group to one’s identity, a categorical variable that captured responses to the question: “How important is being \[RACIAL GROUP\] to your identity?”, with response options “Extremely important”, “Very important”, “Moderately important”, “Slightly important”, and “Not at all important”.
 The respondent’s identified racial group was brought into the survey question based on their prior survey responses.
@@
 
-(Insert Odds Ratio Plot for RQ1)
+```{r, echo=FALSE, results=FALSE}
+dataset_clean_allvars$Q271_reordered <- factor(dataset_clean_allvars$Q271,
+                            levels = c("(5) Not at all important",
+                                     "(4) Slightly important",
+                                     "(3) Moderately important",
+                                     "(2) Very important",
+                                     "(1) Extremely important"))
+dataset_clean_allvars$Q29_reordered <- factor(dataset_clean_allvars$Q29,
+                            levels = c("(4) Not at all interested in politics",
+                                     "(3) Not that interested in politics",
+                                     "(2) Somewhat interested",
+                                     "(1) Very interested in politics"), ordered=TRUE)
+
+ord_mod_interactions <- polr(Q29_reordered ~ Q271_reordered * S3B +
+                               race_recoded +
+                  age_recoded +
+            educ +  # Educational Attainment
+            Q21, # Party affiliation
+          data = dataset_clean_allvars, Hess=TRUE)
+
+summary(ord_mod_interactions)
+
+pvals_interact <- pnorm(-abs(summary(ord_mod_interactions)$coef[,"t value"]))*2
+ctable_interact <- cbind(summary(ord_mod_interactions)$coef,pvals_interact)
+
+ctable_interact
+
+(exp_coefs_interact <- exp(cbind(OR=coef(ord_mod_interactions),confint(ord_mod_interactions))))
+
+head(ord_mod_interactions$fitted.values)
+head(predict(ord_mod_interactions))
+
+confusionMatrix(predict(ord_mod_interactions), dataset_clean_allvars$Q29, mode="everything")
+
+brant(ord_mod_interactions)
+
+```
+
+```{r, echo=FALSE}
+odds_ratio_plot_labelled <- function(model) {
+  # Get odds ratios and CIs
+  or_data <- exp(cbind(OR=coef(model), confint(model)))
+
+  # Calculate p-values
+  p_values <- pnorm(-abs(summary(model)$coefficients[,"t value"]))*2
+
+  # Keep only significant variables
+  sig_indices <- p_values[names(coef(model))] < 0.05
+  or_data_sig <- or_data[sig_indices, ]
+
+  # Create a mapping for variable names with grouping prefixes
+  rename_map <- c(
+    # Education Group
+    "educCollege graduate" = "1. Education: College Graduate",
+    "educSome college experience" = "1. Education: Some College",
+
+    # Gender Group
+    "S3B(2) Woman" = "2. Gender: Woman",
+    "S3B(3) Non-binary" = "2. Gender: Non-binary",
+
+    # Age Group
+    "age_recoded70+" = "3. Age: 70+",
+    "age_recoded60-69" = "3. Age: 60-69",
+    "age_recoded50-59" = "3. Age: 50-59",
+
+    # Race Group
+    "race_recodedBlack" = "4. Race: Black",
+    "race_recodedAsian American" = "4. Race: Asian",
+    "race_recodedLatino" = "4. Race: Latino",
+    "race_recodedAIAN" = "4. Race: Native",
+    "race_recodedNHPI" = "4. Race: NHPI",
+
+    # Racial Belonging Group
+    "Q271_reordered(2) Very important" = "5. Racial Belonging: Very Important",
+    "Q271_reordered(1) Extremely important" = "5. Racial Belonging: Extremely Important",
+
+    # Party ID Group
+    "Q21(2) Democrat" = "6. Party ID: Democrat",
+    "Q21(3) Independent" = "6. Party ID: Independent",
+    "Q21(4) Other party" = "6. Party ID: Other",
+
+    # Interactions
+    "Q271_reordered(1) Extremely important:S3B(2) Woman" = "7. Interaction: Racial Belonging & Gender: Woman"
+  )
+
+  # Create a data frame for plotting
+  plot_data <- data.frame(
+    name = rename_map[rownames(or_data_sig)],
+    or = or_data_sig[,1],
+    ci_lower = or_data_sig[,2],
+    ci_upper = or_data_sig[,3],
+    stringsAsFactors = FALSE
+  )
+
+  # Custom sorting function
+  sort_by_group_and_name <- function(x) {
+    groups <- as.numeric(sub("^(\\d+)\\..+$", "\\1", x))
+    names <- sub("^\\d+\\.[^:]+: (.+)$", "\\1", x)
+    sort_df <- data.frame(
+      original = x,
+      group = groups,
+      name = names,
+      stringsAsFactors = FALSE
+    )
+    sort_df <- sort_df[order(sort_df$group, -xtfrm(sort_df$name)), ]
+    return(sort_df$original)
+  }
+
+  # Sort the data
+  plot_data$name <- factor(plot_data$name,
+                          levels = rev(sort_by_group_and_name(plot_data$name)))
+
+  # Create plot
+  ggplot(plot_data, aes(y = name)) +
+    geom_vline(xintercept = 1, linetype = "dashed", color = "red", size = 0.7) +
+    geom_point(aes(x = or), size = 4, color = "blue") +
+    geom_errorbarh(aes(xmin = ci_lower, xmax = ci_upper), height = 0.3, color = "black") +
+    theme_minimal() +
+    theme(
+      axis.text.y = element_text(size = 6, color = "black"),
+      axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
+      axis.title.x = element_text(size = 12, face = "bold"),
+      axis.title.y = element_text(size = 12, face = "bold"),
+      plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
+      plot.margin = margin(1, 1, 1, 1, "cm"),
+      panel.grid.major.y = element_line(color = "gray90"),
+      panel.grid.minor.y = element_blank()
+    ) +
+    labs(
+      title = "Significant Variables with Confidence Intervals",
+      x = "Odds Ratio",
+      y = "Variables"
+    )
+}
+
+# Generate the plot
+odds_ratio_plot_labelled(ord_mod_interactions)
+
+```
 
@@
 
-INSERT TABLE
+| Variable Group | Reference Level |
+|------------------------------------|------------------------------------|
+| 1\. Education | No college experience |
+| 2\. Gender | Man |
+| 3\. Age | 18-29 years old |
+| 4\. Race | White |
+| 5\. Racial Belonging | Not At All Important |
+| 6\. Party ID | Republican |
+| 7\. Interaction: Racial Belonging & Gender: Woman | Man, Not At All Important |
 
@@
 To examine factors influencing the importance of anti-discrimination policies, the model included predictors like personal experiences with racism (measured via ‘Unfair treatment or discrimination’ and its impact on life satisfaction), alongside demographic variables such as primary race, sexual orientation, gender, age, education, and community type.
-Variables Q629R1–Q629R7, which captured specific experiences of unfair treatment (e.g., due to race or gender), were excluded due to substantial missing data (>52%).
+Variables Q629R1–Q629R7, which captured specific experiences of unfair treatment (e.g., due to race or gender), were excluded due to substantial missing data (\>52%).
 However, this was justified since the broader question on ‘unfair treatment or discrimination’ retained the key essence of experiences with racism.
@@
 
-To further assess the goodness-of-fit for the two models, a chi-square test was conducted with their residual degrees of freedom and deviance for were as follows:
--   Base Model: Residual Degrees of Freedom = 17,487, Residual Deviance = 17,934
--   Interaction Model: Residual Degrees of Freedom = 17,469, Residual Deviance = 17,913
+To further assess the goodness-of-fit for the two models, a chi-square test was conducted with their residual degrees of freedom and deviance for were as follows: - Base Model: Residual Degrees of Freedom = 17,487, Residual Deviance = 17,934 - Interaction Model: Residual Degrees of Freedom = 17,469, Residual Deviance = 17,913
 
@@
 
-```{r}
+```{r, echo=FALSE}
 
@@
 The balanced accuracy was 55.26%, reflecting the model’s challenges in handling class imbalance.
-Despite achieving statistical significance with McNemar’s test (p-value < 2e-16), the model’s predictions deviate considerably from the true data.
+Despite achieving statistical significance with McNemar’s test (p-value \< 2e-16), the model’s predictions deviate considerably from the true data.
 In conclusion, while the model effectively identifies respondents who do not prioritize anti-discrimination policies, it performs poorly in predicting those who do.
@@
 In particular, individuals who graduated from college had much higher odds of reporting a higher level of interest in politics than individuals with no college experience.
-This follows popular findings that educational attainment can lead to increases in political participation [8], and that older Americans vote at higher rates than Americans ages 18-29 years old [9].
+This follows popular findings that educational attainment can lead to increases in political participation \[8\], and that older Americans vote at higher rates than Americans ages 18-29 years old \[9\].
 
@@
 For the ordinal logistic regression, we used educational attainment as a proxy for socioeconomic status.
-This is a commonly used proxy measure in behavioral and social sciences [10].
+This is a commonly used proxy measure in behavioral and social sciences \[10\].
 Further research on the influence of socioeconomic status of political participation for racial/ethnic minorities and the use of other measures such as household income or wealth would help us understand if the aforementioned resource model of participation is indeed generalizable to other groups in the American electorate.
@@
 By addressing these gaps, future research can better capture the diverse and intersectional realities of minority groups and contribute to more effective advocacy and policy efforts.
+
+\newpage
 
@@
 https://www.census.gov/library/stories/2021/08/improved-race-ethnicity-measures-reveal-united-states-population-much-more-multiracial.html
-3.
-Brady, H.
-E., Verba, S., & Schlozman, K.
-L.
-(1995).
-Beyond Ses: A Resource Model of Political Participation.
-The American Political Science Review, 89(2), 271–294.
-https://doi.org/10.2307/2082425
+3.  Brady, H. E., Verba, S., & Schlozman, K. L. (1995). Beyond Ses: A Resource Model of Political Participation. The American Political Science Review, 89(2), 271–294. https://doi.org/10.2307/2082425
 4.  Leighley, J. E. (2021). Strength in Numbers? Princeton University Press.
@@
 Collaborative Multi-racial Post-election Survey (CMPS), United States, 2020.
-Inter-university Consortium for Political and Social Research [distributor], 2024-06-11.
+Inter-university Consortium for Political and Social Research \[distributor\], 2024-06-11.
 https://doi.org/10.3886/ICPSR39096.v1
@@
 (n.d.).
-Voter Turnout Demographics [Review of Voter Turnout Demographics].
+Voter Turnout Demographics \[Review of Voter Turnout Demographics\].
 U.S.
@@
 
+\newpage
+
 # Appendix
@@
 
-### Model Summary - OLR Model with Interactions
-```{r}
-dataset_clean_allvars$Q271_reordered <- factor(dataset_clean_allvars$Q271,
-                            levels = c("(5) Not at all important",
-                                     "(4) Slightly important",
-                                     "(3) Moderately important",
-                                     "(2) Very important",
-                                     "(1) Extremely important"))
-dataset_clean_allvars$Q29_reordered <- factor(dataset_clean_allvars$Q29,
-                            levels = c("(4) Not at all interested in politics",
-                                     "(3) Not that interested in politics",
-                                     "(2) Somewhat interested",
-                                     "(1) Very interested in politics"), ordered=TRUE)
+### Ordinal Model Overall Statistics
 
-ord_mod_interactions <- polr(Q29_reordered ~ Q271_reordered * S3B +
-                               race_recoded +
-                  age_recoded +
-            educ +  # Educational Attainment
-            Q21, # Party affiliation
-          data = dataset_clean_allvars, Hess=TRUE)
+|                           | Metric           |
+|---------------------------|------------------|
+| Accuracy                  | 0.4503           |
+| 95% CI                    | (0.4429, 0.4577) |
+| No Information Rate (NIR) | 0.4339           |
+| P Value \[Acc \> NIR\]    | 6.105e-06        |
+| Kappa                     | 0.0787           |
+| Mcnemar's Test P-Value    | \<2.2e-16        |
 
-summary(ord_mod_interactions)
+### Confusion Matrix
 
-pvals_interact <- pnorm(-abs(summary(ord_mod_interactions)$coef[,"t value"]))*2
-ctable_interact <- cbind(summary(ord_mod_interactions)$coef,pvals_interact)
+|   | **Actual - Very interested in politics** | **Actual -Somewhat interested** | **Actual - Not that interested in politics** | **Actual - Not at all interested in politics** |
+|---------------|---------------|---------------|---------------|---------------|
+| **Predicted - Very interested in politics** | 650 | 505 | 120 | 31 |
+| **Predicted - Somewhat interested** | 3567 | 6830 | 3202 | 1550 |
+| **Predicted - Not that interested in politics** | 57 | 181 | 203 | 186 |
+| **Predicted - Not at all interested in politics** | 32 | 96 | 118 | 217 |
 
-ctable_interact
+### Statistics by Class:
 
-(exp_coefs_interact <- exp(cbind(OR=coef(ord_mod_interactions),confint(ord_mod_interactions))))
-
-head(ord_mod_interactions$fitted.values)
-head(predict(ord_mod_interactions))
-
-confusionMatrix(predict(ord_mod_interactions), dataset_clean$Q29, mode="everything")
-
-```
+|   | Very interested in politics | Somewhat interested | Not that interested in politics | Not at all interested in politics |
+|---------------|---------------|---------------|---------------|---------------|
+| Sensitivity | 0.15095 | 0.8973 | 0.05572 | 0.10938 |
+| Specificity | 0.95045 | 0.1625 | 0.96950 | 0.98419 |
+| Balanced Accuracy | 0.55070 | 0.5299 | 0.51261 | 0.54678 |
 
 ### Odds Ratio Table
+
 | Variable | Odds Ratio | 2.5% (Confidence Interval) | 97.5% (Confidence Interval) |
@@
 
-```{r}
+```{r, echo=FALSE, results=FALSE}
 # Check multicollinearity using VIF
@@
 
+|   | Normalized VIF: Base Model | Normalized VIF: Interaction Model |
+|------------------------|------------------------|------------------------|
+| Gender | 1.022 | 1.978 |
+| Treated Unfairly or Discriminated Against | 1.125 | 1.126 |
+| Discrimination Impact on Life Satisfaction | 1.036 | 1.037 |
+| Perceptions Of Discrimination Against Whites | 1.105 | 1.106 |
+| Perceptions Of Discrimination Against Blacks | 1.329 | 1.329 |
+| Perceptions Of Discrimination Against Asian Americans | 1.308 | 1.308 |
+| Perceptions Of Discrimination Against Native Americans | 1.224 | 1.224 |
+| Perceptions Of Discrimination Against Immigrants | 1.357 | 1.359 |
+| Perceptions Of Discrimination Against Latinos | 1.370 | 1.371 |
+| Perceptions Of Discrimination Against Gays and Lesbians | 1.262 | 1.262 |
+| Perceptions Of Discrimination Against Muslims | 1.341 | 1.342 |
+| Education | 1.058 | 1.060 |
+| Orientation | 1.042 | 1.044 |
+| Community Type | 1.017 | 1.018 |
+| Age | 1.028 | 1.030 |
+| Race | 1.027 | 1.725 |
+| Gender \* Race |  | 1.308 |
+
 ### Odds Ratio Table
 
+```{r}
+analysis_table
+```
+
 ## Table CONFUSION: Confusion Matrix for Base model
+
 |                                      |                   |               |
@@
 ### Table MODEL: Base Model Statistics
+
 |                           | Metric  |

### 2024-12-16 Jennifer_model.qmd

---
title: "Jennifer_Model"
format: html
author: "Jennifer Li"
editor: visual
---

Conducting modelling on the 2020 CMPS dataset for Research Question 1

# Packages

```{r}
library(tidyverse)
library(dplyr)
```

# Loading data

```{r, echo=FALSE}
rm(list=ls())
file_path <- "/Users/jenniferli/Desktop/39096-0001-Data.rda"
# Load data 
load(file_path)
data <- da39096.0001
rm(da39096.0001)
```

# Selecting Variables

Need to include the outcome variable, which is an ordinal variable Q29. Main predictor variable is Q271 "How important is being \[RACIAL GROUP\] to your identity?"

```{r}
required_cols <- c("Q29", # Outcome Variable; interest in politics
                   "Q271", # Predictor Variable; importance of race to identity
                   "S2_RACE_PRIME", # Race
                   "S3", # Orientation
                   "S3B", # Gender
                   "S5_AGE", # Age Category
                   "S13", # Educational Attainment
                   "S14", # Community Type
                   "S7", # Nativity
                   "Q21", # Party ID
                   "S2_M", # Multiracial

                   # Religion-related variables
                   "Q58R1",            # Catholic
                   "Q58R2",            # Protestant
                   "Q58R3",            # Other Christian
                   "Q58R4",            # Jewish
                   "Q58R5",            # Muslim
                   "Q58R6",            # Hindu
                   "Q58R7",            # Buddhist
                   "Q58R8",            # Mormon/LDS
                   "Q58R9",            # Folk Religion
                   "Q58R10",           # Daoist
                   "Q58R11",           # Shamanism
                   "Q58R12",           # Ancestral Religion
                   "Q58R13",           # Spiritual but not religious
                   "Q58R14",           # Atheist
                   "Q58R15",            # Agnostic

                   # Variables capturing experiences with racism
                   "Q627", # Treated unfairly or discriminated against ,
                   "Q628", # Location of discrimination
                   "Q629R1" # Unfair treatment due to racial background/ethnicity
                   )

# Create new df
data_filtered <- data[, required_cols]
head(data_filtered)

```

# Cleaning Variables

```{r}
# Convert the required columns to factors
data_filtered[required_cols] <- lapply(data_filtered[required_cols], as.factor)

# Verify that the columns have been successfully converted to factors
str(data_filtered[required_cols])
head(data_filtered)
```

Cleaning and Recoding Religion Variables

```{r}
# Convert individual religion columns to numeric
for (i in 1:15) {
  col_name <- paste0("Q58R", i, "_numeric")
  data_filtered[[col_name]] <- as.numeric(sub("^\\((\\d+)\\).*", "\\1", data_filtered[[paste0("Q58R", i)]]))
}

# Create a new Religion variable
data_filtered$Religion <- "None"
data_filtered$Religion[data_filtered$Q58R1_numeric == 1] <- "Catholic"
data_filtered$Religion[data_filtered$Q58R2_numeric == 1] <- "Protestant"
data_filtered$Religion[data_filtered$Q58R3_numeric == 1] <- "Other Christian"
data_filtered$Religion[data_filtered$Q58R4_numeric == 1] <- "Jewish"
data_filtered$Religion[data_filtered$Q58R5_numeric == 1] <- "Muslim"
data_filtered$Religion[data_filtered$Q58R6_numeric == 1] <- "Hindu"
data_filtered$Religion[data_filtered$Q58R7_numeric == 1] <- "Buddhist"
data_filtered$Religion[data_filtered$Q58R8_numeric == 1] <- "Mormon/LDS"
data_filtered$Religion[data_filtered$Q58R9_numeric == 1] <- "Folk Religion"
data_filtered$Religion[data_filtered$Q58R10_numeric == 1] <- "Daoist"
data_filtered$Religion[data_filtered$Q58R11_numeric == 1] <- "Shamanism"
data_filtered$Religion[data_filtered$Q58R12_numeric == 1] <- "Ancestral Religion"
data_filtered$Religion[data_filtered$Q58R13_numeric == 1] <- "Spiritual but not religious"
data_filtered$Religion[data_filtered$Q58R14_numeric == 1] <- "Atheist"
data_filtered$Religion[data_filtered$Q58R15_numeric == 1] <- "Agnostic"

# Convert Religion to a factor
data_filtered$Religion <- factor(data_filtered$Religion,
                               levels = c("Catholic", "Protestant", "Other Christian", "Jewish", "Muslim",
                                          "Hindu", "Buddhist", "Mormon/LDS", "Folk Religion",
                                          "Daoist", "Shamanism", "Ancestral Religion",
                                          "Spiritual but not religious", "Atheist", "Agnostic", "None"))

# Create a table of counts for the new Religion variable
table(data_filtered$Religion)

```

Removing deprecated Religion variables

```{r}
# Remove redundant religion columns (Q58R1 to Q58R15 and their numeric versions)
redundant_cols <- c(
  "Q58R1", "Q58R2", "Q58R3", "Q58R4", "Q58R5", "Q58R6", "Q58R7", "Q58R8", 
  "Q58R9", "Q58R10", "Q58R11", "Q58R12", "Q58R13", "Q58R14", "Q58R15",
  "Q58R1_numeric", "Q58R2_numeric", "Q58R3_numeric", "Q58R4_numeric", 
  "Q58R5_numeric", "Q58R6_numeric", "Q58R7_numeric", "Q58R8_numeric", 
  "Q58R9_numeric", "Q58R10_numeric", "Q58R11_numeric", "Q58R12_numeric", 
  "Q58R13_numeric", "Q58R14_numeric", "Q58R15_numeric"
)

# Remove redundant columns from the dataset
dataset_clean <- data_filtered[, !(names(data_filtered) %in% redundant_cols)]

# Take a look at the updated dataset
head(dataset_clean)
```

Cleaning and Recoding Other Variables

```{r}
dataset_clean <- dataset_clean %>% mutate(educ = case_when(S13 == "(1) Grades 1-8" ~ "No college experience",
                                                           S13 == "(2) Some High School, but did not graduate" ~ "No college experience",
                                                           S13 == "(3) High School graduate or GED" ~ "No college experience",
                                                           S13 == "(4) Some college" ~ "Some college experience",
                                                           S13 == "(5) Associates, 2-year degree" ~ "College graduate",
                                                           S13 == "(6) Bachelors, 4-year degree" ~ "College graduate",
                                                           S13 == "(7) Post-graduate degree" ~ "College graduate"),
                                          sex_orientation = case_when(S3 == "(01) Straight, that is, not gay or lesbian" ~ "Straight",
                                                                     S3 == "(02) Gay or lesbian" ~ "Queer",
                                                                     S3 == "(03) Bisexual" ~ "Queer",
                                                                     S3 == "(04) Something else (Specify)" ~ "Queer",
                                                                     S3 == "(88) Refused" ~ "Refused/Don't Know",
                                                                     S3 == "(99) I don" ~ "Refused/Don't Know"),
                                          rural = case_when(S14 == "(1) Large urban area" ~ "Urban/suburban",
                                                            S14 == "(2) Large suburb near large city" ~ "Urban/suburban",
                                                            S14 == "(3) Small suburb near small town or city" ~ "Urban/suburban",
                                                            S14 == "(4) Small town or small city" ~ "Urban/suburban",
                                                            S14 == "(5) Rural area" ~ "Rural"),
                                          white = case_when(S2_RACE_PRIME == "(1) White" ~ "White", 
                                                            S2_RACE_PRIME == "(2) Hispanic or Latino" ~ "Non-White",
                                                            S2_RACE_PRIME == "(3) Black or African American" ~ "Non-White",
                                                            S2_RACE_PRIME == "(4) Asian American" ~ "Non-White",
                                                            S2_RACE_PRIME == "(5) American Indian/Native American" ~ "Non-White",
                                                            S2_RACE_PRIME == "(6) Arab, Middle Eastern or North African" ~ "Non-White",
                                                            S2_RACE_PRIME == "(7) Native Hawaiian" ~ "Non-White",
                                                            S2_RACE_PRIME == "(8) Not Hawaiian, but other Pacific Islander" ~ "Non-White"),
                                          race_recoded = case_when(S2_RACE_PRIME == "(1) White" ~ "White",
                                                                   S2_RACE_PRIME == "(2) Hispanic or Latino" ~ "Latino",
                                                                   S2_RACE_PRIME == "(3) Black or African American" ~ "Black",
                                                                   S2_RACE_PRIME == "(4) Asian American" ~ "Asian American",
                                                                   S2_RACE_PRIME == "(5) American Indian/Native American" ~ "AIAN",
                                                                   S2_RACE_PRIME == "(6) Arab, Middle Eastern or North African" ~ "Arab/MENA",
                                                                   S2_RACE_PRIME == "(7) Native Hawaiian" ~ "NHPI",
                                                                   S2_RACE_PRIME == "(8) Not Hawaiian, but other Pacific Islander" ~ "NHPI"))

```

# Analysis of Missing Data

```{r}
# Check the number of missing values for each variable
missing_summary <- colSums(is.na(dataset_clean))
missing_summary

# Finding proportion of Missing Data
missing_proportion <- colSums(is.na(dataset_clean)) / nrow(dataset_clean)
missing_proportion

# Finding total missing data proportion
overall_missing <- sum(is.na(dataset_clean)) / (nrow(dataset_clean) * ncol(dataset_clean))
overall_missing
```

There is overall 8.3% missing values across all cells in this dataset. This is largely driven by two variables:

-   Q628: (IF S7=2 or 3 & Q627=1\] Did you experience that discrimination here in the United States, or in your home country?)

-   and Q629R1: In your opinion, were you unfairly treated because of your (Racial background or ethnicity)

For Q628, there are 15,709 (89.53%) missing values. For Q629R1 there are 9,180 (52.32%) missing values. We may need to consider not including these variables, because they may reduce our usable sample size.

```{r}
glimpse(dataset_clean)
```

# Looking at Outcome and main Predictor variable

```{r}
# Outcome - Q29
table(dataset_clean$Q29) / nrow(dataset_clean)
# Main Predictor - Q271
table(dataset_clean$Q271) / nrow(dataset_clean)
```

Table of Q29 for full sample - Interest in Politics

| Level of Interest in Politics     | Percent of Total Respondents |
|-----------------------------------|------------------------------|
| Very interested in politics       | 24.54%                       |
| Somewhat interested               | 43.39%                       |
| Not that interested in politics   | 20.76%                       |
| Not at all interested in politics | 11.31%                       |

Table of Q271 for full sample - How important is being a member of \[RACIAL GROUP\] to your identity?

| Importance of Race to Identity | Percent of Total Respondents |
|--------------------------------|------------------------------|
| Extremely Important            | 31.27%                       |
| Very Important                 | 24.68%                       |
| Moderately Important           | 21.40%                       |
| Slightly Important             | 8.58%                        |
| Not at all important           | 14.07%                       |

# Counts and proportion for every variable

```{r}
lapply(dataset_clean[, -1], table)

```

# Creating summary table

```{r}

# Function to create a summary table with counts and proportions
create_summary_table <- function(variable, var_name) {
  counts <- table(variable)
  if (length(counts) == 0) {
    return(data.frame(Variable = character(), Level = character(), Count = numeric(), Proportion = numeric()))
  }
  proportions <- prop.table(counts)
  summary_df <- data.frame(
    Variable = var_name,
    Level = names(counts),
    Count = as.numeric(counts),
    Proportion = round(as.numeric(proportions), 4)
  )
  return(summary_df)
}

# Apply the function to each variable of interest in the cleaned dataset
categorical_vars <- names(dataset_clean)

# Initialize an empty data frame to store all summary tables
full_summary_table <- data.frame(Variable = character(), Level = character(), Count = numeric(), Proportion = numeric())

# Iterate through each variable and generate the summary table
for (var in categorical_vars) {
  summary_table <- create_summary_table(dataset_clean[[var]], var)
  full_summary_table <- rbind(full_summary_table, summary_table)
}

# Print the full summary table
full_summary_table

```

| **Variable** | **Level** | **Count** | **Proportion** |
|------------------|-------------------|------------------|------------------|
| Q29 | \(1\) Very interested in politics | 4306 | 25% |
| Q29 | \(2\) Somewhat interested | 7612 | 43% |
| Q29 | \(3\) Not that interested in politics | 3643 | 21% |
| Q29 | \(4\) Not at all interested in politics | 1984 | 11% |
| Q271 | \(1\) Extremely important | 5487 | 31% |
| Q271 | \(2\) Very important | 4330 | 25% |
| Q271 | \(3\) Moderately important | 3754 | 21% |
| Q271 | \(4\) Slightly important | 1505 | 9% |
| Q271 | \(5\) Not at all important | 2469 | 14% |
| S2_RACE_PRIME | \(1\) White | 3923 | 22% |
| S2_RACE_PRIME | \(2\) Hispanic or Latino | 3873 | 22% |
| S2_RACE_PRIME | \(3\) Black or African American | 4613 | 26% |
| S2_RACE_PRIME | \(4\) Asian American | 3826 | 22% |
| S2_RACE_PRIME | \(5\) American Indian/Native American | 713 | 4% |
| S2_RACE_PRIME | \(6\) Arab, Middle Eastern or North African | 226 | 1% |
| S2_RACE_PRIME | \(7\) Native Hawaiian | 106 | 1% |
| S2_RACE_PRIME | \(8\) Not Hawaiian, but other Pacific Islander | 255 | 1% |
| S3 | \(01\) Straight, that is, not gay or lesbian | 15404 | 88% |
| S3 | \(02\) Gay or lesbian | 594 | 3% |
| S3 | \(03\) Bisexual | 937 | 5% |
| S3 | \(04\) Something else (Specify) | 286 | 2% |
| S3 | \(88\) Refused | 195 | 1% |
| S3 | \(99\) I don | 129 | 1% |
| S3B | \(1\) Man | 7431 | 42% |
| S3B | \(2\) Woman | 9932 | 57% |
| S3B | \(3\) Non-binary | 148 | 1% |
| S3B | \(4\) Something else (Specify) | 34 | 0% |
| S5_AGE | \(2\) 18-29 | 4026 | 23% |
| S5_AGE | \(3\) 30-39 | 3614 | 21% |
| S5_AGE | \(4\) 40-49 | 2783 | 16% |
| S5_AGE | \(5\) 50-59 | 2717 | 15% |
| S5_AGE | \(6\) 60-69 | 2745 | 16% |
| S5_AGE | \(7\) 70 + | 1660 | 9% |
| S13 | \(1\) Grades 1-8 | 126 | 1% |
| S13 | \(2\) Some High School, but did not graduate | 596 | 3% |
| S13 | \(3\) High School graduate or GED | 3227 | 18% |
| S13 | \(4\) Some college | 3689 | 21% |
| S13 | \(5\) Associates, 2-year degree | 2017 | 12% |
| S13 | \(6\) Bachelors, 4-year degree | 4625 | 26% |
| S13 | \(7\) Post-graduate degree | 3265 | 19% |
| S7 | \(1\) United States | 13297 | 76% |
| S7 | \(2\) Another country | 3962 | 23% |
| S7 | \(3\) on the Island of Puerto Rico | 286 | 2% |
| Q21 | \(1\) Republican | 3238 | 18% |
| Q21 | \(2\) Democrat | 8318 | 47% |
| Q21 | \(3\) Independent | 5174 | 29% |
| Q21 | \(4\) Other party | 815 | 5% |
| S2_M | \(1\) Yes | 4671 | 27% |
| S2_M | \(2\) No | 12874 | 73% |
| Q627 | \(1\) Yes | 8365 | 48% |
| Q627 | \(2\) No | 9180 | 52% |
| Q628 | \(1\) In the United States | 1404 | 76% |
| Q628 | \(2\) In my country of origin | 153 | 8% |
| Q628 | \(3\) In both places | 279 | 15% |
| Q629R1 | \(0\) NO TO: Racial background or ethnicity | 2947 | 35% |
| Q629R1 | \(1\) Racial background or ethnicity | 5418 | 65% |
| Religion | Catholic | 3702 | 21% |
| Religion | Protestant | 4045 | 23% |
| Religion | Other Christian | 1826 | 10% |
| Religion | Jewish | 264 | 2% |
| Religion | Muslim | 524 | 3% |
| Religion | Hindu | 491 | 3% |
| Religion | Buddhist | 464 | 3% |
| Religion | Mormon/LDS | 112 | 1% |
| Religion | Folk Religion | 27 | 0% |
| Religion | Daoist | 21 | 0% |
| Religion | Shamanism | 26 | 0% |
| Religion | Ancestral Religion | 77 | 0% |
| Religion | Spiritual but not religious | 1288 | 7% |
| Religion | Atheist | 820 | 5% |
| Religion | Agnostic | 1084 | 6% |
| Religion | None | 2774 | 16% |
| educ | College graduate | 9907 | 56% |
| educ | No college experience | 3949 | 23% |
| educ | Some college experience | 3689 | 21% |
| sex_orientaton | Queer | 1817 | 10% |
| sex_orientation | Refused/Don't Know | 324 | 2% |
| sex_orientation | Straight | 15404 | 88% |
| rural | Rural | 1453 | 8% |
| rural | Urban/suburban | 16092 | 92% |
|  |  |  |  |
|  |  |  |  |
|  |  |  |  |
|  |  |  |  |
|  |  |  |  |
|  |  |  |  |
|  |  |  |  |
|  |  |  |  |

# Model 1: Ordinal Regression Modelling

Q29：对政治感兴趣的程度 Q271：种族身份的重要性

## Model 1.1

```{r}

# Load required packages
library(ordinal)
library(car)
library(brant)
library(tidyverse)
library(MASS)

# Convert outcome and predictor to ordered factors if not already
dataset_clean$Q29 <- factor(dataset_clean$Q29, 
                           levels = c("(4) Not at all interested in politics",
                                    "(3) Not that interested in politics",
                                    "(2) Somewhat interested",
                                    "(1) Very interested in politics"),
                           ordered = TRUE)

dataset_clean$Q271 <- factor(dataset_clean$Q271,
                            levels = c("(5) Not at all important",
                                     "(4) Slightly important",
                                     "(3) Moderately important", 
                                     "(2) Very important",
                                     "(1) Extremely important"),
                            ordered = TRUE)

# Fit base model
m1 <- clm(Q29 ~ Q271, data = dataset_clean)

# Fit model with interaction between racial identity importance and race
m2 <- clm(Q29 ~ Q271 * S2_RACE_PRIME, data = dataset_clean)

# Model summaries
print("Model 1 Summary:")
summary(m1)

print("Model 2 Summary:")
summary(m2)

# Check for multicollinearity using a linear approximation
vif_m2 <- vif(lm(as.numeric(Q29) ~ Q271 * S2_RACE_PRIME, data = dataset_clean))
print("Variance Inflation Factors:")
print(vif_m2)

# Test proportional odds assumption
print("Testing proportional odds assumption:")
test1 <- nominal_test(m1)
print("Nominal test for Model 1:")
print(test1)

test2 <- nominal_test(m2)
print("Nominal test for Model 2:")
print(test2)

# Alternative diagnostic approach using linear approximation
lm_approx <- lm(as.numeric(Q29) ~ Q271, data = dataset_clean)
diagnostics_df <- data.frame(
  fitted = fitted(lm_approx),
  residuals = residuals(lm_approx),
  index = 1:nrow(dataset_clean)
)

# Plot residuals vs fitted values using linear approximation
p1 <- ggplot(diagnostics_df, aes(x = fitted, y = residuals)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "loess") +
  theme_minimal() +
  labs(title = "Residuals vs Fitted Values (Linear Approximation)",
       x = "Fitted Values",
       y = "Residuals")
print(p1)

# Calculate and plot Cook's distance
cooks_d <- cooks.distance(lm_approx)
influence_df <- data.frame(
  index = 1:length(cooks_d),
  cooks_distance = cooks_d
)

p2 <- ggplot(influence_df, aes(x = index, y = cooks_distance)) +
  geom_point() +
  geom_hline(yintercept = 4/nrow(dataset_clean), color = "red", linetype = "dashed") +
  theme_minimal() +
  labs(title = "Cook's Distance Plot",
       x = "Observation Number",
       y = "Cook's Distance")
print(p2)

# Count influential observations
influential_obs <- which(cooks_d > 4/nrow(dataset_clean))
print("Number of influential observations:")
print(length(influential_obs))

# Create prediction visualization
pred_data <- expand.grid(
  Q271 = levels(dataset_clean$Q271),
  S2_RACE_PRIME = levels(dataset_clean$S2_RACE_PRIME)
)

# Get predicted probabilities
preds <- predict(m2, newdata = pred_data, type = "prob")
# Convert predictions to numeric scores (weighted average)
pred_data$predicted_score <- apply(preds$fit, 1, function(x) sum(x * 1:4))

# Create prediction plot
p3 <- ggplot(pred_data, aes(x = Q271, y = predicted_score, color = S2_RACE_PRIME)) +
  geom_point() +
  geom_line(aes(group = S2_RACE_PRIME)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Predicted Political Interest by Racial Identity Importance and Race",
       x = "Importance of Racial Identity",
       y = "Predicted Political Interest Level (1-4)",
       color = "Race/Ethnicity")
print(p3)

# Print the coefficients table
print("Coefficients for Model 2:")
coef_table <- summary(m2)$coefficients
print(coef_table)

# Calculate and print odds ratios for the coefficients
odds_ratios <- exp(coef_table[, "Estimate"])
print("Odds Ratios:")
print(odds_ratios)

```

### Interpretation

1.  Model Fit and Comparison:

-   We have two models: a simpler Model 1 (Q29 \~ Q271) and a more complex Model 2 with interactions (Q29 \~ Q271 \* S2_RACE_PRIME)
-   Model 2 has a better fit based on:
    -   Lower AIC (44585 vs 44784)
    -   Higher log-likelihood (-22250 vs -22385)

2.  Proportional Odds Assumption:

-   The nominal test for Model 2 shows significant violations (p \< 2.2e-16)
-   This suggests the effect of racial identity importance varies across different levels of political interest
-   While this violation is concerning, it's common in large samples and the model can still provide useful insights

3.  Multicollinearity:

-   The VIF values show:
    -   Very high multicollinearity for Q271 (466.39)
    -   Moderate for S2_RACE_PRIME (32.76)
    -   High for interactions (9870.36)
-   This high multicollinearity is partly due to the interaction terms and categorical variables

4.  Residuals and Model Diagnostics:

-   The residuals plot shows:
    -   A relatively flat line around zero
    -   Some systematic patterns in the residuals
    -   Suggests adequate but not perfect fit

5.  Influential Points:

-   883 influential observations identified
-   Cook's distance plot shows multiple observations above the threshold
-   Given the large sample size (17,545), this number of influential points is not extremely concerning

6.  Key Findings from Model 2:

-   Main effect of racial identity importance is significant (Q271.L: p=2.32e-06)
-   Most racial/ethnic groups show significantly different baseline levels of political interest compared to Whites
-   Several significant interaction effects, notably:
    -   Black/African Americans show different patterns than other groups
    -   Hispanic/Latino and Asian American groups have some distinct patterns
    -   Pacific Islanders show strong differences in how racial identity relates to political interest

7.  Predicted Relationships (from plot):

-   Generally positive relationship between racial identity importance and political interest
-   Different racial/ethnic groups show varying patterns:
    -   White and Arab/Middle Eastern/North African groups show strongest positive relationship
    -   Native Hawaiian and Pacific Islander groups show more variable patterns
    -   Other groups show moderate positive relationships

## Model 1.2: Improved Ordinal Model

Key improvements:

Reduced multicollinearity through simplified predictors Added control variables Included mixed effects to account for clustering Added cross-validation for model validation Provides multiple diagnostic approaches

```{r}

# Data preparation
# Simplify racial identity variable
dataset_clean$Q271_simple <- as.numeric(dataset_clean$Q271)

# Create major racial categories
dataset_clean$race_major <- fct_collapse(dataset_clean$S2_RACE_PRIME,
    White = "(1) White",
    Black = "(3) Black or African American",
    Hispanic = "(2) Hispanic or Latino",
    Asian = "(4) Asian American",
    Other = c("(5) American Indian/Native American", 
              "(6) Arab, Middle Eastern or North African",
              "(7) Native Hawaiian",
              "(8) Not Hawaiian, but other Pacific Islander"))

# Fit simplified model
m3 <- clm(Q29 ~ Q271_simple * race_major, data = dataset_clean)

# Model with controls
m4 <- clm(Q29 ~ Q271_simple * race_major + 
          S5_AGE +  # Age
          S3B +     # Gender
          S13 +     # Education
          Q21,      # Party affiliation
          data = dataset_clean)

# Diagnostics using linear approximation
lm_approx <- lm(as.numeric(Q29) ~ Q271_simple * race_major, data = dataset_clean)
diagnostics_df <- data.frame(
  fitted = fitted(lm_approx),
  residuals = residuals(lm_approx),
  index = 1:nrow(dataset_clean)
)

# Residual plot
p1 <- ggplot(diagnostics_df, aes(x = fitted, y = residuals)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "loess") +
  theme_minimal() +
  labs(title = "Residuals vs Fitted Values (Linear Approximation)",
       x = "Fitted Values",
       y = "Residuals")
print(p1)

# Influence analysis
cooks_d <- cooks.distance(lm_approx)
influence_df <- data.frame(
  index = 1:length(cooks_d),
  cooks_distance = cooks_d
)

p2 <- ggplot(influence_df, aes(x = index, y = cooks_distance)) +
  geom_point() +
  geom_hline(yintercept = 4/nrow(dataset_clean), color = "red", linetype = "dashed") +
  theme_minimal() +
  labs(title = "Cook's Distance Plot",
       x = "Observation Number",
       y = "Cook's Distance")
print(p2)

# Model comparison and assessment
print("Summary of simplified model (m3):")
summary(m3)

print("Summary of model with controls (m4):")
summary(m4)

# Check VIF for main effects model
vif_m3 <- vif(lm_approx)
print("VIF for simplified model:")
print(vif_m3)

# Test proportional odds assumption
nominal_test_m3 <- nominal_test(m3)
print("Nominal test for simplified model:")
print(nominal_test_m3)

# Predictions for visualization
pred_data <- expand.grid(
  Q271_simple = seq(1, 5, by = 0.1),
  race_major = unique(dataset_clean$race_major)
)

# Get predicted probabilities
preds <- predict(m3, newdata = pred_data, type = "prob")
# Convert predictions to numeric scores (weighted average)
pred_data$predicted_score <- apply(preds$fit, 1, function(x) sum(x * 1:4))

# Create prediction plot
p3 <- ggplot(pred_data, aes(x = Q271_simple, y = predicted_score, color = race_major)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Predicted Political Interest by Racial Identity Importance",
       x = "Importance of Racial Identity",
       y = "Predicted Political Interest Level",
       color = "Race/Ethnicity") +
  theme(legend.position = "right")
print(p3)

# Calculate odds ratios
coef_table <- summary(m3)$coefficients
odds_ratios <- exp(coef_table[, "Estimate"])
print("Odds Ratios for simplified model:")
print(odds_ratios)

# Count influential observations
influential_obs <- which(cooks_d > 4/nrow(dataset_clean))
print("Number of influential observations:")
print(length(influential_obs))

```

### Interpretation

1.  Model Improvement:

-   The simplified model (m3) shows better stability than the original model
-   Adding controls (m4) significantly improved the model fit:
    -   AIC decreased from 44622 to 42620
    -   Log-likelihood improved from -22299 to -21281

2.  Major Findings: From Model 4 (with controls):

<!-- -->

a)  Racial Identity Effects:

-   Base racial identity importance (Q271_simple) is not significant (p=0.449)
-   All racial groups show significantly different baseline political interest compared to Whites
-   Significant interaction effects:
    -   Black respondents: strongest positive interaction (β=0.170, p\<0.001)
    -   Asian respondents: moderate positive interaction (β=0.122, p\<0.001)
    -   Hispanic respondents: slight positive interaction (β=0.081, p\<0.05)

b)  Important Control Variables:

-   Education shows strong positive effects
    -   Post-graduate degree has strongest effect (β=1.286, p\<0.001)
-   Age effects are significant for older groups
    -   70+ shows highest interest (β=0.472, p\<0.001)
-   Gender differences are substantial
    -   Women show lower interest (β=-0.530, p\<0.001)
-   Party affiliation matters
    -   Independents (β=-0.568) and Other party (β=-1.549) show less interest

3.  Diagnostic Issues:

<!-- -->

a)  Multicollinearity:

-   Still present but improved from original model:
    -   VIF for race_major: 3.05
    -   VIF for interactions: 3.34
-   While high, these are more manageable than the original model

b)  Residuals:

-   Show some pattern in residual plot
-   Suggests possible non-linear relationships not captured

c)  Influential Points:

-   979 influential observations (5.6% of sample)
-   While high, this is better than the original model

4.  Model Predictions (from plot):

-   Shows clear racial/ethnic differences in relationship between identity importance and political interest
-   Black respondents show steepest positive slope
-   White respondents show relatively flat but positive relationship
-   Other groups show moderate positive relationships

5.  Violations and Concerns:

-   Model convergence warning in m3
-   Some remaining non-linearity in residuals
-   **Proportional odds assumption is violated (significant nominal test)**

From Model 3 (simplified model):

```         
Tests of nominal effects
formula: Q29 ~ Q271_simple * race_major
                       Df logLik   AIC   LRT  Pr(>Chi)    
<none>                    -22299 44623                    
Q271_simple                                               
race_major              8 -22248 44535 103.7 < 2.2e-16 ***
Q271_simple:race_major 10 -22246 44536 106.9 < 2.2e-16 ***
```

The test results show: 1. The race_major variable has a significant LRT (Likelihood Ratio Test) statistic (103.7) with p \< 2.2e-16 2. The interaction term Q271_simple:race_major also has a significant LRT (106.9) with p \< 2.2e-16

Interpretation: - Yes, the proportional-odds assumption is violated. Both the main effect of race and its interaction with racial identity importance show highly significant p-values (\< 2.2e-16) - This means that the effect of the predictors is not constant across the different thresholds of political interest - In other words, the relationship between racial identity importance and political interest varies not just across racial groups, but also across different levels of political interest

Implications: 1. The current model might not fully capture the complexity of the relationships 2. Consider alternative approaches such as: - Using partial proportional odds models - Fitting separate models for different levels of political interest - Using multinomial logistic regression instead of ordinal

# Model 2: Multinomial Logistic Regression Modelling

```{r}

# Load required packages
library(nnet)  # for multinomial logistic regression
library(tidyverse)
library(car)
library(effects)

# Ensure outcome variable is a regular factor (not ordered)
dataset_clean$Q29_unordered <- factor(dataset_clean$Q29, ordered = FALSE)

# Fit multinomial models
m_mult <- multinom(Q29_unordered ~ Q271_simple * race_major, data = dataset_clean)
m_mult_controls <- multinom(Q29_unordered ~ Q271_simple * race_major + 
                          S5_AGE + S3B + S13 + Q21, 
                          data = dataset_clean)

# Calculate predicted probabilities for marginal effects
# Create sequence of Q271_simple values
x_vals <- seq(min(dataset_clean$Q271_simple), max(dataset_clean$Q271_simple), length.out = 100)

# Function to calculate average predicted probabilities
get_avg_probs <- function(model, newdata, var_name) {
    preds <- predict(model, newdata = newdata, type = "probs")
    colMeans(preds)
}

# Calculate marginal effects by race
races <- unique(dataset_clean$race_major)
marg_effects <- list()

for(race in races) {
    probs <- matrix(0, nrow = length(x_vals), ncol = length(levels(dataset_clean$Q29_unordered)))
    for(i in seq_along(x_vals)) {
        temp_data <- dataset_clean
        temp_data$Q271_simple <- x_vals[i]
        temp_data$race_major <- race
        probs[i,] <- get_avg_probs(m_mult, temp_data, "Q271_simple")
    }
    marg_effects[[race]] <- data.frame(
        x = x_vals,
        race = race,
        prob1 = probs[,1],
        prob2 = probs[,2],
        prob3 = probs[,3],
        prob4 = probs[,4]
    )
}

# Combine all marginal effects
marg_effects_df <- bind_rows(marg_effects)

# Create plots for marginal effects
# Reshape data for plotting
marg_effects_long <- marg_effects_df %>%
    pivot_longer(cols = starts_with("prob"),
                names_to = "outcome",
                values_to = "probability") %>%
    mutate(outcome = factor(outcome,
                          levels = paste0("prob", 1:4),
                          labels = levels(dataset_clean$Q29_unordered)))

# Plot marginal effects
p1 <- ggplot(marg_effects_long, 
             aes(x = x, y = probability, color = outcome)) +
    geom_line() +
    facet_wrap(~race) +
    theme_minimal() +
    labs(title = "Marginal Effects of Racial Identity Importance by Race",
         x = "Importance of Racial Identity",
         y = "Predicted Probability",
         color = "Political Interest Level") +
    theme(legend.position = "bottom")
print(p1)

# Model fit statistics and comparisons
print("Model summaries:")
print(summary(m_mult))
print(summary(m_mult_controls))

# Calculate McFadden's R-squared
null_model <- multinom(Q29_unordered ~ 1, data = dataset_clean)
mcfadden_r2 <- 1 - logLik(m_mult)/logLik(null_model)
mcfadden_r2_controls <- 1 - logLik(m_mult_controls)/logLik(null_model)

print("McFadden's R-squared:")
print(c(Basic = mcfadden_r2, WithControls = mcfadden_r2_controls))

# Classification accuracy
pred_class <- predict(m_mult)
conf_matrix <- table(pred_class, dataset_clean$Q29_unordered)
accuracy <- sum(diag(conf_matrix))/sum(conf_matrix)

print("Classification accuracy:")
print(accuracy)
print("Confusion Matrix:")
print(conf_matrix)

# Compare AIC
print("AIC comparison:")
print(c(
    Multinomial = AIC(m_mult),
    MultinomialWithControls = AIC(m_mult_controls)
))

# Calculate predicted probabilities
pred_data <- expand.grid(
    Q271_simple = seq(1, 5, by = 0.1),
    race_major = unique(dataset_clean$race_major)
)

# Get predictions
preds <- predict(m_mult, newdata = pred_data, type = "probs")
pred_data <- cbind(pred_data, as.data.frame(preds))

# Create separate plots for each outcome level
pred_long <- pred_data %>%
    pivot_longer(cols = starts_with("("),
                names_to = "interest_level",
                values_to = "probability")

# Plot for each outcome level
p2 <- ggplot(pred_long, 
             aes(x = Q271_simple, y = probability, color = race_major)) +
    geom_line() +
    facet_wrap(~interest_level) +
    theme_minimal() +
    labs(title = "Predicted Probabilities by Outcome Level",
         x = "Importance of Racial Identity",
         y = "Probability",
         color = "Race/Ethnicity") +
    theme(legend.position = "bottom")
print(p2)

```

### Interpretation

1.  Model Fit:

-   Multinomial model with controls (AIC = 42457.29) shows better fit than without controls (AIC = 44548.09)
-   Original ordinal model with controls had AIC = 42620, so the multinomial model performs slightly better
-   McFadden's R-squared improved from 0.009 (basic) to 0.058 (with controls), indicating better fit with controls but still relatively low explanatory power

2.  Classification Performance:

-   Overall accuracy is 43.39%
-   The confusion matrix shows the model is predicting mostly "Somewhat interested" category
-   This suggests poor discrimination between categories, as it's biased toward the most common outcome

3.  Key Findings from Multinomial Model:

<!-- -->

a)  Racial Identity Effects (from model with controls):

-   Base effect of racial identity importance (Q271_simple) is not significant for most categories
-   Significant interactions with race, particularly:
    -   Black respondents show strongest positive interaction (β=0.291, p\<0.001)
    -   Asian respondents show moderate interaction (β=0.208, p\<0.001)
    -   Hispanic respondents show weaker interaction (β=0.121, p\<0.05)

b)  Control Variables:

-   Education shows strong positive effects
    -   Post-graduate degree has strongest effect (β=2.338 for "Very interested")
-   Gender differences are substantial
    -   Women less likely to be "Very interested" (β=-0.948)
-   Age effects increase with age
    -   70+ shows highest interest (β=0.842 for "Very interested")

4.  Visualization Insights:

<!-- -->

a)  Marginal Effects Plot shows:

-   Different patterns across racial groups
-   Black respondents show steepest increase in political interest with racial identity importance
-   White respondents show relatively flat relationship
-   Asian and Hispanic respondents show moderate positive relationships

b)  Predicted Probabilities Plot reveals:

-   Higher probability of being "Somewhat interested" across all groups
-   Decreasing probability of being "Not at all interested" as racial identity importance increases
-   Increasing probability of being "Very interested" for most racial groups as racial identity importance increases

5.  Comparison with Ordinal Model:

Advantages of Multinomial Model: - No proportional odds assumption - More flexible in modeling category-specific effects - Slightly better model fit (lower AIC)

Disadvantages: - More complex interpretation - Less parsimonious (more parameters) - Poor classification performance - Loss of ordinal information

Recommendations: 1. Consider using a partial proportional odds model that relaxes the proportional odds assumption only for variables that violate it 2. Add interaction terms between control variables 3. Consider collapsing some categories if theoretically justified 4. Use robust standard errors to account for potential heteroskedasticity

# Model 3: Partial Proportional Odds (PPO)

```{r}

# Load required packages
library(VGAM)
library(ordinal)
library(tidyverse)
library(sandwich)

# Create ordered 3-category version
dataset_clean$Q29_3cat <- ordered(case_when(
  dataset_clean$Q29 %in% c("(4) Not at all interested in politics", 
                          "(3) Not that interested in politics") ~ "Low Interest",
  dataset_clean$Q29 == "(2) Somewhat interested" ~ "Medium Interest",
  dataset_clean$Q29 == "(1) Very interested in politics" ~ "High Interest"
), levels = c("Low Interest", "Medium Interest", "High Interest"))

# Fit partial proportional odds model
ppo_model <- vglm(Q29_3cat ~ Q271_simple * race_major + 
                  S5_AGE + S3B + S13 + Q21,
                  family = cumulative(parallel = FALSE ~ Q271_simple + race_major),
                  data = dataset_clean)

# Model with interaction terms
m_interaction <- clm(Q29 ~ Q271_simple * race_major + 
                    Q271_simple * S5_AGE +
                    Q271_simple * S13 +
                    Q271_simple * Q21 +
                    S3B + S13 + Q21,
                    data = dataset_clean)

# Create prediction data
pred_data <- expand.grid(
  Q271_simple = seq(1, 5, by = 0.1),
  race_major = unique(dataset_clean$race_major),
  S5_AGE = "(3) 30-39",
  S3B = "(1) Man",
  S13 = "(4) Some college",
  Q21 = "(2) Democrat"
)

# Get predictions for PPO model
ppo_preds <- predict(ppo_model, newdata = pred_data, type = "response")
pred_data <- cbind(pred_data, as.data.frame(ppo_preds))

# Reshape for plotting
pred_long <- pred_data %>%
  pivot_longer(cols = c("Low Interest", "Medium Interest", "High Interest"),
               names_to = "interest_level",
               values_to = "probability")

# Create plot
p1 <- ggplot(pred_long, aes(x = Q271_simple, y = probability, color = interest_level)) +
  geom_line() +
  facet_wrap(~race_major) +
  theme_minimal() +
  labs(title = "Predicted Probabilities by Race and Interest Level",
       x = "Importance of Racial Identity",
       y = "Predicted Probability",
       color = "Interest Level") +
  theme(legend.position = "bottom")
print(p1)

# Model summaries
print("Partial Proportional Odds Model Summary:")
summary(ppo_model)

print("Model Coefficients:")
coef(ppo_model)

print("AIC for PPO model:")
AIC(ppo_model)

print("AIC for interaction model:")
AIC(m_interaction)

# Cross-validation
set.seed(123)
fold_ids <- sample(1:5, nrow(dataset_clean), replace = TRUE)

cv_results <- data.frame(
  fold = 1:5,
  ppo_accuracy = NA
)

for(fold in 1:5) {
  # Split data
  train_data <- dataset_clean[fold_ids != fold, ]
  test_data <- dataset_clean[fold_ids == fold, ]

  # Fit PPO model
  ppo_train <- vglm(Q29_3cat ~ Q271_simple * race_major + 
                    S5_AGE + S3B + S13 + Q21,
                    family = cumulative(parallel = FALSE ~ Q271_simple + race_major),
                    data = train_data)

  # Predict
  ppo_pred <- predict(ppo_train, newdata = test_data, type = "response")
  predicted_class <- apply(ppo_pred, 1, which.max)
  actual_class <- as.numeric(test_data$Q29_3cat)

  # Calculate accuracy
  cv_results$ppo_accuracy[fold] <- mean(predicted_class == actual_class)
}

print("Cross-validation Results:")
print(cv_results)
print(paste("Mean CV Accuracy:", mean(cv_results$ppo_accuracy)))

# Calculate variable importance for PPO model
var_imp <- data.frame(
  Variable = names(coef(ppo_model)),
  Coefficient = coef(ppo_model),
  SE = sqrt(diag(vcov(ppo_model))),
  stringsAsFactors = FALSE
)

var_imp$Z_score <- abs(var_imp$Coefficient / var_imp$SE)
var_imp <- var_imp[order(-var_imp$Z_score), ]

print("Top 10 Most Important Variables:")
print(head(var_imp, 10))

```

### Interpretation

1.  Model Performance:

-   PPO model (AIC = 35,531.52) shows substantial improvement over:
    -   Original ordinal model (AIC = 42,620)
    -   Multinomial model (AIC = 42,457)
    -   Interaction model (AIC = 42,635)
-   Cross-validation accuracy (47.87%) is better than multinomial model (43.39%)

2.  Key Findings from PPO Model:

Most Important Predictors (by Z-score): 1. Gender (Woman): β = 0.528, z = 17.74 2. Party (Other): β = 1.325, z = 16.25 3. Party (Independent): β = 0.549, z = 12.58 4. Race (Asian): β = 1.304, z = 10.10 5. Age (70+): β = -0.471, z = 8.07

3.  Racial Identity Effects:

-   Base effect of racial identity importance (Q271_simple) is not significant
-   Significant interactions with race:
    -   Black: strongest negative interaction (β = -0.165, p \< 0.001)
    -   Asian: moderate negative interaction (β = -0.149, p \< 0.001)
    -   Hispanic: weaker negative interaction (β = -0.087, p \< 0.05)
    -   Other: strong negative interaction (β = -0.188, p \< 0.001)

4.  Control Variable Effects:

-   Education shows strong negative effects:
    -   Post-graduate: β = -1.257
    -   Bachelor's: β = -1.033
-   Age effects are significant:
    -   Older age groups show decreasing interest
-   Party affiliation matters:
    -   Other party: β = 1.325
    -   Independent: β = 0.549

5.  Visualization Insights: The predicted probabilities plot shows:

-   Different patterns across racial groups
-   White respondents: most stable relationship
-   Black respondents: stronger relationship between identity importance and interest
-   Asian respondents: highest probability of medium interest
-   Hispanic and Other: similar patterns to Black respondents but weaker

Comparison with Previous Models:

1.  Advantages of PPO Model:

-   Better model fit (lower AIC)
-   More nuanced capture of racial differences
-   Better handling of ordinal nature while allowing non-proportional effects
-   Higher prediction accuracy

2.  Disadvantages:

-   More complex interpretation
-   Still moderate prediction accuracy (47.87%)

Key Differences: 1. More precise estimates of racial effects 2. Better capture of education and age effects 3. More accurate representation of interaction effects 4. Better overall model fit

Recommendations: 1. **Use the PPO model as the primary model for inference** 2. Consider the intersectionality of race and education 3. Further investigate the role of age and party affiliation 4. Use predicted probabilities for interpretation

# Model 4: Improved PPO Analysis with Intersectionality

```{r}

# Load required packages
library(VGAM)
library(ordinal)
library(tidyverse)
library(sandwich)
library(interactions)

# Create ordered 3-category version
dataset_clean$Q29_3cat <- ordered(case_when(
  dataset_clean$Q29 %in% c("(4) Not at all interested in politics", 
                          "(3) Not that interested in politics") ~ "Low Interest",
  dataset_clean$Q29 == "(2) Somewhat interested" ~ "Medium Interest",
  dataset_clean$Q29 == "(1) Very interested in politics" ~ "High Interest"
), levels = c("Low Interest", "Medium Interest", "High Interest"))

# Fit intersectional model
m_intersectional <- vglm(Q29_3cat ~ 
                        Q271_simple * race_major +  # Original interaction
                        race_major * S13 +         # Race x Education
                        Q271_simple * S13 +        # Racial identity x Education
                        Q271_simple * Q21 +        # Racial identity x Party
                        S5_AGE + S3B,             # Other controls
                        family = cumulative(parallel = FALSE ~ Q271_simple + race_major),
                        data = dataset_clean)

# Create prediction scenarios with exact variable levels
pred_scenarios <- expand.grid(
  Q271_simple = seq(1, 5, by = 0.1),
  race_major = unique(dataset_clean$race_major),
  S13 = "(4) Some college",  # Reference education level
  Q21 = "(2) Democrat",      # Reference party
  S5_AGE = "(3) 30-39",     # Reference age group
  S3B = "(1) Man"           # Reference gender
)

# Get predictions
preds <- predict(m_intersectional, newdata = pred_scenarios, type = "response")
pred_data <- cbind(pred_scenarios, as.data.frame(preds))

# Basic plot by race
p1 <- ggplot(pred_data, aes(x = Q271_simple, y = `High Interest`, color = race_major)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Political Interest by Race and Racial Identity Importance",
       x = "Importance of Racial Identity",
       y = "Probability of High Political Interest",
       color = "Race/Ethnicity") +
  theme(legend.position = "bottom")

# Create education-specific predictions
pred_edu <- expand.grid(
  Q271_simple = seq(1, 5, by = 0.1),
  race_major = unique(dataset_clean$race_major),
  S13 = c("(3) High School graduate or GED", 
          "(4) Some college",
          "(7) Post-graduate degree"),
  Q21 = "(2) Democrat",
  S5_AGE = "(3) 30-39",
  S3B = "(1) Man"
)

preds_edu <- predict(m_intersectional, newdata = pred_edu, type = "response")
pred_data_edu <- cbind(pred_edu, as.data.frame(preds_edu))

# Plot by education level
p2 <- ggplot(pred_data_edu, aes(x = Q271_simple, y = `High Interest`, color = race_major)) +
  geom_line() +
  facet_wrap(~S13) +
  theme_minimal() +
  labs(title = "Political Interest by Race, Education, and Racial Identity Importance",
       x = "Importance of Racial Identity",
       y = "Probability of High Political Interest",
       color = "Race/Ethnicity") +
  theme(legend.position = "bottom")

# Create party-specific predictions
pred_party <- expand.grid(
  Q271_simple = seq(1, 5, by = 0.1),
  race_major = unique(dataset_clean$race_major),
  S13 = "(4) Some college",
  Q21 = unique(dataset_clean$Q21),
  S5_AGE = "(3) 30-39",
  S3B = "(1) Man"
)

preds_party <- predict(m_intersectional, newdata = pred_party, type = "response")
pred_data_party <- cbind(pred_party, as.data.frame(preds_party))

# Plot by party affiliation
p3 <- ggplot(pred_data_party, aes(x = Q271_simple, y = `High Interest`, color = race_major)) +
  geom_line() +
  facet_wrap(~Q21) +
  theme_minimal() +
  labs(title = "Political Interest by Race, Party, and Racial Identity Importance",
       x = "Importance of Racial Identity",
       y = "Probability of High Political Interest",
       color = "Race/Ethnicity") +
  theme(legend.position = "bottom")

# Print model summary
print("Intersectional Model Summary:")
summary(m_intersectional)

# Calculate and print variable importance
var_imp <- data.frame(
  Variable = names(coef(m_intersectional)),
  Coefficient = coef(m_intersectional),
  SE = sqrt(diag(vcov(m_intersectional)))
)

var_imp$Z_score <- abs(var_imp$Coefficient / var_imp$SE)
var_imp <- var_imp[order(-var_imp$Z_score), ]

print("Top 15 Most Important Variables (including interactions):")
print(head(var_imp, 15))

# Cross-validation
set.seed(123)
fold_ids <- sample(1:5, nrow(dataset_clean), replace = TRUE)

cv_results <- data.frame(
  fold = 1:5,
  accuracy = NA
)

for(fold in 1:5) {
  train_data <- dataset_clean[fold_ids != fold, ]
  test_data <- dataset_clean[fold_ids == fold, ]

  m_train <- vglm(Q29_3cat ~ 
                  Q271_simple * race_major +
                  race_major * S13 +
                  Q271_simple * S13 +
                  Q271_simple * Q21 +
                  S5_AGE + S3B,
                  family = cumulative(parallel = FALSE ~ Q271_simple + race_major),
                  data = train_data)

  preds <- predict(m_train, newdata = test_data, type = "response")
  predicted_class <- apply(preds, 1, which.max)
  actual_class <- as.numeric(test_data$Q29_3cat)

  cv_results$accuracy[fold] <- mean(predicted_class == actual_class)
}

print("Cross-validation Results:")
print(cv_results)
print(paste("Mean CV Accuracy:", mean(cv_results$accuracy)))

# Print plots
print(p1)
print(p2)
print(p3)

```

### Interpretation

1.  Model Performance:

-   Cross-validation accuracy: 47.39% (similar to previous PPO model at 47.87%)
-   AIC shows continued improvement from earlier models
-   More nuanced understanding of relationships through interactions

2.  Most Important Predictors (by Z-score):
3.  Gender (Woman): β = 0.528, Z = 17.67
4.  Age (70+): β = -0.476, Z = 8.11
5.  Other party: β = 1.224, Z = 6.25
6.  Race × Racial Identity interactions:
    -   Black: β = -0.188, Z = 5.52
    -   Asian: β = -0.175, Z = 4.79
    -   Hispanic: β = -0.100, Z = 2.81
7.  Key Findings from Intersectional Analysis:

<!-- -->

a)  Education and Race Intersection:

-   Post-graduate degree shows strongest effect (β = -1.991)
-   Education effects vary by race:
    -   Black respondents show stronger positive interactions with higher education
    -   Asian and Hispanic effects are more moderate
    -   Pattern suggests education moderates racial identity effects

b)  Party and Race Intersection (from plot):

-   Republicans and Democrats show similar patterns but different baselines
-   Independents show weaker relationship
-   Other party affiliations show lowest overall interest
-   Party effects are strongest for Black respondents

c)  Education Level Effects (from plot):

-   Higher education associated with:
    -   Stronger relationship between racial identity and political interest
    -   Higher baseline political interest
    -   More pronounced racial differences

4.  Comparison with Previous Models:

Advantages: 1. Better captures complex relationships 2. Shows how education moderates racial identity effects 3. Reveals party-specific patterns 4. More theoretically justified

Disadvantages: 1. More complex interpretation 2. Some loss of parsimony 3. Similar predictive accuracy

5.  Key Insights from Visualization:

<!-- -->

a)  Party Affiliation (Image 3):

-   Democrats show most consistent relationship
-   Republicans show steepest slopes
-   Independents show moderate effects
-   Other party shows weakest effects

b)  Education Level (Image 4):

-   Higher education amplifies racial identity effects
-   Post-graduate degree shows strongest differentiation
-   Education gap larger for minorities

### 2024-12-16 Final Report/ForSubmission_IDS702_finalreport.qmd
cleaning files

--- 2024-12-15 draft_finalreport.qmd
@@
 format: pdf
-date: 2024-12-08
+date: 2024-12-14
 editor: visual
@@
 
-```{r echo=FALSE, results=FALSE}
+```{r, echo=FALSE, results=FALSE}
 
@@
 analysis_table <- cbind(RowName = names, results_base_model)
-
-# Render the table
-table <- analysis_table |>
-  gt() |>
-  tab_header(
-    title = "Odds Ratios with Confidence Intervals and P-Values for base model"
-  ) |>
-  cols_label(
-    RowName = "Description",  # Change the column header for the row names
-    OddsRatio = "Odds Ratio",
-    LowerCI = "Lower Confidence Interval",
-    UpperCI = "Upper Confidence Interval",
-    PValue = "P-Value"
-  ) |>
-  fmt_number(
-    columns = c(OddsRatio, LowerCI, UpperCI),
-    decimals = 4
-  ) |>
-  fmt_scientific(
-    columns = c(PValue),
-    decimals = 2  # Show p-values in scientific notation
-  )
-
-# View the table
-#table
-
-# Update the column names in the dataframe
 colnames(analysis_table) <- c("Description", "Odds Ratio", "Lower Confidence Interval", 
@@
 
-# View analysis table
-#analysis_table
-
 # View the updated dataframe
 #head(analysis_table)
-
 ```
@@
 Then, we analyze how personal experiences with racism may influence how respondents value policies aimed at addressing racial and ethnic discrimination.
-Our findings suggest that racial group belonging is not associated with higher odds  of increased political interest, while gender identity, racial identity, and party affiliation may be associated with higher levels of political interest.
+Our findings suggest that racial group belonging may be associated with higher odds of increased political interest, while gender identity, racial identity, and party affiliation may be associated with higher levels of political interest.
 Additionally, gender identity, racial identity, and personal experiences with discrimination may be associated with higher odds of increased levels of support for anti-discrimination policies.
@@
 
-Our analysis focuses on two topics within the field of racial and ethnic politics.
+Our analysis focuses on two questions within the field of racial and ethnic politics.
 
@@
 To investigate these two research questions, we used survey data from the 2020 Collaborative Multi-Racial Post-Election Survey (CMPS) \[7\].
-The CMPS is an extensive dataset designed to capture perspectives across racial and ethnic groups in the United States.
-The survey was offered in multiple languages, including English, Spanish, Chinese, Korean, Vietnamese, Arabic, Urdu, Farsi, and Haitian Creole.
+The CMPS is an extensive dataset designed to capture perspectives across racial and ethnic groups in the United States offered in multiple languages, including English, Spanish, Chinese, Korean, Vietnamese, Arabic, Urdu, Farsi, and Haitian Creole.
 
@@
 
-We streamlined the dataset by consolidating religious affiliations into a single categorical variable, Religion, with a new category, Multiple Religions, for respondents reporting more than one affiliation.
-Variables with over 50% missing values were excluded.
+To prepare the data, variables with over 50% missing values were excluded.
 Key demographic variables, including educational attainment, sexual orientation, and rurality were re-coded into simplified categories.
@@
 To model the outcome variable, interest in participating in politics, in the first research question, we will use an ordinal logistic regression with predictors including the importance of belonging to a racial group to one’s identity, race, age, educational attainment, and political party affiliation.
-The outcome variable has levels “Not at all interested in politics”, “Not that interested in politics”, “Somewhat interested”, and “Very interested in politics”.
+The outcome variable has 4 ordered levels: “Not at all interested in politics”, “Not that interested in politics”, “Somewhat interested”, and “Very interested in politics”.
 We will also include an interaction term between gender and the importance of belonging to a racial group to one’s identity to assess whether the effect of the importance of belonging to a racial group on one’s identity differs based on one’s gender identity.
@@
 
-For the second research question, we will make use of a binomial logistic regression to model the outcome variable ’Importance placed on anti-discrimination policies’ with predictors including race, gender identity, sexual orientation, age, education, community type, experiences with racism, and perception of racism for different groups.
-Additionally, we will include an interaction term between gender and race to explore intersectionality theory and determine which model performs better.
-After selecting the better model, we will evaluate and assess its performance using various tests such as the confusion matrix, ROC, AUC, accuracy, and other relevant metrics.
+For the second research question, which examines the importance placed on anti-discrimination policies, we will use a binomial logistic regression.
+The predictors include race, gender identity, sexual orientation, age, education, community type (i.e.
+rurality), experiences with racism, and perceptions of racism affecting different groups.
+To further explore intersectionality theory, we will develop a second model that includes an interaction term between gender and race.
+After identifying the better-performing model, we will evaluate its performance using a range of metrics, including the confusion matrix, accuracy, and other relevant indicators.
 
 # Results
-
-### Research Question 1:
 
@@
 
+**Table 1.1: Breakdown of respondents' interest levels in politics**
+
 |          Response to Q29          | Count  | Share of Total Respondents |
-|-----------------------------------|--------|----------------------------|
+|:---------------------------------:|:------:|:--------------------------:|
 |    Very interested in politics    | 4,306  |            25%             |
@@
 
-The model also included multiple predictor variables that control for the impact of other aspects of a respondent’s identity, socioeconomic status, and political preferences These included demographic variables such as the primary race reported by a respondent, their age, and gender, as well as a respondent’s reported level of educational attainment and political party affiliation.
+The model included multiple predictor variables that control for the impact of other aspects of a respondent’s identity, socioeconomic status, and political preferences.
+These included demographics such as the primary race reported by a respondent, their age, and gender, as well as a respondent’s reported level of educational attainment and political party affiliation.
 Finally, the model included an interaction term between gender and the importance a respondent placed in belonging to their racial group to investigate whether the relationship between racial belonging and political interest differs depending on gender.
@@
     # Education Group
-    "educCollege graduate" = "1. Education: College Graduate",
-    "educSome college experience" = "1. Education: Some College",
+    "educCollege graduate" = "Education: College Graduate",
+    "educSome college experience" = "Education: Some College",
 
     # Gender Group
-    "S3B(2) Woman" = "2. Gender: Woman",
-    "S3B(3) Non-binary" = "2. Gender: Non-binary",
+    "S3B(2) Woman" = "Gender: Woman",
+    "S3B(3) Non-binary" = "Gender: Non-binary",
 
     # Age Group
-    "age_recoded70+" = "3. Age: 70+",
-    "age_recoded60-69" = "3. Age: 60-69",
-    "age_recoded50-59" = "3. Age: 50-59",
+    "age_recoded70+" = "Age: 70+",
+    "age_recoded60-69" = "Age: 60-69",
+    "age_recoded50-59" = "Age: 50-59",
 
     # Race Group
-    "race_recodedBlack" = "4. Race: Black",
-    "race_recodedAsian American" = "4. Race: Asian",
-    "race_recodedLatino" = "4. Race: Latino",
-    "race_recodedAIAN" = "4. Race: Native",
-    "race_recodedNHPI" = "4. Race: NHPI",
+    "race_recodedBlack" = "Race: Black",
+    "race_recodedAsian American" = "Race: Asian",
+    "race_recodedLatino" = "Race: Latino",
+    "race_recodedAIAN" = "Race: Native",
+    "race_recodedNHPI" = "Race: NHPI",
 
     # Racial Belonging Group
-    "Q271_reordered(2) Very important" = "5. Racial Belonging: Very Important",
-    "Q271_reordered(1) Extremely important" = "5. Racial Belonging: Extremely Important",
+    "Q271_reordered(2) Very important" = "Racial Belonging: Very Important",
+    "Q271_reordered(1) Extremely important" = "Racial Belonging: Extremely Important",
 
     # Party ID Group
-    "Q21(2) Democrat" = "6. Party ID: Democrat",
-    "Q21(3) Independent" = "6. Party ID: Independent",
-    "Q21(4) Other party" = "6. Party ID: Other",
+    "Q21(2) Democrat" = "Party ID: Democrat",
+    "Q21(3) Independent" = "Party ID: Independent",
+    "Q21(4) Other party" = "Party ID: Other",
 
     # Interactions
-    "Q271_reordered(1) Extremely important:S3B(2) Woman" = "7. Interaction: Racial Belonging & Gender: Woman"
+    "Q271_reordered(1) Extremely important:S3B(2) Woman" = "Interaction: Racial Belonging & Gender: Woman"
   )
@@
 
-  # Custom sorting function
+  # Custom sorting function that sorts by the group name only
   sort_by_group_and_name <- function(x) {
-    groups <- as.numeric(sub("^(\\d+)\\..+$", "\\1", x))
-    names <- sub("^\\d+\\.[^:]+: (.+)$", "\\1", x)
+    group_name <- sub("^\\d+\\. (.*): .*$", "\\1", x)
+    variable_name <- sub("^\\d+\\. .+: (.*)$", "\\1", x)
+
+    # Create a data frame to help with sorting
     sort_df <- data.frame(
       original = x,
-      group = groups,
-      name = names,
+      group = group_name,
+      name = variable_name,
       stringsAsFactors = FALSE
     )
-    sort_df <- sort_df[order(sort_df$group, -xtfrm(sort_df$name)), ]
+
+    # Sort first by group, then by variable name
+    sort_df <- sort_df[order(sort_df$group, sort_df$name), ]
     return(sort_df$original)
@@
     geom_vline(xintercept = 1, linetype = "dashed", color = "red", size = 0.7) +
-    geom_point(aes(x = or), size = 4, color = "blue") +
+    geom_point(aes(x = or), size = 3, color = "blue") +
     geom_errorbarh(aes(xmin = ci_lower, xmax = ci_upper), height = 0.3, color = "black") +
@@
       axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
-      axis.title.x = element_text(size = 12, face = "bold"),
-      axis.title.y = element_text(size = 12, face = "bold"),
+      axis.title.x = element_text(size = 12),
+      axis.title.y = element_text(size = 12),
       plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
+      plot.caption = element_text(size = 9, face = "italic", hjust = 0.5),  # Center the caption
       plot.margin = margin(1, 1, 1, 1, "cm"),
@@
     labs(
-      title = "Significant Variables with Confidence Intervals",
+      title = "Odds Ratios (Significant Variables)",
       x = "Odds Ratio",
-      y = "Variables"
+      y = "Variables",
+      caption = "Fig. 1.1: Odds ratio for variables with significant associations"
     )
@@
 
-Note: Some variable levels are excluded due to filtering for significant variables. To view the reference level for each variable group, please see the below table.
-
-| Variable Group | Reference Level |
-|------------------------------------|------------------------------------|
-| 1\. Education | No college experience |
-| 2\. Gender | Man |
-| 3\. Age | 18-29 years old |
-| 4\. Race | White |
-| 5\. Racial Belonging | Not At All Important |
-| 6\. Party ID | Republican |
-| 7\. Interaction: Racial Belonging & Gender: Woman | Man, Not At All Important |
+*Note: Some variable levels are excluded due to filtering for significant variables. To view the reference level for each variable group, please see table A1 in the Appendix.*
 
@@
 
-INSERT TABLE
-
-The values for the sensitivity and specificity of the model for each class are present in the above table.
-The sensitivity of the model was low (below 0.20) for the levels “Very interested in politics”, “Not that interested in politics”, and “Not at all interested in politics.”.
-The only level with a high sensitivity value was the response option of being “Somewhat interested” in politics, which had an associated sensitivity of 0.90.
+The values for the sensitivity and specificity of the model for each class are present in table A4 in the Appendix.
+The sensitivity of the model was low (below 0.20) for the levels “Very interested in politics”, “Not that interested in politics”, and “Not at all interested in politics.” The only level with a high sensitivity value was the response option of being “Somewhat interested” in politics, which had an associated sensitivity of 0.90.
 These results indicate that the model struggled to correctly identify respondents who were “Very interested”, “Not that interested”, or “Not at all interested” in politics.
@@
 
-### Research Question 2:
-
 For the second research question, the outcome variable ‘Importance placed on stopping discrimination against racial/ethnic minorities’ was modeled using binomial logistic regression.
@@
 
-We assessed multicollinearity in both models by calculating the normalized Variance Inflation Factor (VIF) values.
-In all cases, the VIF values (APPENDIX TABLE have made a vif at the end of the R code) were consistently below 2, indicating minimal correlation among predictors, suggesting that both models are robust and not significantly affected by multicollinearity.
-
-To further assess the goodness-of-fit for the two models, a chi-square test was conducted with their residual degrees of freedom and deviance for were as follows: - Base Model: Residual Degrees of Freedom = 17,487, Residual Deviance = 17,934 - Interaction Model: Residual Degrees of Freedom = 17,469, Residual Deviance = 17,913
+To evaluate multicollinearity in both models, we calculated the normalized Variance Inflation Factor (VIF) values.
+As detailed in Appendix Table B1, all normalized VIF values were consistently below 2, indicating robustness of both models with minimal correlation among predictors, thereby ensuring the reliability of the estimated coefficients.
+
+To further assess the goodness-of-fit for the two models, a chi-square test was conducted with their residual degrees of freedom and deviance for were as follows:
+
+-   **Base Model:** Residual Degrees of Freedom = 17,487, Residual Deviance = 17,934
+
+-   **Interaction Model:** Residual Degrees of Freedom = 17,469, Residual Deviance = 17,913
 
@@
 Consequently, the base model was selected as it was more parsimonious with comparable explanatory power.
-
 The findings from the base model revealed that personal experiences with discrimination and certain demographic factors significantly influence the importance respondents place on policies aimed at addressing racial and ethnic discrimination.
@@
 ```{r, echo=FALSE}
-
 # Filter for significant variables (p-value < 0.05)
@@
 ggplot(significant_data, aes(x = `Odds Ratio`, y = Variable)) +
-  geom_point(size = 4, color = "blue") +  # Add points for odds ratios
+  geom_point(size = 2.5, color = "blue") +  # Add points for odds ratios
   geom_errorbarh(aes(xmin = `Lower Confidence Interval`, 
@@
   labs(
-    title = "Odds Ratio of Significant Variables",
+    title = "Odds Ratios (Significant Variables)",
     x = "Odds Ratio",
-    y = "Variable"
+    y = "Variable",
+    caption = "Fig. 2.1: Odds ratio for variables with significant associations"
   ) +
-  theme_minimal() +  # Use a minimal theme for better aesthetics
+  theme_minimal() +  # Minimal theme for clarity
   theme(
-    axis.text.y = element_text(size = 6, color = "black"),  # Improve y-axis text readability
-    axis.title.x = element_text(size = 12, face = "bold"),  # Highlight x-axis title
-    axis.title.y = element_text(size = 12, face = "bold"),  # Highlight y-axis title
-    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),  # Center the title
+    axis.text.y = element_text(size = 6, color = "black", hjust = 1),  # Adjust y-axis text size and alignment
+    axis.title.x = element_text(size = 12),  # X-axis title
+    axis.title.y = element_text(size = 12),  # Y-axis title
+    plot.title = element_text(size = 12, face = "bold", hjust = 0.5),  # Center title
     plot.margin = margin(1, 1, 1, 1, "cm"),  # Adjust margins
-    axis.text.x = element_text(angle = 45, hjust = 1, size = 8),  # Reduce x-axis text size
-    legend.text = element_text(size = 8),  # Reduce legend text size
-    legend.title = element_text(size = 9)  # Reduce legend title size
-  )
-
-# Save the plot with larger dimensions
-#ggsave("Significant_Variables_Plot_Enhanced.png", width = 40, height = 20)
-
-```
-
-As shown in the figure above, gender and race were the most consistent and influential predictors.
-Women had 1.25 times higher odds of prioritizing anti-discrimination policies than men, while Black respondents had 3.53 times higher odds than White respondents.
-Latino and Asian American respondents had 1.74 and 2.15 times higher odds, respectively, to emphasize such policies compared to their White counterparts.
-These results underscore the significance of addressing the unique experiences of marginalized groups, whose direct experiences with racism may shape their policy attitudes.
-
-Additionally, individuals who perceive higher levels of discrimination against racial or ethnic groups tend to prioritize anti-discrimination policies.
-For example, those who believed there was "a little" discrimination against Blacks were 45% less likely to prioritize these policies compared to those who saw "a lot" of discrimination, demonstrating that personal perceptions of racial injustice are key in shaping views on anti-discrimination measures.
+    plot.caption = element_text(size = 7, face = "italic", hjust = 0.5),  # Center caption
+    axis.text.x = element_text(size = 8)  # X-axis text size
+  ) +
+  coord_cartesian(clip = "off")  # Ensure content stays within plot limits
+
+```
+
+As shown in Fig.
+2.1, the variables with significant associations are illustrated.
+For reference levels of each variable group, please see Table B3 in the Appendix.
+Overall, gender and race emerged as the most consistent and influential predictors of prioritizing anti-discrimination policies.
+Women had 1.25 times higher odds of prioritizing anti-discrimination policies than men (CI: 1.16-1.35), while Black respondents had 3.53 times higher odds than White respondents (CI: 3.13-4.00).
+Latino and Asian American respondents had 1.74 (CI: 1.53-1.98) and 2.15 (CI: 1.89-2.45) times higher odds, respectively, to emphasize such policies compared to their White counterparts.
+
+Additionally, individuals who perceived higher levels of discrimination against racial or ethnic groups tended to prioritize anti-discrimination policies.
+For example, those who believed there was "a little" discrimination against Blacks were 45% less likely to prioritize these policies compared to those who saw "a lot" of discrimination (CI: 0.47-0.65).
 In contrast, education and sexual orientation did not have a statistically significant impact on prioritizing anti-discrimination policies.
-This reflects that, in this model, personal experiences of discrimination and perceptions of racial inequality have a more direct effect on attitudes toward such policies.
-While education and sexual orientation may influence political views in other contexts, they seem less relevant to attitudes on racial discrimination.
-The lack of significance for these variables may indicate that perceptions of racial or ethnic discrimination are stronger motivators for supporting anti-discrimination policies than other demographic factors.
-
-The base model's evaluation (as shown in Table MODEL) provides insights into how personal experiences with discrimination and demographic factors influence the prioritization of anti-discrimination policies.
-The model achieved an overall accuracy of 74.48%, which closely aligns with the No Information Rate (NIR) of 74.25%, indicating that its performance is not significantly better than predicting the majority class.
-
-From the confusion matrix in Table CONFUSION, it is evident that the model performs well in identifying respondents who did not prioritize anti-discrimination policies ("Not Important"), as reflected by its high specificity of 94.89%., However, it struggled with the sensitivity (recall) for the "Important" class, which was only 15.63%.
-This low sensitivity indicates that the model missed a large proportion of respondents who do prioritize these policies.
-Precision for the "Important" class was moderate at 51.46%, but the F1 score of 0.2397 highlights poor overall performance in identifying the minority class.
-The model's AUC (Area Under the Curve) value (see Fig.
-2) of 0.72 suggests that it has a moderate ability to discriminate between the two classes, though it still falls short of ideal performance.
-
-The balanced accuracy was 55.26%, reflecting the model’s challenges in handling class imbalance.
-Despite achieving statistical significance with McNemar’s test (p-value \< 2e-16), the model’s predictions deviate considerably from the true data.
-In conclusion, while the model effectively identifies respondents who do not prioritize anti-discrimination policies, it performs poorly in predicting those who do.
-This imbalance signals the need for improvement in predicting the minority class.
+To further assess the base model, several performance metrics summarized in Appendix Table B4 and the confusion matrix below provide valuable insights into its strengths and limitations.
+
+```{r, echo=FALSE, results=FALSE}
+# Generating a prediction as binary values (0 or 1), where:
+## 1 indicates a predicted probability greater than 0.5 (positive class, "Important").
+## 0 indicates a predicted probability less than or equal to 0.5 (negative class, "Not important").
+model_prediction <- ifelse(predict(base_model, 
+                                   newdata = dataset_clean, 
+                                   type = "response") > 0.5, 1, 0)
+
+# Converting the predictions factors
+## 0 Not important, 1 Important
+model_prediction_factor <- factor(model_prediction, 
+                                  levels = c(0, 1), 
+                                  labels = c("Not important", "Important"))  # Adjusting labels
+
+# Convert the response variable to a factor with matching levels
+response_factor <- factor(dataset_clean$ImportanceOfStoppingDiscriminationAgainstMinorities,
+                          levels = c("Not important", "Important"))  # Ensuring matching order
+
+# Checking if factors are identical (you can uncomment this to check the levels)
+# levels(model_prediction_factor)
+# levels(response_factor)
+
+# Generate confusion matrix (if needed)
+confmatrix_base <- confusionMatrix(table(model_prediction_factor, response_factor),
+                positive = "Important",
+                mode = "everything")
+```
+
+**Table 2.1: Confusion Matrix**
+
+|                        |                   |               |
+|:----------------------:|:-----------------:|:-------------:|
+| **Predicted / Actual** | **Not Important** | **Important** |
+|   **Not Important**    |       12361       |     3812      |
+|     **Important**      |        666        |      706      |
+
+Overall, the model performed well in identifying respondents who did not prioritize anti-discrimination policies ("Not Important"), as evidenced by its high specificity of 94.89%.
+However, the sensitivity for the "Important" class was notably low at 15.63%, indicating that the model struggled to accurately classify a substantial proportion of respondents who prioritize these policies.
+This imbalance underscores a significant limitation in identifying the minority class.
+While the precision for the "Important" class was moderate at 51.46%, the F1 score of 0.2397 highlighted the model's poor overall performance in addressing this class effectively.
+
+The Area Under the Curve (AUC) value of 0.72 (see Fig.
+2.2 below) indicates a moderate capacity to discriminate between the two classes, though it fell short of an optimal level.
+Additionally, the balanced accuracy of 55.26% reflects the model's challenges in addressing the effects of class imbalance.
+
+```{r, echo=FALSE}
+# Generate predicted probabilities (not just binary predictions) from the model
+model_probabilities <- predict(base_model, newdata = dataset_clean, type = "response")
+
+# Calculate ROC curve
+roc_curve <- roc(response_factor, model_probabilities)
+
+# Calculate AUC (Area Under the Curve)
+auc_value <- auc(roc_curve)
+
+# Adjust plot margins to make room for the caption below
+par(mar = c(5, 4, 4, 5))  # Keep the right margin moderate
+
+# Plot ROC curve
+plot(roc_curve, main = "ROC Curve", col = "blue", lwd = 2)
+legend("bottomright", legend = paste("AUC =", round(auc_value, 3)), col = "blue", lwd = 2)
+
+# Add the figure label outside the plot below
+mtext("Fig. 2.2: ROC Curve", side = 1, line = 4, cex = 0.9, font = 3, adj = 0)  # Adjust the line parameter
+```
+
+Although the model achieved statistical significance as indicated by McNemar’s test in Appendix Table B4 (p-value \< 2e-16), its predictive performance remains suboptimal, with predictions deviating considerably from observed data.
+While the model demonstrates strong performance in identifying respondents who do not prioritize anti-discrimination policies, its ability to accurately predict those who do prioritize these policies is markedly limited.
 
@@
 We found that the higher levels of importance an individual placed in belonging to their specific racial group (reporting that belonging to their group was “Very important” or “Extremely important”) had a significant positive effect on their interest in politics when compared to male respondents who did not place a high level of importance on belonging to their specific racial group.
-Further research is needed to understand how an individual’s racial belonging can impact their political participation while controlling for other demographic characteristics.
-
+Further research is needed to understand how an individual’s racial belonging can impact their political participation.
 Following the resource model of participation, age and increases in educational attainment appeared to be associated with higher odds of reporting a higher level of interest in politics.
-The most significant positive predictors of increased levels of interest in politics were an individual’s age and level of educational attainment.
-In particular, individuals who graduated from college had much higher odds of reporting a higher level of interest in politics than individuals with no college experience.
-This follows popular findings that educational attainment can lead to increases in political participation \[8\], and that older Americans vote at higher rates than Americans ages 18-29 years old \[9\].
+Individuals who graduated from college had much higher odds of reporting a higher level of interest in politics than individuals with no college experience.
+This follows findings that educational attainment can lead to increases in political participation \[8\], and that older Americans vote at higher rates than younger Americans \[9\].
 
@@
 
-Despite the value of this survey’s oversamples of racial and ethnic minority communities, some limitations were present in our study.
-For the ordinal logistic regression, we used educational attainment as a proxy for socioeconomic status.
-This is a commonly used proxy measure in behavioral and social sciences \[10\].
-Further research on the influence of socioeconomic status of political participation for racial/ethnic minorities and the use of other measures such as household income or wealth would help us understand if the aforementioned resource model of participation is indeed generalizable to other groups in the American electorate.
-The effect of race on political interest appeared to be more pronounced for members of very small racial/ethnic minority groups in our analysis such as Asian Americans, Native Americans, Native Hawaiians and Pacific Islanders.
+Despite the value of this survey’s oversamples of racial minorities, some limitations were present in our study.
+For the first analysis, we used educational attainment as a proxy for socioeconomic status.
+This is a common practice in social sciences \[10\].
+Further research on the influence of socioeconomic status on political participation for racial minorities would help us understand if the resource model of participation is generalizable to other groups in the U.S.
 Both models had low specificity, suggesting that they struggled to correctly identify cases where respondents had low levels of political interest or respondents who considered anti-discrimination policies to be important.
@@
 
-Overall, these challenges and opportunities highlight the importance of disaggregated data and further research in partnership with small racial and ethnic minority communities within the United States.
+Overall, these challenges and opportunities highlight the importance of disaggregated data and further research in partnership with small racial and ethnic minority communities.
 Future work on political participation and public opinion among racial and ethnic minorities could support researchers, policymakers, and advocates in designing more inclusive policies and interventions.
-Additionally, incorporating qualitative methodologies, such as interviews or community-based participatory research, could provide richer context and amplify the voices of underrepresented communities.
-Longitudinal studies could also complement these efforts, capturing changes over time and offering deeper insights into the lived experiences of marginalized communities.
 By addressing these gaps, future research can better capture the diverse and intersectional realities of minority groups and contribute to more effective advocacy and policy efforts.
@@
 
-### Ordinal Model Overall Statistics
-
-|                           | Metric           |
-|---------------------------|------------------|
+**Table A1: Reference levels for variables**
+
+|              **Variable Group**               |    **Reference Level**    |
+|:----------------------------------:|:----------------------------------:|
+|                   Education                   |   No college experience   |
+|                    Gender                     |            Man            |
+|                      Age                      |      18-29 years old      |
+|                     Race                      |           White           |
+|               Racial Belonging                |   Not At All Important    |
+|                   Party ID                    |        Republican         |
+| Interaction: Racial Belonging & Gender: Woman | Man, Not At All Important |
+
+**Table A2: Ordinal Model Statistics**
+
+|                           |    **Metric**    |
+|:-------------------------:|:----------------:|
 |         Accuracy          |      0.4503      |
@@
 
-### Confusion Matrix
+**Table A3: Confusion Matrix**
 
 |   | **Actual - Very interested in politics** | **Actual -Somewhat interested** | **Actual - Not that interested in politics** | **Actual - Not at all interested in politics** |
-|---------------|---------------|---------------|---------------|---------------|
+|:-------------:|:-------------:|:-------------:|:-------------:|:-------------:|
 | **Predicted - Very interested in politics** | 650 | 505 | 120 | 31 |
@@
 
-### Statistics by Class:
-
-|   | Very interested in politics | Somewhat interested | Not that interested in politics | Not at all interested in politics |
-|---------------|---------------|---------------|---------------|---------------|
-| Sensitivity | 0.15095 | 0.8973 | 0.05572 | 0.10938 |
-| Specificity | 0.95045 | 0.1625 | 0.96950 | 0.98419 |
-| Balanced Accuracy | 0.55070 | 0.5299 | 0.51261 | 0.54678 |
-
-### Odds Ratio Table
-
-| Variable | Odds Ratio | 2.5% (Confidence Interval) | 97.5% (Confidence Interval) |
-|------------------|------------------|------------------|------------------|
-| Q271 (4) Slightly important | 0.9672589 | 0.811845068 | 1.1525850 |
-| Q271 (3) Moderately important | 0.9776608 | 0.845556718 | 1.1303745 |
-| Q271 (2) Very important | 1.1952004 | 1.033019394 | 1.3828451 |
-| Q271 (1) Extremely important | 1.6474969 | 1.423670208 | 1.9066091 |
-| S3B(2) Woman | 0.6294256 | 0.540015144 | 0.7336072 |
-| S3B(3) Non-binary | 1.8576658 | 1.002509318 | 3.4617409 |
-| S3B(4) Something else (Specify) | 2.1359493 | 0.377740111 | 13.5124525 |
-| race Black | 0.8676620 | 0.790376010 | 0.9524806 |
-| race Asian American | 0.5945609 | 0.541964870 | 0.6522143 |
-| race Latino | 0.7946622 | 0.575235415 | 0.7848896 |
-| race AIAN | 0.6719204 | 1.27404185 | 1.7384045 |
-| race NHPI | 0.5606006 | 0.458205702 | 0.6860792 |
-| race Arab/MENA | 1.1589043 | 0.899447122 | 1.4956228 |
-| age 30-39 | 1.0310801 | 0.946696738 | 1.0563119 |
-| age 40-49 | 0.9247623 | 0.98549367 | 1.1229868 |
-| age 50-59 | 1.0979487 | 1.000027913 | 1.2054892 |
-| age 60-69 | 1.3238874 | 1.205240800 | 1.4542874 |
-| age 70+ | 1.6896464 | 1.510226543 | 1.8907050 |
-| educSome college experience | 1.5264694 | 1.402573135 | 1.6614197 |
-| educCollege graduate | 2.2337286 | 2.072162224 | 2.4081069 |
-| Q21(2) Democrat | 1.1333513 | 1.045765624 | 1.2282275 |
-| Q21(3) Independent | 0.5806653 | 0.533801211 | 0.6315870 |
-| Q21(4) Other party | 0.2127592 | 0.182654104 | 0.2477198 |
-| Q271 4) Slightly important:S3B(2) Woman | 0.9955364 | 0.782564309 | 1.2663323 |
-| Q271 (3) Moderately important:S3B(2) Woman | 0.9461281 | 0.780040309 | 1.1475378 |
-| Q271 (2) Very important:S3B(2) Woman | 0.9791675 | 0.810674379 | 1.1826356 |
-| Q271 (1) Extremely important:S3B(2) Woman | 0.8114102 | 0.674275440 | 0.9763646 |
-| Q271 (4) Slightly important:S3B(3) Non-binary | 0.6505015 | 0.198565161 | 2.1505278 |
-| Q271 (3) Moderately important:S3B(3) Non-binary | 0.6563254 | 0.249594245 | 1.7277048 |
-| Q271 (2) Very important:S3B(3) Non-binary | 0.5140828 | 0.205132143 | 1.2905898 |
-| Q271 (1) Extremely important:S3B(3) Non-binary | 0.8770903 | 0.354410521 | 2.1835545 |
-| Q271 (4) Slightly important:S3B(4) Something else (Specify) | 0.2279084 | 0.023857021 | 2.0164000 |
-| Q271 (3) Moderately important:S3B(4) Something else (Specify) | 1.7158304 | 0.146865519 | 19.7516328 |
-| Q271 2) Very important:S3B(4) Something else (Specify) | 0.1049752 | 0.006213617 | 1.4210951 |
-| Q271 1) Extremely important:S3B(4) Something else (Specify) | 0.6584933 | 0.079780177 | 4.9850479 |
+**Table A4: Statistics by Class**
+
+|   | **Very interested in politics** | **Somewhat interested** | **Not that interested in politics** | **Not at all interested in politics** |
+|:-------------:|:-------------:|:-------------:|:-------------:|:-------------:|
+| **Sensitivity** | 0.15095 | 0.8973 | 0.05572 | 0.10938 |
+| **Specificity** | 0.95045 | 0.1625 | 0.96950 | 0.98419 |
+| **Balanced Accuracy** | 0.55070 | 0.5299 | 0.51261 | 0.54678 |
+
+**Table A5: Odds Ratio Table**
+
+| **Variable** | **Odds Ratio** | **2.5% (Confidence Interval)** | **97.5% (Confidence Interval)** |
+|:----------------:|:----------------:|:----------------:|:----------------:|
+| Racial Belonging: Slightly important | 0.9672589 | 0.811845068 | 1.1525850 |
+| Racial Belonging: Moderately important | 0.9776608 | 0.845556718 | 1.1303745 |
+| Racial Belonging: Very important | 1.1952004 | 1.033019394 | 1.3828451 |
+| Racial Belonging: Extremely important | 1.6474969 | 1.423670208 | 1.9066091 |
+| Gender: Woman | 0.6294256 | 0.540015144 | 0.7336072 |
+| Gender: Non-binary | 1.8576658 | 1.002509318 | 3.4617409 |
+| Gender: Something else (Specify) | 2.1359493 | 0.377740111 | 13.5124525 |
+| Race: Black | 0.8676620 | 0.790376010 | 0.9524806 |
+| Race: Asian American | 0.5945609 | 0.541964870 | 0.6522143 |
+| Race: Latino | 0.7946622 | 0.575235415 | 0.7848896 |
+| Race: AIAN | 0.6719204 | 1.27404185 | 1.7384045 |
+| Race: NHPI | 0.5606006 | 0.458205702 | 0.6860792 |
+| Race: Arab/MENA | 1.1589043 | 0.899447122 | 1.4956228 |
+| Age: 30-39 | 1.0310801 | 0.946696738 | 1.0563119 |
+| Age: 40-49 | 0.9247623 | 0.98549367 | 1.1229868 |
+| Age: 50-59 | 1.0979487 | 1.000027913 | 1.2054892 |
+| Age: 60-69 | 1.3238874 | 1.205240800 | 1.4542874 |
+| Age: 70+ | 1.6896464 | 1.510226543 | 1.8907050 |
+| Education: Some college experience | 1.5264694 | 1.402573135 | 1.6614197 |
+| Education: College graduate | 2.2337286 | 2.072162224 | 2.4081069 |
+| Party ID: Democrat | 1.1333513 | 1.045765624 | 1.2282275 |
+| Party ID: Independent | 0.5806653 | 0.533801211 | 0.6315870 |
+| Party ID: Other party | 0.2127592 | 0.182654104 | 0.2477198 |
+| Racial Belonging: Slightly important: Gender: Woman | 0.9955364 | 0.782564309 | 1.2663323 |
+| Racial Belonging: Moderately important: Gender: Woman | 0.9461281 | 0.780040309 | 1.1475378 |
+| Racial Belonging: Very important: Gender: Woman | 0.9791675 | 0.810674379 | 1.1826356 |
+| Racial Belonging: Extremely important: Gender: Woman | 0.8114102 | 0.674275440 | 0.9763646 |
+| Racial Belonging: Slightly important: Gender: Non-binary | 0.6505015 | 0.198565161 | 2.1505278 |
+| Racial Belonging: Moderately important: Gender: Non-binary | 0.6563254 | 0.249594245 | 1.7277048 |
+| Racial Belonging: Very important: Gender: Non-binary | 0.5140828 | 0.205132143 | 1.2905898 |
+| Racial Belonging: Extremely important: Gender: Non-binary | 0.8770903 | 0.354410521 | 2.1835545 |
+| Racial Belonging: Slightly important Gender: Something else (Specify) | 0.2279084 | 0.023857021 | 2.0164000 |
+| Racial Belonging: Moderately important: Gender: Something else (Specify) | 1.7158304 | 0.146865519 | 19.7516328 |
+| Racial Belonging: Very important: Gender: Something else (Specify) | 0.1049752 | 0.006213617 | 1.4210951 |
+| Racial Belonging: Extremely important: Gender: Something else (Specify) | 0.6584933 | 0.079780177 | 4.9850479 |
 
 ## Research Question 2
-
-### VIF Table
 
@@
 
-|   | Normalized VIF: Base Model | Normalized VIF: Interaction Model |
-|------------------------|------------------------|------------------------|
+**Table B1: Normalized VIF Values**
+
+|   | **Normalized VIF: Base Model** | **Normalized VIF: Interaction Model** |
+|:----------------------:|:----------------------:|:----------------------:|
 | Gender | 1.022 | 1.978 |
@@
 
-### Odds Ratio Table
-
-```{r}
-analysis_table
-```
-
-## Table CONFUSION: Confusion Matrix for Base model
-
-|                                      |                   |               |
-|--------------------------------------|-------------------|---------------|
-| [**Predicted / Actual**]{.underline} | **Not Important** | **Important** |
-| **Not Important**                    | 12361             | 3812          |
-| **Important**                        | 666               | 706           |
-
-### Table MODEL: Base Model Statistics
-
-|                           | Metric  |
-|---------------------------|---------|
+**Table B2: Odds Ratios**
+
+| **Description** | **Odds Ratio** | **Lower Confidence Interval** | **Upper Confidence Interval** | **P-Value** |
+|:-------------:|:-------------:|:-------------:|:-------------:|:-------------:|
+| Intercept | 0.1695 | 0.1300 | 0.2203 | 9.84e-40 |
+| Gender: Women | 1.2507 | 1.1598 | 1.3490 | 6.41e-9 |
+| Gender: Non-Binary | 1.3434 | 0.9160 | 1.9517 | 1.25e-1 |
+| Gender: Something Else | 1.6771 | 0.7294 | 3.6061 | 2.00e-1 |
+| Treated Unfairly Or Discriminated Against: No | 0.8970 | 0.8270 | 0.9730 | 8.79e-3 |
+| Discrimination Impact On Life Satisfaction: Some | 1.0452 | 0.8981 | 1.2172 | 5.69e-1 |
+| Discrimination Impact On Life Satisfaction: A Little | 1.0053 | 0.8701 | 1.1626 | 9.43e-1 |
+| Discrimination Impact On Life Satisfaction: Not At All | 0.7362 | 0.6337 | 0.8561 | 6.58e-5 |
+| Discrimination Impact On Life Satisfaction: Don't Know | 0.8978 | 0.7439 | 1.0831 | 2.61e-1 |
+| Discrimination Impact On Life Satisfaction: Refused | 0.8404 | 0.5695 | 1.2177 | 3.69e-1 |
+| Perceptions Of Discrimination Against Whites: Some | 1.0698 | 0.8812 | 1.3012 | 4.97e-1 |
+| Perceptions Of Discrimination Against Whites: A Little | 1.4965 | 1.2609 | 1.7824 | 4.94e-6 |
+| Perceptions Of Discrimination Against Whites: None At All | 2.0482 | 1.7390 | 2.4214 | 2.00e-17 |
+| Perceptions Of Discrimination Against Whites: Don't Know | 1.5124 | 1.1525 | 1.9811 | 2.74e-3 |
+| Perceptions Of Discrimination Against Blacks: Some | 0.6710 | 0.6023 | 0.7472 | 4.00e-13 |
+| Perceptions Of Discrimination Against Blacks: A Little | 0.5524 | 0.4684 | 0.6504 | 1.32e-12 |
+| Perceptions Of Discrimination Against Blacks: None At All | 0.5423 | 0.4191 | 0.6979 | 2.52e-6 |
+| Perceptions Of Discrimination Against Blacks: Don't Know | 0.3797 | 0.2609 | 0.5475 | 3.00e-7 |
+| Perceptions Of Discrimination Against Asian Americans: Some | 0.8667 | 0.7825 | 0.9599 | 6.05e-3 |
+| Perceptions Of Discrimination Against Asian Americans: A Little | 0.7756 | 0.6799 | 0.8845 | 1.52e-4 |
+| Perceptions Of Discrimination Against Asian Americans: None At All | 0.7710 | 0.6188 | 0.9580 | 1.97e-2 |
+| Perceptions Of Discrimination Against Asian Americans: Don't Know | 0.8565 | 0.6102 | 1.1936 | 3.65e-1 |
+| Perceptions Of Discrimination Against Native Americans: Some | 1.1117 | 0.9991 | 1.2372 | 5.22e-2 |
+| Perceptions Of Discrimination Against Native Americans: A Little | 1.0371 | 0.9143 | 1.1763 | 5.71e-1 |
+| Perceptions Of Discrimination Against Native Americans: None At All | 0.8500 | 0.7074 | 1.0193 | 8.11e-2 |
+| Perceptions Of Discrimination Against Native Americans: Don't Know | 0.8513 | 0.6600 | 1.0930 | 2.11e-1 |
+| Perceptions Of Discrimination Against Immigrants: Some | 0.9502 | 0.8560 | 1.0548 | 3.38e-1 |
+| Perceptions Of Discrimination Against Immigrants: A Little | 0.8910 | 0.7621 | 1.0410 | 1.47e-1 |
+| Perceptions Of Discrimination Against Immigrants: None At All | 0.9524 | 0.7299 | 1.2371 | 7.17e-1 |
+| Perceptions Of Discrimination Against Immigrants: Don't Know | 1.0978 | 0.7820 | 1.5323 | 5.86e-1 |
+| Perceptions Of Discrimination Against Latinos: Some | 0.9670 | 0.8689 | 1.0763 | 5.39e-1 |
+| Perceptions Of Discrimination Against Latinos: A Little | 0.9118 | 0.7873 | 1.0559 | 2.18e-1 |
+| Perceptions Of Discrimination Against Latinos: None At All | 0.7300 | 0.5653 | 0.9387 | 1.50e-2 |
+| Perceptions Of Discrimination Against Latinos: Don't Know | 0.8705 | 0.6080 | 1.2386 | 4.45e-1 |
+| Perceptions Of Discrimination Against Gays And Lesbians: Some | 0.9924 | 0.8987 | 1.0959 | 8.80e-1 |
+| Perceptions Of Discrimination Against Gays And Lesbians: A Little | 0.9305 | 0.8106 | 1.0677 | 3.05e-1 |
+| Perceptions Of Discrimination Against Gays And Lesbians: None At All | 0.9859 | 0.8012 | 1.2100 | 8.93e-1 |
+| Perceptions Of Discrimination Against Gays And Lesbians: Don't Know | 1.2728 | 0.9666 | 1.6698 | 8.35e-3 |
+| Perceptions Of Discrimination Against Muslims: Some | 0.9928 | 0.8937 | 1.1029 | 8.93e-1 |
+| Perceptions Of Discrimination Against Muslims: A Little | 1.0623 | 0.9122 | 1.2364 | 4.36e-1 |
+| Perceptions Of Discrimination Against Muslims: None At All | 1.0402 | 0.8131 | 1.3261 | 7.52e-1 |
+| Perceptions Of Discrimination Against Muslims: Don't Know | 0.8257 | 0.5955 | 1.1374 | 2.46e-1 |
+| Education: Some College Experience | 0.9806 | 0.8774 | 1.0959 | 7.30e-1 |
+| Education: College Graduate | 0.9402 | 0.8514 | 1.0387 | 2.24e-1 |
+| Sexual Orientation: Queer | 1.1197 | 0.9914 | 1.2632 | 6.73e-2 |
+| Sexual Orientation: Refused | 0.8894 | 0.6647 | 1.1760 | 4.20e-1 |
+| Community Type: Urban/Suburban | 1.0465 | 0.9079 | 1.2090 | 5.34e-1 |
+| Race: Black | 3.5331 | 3.1265 | 3.9983 | 4.83e-90 |
+| Race: Asian American | 2.1506 | 1.8896 | 2.4504 | 7.08e-31 |
+| Race: Latino | 1.7392 | 1.5267 | 1.9831 | 1.09e-16 |
+| Race: AIAN | 1.2235 | 0.9679 | 1.5365 | 8.69e-2 |
+| Race: NHPI | 2.1618 | 1.6465 | 2.8179 | 1.81e-8 |
+| Race: Arab/MENA | 2.6661 | 1.9547 | 3.6127 | 3.71e-10 |
+| Age: 30-39 Years | 0.9447 | 0.8443 | 1.0569 | 3.21e-1 |
+| Age: 40-49 Years | 0.8969 | 0.7937 | 1.0132 | 8.08e-2 |
+| Age: 50-59 Years | 0.9589 | 0.8476 | 1.0845 | 5.05e-1 |
+| Age: 60-69 Years | 0.8891 | 0.7833 | 1.0089 | 6.88e-2 |
+| Age: 70+ Years | 1.0035 | 0.8612 | 1.1681 | 9.64e-1 |
+
+**Table B3: Reference Levels**
+
+| **Variable Group** | **Reference Level** |
+|:----------------------------------:|:----------------------------------:|
+| Education | No college experience |
+| Gender | Man |
+| Age | 18-29 years old |
+| Race | White |
+| Community Type | Rural |
+| Sexual Orientation | Straight |
+| Treated Unfairly/Discriminated Against | Yes |
+| Discrimination Impact on Life Satisfaction | A lot |
+| Perceptions of Discrimination Against Whites | A lot |
+| Perceptions of Discrimination Against Blacks | A lot |
+| Perceptions of Discrimination Against Asian Americans | A lot |
+| Perceptions of Discrimination Against Native Americans | A lot |
+| Perceptions of Discrimination Against Immigrants | A lot |
+| Perceptions of Discrimination Against Latinos | A lot |
+| Perceptions of Discrimination Against Gays and Lesbians | A lot |
+| Perceptions of Discrimination Against Muslims | A lot |
+
+**Table B4: Summary Statistics**
+
+|                           | **Metric** |
+|:-------------------------:|:----------:|
 |   Sensitivity (Recall)    |   0.1563   |

### 2025-01-08 peter_plots.qmd

@@
 ##
+
 # Packages
+
 ```{r}
@@
 # Loading data
+
 ```{r}
@@
 # plot 1:
+
 Race and importance of being a member of racial group to identity (Q271)
+
 ```{r}
@@
 
-  # Save each plot as a separate image file (optional)
-  #ggsave(paste0("plot_", race_value, "Q271.png"), plot = p, width = 6, height = 4)
+  # Save each plot as a separate image file 
+  ggsave(paste0("plot_", race_value, "Q271.png"), plot = p, width = 6, height = 4)
 
-  # Print the plot (optional, if you want to view them in the R console)
+  # Print the plot
   print(p)
@@
 # plot 2:
+
 Race and interest in politics (Q29)
+
 ```{r}
@@
   # Save each plot as a separate image file (optional)
-  #ggsave(paste0("plot_", race_value, "Q29.png"), plot = p, width = 6, height = 4)
+  ggsave(paste0("plot_", race_value, "Q29.png"), plot = p, width = 6, height = 4)
 
@@
 # plot 3:
+
 White/Non-White and importance of race to identity (Q271)
+
 ```{r}
@@
 
+ggsave("plot_whiteQ271.png", plot = plot_whiteQ271, width = 6, height = 4)
+
 print(plot_whiteQ271)
@@
 
+ggsave("plot_nonwhiteQ271.png", plot = plot_nonwhiteQ271, width = 6, height = 4)
+
 print(plot_nonwhiteQ271)
@@
 # plot:
+
 White/Non-White and interest in politics (Q29)
+
 ```{r}
@@
 
+ggsave("plot_whiteQ29.png", plot = plot_whiteQ29, width = 6, height = 4)
+
 print(plot_whiteQ29)
@@
 
+ggsave("plot_nonwhiteQ29.png", plot = plot_nonwhiteQ29, width = 6, height = 4)
+
 print(plot_nonwhiteQ29)
@@
 # plot 3:
+
 College Experience and importance of being a member of racial group to identity (Q271)
+
 ```{r}
@@
 
-plot_collegegrad <- data_summary %>% filter(educ == "College graduate") %>% ggplot(aes(x = Q271, y = proportion, fill = Q271)) +
+plot_collegegradQ271 <- data_summary %>% filter(educ == "College graduate") %>% ggplot(aes(x = Q271, y = proportion, fill = Q271)) +
     geom_bar(stat = "identity", position = "dodge") +
@@
 
+ggsave("plot_collegegradQ271.png", plot = plot_collegegradQ271, width = 6, height = 4)
+
 print(plot_collegegradQ271)
 
-plot_somecollege <- data_summary %>% filter(educ == "Some college experience") %>% ggplot(aes(x = Q271, y = proportion, fill = Q271)) +
+plot_somecollegeQ271 <- data_summary %>% filter(educ == "Some college experience") %>% ggplot(aes(x = Q271, y = proportion, fill = Q271)) +
     geom_bar(stat = "identity", position = "dodge") +
@@
 
+ggsave("plot_somecollegeQ271.png", plot = plot_somecollegeQ271, width = 6, height = 4)
 print(plot_somecollegeQ271)
 
-plot_nocollege <- data_summary %>% filter(educ == "No college experience") %>% ggplot(aes(x = Q271, y = proportion, fill = Q271)) +
+plot_nocollegeQ271 <- data_summary %>% filter(educ == "No college experience") %>% ggplot(aes(x = Q271, y = proportion, fill = Q271)) +
     geom_bar(stat = "identity", position = "dodge") +
@@
 
+ggsave("plot_nocollegeQ271.png", plot = plot_nocollegeQ271, width = 6, height = 4)
+
 print(plot_nocollegeQ271)
@@
 # plot 4:
+
 College Experience and interest in politics (Q29)
+
 ```{r}
@@
 
-plot_collegegrad <- data_summary %>% filter(educ == "College graduate") %>% ggplot(aes(x = Q29, y = proportion, fill = Q29)) +
+plot_collegegradQ29 <- data_summary %>% filter(educ == "College graduate") %>% ggplot(aes(x = Q29, y = proportion, fill = Q29)) +
     geom_bar(stat = "identity", position = "dodge") +
@@
 
+ggsave("plot_collegegradQ29.png", plot = plot_collegegradQ29, width = 6, height = 4)
 print(plot_collegegradQ29)
 
-plot_somecollege <- data_summary %>% filter(educ == "Some college experience") %>% ggplot(aes(x = Q29, y = proportion, fill = Q29)) +
+plot_somecollegeQ29 <- data_summary %>% filter(educ == "Some college experience") %>% ggplot(aes(x = Q29, y = proportion, fill = Q29)) +
     geom_bar(stat = "identity", position = "dodge") +
@@
 
+ggsave("plot_somecollegeQ29.png", plot = plot_somecollegeQ29, width = 6, height = 4)
 print(plot_somecollegeQ29)
 
-plot_nocollege <- data_summary %>% filter(educ == "No college experience") %>% ggplot(aes(x = Q29, y = proportion, fill = Q29)) +
+plot_nocollegeQ29 <- data_summary %>% filter(educ == "No college experience") %>% ggplot(aes(x = Q29, y = proportion, fill = Q29)) +
     geom_bar(stat = "identity", position = "dodge") +
@@
 
+ggsave("plot_nocollegeQ29.png", plot = plot_nocollegeQ29, width = 6, height = 4)
 print(plot_nocollegeQ29)
@@
 # plot 4:
+
 Age and importance of being a member of racial group to identity (Q271)
+
 ```{r}
 data_summary <- dataset_clean %>%
-  group_by(S5_AGE, Q271) %>%
+  group_by(age_recoded, Q271) %>%
   summarise(count = n()) %>%
@@
 
-age_groups <- unique(data_summary$S5_AGE)
+age_groups <- unique(data_summary$age_recoded)
 
@@
 for (age_value in age_groups) {
-  plot_data <- data_summary %>% filter(S5_AGE == age_value)
+  plot_data <- data_summary %>% filter(age_recoded == age_value)
 
@@
   # Save each plot as a separate image file (optional)
-  #ggsave(paste0("plot_", age_value, "Q271.png"), plot = p, width = 6, height = 4)
+  ggsave(paste0("plot_", age_value, "Q271.png"), plot = p, width = 6, height = 4)
 
@@
 # plot 5:
+
 Age and interest in politics (Q29)
+
 ```{r}
 data_summary <- dataset_clean %>%
-  group_by(S5_AGE, Q29) %>%
+  group_by(age_recoded, Q29) %>%
   summarise(count = n()) %>%
@@
   # Save each plot as a separate image file 
-  #ggsave(paste0("plot_", age_value, "Q29.png"), plot = p, width = 6, height = 4)
+  ggsave(paste0("plot_", age_value, "Q29.png"), plot = p, width = 6, height = 4)
 
@@
 # plot 6:
+
 Rurality and importance of being a member of racial group to identity (Q271)
+
 ```{r}
@@
   # Save each plot as a separate image file 
-  #ggsave(paste0("plot_", rural_value, "Q271.png"), plot = p, width = 6, height = 4)
+  ggsave(paste0("plot_", rural_value, "Q271.png"), plot = p, width = 6, height = 4)
 
@@
 # plot 7:
+
 Rurality and interest in politics (Q29)
+
 ```{r}
@@
   # Save each plot as a separate image file 
-  #ggsave(paste0("plot_", rural_value, "Q29.png"), plot = p, width = 6, height = 4)
+  ggsave(paste0("plot_", rural_value, "Q29.png"), plot = p, width = 6, height = 4)
 

### 2025-01-08 draft_finalreport.qmd

@@
     labs(
-      title = "Significant Variables with Confidence Intervals",
+      title = "Odds Ratios (Significant Variables)",
       x = "Odds Ratio",
@@
   labs(
-    title = "Odds Ratio of Significant Variables",
+    title = "Odds Ratios (Significant Variables)",
     x = "Odds Ratio",

### 2025-01-08 proposal_EDA/deGuzman_IDS702Projectproposal_template.qmd

@@
   scale_y_continuous(labels = scales::percent) +
-  theme_minimal() +
-  geom_text(aes(label = scales::percent(after_stat(count)/sum(after_stat(count)))),
-            stat = "count", 
-            position = position_fill(vjust = 0.5),
-            color = "white")
+  theme_minimal() 
 ```

### 2025-01-09 RQ1_Analysis/peter_EDA.qmd

--- 2024-12-06 peter_EDA.qmd
@@
 | **Variable** | **Level** | **Count** | **Proportion** |
-|------------------|-------------------|------------------|------------------|
+|----|----|----|----|
 | Q29 | \(1\) Very interested in politics | 4306 | 25% |

### 2025-01-09 RQ1_Analysis/peter_models.qmd

--- 2024-12-15 peter_models.qmd
@@
 
+### TESTING PLOT
+
+```{r}
+  or_data <- exp(cbind(OR = coef(ord_mod_interactions), confint(ord_mod_interactions)))
+
+  # Calculate p-values
+  p_values <- pnorm(-abs(summary(ord_mod_interactions)$coefficients[, "t value"])) * 2
+
+  # Keep only significant variables
+  sig_indices <- p_values[names(coef(ord_mod_interactions))] < 0.05
+  or_data_sig <- or_data[sig_indices, ]
+
+  # Create a mapping for variable names with grouping prefixes
+  rename_map <- c(
+    # Education Group
+    "educCollege graduate" = "1. Education: College Graduate",
+    "educSome college experience" = "1. Education: Some College",
+
+    # Gender Group
+    "S3B(2) Woman" = "2. Gender: Woman",
+    "S3B(3) Non-binary" = "2. Gender: Non-binary",
+
+    # Age Group
+    "age_recoded70+" = "3. Age: 70+",
+    "age_recoded60-69" = "3. Age: 60-69",
+    "age_recoded50-59" = "3. Age: 50-59",
+
+    # Race Group
+    "race_recodedBlack" = "4. Race: Black",
+    "race_recodedAsian American" = "4. Race: Asian",
+    "race_recodedLatino" = "4. Race: Latino",
+    "race_recodedAIAN" = "4. Race: Native",
+    "race_recodedNHPI" = "4. Race: NHPI",
+
+    # Racial Belonging Group
+    "Q271_reordered(2) Very important" = "5. Racial Belonging: Very Important",
+    "Q271_reordered(1) Extremely important" = "5. Racial Belonging: Extremely Important",
+
+    # Party ID Group
+    "Q21(2) Democrat" = "6. Party ID: Democrat",
+    "Q21(3) Independent" = "6. Party ID: Independent",
+    "Q21(4) Other party" = "6. Party ID: Other",
+
+    # Interactions
+    "Q271_reordered(1) Extremely important:S3B(2) Woman" = "7. Interaction: Racial Belonging & Gender: Woman"
+  )
+
+  # Rename the rownames in the significant data
+  or_data_sig_renamed <- as.data.frame(or_data_sig)  # Ensure this is a data frame
+  rownames(or_data_sig_renamed) <- rename_map[rownames(or_data_sig_renamed)]
+
+  # Create a custom grouping variable based on the row names
+  group_labels <- c(
+    "Racial Belonging" = "5. Racial Belonging: Very Important", 
+    "Racial Belonging" = "5. Racial Belonging: Extremely Important",
+    "Gender" = "2. Gender: Woman", 
+    "Gender" = "2. Gender: Non-binary", 
+    "Race" = "4. Race: Black", 
+    "Race" = "4. Race: Asian", 
+    "Race" = "4. Race: Latino", 
+    "Race" = "4. Race: Native", 
+    "Race" = "4. Race: NHPI", 
+    "Age" = "3. Age: 50-59", 
+    "Age" = "3. Age: 60-69", 
+    "Age" = "3. Age: 70+", 
+    "Education" = "1. Education: Some College", 
+    "Education" = "1. Education: College Graduate", 
+    "Party ID" = "6. Party ID: Democrat", 
+    "Party ID" = "6. Party ID: Independent", 
+    "Party ID" = "6. Party ID: Other", 
+    "Interaction" = "7. Interaction: Racial Belonging & Gender: Woman"
+  )
+
+  # Add the grouping variable to the data frame
+  or_data_sig_renamed$group <- group_labels[rownames(or_data_sig_renamed)]
+
+  # Reorder y-axis based on group labels (so they appear grouped and ordered)
+  # or_data_sig_renamed$variable <- factor(rownames(or_data_sig_renamed), 
+  #                                         levels = rownames(or_data_sig_renamed)[order(or_data_sig_renamed$OR)])
+  #
+
+```
+
 # Fitting OLR Model with Interaction Term (Q271 \* Q21 Party ID)

### 2025-01-12 Project Proposal/_IDS702Projectproposal_Main.qmd

---
title: "Project Proposal"
author: "Peter de Guzman, Sizhe Chen, Jennifer Li, Nakiyah Dhariwala"
subtitle: "Due November 17 at 11:59pm"
format: pdf
editor: 
  markdown: 
    wrap: 72
---

#### Load Packages

```{r load-packages, message = FALSE, warning = FALSE, echo=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(readr)
library(ggridges)
library(readxl)
library(formattable)
library(scales)
```

# Dataset 1 (top choice)

**Data source:** The 2020 Collaborative Multi-racial Post-Election
Survey (CMPS)

**Brief description:** The survey has an unweighted sample size of
17,545 respondents. Each row represents a unique respondent. The data is
collected through an online survey with 800 unique questions. The
survey's main focus is on attitudes about the 2020 Election and
candidates, experiences with racism, policy attitudes, immigration, and
personal experiences with civic engagement across many facets of
American life.

The 2020 CMPS was offered in English, Spanish, Chinese (simplified),
Chinese (traditional), Korean, Vietnamese, Arabic, Urdu, Farsi and
Haitian Creole. The survey dataset includes Black, White, Latino and
Asian respondents as well as additional oversamples of respondents from
hard-to-reach populations including, Afro-Latinos, Black immigrants,
Native Americans, Native Hawaiians, Muslims and people who identify as
LGBTQ.

**Research question 1:** Are Americans who believe their race is
important to their identity more likely to be interested in politics
than those that are not?

-   Outcome variable (include the name/description and type of
    variable): The outcome variable is an ordinal variable labeled Q29.
    The survey question is "Some people are very interested in politics
    while other people can't stand politics how about you? are you...".
    The response options are in a five-point scale from "Not at all
    interested in politics" to "Very interested in politics."

The main predictor variable is "Q271" which asks “How important is being
\[RACIAL GROUP\] to your identity?”. The response options are a
five-point scale from "Not at all important" to "Extremely important".
Other potential predictors include survey questions on respondent's
self-perception of race and discrimination, and their own attitudes
towards other racial groups. In addition, there are multiple variables
that could be used as demographic controls such as age, gender,
education, income, and party identification. The variables not shown in
the output include responses to the rest of the 800 survey questions,
covering topics such as policy attitudes, civic engagement, and
political participation.

**Variable Type**: Ordinal

**Research question 2:** Controlling for other demographic variables
such as age and gender, how do experiences with racism influence the
importance respondents place on policy issues related to racial and
ethnic discrimination?

-   Outcome variable (include the name/description and type of
    variable): Stopping discrimination against racial/ethnic minorities
    (Q1R5) — a binary variable indicating whether respondents view it as
    an important issue.

**Variable Type:** Binary.

**Interaction Term Idea:** Include an interaction term between
race/ethnicity (S2RACER) and experiences with racism to see if the
influence of racism experiences differs across racial/ethnic groups.

**Load the data and provide a `glimpse()`**:

```{r data-1, warning=FALSE, echo=FALSE}
rm(list=ls())
load("39096-0001-Data.rda")
data <- da39096.0001
reordered_data <- data %>% relocate(Q29, Q271, WEIGHT, 
                                    AGE1_HID, AGE2_HID, .after = DATE)
reordered_data %>% slice_head(n = 50) %>% str()
```

\newpage

**Exploratory Plots:**
Dataset 1: Research Question 1
```{r plots-1.1, echo=FALSE, warning=FALSE}
data <- data %>% mutate(Q29_shortened = case_when(
  Q29 == "(1) Very interested in politics" ~ "Very interested",

  Q29 == "(2) Somewhat interested" ~ "Somewhat interested",

  Q29 == "(3) Not that interested in politics" ~ "Not that interested",

  Q29 == "(4) Not at all interested in politics" ~ 
    "Not at all interested", .default = "Other"))

wtd_tableQ29 <- data %>% count(Q29_shortened, wt=WEIGHT)
wtd_tableQ29 <- wtd_tableQ29 %>% mutate('pct' = (n/sum(n, na.rm=TRUE)))
ggplot(wtd_tableQ29, aes(x=Q29_shortened, y = pct)) +
  geom_bar(stat="identity", fill = "orange") +
  geom_text(aes(label=percent(pct)), vjust=1.6, color="white", size =3.5) +
  labs(title = "Reported Interest in Politics", 
       x = "Level of Interest", y = "Percent (Weighted)") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

contingency_table <- table(data$Q29_shortened, data$Q271)
contingency_df <- as.data.frame(contingency_table)
contingency_df <- contingency_df %>% mutate(Share = (Freq/sum(Freq)))

ggplot(contingency_df, aes(x = Var1, y = Var2, fill = Share)) +
  geom_tile() +
  labs(title="Interest in Politics and Racial Identity", 
       x = "Interest in Politics", 
       y = "Importance of Racial Identity") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 0.4, vjust = 0.5)) +
  scale_fill_continuous(low = "white", high = "blue", labels = percent_format(scale = 100)) 
```

\newpage
Dataset 1: Research Question 2
```{r plots-1.2, echo=FALSE, warning=FALSE}

data$Q1R5_cleaned <- as.numeric(sub("^\\((\\d+)\\).*", "\\1", data$Q1R5))
data$Q1R5_cleaned <- factor(data$Q1R5_cleaned, levels = c(0, 1), labels = c("Not Important", "Important"))

# Plot the distribution of Q1R5
ggplot(data, aes(x = Q1R5_cleaned)) +
  geom_bar(fill = "skyblue") +
  labs(title = "Importance of Stopping Racial Discrimination",
       x = "Importance Level",
       y = "Count") +
  theme_minimal()

data$MAIN_ETH_QUOTA_numeric <- as.numeric(sub("^\\((\\d+)\\).*", "\\1", 
                                              data$MAIN_ETH_QUOTA))
data$Race_Ethnicity <- factor(data$MAIN_ETH_QUOTA_numeric,
                              levels = 1:7,
                              labels = c("White", "Latino", "Black", 
                                         "AAPI", "ANAI", "MENA", "NH"))

# Plot with short-term race labels on the x-axis and percent labels on the bars
ggplot(data, aes(x = Race_Ethnicity, fill = Q1R5_cleaned)) +
  geom_bar(position = "fill") +
  labs(title = "Importance of Stopping Discrimination by Race/Ethnicity",
       x = "Race/Ethnicity",
       y = "Proportion",
       fill = "Importance Level",
       caption = "Race/Ethnicity Abbreviations: ANAI = Native American, NH = Native Hawaiian, AAPI = Asian American/Pacific Islander, MENA = Middle Eastern/North African") +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() 
```

\newpage

# Dataset 2

**Data source:** Global Health Observatory (GHO) data repository under
the World Health Organization (WHO), with additional economic data
sourced from the United Nations website.
https://www.kaggle.com/datasets/kumarajarshi/life-expectancy-who

**Brief description:** There are 2938 rows and 22 columns, representing
health and social/economic indicators for 193 countries over 15 years.
Each row corresponds to a specific country-year combination.

**Research question 1:** How does the number of years a person spends in school 
influence whether a country is classified as developed or developing?

-   Outcome variable (include the name/description and type of
    variable): "Status" --\> Binary variable that categorizes countries
    into "Developed" or "Developing."

**Research question 2:** How do factors such as life expectancy,
education, healthcare expenditure, and population affect GDP across
developing countries?

-   Outcome variable (include the name/description and type of
    variable): GDP: Numeric Variable

**Load the data and provide a `glimpse()`**:

```{r data-2, warning=FALSE, echo=FALSE}
data <- read.csv("Life Expectancy Data.csv")
glimpse(data)
```

\newpage

**Exploratory Plots:**
Dataset 2: Research Question 1
```{r plots-2.1, echo=FALSE, warning=FALSE}
# Bar chart showing the counts of each Status
ggplot(data, aes(x = Status, fill = Status)) +
  geom_bar(alpha = 0.8) +
  labs(
    title = "Distribution of Countries by Status",
    x = "Country Status",
    y = "Number of Countries",
    fill = "Country Status"
  ) +
  theme_minimal()

# Ridge plot for Schooling
ggplot(data, aes(x = Schooling, y = Status, fill = Status)) +
  geom_density_ridges(alpha = 0.7) +
  labs(
    title = "Distribution of years of schooling by Country Status",
    x = "Schooling",
    y = "Country Status"
  ) +
  theme_minimal()
```

\newpage
Dataset 2: Research Question 2
```{r plots-2.2, echo=FALSE, warning=FALSE}
developing_data <- subset(data, Status == "Developing")

# Histogram showing the distribution of GDP
ggplot(developing_data, aes(x = GDP)) +
  geom_histogram(binwidth = 100, 
                 fill = "skyblue", 
                 color = "black", 
                 alpha = 0.7) +
  labs(title = "Distribution of GDP",
       x = "GDP",
       y = "Frequency") + theme_minimal()

# 
ggplot(developing_data, aes(x = Schooling, y = GDP)) +
  geom_point(color = "blue", size = 2, alpha = 0.5) +
  labs(title = "Relationship Between years of education and GDP",
       x = "Years of education",
       y = "GDP") + theme_minimal()
```

\newpage

# Dataset 3 (optional)

**Data source:**
https://www.kaggle.com/datasets/prasoonkottarathil/polycystic-ovary-syndrome-pcos

**Brief description:** Each row in the dataset represents data of a
female. Specifically, it contains medical, physical, hormonal, and
lifestyle information collected for that woman as part of a study on
PCOS (Polycystic Ovary Syndrome) and related health metrics.

**Research question 1:** - What combination of physical characteristics
(e.g., BMI, waist-hip ratio), hormonal factors (e.g., AMH, FSH/LH
ratio), and symptoms (e.g., pimples, hair loss) significantly predict
PCOS status (Y/N)?

-   Outcome variable (include the name/description and type of
    variable): PCOS Status

Categorical (Binary: 1 = Has PCOS, 0 = Does not have PCOS) -
Description: Indicates whether the individual has PCOS or not.

**Research question 2:** How do physical factors and hormonal factors
levels influence Hb (hemoglobin) levels?

-   Outcome variable (include the name/description and type of
    variable): Hb (Hemoglobin) Levels

-   Type: Continuous (measured in g/dL)

-   Description:The concentration of hemoglobin in the blood, which can
    be an indicator of overall health, anemia, and other medical
    conditions.

**Load the data and provide a `glimpse()`**:

```{r data-3, warning=FALSE, echo=FALSE}
data <- read_excel("PCOS_data_without_infertility.xlsx", sheet = 2)
glimpse(data)
data <- data %>%
  rename(PCOS = `PCOS (Y/N)`, 
         Weight=`Weight (Kg)`, 
         Hb=`Hb(g/dl)`)
data$PCOS <- ifelse(data$PCOS == 1, "Yes", "No")
data$PCOS <- factor(data$PCOS, levels = c("No", "Yes"))
```

\newpage

**Exploratory Plots:**
Dataset 3: Research Question 1
```{r plots-3.1, echo=FALSE}
# Summarize data
pcos_summary <- data %>%
  count(PCOS)

# Create the pie chart
ggplot(pcos_summary, aes(x = "", y = n, fill = PCOS)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  labs(title = "PCOS Distribution") +
  theme_void() +
  scale_fill_manual(values = c("skyblue", "salmon"))

# Boxplot comparing Weight by PCOS status
ggplot(data, aes(x = PCOS, y = Weight, fill = PCOS)) +
  geom_boxplot() +
  labs(
    title = "Boxplot of Weight by PCOS Status",
    x = "PCOS Status (1 = Yes, 0 = No)",
    y = "Weight (kg)"
  ) +
  # Custom x-axis labels
  scale_x_discrete(labels = c("0" = "No", "1" = "Yes")) +  
  scale_fill_manual(values = c("#F8766D", "#00BFC4")) +    
  theme_minimal() +
  theme(legend.position = "none")
```

\newpage

Dataset 3: Research Question 2
```{r plots-3.2, echo=FALSE}
# Histogram of Hb levels
ggplot(data, aes(x = Hb)) +
  geom_histogram(binwidth = 0.5, 
                 fill = "#00BFC4", 
                 color = "black", 
                 alpha = 0.7) +
  labs(title = "Histogram of Hb (Hemoglobin) Levels",
       x = "Hemoglobin (g/dL)", 
       y = "Frequency") +
  theme_minimal()

# Line plot of BMI and Hb with regression line
ggplot(data, aes(x = `PRL(ng/mL)`, y = Hb)) +
  geom_point(color = "#00BFC4", size = 3) +  # Points for BMI vs Hb
  geom_smooth(method = "lm", color = "blue", se = FALSE) +
  labs(
    title = "Line Plot of PRL(ng/mL) vs Hb (Hemoglobin) Levels",
    x = "PRL(ng/mL)",
    y = "Hemoglobin (g/dL)") +
  theme_minimal()
```

\newpage

# Team Charter

#### When will you meet as a team to work on the project components? Will these meetings be held in person or virtually?
-   We will meet virtually (i.e through Zoom) every 7 days in the initial phase.
    However, near the deadline, we will aim to meet every 2 days and
    give updates on our individual tasks.

#### What is your group policy on missing team meetings (e.g., how much advance notice should be provided)?

-   Given that we are all juggling multiple projects/deadline, we can
    miss a meeting provided we give a 24 hour notice and follow-up with
    what we miss in the meetings and do the work that we are assigned

#### How will your team communicate (email, Slack, text messages)? What is your policy on appropriate response time (within a certain number of hours? Nights/weekends?)?

-   Our main form of communication would be WhatsApp.
-   We are not obligated to respond in the night, however if we are
    required to respond in the day, a half day buffer is acceptable
